### 2022-03-24 paper/paper.Rmd
reorganized

---
title: |
  A new simplified manual tour, with examples in mathematica
type: Short Technical Note
author:
  - name: Alex Aumann
    affil: a
    email: aaum0002@student.monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
  - name: Ursula Laa
    affil: b
    email: ursula.laa@boku.ac.at
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Something here
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

%\doublespacing
# Introduction

<!-- Background -->

# Manual tour {#sec:method}

## Background

## New definition

# Implementation {#sec:implementation}

# Applications {#sec:examples}

# Discussion {#sec:discussion}

# Acknowledgements {-}

The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].

# Supplementary material {-}

The source material and animated gifs for this paper are available at

### 2022-03-24 paper/paper.tex
reorganized

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}

\begin{document}

\articletype{Short Technical Note}

\title{A new simplified manual tour, with examples in mathematica}

\author{\name{Alex Aumann$^{a}$, German Valencia$^{a}$, Ursula
Laa$^{b}$, Dianne Cook$^{c}$}
\affil{$^{a}$School of Physics and Astronomy, Monash
University; $^{b}$Institute of Statistics, University of Natural
Resources and Life Sciences, Vienna; $^{b}$Department of Econometrics
and Business Statistics, Monash University}
}

\thanks{CONTACT Alex
Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, German
Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}, Ursula
Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}}

\maketitle

\begin{abstract}
Something here
\end{abstract}

\begin{keywords}
data visualisation; grand tour; statistical computing; statistical
graphics; multivariate data; dynamic graphics
\end{keywords}

\%\doublespacing \# Introduction

\hypertarget{sec:method}{%
\section{Manual tour}\label{sec:method}}

\hypertarget{background}{%
\subsection{Background}\label{background}}

\hypertarget{new-definition}{%
\subsection{New definition}\label{new-definition}}

\hypertarget{sec:implementation}{%
\section{Implementation}\label{sec:implementation}}

\hypertarget{sec:examples}{%
\section{Applications}\label{sec:examples}}

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

The authors gratefully acknowledge the support of the Australian
Research Council. The paper was written in \texttt{rmarkdown}
\citep{rmarkdown} using \texttt{knitr} \citep{knitr}.

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

The source material and animated gifs for this paper are available at

\bibliographystyle{tfcad}
\bibliography{biblio.bib}

\end{document}

### 2022-03-29 manual.R
code to make manual tour plots

@@ -1,0 +1,51 @@
+# Load libraries
+library(tourr)
+library(plotrix)
+
+# Code to re-produce mathematica manual control
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+
+# Plot basis
+png("mtour_plots/mtour0.png", width=250, height=250)
+plot_basis(A)
+dev.off()
+
+# Now set a new position
+vchange <- 3
+eps <- 0.01
+#A[vchange,] <- A[vchange,]*(1-eps)
+#A <- tourr::orthonormalise(A)
+#plot_basis(A)
+
+# Iterate and save images
+for (i in 1:20) {
+  flname <- paste0("mtour_plots/mtour", i, ".png")
+  A[vchange,] <- A[vchange,]*(1-i*eps)
+  A <- tourr::orthonormalise(A)
+  png(flname, width=250, height=250)
+  plot_basis(A)
+  dev.off()
+}
+
+# Utility function to plot a basis
+plot_basis <- function(A) {
+  # draw circle
+  plot.new()
+  par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0))
+  plot(c(-1.1, 1.1), c(-1.1, 1.1), type="n", xlab="", ylab="")
+  plotrix::draw.circle(0,0,1)
+
+  # label axes
+  text(A[,1], A[,2], rownames(A), col="grey50")
+
+  # draw axes
+  A <- cbind(A, xctr=rep(0, nrow(A)), yctr=rep(0, nrow(A)))
+  segments(A[,3], A[,4], A[,1], A[,2], lwd=3)
+}

### 2022-03-29 paper/paper.Rmd

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE)
+knitr::opts_chunk$set(
+  echo = FALSE, 
+  message = FALSE,
+  warning = FALSE)
+```
+
+```{r libraries}
+# Load libraries
+library(tourr)
+library(plotrix)
 ```
@@
 # Manual tour {#sec:method}
+
+```{r manualsequence, outwidth="80%", width=12, height=8, fig.cap="Sequence of projections where V3 contribution is changed."}
+source("../code/manual.R")
+
+# Base plot set up
+plot.new()
+par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0),
+    mfrow=c(3,4))
+
+# Create manual tour example
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+plot_basis(A)
+
+# Now set a new position
+vchange <- 3
+eps <- 0.02
+
+# Iterate and save images
+for (i in 1:11) {
+  flname <- paste0("mtour_plots/mtour", i, ".png")
+  A[vchange,] <- A[vchange,]*(1-i*eps)
+  A <- tourr::orthonormalise(A)
+  #png(flname, width=250, height=250)
+  plot_basis(A)
+  #dev.off()
+}
+```
 

### 2022-04-01 paper/paper.Rmd
started intro

@@
 
-%\doublespacing
+<!-- \doublespacing-->
 # Introduction
@@
 <!-- Background -->
+From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space.
+These are dynamic data visualizations methods and are collected under the term *tours*.
+Tours involve views of high-dimensional ($p$) data in low-dimensional ($d$) projections.
+In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
+Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
+PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
+It is impractical to impossible to steer through all possible projections, unlike Asimov's tours which allows one to quickly see many, many different projections.
+After Asimov there have been many, many tour developments, as summarized in @lee2021.
+
+<!-- Manual tours -->
+One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour.
+@cook_manual_1997 describes controls for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into or out of or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
+@spyrison_spinifex_2020 refines this algorithm and implements them to generate animations.
+
+<!-- Purpose -->
+Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection.
+There are many places where it is useful.
+In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering.
+For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
+These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
 

### 2022-04-01 paper/paper.tex
started intro

@@
 
-\%\doublespacing \# Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+From a statistical perspective it is rare to have data that are strictly
+3D, and so unlike in most computer graphics applications, the more
+useful methods for data analysis show projections from an arbitrary
+dimensional space. These are dynamic data visualizations methods and are
+collected under the term \emph{tours}. Tours involve views of
+high-dimensional (\(p\)) data in low-dimensional (\(d\)) projections. In
+his original paper on the grand tour, \citet{As85} provided several
+algorithms for tour paths that could theoretically show the viewer the
+data \emph{from all sides}. Prior to Asimov's work, there were numerous
+preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
+user-controlled rotations on coordinate axes, allowing one to manually
+tour through low-dimensional projections. It is impractical to
+impossible to steer through all possible projections, unlike Asimov's
+tours which allows one to quickly see many, many different projections.
+After Asimov there have been many, many tour developments, as summarized
+in \citet{lee2021}.
+
+One such direction of work develops the ideas from PRIM-9, to provide
+manual control of a tour. \citet{cook_manual_1997} describes controls
+for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing
+the user to select any variable axis, and rotate it into or out of or
+around the projection through horizontal, vertical, oblique, radial or
+angular changes in value. \citet{spyrison_spinifex_2020} refines this
+algorithm and implements them to generate animations.
+
+Manual controls are especially useful for assessing sensitivity of
+structure to particular elements of the projection. There are many
+places where it is useful. In exploratory data analysis, where one sees
+clusters in a projection, can some variables be removed from the
+projection without affecting the clustering. For interpreting models,
+one can reduce or increase a variable's contribution to examine the
+variable importance. These controls can also be used to interactively
+generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
+\citep{XXX}.
 
@@
 \section{Manual tour}\label{sec:method}}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/manualsequence-1.pdf}
+\caption{Sequence of projections where V3 contribution is changed.}
+\end{figure}
 

### 2022-04-01 paper/paper.Rmd
colour of selected variable

@@
 These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
+Having the user interact with a projection is extremely valuable for understanding high-dimensional data.
+However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, (2) extending to higher dimensional control is difficult.
+
+Through experiments with the relatively new interactive graphics capabilities in mathematica(?), we have realized that there is a simpler approach, which is more direct, and extensible for generating user interaction.
+This paper explains this, and is organized as follows.
+The next section describes the new algorithm for manual control.
+This is followed by details on implementation.
+The applications section illustrate how these can be used.
 
 # Manual tour {#sec:method}
+
+## Background
+
+Describe the manip space, out of manual tour
+
+## New definition
+
+1. Provide a k-D projection
+2. Select variable to control
+3. Detect or provide direction of change
+4. Update projection, and orthonormalise
+
+Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+
+Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?
 
@@
 A <- tourr::orthonormalise(A)
-plot_basis(A)
 
@@
 eps <- 0.02
+plot_basis(A, vchange)
 
@@
   #png(flname, width=250, height=250)
-  plot_basis(A)
+  plot_basis(A, vchange)
   #dev.off()
@@
 ```
-
-## Background
-
-## New definition
 

### 2022-04-01 paper/paper.tex
colour of selected variable

@@
 generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
-\citep{XXX}.
+\citep{XXX}. Having the user interact with a projection is extremely
+valuable for understanding high-dimensional data. However, these
+algorithms have two problems: (1) the pre-processing of creating a
+manipulation space overly complicates the algorithm, (2) extending to
+higher dimensional control is difficult.
+
+Through experiments with the relatively new interactive graphics
+capabilities in mathematica(?), we have realized that there is a simpler
+approach, which is more direct, and extensible for generating user
+interaction. This paper explains this, and is organized as follows. The
+next section describes the new algorithm for manual control. This is
+followed by details on implementation. The applications section
+illustrate how these can be used.
 
@@
 \section{Manual tour}\label{sec:method}}
+
+\hypertarget{background}{%
+\subsection{Background}\label{background}}
+
+Describe the manip space, out of manual tour
+
+\hypertarget{new-definition}{%
+\subsection{New definition}\label{new-definition}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Provide a k-D projection
+\item
+  Select variable to control
+\item
+  Detect or provide direction of change
+\item
+  Update projection, and orthonormalise
+\end{enumerate}
+
+Need to think about how the values for selected variable are made exact,
+that is constrained orthonormalisation.
+
+Need to think about checks, and error catching, maybe in implementation
+section. If selected variable has 0 coefficient, will that generate an
+orthonormalisation error?
 
@@
 \end{figure}
-
-\hypertarget{background}{%
-\subsection{Background}\label{background}}
-
-\hypertarget{new-definition}{%
-\subsection{New definition}\label{new-definition}}
 

### 2022-04-01 code/manual.R

@@ -12,7 +12,8 @@
 
 
 # Utility function to plot a basis
-plot_basis <- function(A) {
+# Need to generalise to arbitrary pxd
+plot_basis <- function(A, v=NULL) {
   # Set plot canvas
   plot(c(-1.1, 1.1), c(-1.1, 1.1),
        type="n", xlab="", ylab="", axes=FALSE)
@@ -22,8 +23,15 @@ plot_basis <- function(A) {
 
   # label axes
   text(A[,1], A[,2], rownames(A), col="grey50")
+  if (!is.null(v))
+    text(A[v,1], A[v,2], rownames(A), col="orange")
+
 
   # draw axes
   A <- cbind(A, xctr=rep(0, nrow(A)), yctr=rep(0, nrow(A)))
   segments(A[,3], A[,4], A[,1], A[,2], lwd=3)
+  if (!is.null(v))
+    segments(A[v,3], A[v,4], A[v,1], A[v,2],
+             col="orange", lwd=3)
+
 }

### 2022-04-05 paper/paper.Rmd
done with writing out the algorithms

@@
   \usepackage{setspace}
+  \usepackage{graphicx}
 output: rticles::tf_article
@@
 
+An orthonormal basis ($A_{p\times d}$) and a variable id ($m \in \{1, ..., p\}$) to control are provided to initialise a manual tour.
+A method to update the values of the component of the controlled variable $V_m$ is needed.
+
 ## Background
 
-Describe the manip space, out of manual tour
+In the original work, the method for updating component values, for a 2D projection, was built trackball controls in 3D.
+A 3D manipulation space is created, as illustrated in Figure \ref{manipspace}, where the controlled variable has full range of motion from -1 to 1.
+Movements of a cursor are recorded and converted into changes in the values of $V_m$ to change it's values in the displayed 2D projection.
+Movement could also be constrained to be only in horizontal, vertical, radial or angular motions.
 
-## New definition
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
+\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
+(Figure 3 from Cook and Buja (1997).)}
+\label{manipspace}
+\end{figure*}
 
-1. Provide a k-D projection
-2. Select variable to control
-3. Detect or provide direction of change
-4. Update projection, and orthonormalise
+## New simpler definition
 
-Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+The new approach emerged from experiments in mathematica.
+The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
+The updated matrix is then orthonormalised.
 
-Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?
+### Algorithm 1
 
-```{r manualsequence, outwidth="80%", width=12, height=8, fig.cap="Sequence of projections where V3 contribution is changed."}
-source("../code/manual.R")
+1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
+2.
+Change values in row $m$, giving $A^*$.
+A large change in these values would correspond to making a large jump from the current projection.
+Small changes would correspond to tracking a cursor, making small jumps from the current projection.
+3. Orthonormalise $A^*$, using Gram-Schmidt. For $d=2$, and $A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]$, the steps are:
+
+    i. Normalise ${\boldmath a}_{.1}$, and ${\boldmath a}_{.2}$.
+    ii. ${\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}$.
+    iii. Normalise ${\boldmath a}^*_{.2}$.
+
+This algorithm will produce the changes to a projection as illustrated in Figure \ref{fig:manualsequence} (top row).
+The controlled variable, $V_m$, corresponds to the black line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (orange dot).
+Changes in the other components happen as a result of the orthonormalisation, but are uncontrolled.
+
+### Algorithm 2
+
+The problem with Algorithm 1 is that the precise values for $V_m$ cannot be specified because the orthonormalisation wil change them.
+This modification will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)).
+The algorithm is as follows:
+
+1. Provide $A$, and $m$.
+2. Change values in row $m$, giving $A^*$.
+3. Store row $m$ separately, and zero the values of row $m$ in $A^*$, giving $A^{*0}$. 
+4. Orthonormalise $A^{*0}$, using Gram-Schmidt.
+5. Replace row $m$ with the original values, giving $A^{**}$. 
+6. For $d=2$, adjust the values of ${\boldmath a}^{**}_{.2}$ using
+
+$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m$$.
+
+which ensures that
+
+$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0$$.
+
+If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
+If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalised after each adjustment.
+
+<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+
+Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
+
+```{r manualsequence, outwidth="80%", width=12, height=5, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+The dot (orange) indicates the chosen values for the controlled variable.
+For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation."}
+source("../code/plot_basis.R")
+source("../code/linear_alg.R")
 
@@
       omi=c(0,0,0,0), mar=c(0,0,0,0),
-    mfrow=c(3,4))
+    mfrow=c(2,4))
 
@@
 vchange <- 3
-eps <- 0.02
-plot_basis(A, vchange)
+eps <- 0.1
+plot_basis(A, vchange, vcol="black")
 
 # Iterate and save images
-for (i in 1:11) {
-  flname <- paste0("mtour_plots/mtour", i, ".png")
-  A[vchange,] <- A[vchange,]*(1-i*eps)
-  A <- tourr::orthonormalise(A)
-  #png(flname, width=250, height=250)
-  plot_basis(A, vchange)
-  #dev.off()
+# Unconstrained orthonormalisation
+Anew <- A
+for (i in 1:3) {
+  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+  exact <- Anew[vchange,]
+  Anew <- tourr::orthonormalise(Anew)
+  plot_basis(Anew, vchange, vcol="black")
+  points(exact[1], exact[2], pch=16, col="orange")
+}
+
+# Iterate and save images
+# Constrained orthonormalisation
+Anew <- A
+plot_basis(A, vchange, vcol="black")
+for (i in 1:3) {
+  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+  exact <- Anew[vchange,]
+  Anew <- orthonormalise_frozen(Anew, vchange)
+  plot_basis(Anew, vchange, vcol="black")
+  points(exact[1], exact[2], pch=16, col="orange")
 }

### 2022-04-05 paper/paper.tex
done with writing out the algorithms

@@
 \usepackage{setspace}
+\usepackage{graphicx}
 
@@
 
+An orthonormal basis (\(A_{p\times d}\)) and a variable id
+(\(m \in \{1, ..., p\}\)) to control are provided to initialise a manual
+tour. A method to update the values of the component of the controlled
+variable \(V_m\) is needed.
+
 \hypertarget{background}{%
@@
 
-Describe the manip space, out of manual tour
+In the original work, the method for updating component values, for a 2D
+projection, was built trackball controls in 3D. A 3D manipulation space
+is created, as illustrated in Figure \ref{manipspace}, where the
+controlled variable has full range of motion from -1 to 1. Movements of
+a cursor are recorded and converted into changes in the values of
+\(V_m\) to change it's values in the displayed 2D projection. Movement
+could also be constrained to be only in horizontal, vertical, radial or
+angular motions.
 
-\hypertarget{new-definition}{%
-\subsection{New definition}\label{new-definition}}
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
+\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
+(Figure 3 from Cook and Buja (1997).)}
+\label{manipspace}
+\end{figure*}
+
+\hypertarget{new-simpler-definition}{%
+\subsection{New simpler definition}\label{new-simpler-definition}}
+
+The new approach emerged from experiments in mathematica. The components
+corresponding to \(V_m\) are directly controlled by cursor movement,
+which updates row \(m\) of \(A\). The updated matrix is then
+orthonormalised.
+
+\hypertarget{algorithm-1}{%
+\subsubsection{Algorithm 1}\label{algorithm-1}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Provide \(A\), and \(m\). (Note that \(m\) could also be automatically
+  chosen as the component that is closest to the cursor position.)
+\item
+  Change values in row \(m\), giving \(A^*\). A large change in these
+  values would correspond to making a large jump from the current
+  projection. Small changes would correspond to tracking a cursor,
+  making small jumps from the current projection.
+\item
+  Orthonormalise \(A^*\), using Gram-Schmidt. For \(d=2\), and
+  \(A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]\), the
+  steps are:
+
+  \begin{enumerate}
+  \def\labelenumii{\roman{enumii}.}
+  \tightlist
+  \item
+    Normalise \({\boldmath a}_{.1}\), and \({\boldmath a}_{.2}\).
+  \item
+    \({\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}\).
+  \item
+    Normalise \({\boldmath a}^*_{.2}\).
+  \end{enumerate}
+\end{enumerate}
+
+This algorithm will produce the changes to a projection as illustrated
+in Figure \ref{fig:manualsequence} (top row). The controlled variable,
+\(V_m\), corresponds to the black line, and sequential changes to row
+\(m\) of \(A\) can be seen to roughly follow a specified position
+(orange dot). Changes in the other components happen as a result of the
+orthonormalisation, but are uncontrolled.
+
+\hypertarget{algorithm-2}{%
+\subsubsection{Algorithm 2}\label{algorithm-2}}
+
+The problem with Algorithm 1 is that the precise values for \(V_m\)
+cannot be specified because the orthonormalisation wil change them. This
+modification will maintain the components of \(V_m\) precisely (Figure
+\ref{fig:manualsequence} (bottom row)). The algorithm is as follows:
 
@@
 \item
-  Provide a k-D projection
+  Provide \(A\), and \(m\).
 \item
-  Select variable to control
+  Change values in row \(m\), giving \(A^*\).
 \item
-  Detect or provide direction of change
+  Store row \(m\) separately, and zero the values of row \(m\) in
+  \(A^*\), giving \(A^{*0}\).
 \item
-  Update projection, and orthonormalise
+  Orthonormalise \(A^{*0}\), using Gram-Schmidt.
+\item
+  Replace row \(m\) with the original values, giving \(A^{**}\).
+\item
+  For \(d=2\), adjust the values of \({\boldmath a}^{**}_{.2}\) using
 \end{enumerate}
 
-Need to think about how the values for selected variable are made exact,
-that is constrained orthonormalisation.
+\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m\].
 
-Need to think about checks, and error catching, maybe in implementation
-section. If selected variable has 0 coefficient, will that generate an
-orthonormalisation error?
+which ensures that
+
+\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0\].
+
+If \(d>2\) the process would be sequentially repeated in the same manner
+that Gram-Schmidt is applied sequentially to orthormalise the columns of
+a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
+vector would simply need to be normalised after each adjustment.
 
@@
 \includegraphics{paper_files/figure-latex/manualsequence-1.pdf}
-\caption{Sequence of projections where V3 contribution is changed.}
+\caption{Sequence of projections where contribution of one variable is
+controlled (black) is changed: (top) unconstrained orthonormalisation,
+(bottom) constrained as specified. The dot (orange) indicates the chosen
+values for the controlled variable. For the constrained
+orthonormalisation it can be seen to precisely match the axis, but not
+so for the unconstrained orthonormalisation.}
 \end{figure}

### 2022-04-05 code/linear_alg.R

@@ -1,0 +1,31 @@
+# Orthonormalise with frozen tour, and post-process to
+# constrain second vector with frozen components
+orthonormalise_frozen <- function(x, v) {
+  # x is a pxd basis, not orthonormal
+  # v is the component that is frozen/fixed
+
+  # Store frozen/fixed components
+  p <- nrow(x)
+  d <- ncol(x)
+  frozen <- matrix(NA, nrow = p, ncol = d)
+  frozen[v,] <- x[v,]
+
+  # Remove from basis
+  x_froz <- tourr::freeze(x, frozen)
+
+  # Orthonormalise non-frozen matrix
+  x_froz <- tourr::orthonormalise(x_froz)
+
+  # Put back frozen components, and adjust second vector
+  # to maintain orthogonality
+  x_thaw <- thaw(x_froz, frozen)
+  x_out <- x_thaw
+  for (i in 1:p) {
+    if (i != v) {
+      x_out[i, 2] <- x_out[i, 2] + x_out[v, 1]*x_out[v, 2]/(p-1)
+    }
+  }
+
+  # Return changed matrix
+  x_out
+}

### 2022-04-05 code/plot_basis.R

@@ -1,0 +1,33 @@
+# Utility function to plot a basis
+# Need to generalise to arbitrary d
+plot_basis <- function(A, v=NULL, vcol = "orange", labels=FALSE) {
+  # A is an orthonormal basis
+  # v needs to be an integer between 1, and nrow(A)
+
+  # Set plot canvas
+  plot(c(-1.1, 1.1), c(-1.1, 1.1),
+       type="n", xlab="", ylab="", axes=FALSE)
+
+  # draw circle
+  plotrix::draw.circle(0,0,1)
+
+  # label axes
+  if (labels) {
+    if (is.null(v)) {
+      text(A[,1], A[,2], rownames(A), col="grey50")
+    } else {
+      text(A[-v,1], A[-v,2], rownames(A)[-v], col="grey80")
+      text(A[v,1], A[v,2], rownames(A)[v], col=vcol)
+    }
+  }
+
+  # draw axes
+  Aaxes <- cbind(A, xctr=rep(0, nrow(A)), yctr=rep(0, nrow(A)))
+  if (is.null(v)) {
+    segments(Aaxes[,3], Aaxes[,4], Aaxes[,1], Aaxes[,2], lwd=3)
+  } else {
+    segments(Aaxes[-v,3], Aaxes[-v,4], Aaxes[-v,1], Aaxes[-v,2], lwd=3, col="grey80")
+    segments(Aaxes[v,3], Aaxes[v,4], Aaxes[v,1], Aaxes[v,2],
+             col=vcol, lwd=3)
+  }
+}

### 2022-04-05 paper/paper.Rmd
title change

@@
 title: |
-  A new simplified manual tour, with examples in mathematica
+  A new simplified manual tour, with examples in mathematica and R
 type: Short Technical Note

### 2022-04-06 paper/paper.Rmd
Sketching a different algorithm (not yet tested)

@@
 
+### Potential Algorithm 3
+
+For now just sketching the idea:
+
+- first step is to click on the axis display to change the contribution of one variable $m$
+- capture that position and replace the corresponding row in the projection matrix
+- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
+- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
+- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
+- rotated back to the original xy basis to update the plots
+
 # Implementation {#sec:implementation}

### 2022-04-12 paper/paper.Rmd
In-plane rotation idea seems to be working

@@
 
+```{r }
+source("../code/plot_basis.R")
+source("../code/linear_alg.R")
+
+# Base plot set up
+plot.new()
+par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0),
+    mfrow=c(1,4))
+
+# Create manual tour example
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+
+# Now set a new position
+vchange <- 3
+eps <- 0.1
+plot_basis(A, vchange, vcol="black")
+
+# j is the component that should be kept fixed
+orthonormalize_rotated <- function(A, j){
+  m1 <- A[j, 1]
+  m2 <- A[j, 2]
+  c <- 1 / (sqrt(1 + (m2/m1)^2))
+  s <- (m2/m1) / (sqrt(1 + (m2/m1)^2))
+  rot <- matrix(c(c, -s, s, c), nrow = 2)
+  rot_inv <- matrix(c(c, s, -s, c), nrow = 2)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot %*% A[i,]
+  }
+  A[,1] <- normalize_w_fixed(A[,1], j)
+  A <- tourr::orthonormalise(A)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot_inv %*% A[i,]
+    }
+  A
+}
+
+# normalize a vector while keeping component j fixed
+normalize_w_fixed <- function(v, j){
+  c <- v[j]
+  k <- sqrt((1-c^2) / sum(v[-j]^2))
+  v[-j] <- k * v[-j]
+  v
+}
+
+# Iterate and save images
+# Unconstrained orthonormalisation
+Anew <- A
+for (i in 1:3) {
+  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+  exact <- Anew[vchange,]
+  Anew <- orthonormalize_rotated(Anew, vchange)
+  plot_basis(Anew, vchange, vcol="black")
+  points(exact[1], exact[2], pch=16, col="orange")
+}
+
+```
+
 # Implementation {#sec:implementation}

### 2022-04-12 paper/paper.tex
title change
In-plane rotation idea seems to be working

@@
 
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% see https://stackoverflow.com/a/47122900
+
+% Pandoc citation processing
 
@@
 
-\title{A new simplified manual tour, with examples in mathematica}
+\title{A new simplified manual tour, with examples in mathematica and R}
 
@@
 
+\hypertarget{potential-algorithm-3}{%
+\subsubsection{Potential Algorithm 3}\label{potential-algorithm-3}}
+
+For now just sketching the idea:
+
+\begin{itemize}
+\tightlist
+\item
+  first step is to click on the axis display to change the contribution
+  of one variable \(m\)
+\item
+  capture that position and replace the corresponding row in the
+  projection matrix
+\item
+  for 2D projection this gives components \(m_1\) along ``x'' direction
+  and \(m_2\) along \(y\) direction
+\item
+  next step: rotate basis such that direction of row m now corresponds
+  to the first basis vector, this means we apply 2x2 rotation matrix to
+  each row of the projection matrix, where the rotation angle is
+  \(tan(\theta) = m_2/m_1\) (that matrix can be written just in terms of
+  that ratio, \(x= m_2/m_1\), when \(m_1<0\) I think we just need to
+  translate \(\theta \rightarrow \theta + \pi\))
+\item
+  take the rotated basis, apply Gram-Schmidt, now the direction of \(m\)
+  will not change (but the length will change during normalization)
+\item
+  rotated back to the original xy basis to update the plots
+\end{itemize}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+
 \hypertarget{sec:implementation}{%

### 2022-04-19 paper/paper.Rmd
Comment added
added section

@@
   \usepackage{graphicx}
+  \usepackage{nicematrix}
+  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
 output: rticles::tf_article
@@
 
+### Algorithm 3
+
+For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
+Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
+Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+One possible algorithm to achieve this is
+\begin{enumerate}
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
+\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
+\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalised), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\end{enumerate}
+\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\end{enumerate}
+The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
+
+\begin{align}
+A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
+e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
+e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
+\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
+\end{align}
+
 ```{r manualsequence, outwidth="80%", width=12, height=5, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
@@
 - rotated back to the original xy basis to update the plots
+
+XXX need to check calculation of angle when $m_1 < 0$
 

### 2022-05-01 paper/paper.Rmd
Update paper.Rmd

@@
 
+The following functions are implemented in the Mathematica package {\tt mmtour.wl}
+
+\begin{itemize}
+
+\item ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
+colour(=Automatic)] with arguments
+
+\begin{itemize}
+\item data: A list of data matrices (they must all have the same dimensionality)
+\item projmat: A projection matrix which describes the projection plane. This is the 
+initial projection and does not neet by orthonormal.
+\item xRange: A list which details the range of the horizontal axis.
+\item yRange: A list which details the range of the vertical axis.
+\item colour: A list of graphics directives to be applied to each data \
+matrix.
+\end{itemize}
+
+There are problems with the default values for x and y ranges and colour. It would be 
+better to fix the ranges and colour based on the data as well as the origin of axes 
+and the size of the points.
+
+The format for the printed projection matrix can be improved also. Maybe fix the 
+number of digits and the position of the label.
+
+\item SlicePlot[data, projmat, centrePoint, height, heightRange, 
+ptSize1(=0.005), ptSize2(=0.004), minDist(=0)] with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
+
+\begin{itemize}
+\item data: A data matrix
+\item projmat: The inital projection matrix
+\item centrepoint: The initial position of the slice
+\item height: The initial height of the slice
+\item heightRange: The range of heights
+\item ptSize1: The size of the points that exist within the slice
+\item ptSize2: The size of the points that exist outside of the slice
+\end{itemize}
+
+This function requires some auxiliary functions. It does not accept more than one data set.
+
+\end{itemize}
+
 # Applications {#sec:examples}

### 2022-05-13 paper/paper.Rmd
put info in tex file

@@
 
+It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?
+
 The format for the printed projection matrix can be improved also. Maybe fix the 
@@
 
+It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?
+
+It may be good to fix the size of the points, the centre of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+
+\item SliceDynamic[data, projmat, centrePoint, height, heightRange, 
+ ptSize1 : 0.005, ptSize2 : 0.004] with arguments
+
+\begin{itemize}
+\item data: A data matrix with an extra groupling or clustering column (numerical) 
+\item projmat: The inital projection matrix
+\item centrepoint: The initial position of the slice
+\item height: The initial height of the slice
+\item heightRange: The range of heights
+\item ptSize1: The size of the points that exist within the slice
+\item ptSize2: The size of the points that exist outside of the slice
+\end{itemize}
+
+Does this need to be a separate function from SlicePlot?
+
+It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?
+
+It may be good to fix the size of the points, the centre of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+
+I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
+
+\item Projected2DSliderPlot[]
+
+This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
+
 \end{itemize}
@@
 # Discussion {#sec:discussion}
+
+- As was said before, we need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
+- Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.
+- Can we have some heuristic to decide the range for the slice thickness slider?
+In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
+My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.
+- For the olives example I think the relation between x2 and the blue region could be interesting.
+- pdfsense data: can we use random samples instead of the first 10k rows?
+Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).
+
+I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?
+
+Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:
+
+- In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
+Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
+We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
+Much less clear, maybe some indication of smaller x2 and larger x3 for this.
+- The second example is too busy for manual exploration I would say (too many axes to move around separately).
+- Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
+Now interpretation becomes tricky.
+Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
+For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.
+- For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)
 

### 2022-06-14 paper/paper.Rmd
fixed lists into Rmarkdown

@@
 
-The following functions are implemented in the Mathematica package {\tt mmtour.wl}
+The following functions are implemented in the Mathematica package `mmtour.wl`:
 
-\begin{itemize}
+- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
+colour(=Automatic)]` with arguments
 
-\item ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
-colour(=Automatic)] with arguments
-
-\begin{itemize}
-\item data: A list of data matrices (they must all have the same dimensionality)
-\item projmat: A projection matrix which describes the projection plane. This is the 
+    - `data`: A list of data matrices (they must all have the same dimensionality)
+    - ``projmat`: A projection matrix which describes the projection plane. This is the 
 initial projection and does not neet by orthonormal.
-\item xRange: A list which details the range of the horizontal axis.
-\item yRange: A list which details the range of the vertical axis.
-\item colour: A list of graphics directives to be applied to each data \
+    - ``xRange`: A list which details the range of the horizontal axis.
+    - ``yRange`: A list which details the range of the vertical axis.
+    - ``colour`: A list of graphics directives to be applied to each data \
 matrix.
-\end{itemize}
 
@@
 
-\item SlicePlot[data, projmat, centrePoint, height, heightRange, 
-ptSize1(=0.005), ptSize2(=0.004), minDist(=0)] with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
+- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
 
-\begin{itemize}
-\item data: A data matrix
-\item projmat: The inital projection matrix
-\item centrepoint: The initial position of the slice
-\item height: The initial height of the slice
-\item heightRange: The range of heights
-\item ptSize1: The size of the points that exist within the slice
-\item ptSize2: The size of the points that exist outside of the slice
-\end{itemize}
+    - `data`: A data matrix
+    - `projmat`: The inital projection matrix
+    - `centrepoint`: The initial position of the slice
+    - `height`: The initial height of the slice
+    - `heightRange`: The range of heights
+    - `ptSize1`: The size of the points that exist within the slice
+    - `ptSize2`: The size of the points that exist outside of the slice
 
@@
 
-\item SliceDynamic[data, projmat, centrePoint, height, heightRange, 
- ptSize1 : 0.005, ptSize2 : 0.004] with arguments
+- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments
 
-\begin{itemize}
-\item data: A data matrix with an extra groupling or clustering column (numerical) 
-\item projmat: The inital projection matrix
-\item centrepoint: The initial position of the slice
-\item height: The initial height of the slice
-\item heightRange: The range of heights
-\item ptSize1: The size of the points that exist within the slice
-\item ptSize2: The size of the points that exist outside of the slice
-\end{itemize}
+- `data`: A data matrix with an extra groupling or clustering column (numerical) 
+- `projmat`: The inital projection matrix
+- `centrepoint`: The initial position of the slice
+- `height`: The initial height of the slice
+- `heightRange`: The range of heights
+- `ptSize1`: The size of the points that exist within the slice
+- `ptSize2`: The size of the points that exist outside of the slice
 
@@
 
-\item Projected2DSliderPlot[]
+- `Projected2DSliderPlot[]`
 
 This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
-
-\end{itemize}
 

### 2022-06-14 paper/paper.tex
fixed lists into Rmarkdown

@@
 
-% see https://stackoverflow.com/a/47122900
-
-% Pandoc citation processing
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \usepackage{graphicx}
+\usepackage{nicematrix}
+\NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
 
@@
 
+\hypertarget{algorithm-3}{%
+\subsubsection{Algorithm 3}\label{algorithm-3}}
+
+For \(d=2\) projections, the projection matrix is the sub-matrix of
+\(A\) formed by its first two columns. Whereas orthonormality of the
+basis for the p-dimensional space is given by
+\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}\), orthonormality of the
+projection matrix is expressed as
+\(P_i\cdot P_j=\delta_{ij}, {i,j=1,2}\). Movement of the cursor takes
+the two components \({x_{m1},x_{m2}}\) into a selected new value
+\({a,b}\). Although the motion is constrained by \(a^2+b^2\leq 1\), this
+is not sufficient to guarantee orthonormality of the new projection
+matrix. One possible algorithm to achieve this is
+
+\begin{enumerate}
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
+\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
+\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalised), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\end{enumerate}
+\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\end{enumerate}
+
+The random choice for the components of \(A\) not in the projection
+matrix allows the exploration of dimensions perpendicular to the
+projection plane.
+
+\begin{align}
+A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
+e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
+e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
+\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
+\end{align}
+
 \begin{figure}
@@
 
+XXX need to check calculation of angle when \(m_1 < 0\)
+
 \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
@@
 \section{Implementation}\label{sec:implementation}}
+
+The following functions are implemented in the Mathematica package
+\texttt{mmtour.wl}:
+
+\begin{itemize}
+\item
+  \texttt{ProjectionPlot{[}data,\ projmat,\ xRange(=Automatic),\ yRange(=Automatic),\ \ colour(=Automatic){]}}
+  with arguments
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{data}: A list of data matrices (they must all have the same
+    dimensionality)
+  \item
+    `\texttt{projmat}: A projection matrix which describes the
+    projection plane. This is the initial projection and does not neet
+    by orthonormal.
+  \item
+    `\texttt{xRange}: A list which details the range of the horizontal
+    axis.
+  \item
+    `\texttt{yRange}: A list which details the range of the vertical
+    axis.
+  \item
+    `\texttt{colour}: A list of graphics directives to be applied to
+    each data\\
+    matrix.
+  \end{itemize}
+\end{itemize}
+
+There are problems with the default values for x and y ranges and
+colour. It would be better to fix the ranges and colour based on the
+data as well as the origin of axes and the size of the points.
+
+It would be good if the initial projection matrix (and maybe the centre)
+had default values. Perhaps a random matrix with the correct dimensions
+read from the data?
+
+The format for the printed projection matrix can be improved also. Maybe
+fix the number of digits and the position of the label.
+
+\begin{itemize}
+\item
+  \texttt{SlicePlot{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1(=0.005),\ ptSize2(=0.004),\ minDist(=0){]}}
+  with arguments (used to be called SliceDynamic in Alex's first
+  version, the name SliceDynamic was taken by something else below)
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{data}: A data matrix
+  \item
+    \texttt{projmat}: The inital projection matrix
+  \item
+    \texttt{centrepoint}: The initial position of the slice
+  \item
+    \texttt{height}: The initial height of the slice
+  \item
+    \texttt{heightRange}: The range of heights
+  \item
+    \texttt{ptSize1}: The size of the points that exist within the slice
+  \item
+    \texttt{ptSize2}: The size of the points that exist outside of the
+    slice
+  \end{itemize}
+\end{itemize}
+
+This function requires some auxiliary functions. It does not accept more
+than one data set.
+
+It would be good if the initial projection matrix (and maybe the centre)
+had default values. Perhaps a random matrix with the correct dimensions
+read from the data?
+
+It may be good to fix the size of the points, the centre of the axes and
+the scales of the axes rather than allow them to change as the
+projection and slicing changes
+
+\begin{itemize}
+\item
+  \texttt{SliceDynamic{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1\ :\ 0.005,\ ptSize2\ :\ 0.004{]}}
+  with arguments
+\item
+  \texttt{data}: A data matrix with an extra groupling or clustering
+  column (numerical)
+\item
+  \texttt{projmat}: The inital projection matrix
+\item
+  \texttt{centrepoint}: The initial position of the slice
+\item
+  \texttt{height}: The initial height of the slice
+\item
+  \texttt{heightRange}: The range of heights
+\item
+  \texttt{ptSize1}: The size of the points that exist within the slice
+\item
+  \texttt{ptSize2}: The size of the points that exist outside of the
+  slice
+\end{itemize}
+
+Does this need to be a separate function from SlicePlot?
+
+It would be good if the initial projection matrix (and maybe the centre)
+had default values. Perhaps a random matrix with the correct dimensions
+read from the data?
+
+It may be good to fix the size of the points, the centre of the axes and
+the scales of the axes rather than allow them to change as the
+projection and slicing changes
+
+I have found it useful to input different data sets rather than have
+this function sort them automatically using the last column.
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{Projected2DSliderPlot{[}{]}}
+\end{itemize}
+
+This function uses a different display for the projection that I find
+very useful. However as it stands it is specific to one data set. Needs
+to be generalised to arbitrary data if we want to keep it.
 
@@
 \section{Discussion}\label{sec:discussion}}
+
+\begin{itemize}
+\tightlist
+\item
+  As was said before, we need to make sure the drawing range is fixed
+  (no resizing of the axis when changing projection or slicing).
+\item
+  Should also make sure the size of the points is fixed (maybe this will
+  be automatic if the drawing range is fixed?), bigger points would be
+  preferred, currently the size is sometimes really tiny and it becomes
+  difficult to see anything.
+\item
+  Can we have some heuristic to decide the range for the slice thickness
+  slider? In the second graph there is a tiny part of the slider where
+  the changes are interesting, difficult to navigate since this makes it
+  super sensitive to any small movement, while most of the options are
+  not interesting. My suggestion would be to keep (as I understand is
+  the case now) the maximum thickness (all points in) as the upper
+  bound, but instead of zero use a thickness below which slices tend to
+  be empty as the lower bound.
+\item
+  For the olives example I think the relation between x2 and the blue
+  region could be interesting.
+\item
+  pdfsense data: can we use random samples instead of the first 10k
+  rows? Then we should still be able to see the bigger picture (with
+  picking the first ones we get some artificial structure from how the
+  data was sampled initially).
+\end{itemize}
+
+I was playing around with this today. Looks like the axes are fixed when
+changing slice thickness, but not when rotating?
+
+Anyways, some observations that might be useful, but all of this seems
+difficult for including in a paper:
+
+\begin{itemize}
+\tightlist
+\item
+  In the first example we can show how the yellow group is associated
+  with the first variable (x1, I guess this is C9?) - when slicing
+  through the center it only appears when x1 has a bigger component on
+  the projection, it is associated with low values of x1. Since they are
+  not near the average value of x1 the points do not get captured in a
+  slice where x1 is not important in the projection. We can move the
+  center point to -1 in the first component to then explore how the
+  yellow group relates to the other three variables. Much less clear,
+  maybe some indication of smaller x2 and larger x3 for this.
+\item
+  The second example is too busy for manual exploration I would say (too
+  many axes to move around separately).
+\item
+  Third example could be similar to the first: blue region requires
+  small values of x2, can be understood from the exploration, then move
+  center point to -0.5 in x2 to explore further. Now interpretation
+  becomes tricky. Sometimes blue and yellow seem just randomly
+  everywhere, sometimes blue more contained in an ellipsoid around the
+  center (indicating some correlation between variables). For example
+  try zeroing out components 2 and 3 to see this slice in x1 vs x4.
+\item
+  For the last example x4 seems to be the one to work with. I feel like
+  this is again very similar, maybe better to instead explore other
+  examples (maybe something without slicing?)
+\end{itemize}
 

### 2022-06-14 paper/paper.Rmd
make sure changes are committed

@@
 
-```{r manualsequence, outwidth="80%", width=12, height=5, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
 The dot (orange) indicates the chosen values for the controlled variable.
@@
 
-```{r }
+```{r othermethod, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
 source("../code/plot_basis.R")
@@
     - ``yRange`: A list which details the range of the vertical axis.
-    - ``colour`: A list of graphics directives to be applied to each data \
-matrix.
+    - ``colour`: A list of graphics directives to be applied to each data matrix.
 
@@
 
-It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?
+**It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
 
@@
 
-Does this need to be a separate function from SlicePlot?
+**Does this need to be a separate function from SlicePlot?**
 
@@
 
-I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
+**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.**
 
@@
 This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
+
+**We would need a screenshot (and video) for each of these.**
 

### 2022-06-14 paper/paper.tex
make sure changes are committed

@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/manualsequence-1.pdf}
-\caption{Sequence of projections where contribution of one variable is
-controlled (black) is changed: (top) unconstrained orthonormalisation,
-(bottom) constrained as specified. The dot (orange) indicates the chosen
-values for the controlled variable. For the constrained
-orthonormalisation it can be seen to precisely match the axis, but not
-so for the unconstrained orthonormalisation.}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+The dot (orange) indicates the chosen values for the controlled variable.
+For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation.}\label{fig:manualsequence}
 \end{figure}
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/othermethod-1} \caption{Manual controls with algorithm 2. The precise location of the axis is maintained.}\label{fig:othermethod}
+\end{figure}
 
@@
     `\texttt{colour}: A list of graphics directives to be applied to
-    each data\\
-    matrix.
+    each data matrix.
   \end{itemize}
@@
 
-It would be good if the initial projection matrix (and maybe the centre)
-had default values. Perhaps a random matrix with the correct dimensions
-read from the data?
+\textbf{It would be good if the initial projection matrix (and maybe the
+centre) had default values. Perhaps a random matrix with the correct
+dimensions read from the data?}
 
@@
 
-Does this need to be a separate function from SlicePlot?
+\textbf{Does this need to be a separate function from SlicePlot?}
 
@@
 
-I have found it useful to input different data sets rather than have
-this function sort them automatically using the last column.
+\textbf{I have found it useful to input different data sets rather than
+have this function sort them automatically using the last column.}
 
@@
 to be generalised to arbitrary data if we want to keep it.
+
+\textbf{We would need a screenshot (and video) for each of these.}
 

### 2022-07-05 code/penguins.R
Setting up the penguins example.

@@ -1,0 +1,30 @@
+library(tidyverse)
+library(palmerpenguins)
+library(classifly)
+library(MASS)
+library(tourr)
+library(randomForest)
+
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x)))
+
+
+penguins_lda <- classifly(penguins_data, species ~ . , lda)
+penguins_lda[,1:4] <- apply(penguins_lda[,1:4], 2,  function(x) (x-mean(x))/sd(x))
+lda_samples <- penguins_lda %>%
+  as_tibble() %>%
+  filter(.TYPE == "simulated")
+
+animate_slice(lda_samples[,1:4], col = lda_samples$species, v_rel = 0.01)
+
+set.seed(21)
+penguins_rf <- classifly(penguins_data, species ~ . , randomForest)
+penguins_rf[,1:4] <- apply(penguins_rf[,1:4], 2,  function(x) (x-mean(x))/sd(x))
+rf_samples <- penguins_rf %>%
+  as_tibble() %>%
+  filter(.TYPE == "simulated")
+
+animate_slice(rf_samples[,1:4], col = rf_samples$species, v_rel = 0.01)
+
+

### 2022-07-05 code/penguins.R
Penguin data written.

@@ -18,6 +18,8 @@ lda_samples <- penguins_lda %>%
 
 animate_slice(lda_samples[,1:4], col = lda_samples$species, v_rel = 0.01)
 
+write_csv(penguins_lda[, c(1:4,6)], file = "data/pengins_lda.csv")
+
 set.seed(21)
 penguins_rf <- classifly(penguins_data, species ~ . , randomForest)
 penguins_rf[,1:4] <- apply(penguins_rf[,1:4], 2,  function(x) (x-mean(x))/sd(x))
@@ -27,4 +29,5 @@ rf_samples <- penguins_rf %>%
 
 animate_slice(rf_samples[,1:4], col = rf_samples$species, v_rel = 0.01)
 
+write_csv(penguins_rf[, c(1:4,6)], file = "data/pengins_rf.csv")
 

### 2022-07-05 code/penguins.R
Updated code penguins data

@@ -14,20 +14,22 @@ penguins_lda <- classifly(penguins_data, species ~ . , lda)
 penguins_lda[,1:4] <- apply(penguins_lda[,1:4], 2,  function(x) (x-mean(x))/sd(x))
 lda_samples <- penguins_lda %>%
   as_tibble() %>%
-  filter(.TYPE == "simulated")
+  filter(.TYPE == "simulated") %>%
+  mutate(species_flag = as.numeric(as.factor(species)))
 
 animate_slice(lda_samples[,1:4], col = lda_samples$species, v_rel = 0.01)
 
-write_csv(penguins_lda[, c(1:4,6)], file = "data/pengins_lda.csv")
+write_csv(lda_samples[, c(1:4,6,12)], file = "data/penguins_lda.csv")
 
 set.seed(21)
 penguins_rf <- classifly(penguins_data, species ~ . , randomForest)
 penguins_rf[,1:4] <- apply(penguins_rf[,1:4], 2,  function(x) (x-mean(x))/sd(x))
 rf_samples <- penguins_rf %>%
   as_tibble() %>%
-  filter(.TYPE == "simulated")
+  filter(.TYPE == "simulated") %>%
+  mutate(species_flag = as.numeric(as.factor(species)))
 
 animate_slice(rf_samples[,1:4], col = rf_samples$species, v_rel = 0.01)
 
-write_csv(penguins_rf[, c(1:4,6)], file = "data/pengins_rf.csv")
+write_csv(rf_samples[, c(1:4,6, 12)], file = "data/penguins_rf.csv")
 

### 2022-07-06 paper/paper.tex
trying to map out the paper

@@
 variable importance. These controls can also be used to interactively
-generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
+generate faceted plots \citep{XXX}, or spatiotemporal glyphmaps
 \citep{XXX}. Having the user interact with a projection is extremely
@@
 \hypertarget{sec:method}{%
-\section{Manual tour}\label{sec:method}}
+\section{What is a manual tour}\label{sec:method}}
 
 An orthonormal basis (\(A_{p\times d}\)) and a variable id
-(\(m \in \{1, ..., p\}\)) to control are provided to initialise a manual
+(\(m \in \{1, ..., p\}\)) to control are provided to initialize a manual
 tour. A method to update the values of the component of the controlled
@@
 
-\hypertarget{background}{%
-\subsection{Background}\label{background}}
+\hypertarget{existing-methods}{%
+\subsection{Existing methods}\label{existing-methods}}
 
@@
 
-\hypertarget{new-simpler-definition}{%
-\subsection{New simpler definition}\label{new-simpler-definition}}
+\hypertarget{the-new-simpler-definition}{%
+\subsection{The new simpler
+definition}\label{the-new-simpler-definition}}
 
@@
 The problem with Algorithm 1 is that the precise values for \(V_m\)
-cannot be specified because the orthonormalisation wil change them. This
-modification will maintain the components of \(V_m\) precisely (Figure
-\ref{fig:manualsequence} (bottom row)). The algorithm is as follows:
+cannot be specified because the orthonormalisation will change them.
+This modification will maintain the components of \(V_m\) precisely
+(Figure \ref{fig:manualsequence} (bottom row)). The algorithm is as
+follows:
 
@@
 a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
-vector would simply need to be normalised after each adjustment.
+vector would simply need to be normalized after each adjustment.
 
@@
 \item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalised), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
 \end{enumerate}
@@
 \hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
+\section{Experimenting with new techniques using
+Mathematica}\label{sec:implementation}}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Why is this a good sandbox}
+\item
+  \textbf{Explain the functionality available in the notebooks}
+\end{itemize}
 
@@
     `\texttt{projmat}: A projection matrix which describes the
-    projection plane. This is the initial projection and does not neet
-    by orthonormal.
+    projection plane. This is the initial projection and does not need
+    to be orthonormal (XXX is this right?).
   \item
@@
 \textbf{It would be good if the initial projection matrix (and maybe the
-centre) had default values. Perhaps a random matrix with the correct
-dimensions read from the data?}
+center) had default values. Perhaps a random matrix with the correct
+dimensions read from the data? Fix 1}
 
@@
 \textbf{It would be good if the initial projection matrix (and maybe the
-centre) had default values. Perhaps a random matrix with the correct
+center) had default values. Perhaps a random matrix with the correct
 dimensions read from the data?}
 
-It may be good to fix the size of the points, the centre of the axes and
+It may be good to fix the size of the points, the center of the axes and
 the scales of the axes rather than allow them to change as the
@@
 \item
-  \texttt{data}: A data matrix with an extra groupling or clustering
+  \texttt{data}: A data matrix with an extra grouping or clustering
   column (numerical)
 \item
-  \texttt{projmat}: The inital projection matrix
+  \texttt{projmat}: The initial projection matrix
 \item
@@
 
-\textbf{Does this need to be a separate function from SlicePlot?}
+\textbf{Does this need to be a separate function from SlicePlot? Fix 2}
 
-It would be good if the initial projection matrix (and maybe the centre)
-had default values. Perhaps a random matrix with the correct dimensions
-read from the data?
+\textbf{It would be good if the initial projection matrix (and maybe the
+center) had default values. Perhaps a random matrix with the correct
+dimensions read from the data? Fix 3}
 
-It may be good to fix the size of the points, the centre of the axes and
-the scales of the axes rather than allow them to change as the
-projection and slicing changes
+\textbf{It may be good to fix the size of the points, the center of the
+axes and the scales of the axes rather than allow them to change as the
+projection and slicing changes. Fix 4}
 
 \textbf{I have found it useful to input different data sets rather than
-have this function sort them automatically using the last column.}
+have this function sort them automatically using the last column. One
+data set with a flag column at the end, is best! Is the code too
+complicated? Fix 5}
 
@@
 
+\textbf{For Alex to do}
+
+\begin{itemize}
+\tightlist
+\item
   \textbf{We would need a screenshot (and video) for each of these.}
+\item
+  \textbf{Fix 1, 2, 3, 4, 5}
+\item
+  \textbf{Code consistent for applications to work - see
+  slice\_tour\_compare\_sets.nb. See below}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \textbf{We need to make sure the drawing range is fixed (no resizing
+    of the axis when changing projection or slicing). It could be nice
+    if there was a manual zoom in/zoom out to change scale
+    occasionally.}
+  \item
+    \textbf{Should also make sure the size of the points is fixed (maybe
+    this will be automatic if the drawing range is fixed?), bigger
+    points would be preferred, currently the size is sometimes really
+    tiny and it becomes difficult to see anything.}
+  \item
+    \textbf{Can we have some heuristic to decide the range for the slice
+    thickness slider? In the second graph there is a tiny part of the
+    slider where the changes are interesting, difficult to navigate
+    since this makes it super sensitive to any small movement, while
+    most of the options are not interesting. My suggestion would be to
+    keep (as I understand is the case now) the maximum thickness (all
+    points in) as the upper bound, but instead of zero use a thickness
+    below which slices tend to be empty as the lower bound.}
+  \item
+    \textbf{For the olives example I think the relation between x2 and
+    the blue region could be interesting.}
+  \item
+    \textbf{pdfsense data: can we use random samples instead of the
+    first 10k rows? Then we should still be able to see the bigger
+    picture (with picking the first ones we get some artificial
+    structure from how the data was sampled initially).}
+  \item
+    \textbf{I was playing around with this today. Looks like the axes
+    are fixed when changing slice thickness, but not when rotating?}
+  \item
+    \textbf{Anyways, some observations that might be useful, but all of
+    this seems difficult for including in a paper:}
+  \item
+    \textbf{In the first example we can show how the yellow group is
+    associated with the first variable (x1, I guess this is C9?) - when
+    slicing through the center it only appears when x1 has a bigger
+    component on the projection, it is associated with low values of x1.
+    Since they are not near the average value of x1 the points do not
+    get captured in a slice where x1 is not important in the projection.
+    We can move the center point to -1 in the first component to then
+    explore how the yellow group relates to the other three variables.
+    Much less clear, maybe some indication of smaller x2 and larger x3
+    for this.}
+  \item
+    \textbf{The second example is too busy for manual exploration I
+    would say (too many axes to move around separately).}
+  \item
+    \textbf{Third example could be similar to the first: blue region
+    requires small values of x2, can be understood from the exploration,
+    then move center point to -0.5 in x2 to explore further. Now
+    interpretation becomes tricky. Sometimes blue and yellow seem just
+    randomly everywhere, sometimes blue more contained in an ellipsoid
+    around the center (indicating some correlation between variables).
+    For example try zeroing out components 2 and 3 to see this slice in
+    x1 vs x4.}
+  \item
+    \textbf{For the last example x4 seems to be the one to work with. I
+    feel like this is again very similar, maybe better to instead
+    explore other examples (maybe something without slicing?)}
+  \end{itemize}
+\end{itemize}
 
@@
 
+\hypertarget{extension-added-in-the-r-package-tourr}{%
+\section{\texorpdfstring{Extension added in the R package
+\texttt{tourr}}{Extension added in the R package tourr}}\label{extension-added-in-the-r-package-tourr}}
+
+\hypertarget{what-would-be-desirable-for-implementations-in-r}{%
+\section{What would be desirable for implementations in
+R?}\label{what-would-be-desirable-for-implementations-in-r}}
+
 \hypertarget{sec:discussion}{%
 \section{Discussion}\label{sec:discussion}}
-
-\begin{itemize}
-\tightlist
-\item
-  As was said before, we need to make sure the drawing range is fixed
-  (no resizing of the axis when changing projection or slicing).
-\item
-  Should also make sure the size of the points is fixed (maybe this will
-  be automatic if the drawing range is fixed?), bigger points would be
-  preferred, currently the size is sometimes really tiny and it becomes
-  difficult to see anything.
-\item
-  Can we have some heuristic to decide the range for the slice thickness
-  slider? In the second graph there is a tiny part of the slider where
-  the changes are interesting, difficult to navigate since this makes it
-  super sensitive to any small movement, while most of the options are
-  not interesting. My suggestion would be to keep (as I understand is
-  the case now) the maximum thickness (all points in) as the upper
-  bound, but instead of zero use a thickness below which slices tend to
-  be empty as the lower bound.
-\item
-  For the olives example I think the relation between x2 and the blue
-  region could be interesting.
-\item
-  pdfsense data: can we use random samples instead of the first 10k
-  rows? Then we should still be able to see the bigger picture (with
-  picking the first ones we get some artificial structure from how the
-  data was sampled initially).
-\end{itemize}
-
-I was playing around with this today. Looks like the axes are fixed when
-changing slice thickness, but not when rotating?
-
-Anyways, some observations that might be useful, but all of this seems
-difficult for including in a paper:
-
-\begin{itemize}
-\tightlist
-\item
-  In the first example we can show how the yellow group is associated
-  with the first variable (x1, I guess this is C9?) - when slicing
-  through the center it only appears when x1 has a bigger component on
-  the projection, it is associated with low values of x1. Since they are
-  not near the average value of x1 the points do not get captured in a
-  slice where x1 is not important in the projection. We can move the
-  center point to -1 in the first component to then explore how the
-  yellow group relates to the other three variables. Much less clear,
-  maybe some indication of smaller x2 and larger x3 for this.
-\item
-  The second example is too busy for manual exploration I would say (too
-  many axes to move around separately).
-\item
-  Third example could be similar to the first: blue region requires
-  small values of x2, can be understood from the exploration, then move
-  center point to -0.5 in x2 to explore further. Now interpretation
-  becomes tricky. Sometimes blue and yellow seem just randomly
-  everywhere, sometimes blue more contained in an ellipsoid around the
-  center (indicating some correlation between variables). For example
-  try zeroing out components 2 and 3 to see this slice in x1 vs x4.
-\item
-  For the last example x4 seems to be the one to work with. I feel like
-  this is again very similar, maybe better to instead explore other
-  examples (maybe something without slicing?)
-\end{itemize}
 

### 2022-07-07 paper/paper.Rmd

@@
 For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
-These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
+These controls can also be used to interactively generate faceted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
 Having the user interact with a projection is extremely valuable for understanding high-dimensional data.
@@
 
-# Manual tour {#sec:method}
+# What is a manual tour {#sec:method}
 
-An orthonormal basis ($A_{p\times d}$) and a variable id ($m \in \{1, ..., p\}$) to control are provided to initialise a manual tour.
+An orthonormal basis ($A_{p\times d}$) and a variable id ($m \in \{1, ..., p\}$) to control are provided to initialize a manual tour.
 A method to update the values of the component of the controlled variable $V_m$ is needed.
 
-## Background
+## Existing methods
 
@@
 
-## New simpler definition
+## The new simpler definition
 
@@
 
-The problem with Algorithm 1 is that the precise values for $V_m$ cannot be specified because the orthonormalisation wil change them.
+The problem with Algorithm 1 is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
 This modification will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)).
@@
 If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
-If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalised after each adjustment.
+If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
 
@@
 
-# Implementation {#sec:implementation}
+# Experimenting with new techniques using Mathematica {#sec:implementation}
+
+- **Why is this a good sandbox**
+- **Explain the functionality available in the notebooks**
 
@@
     - ``projmat`: A projection matrix which describes the projection plane. This is the 
-initial projection and does not neet by orthonormal.
+initial projection and does not need to be orthonormal (XXX is this right?).
     - ``xRange`: A list which details the range of the horizontal axis.
@@
 
-**It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**
 
@@
 
-**It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
 
-It may be good to fix the size of the points, the centre of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
 
@@
 
-- `data`: A data matrix with an extra groupling or clustering column (numerical) 
-- `projmat`: The inital projection matrix
+- `data`: A data matrix with an extra grouping or clustering column (numerical) 
+- `projmat`: The initial projection matrix
 - `centrepoint`: The initial position of the slice
@@
 
-**Does this need to be a separate function from SlicePlot?**
+**Does this need to be a separate function from SlicePlot? Fix 2**
 
-It would be good if the initial projection matrix (and maybe the centre) had default values. Perhaps a random matrix with the correct dimensions read from the data?
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**
 
-It may be good to fix the size of the points, the centre of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**
 
-**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.**
+**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
+One data set with a flag column at the end, is best!
+Is the code too complicated?
+Fix 5**
 
@@
 
-**We would need a screenshot (and video) for each of these.**
+**For Alex to do**
+
+- **We would need a screenshot (and video) for each of these.**
+- **Fix 1, 2, 3, 4, 5**
+- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
+    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
+It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
+    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
+    - **Can we have some heuristic to decide the range for the slice thickness slider?
+In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
+My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
+    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
+    - **pdfsense data: can we use random samples instead of the first 10k rows?
+Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
+    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
+    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
+    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
+Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
+We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
+Much less clear, maybe some indication of smaller x2 and larger x3 for this.**
+    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
+    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
+Now interpretation becomes tricky.
+Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
+For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
+    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
 
@@
 
+# Extension added in the R package `tourr`
+
+# What would be desirable for implementations in R?
+
 # Discussion {#sec:discussion}
-
-- As was said before, we need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
-- Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.
-- Can we have some heuristic to decide the range for the slice thickness slider?
-In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
-My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.
-- For the olives example I think the relation between x2 and the blue region could be interesting.
-- pdfsense data: can we use random samples instead of the first 10k rows?
-Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).
-
-I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?
-
-Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:
-
-- In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
-Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
-We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
-Much less clear, maybe some indication of smaller x2 and larger x3 for this.
-- The second example is too busy for manual exploration I would say (too many axes to move around separately).
-- Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
-Now interpretation becomes tricky.
-Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
-For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.
-- For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)
 

### 2022-07-07 code/penguins.R
Writing out the penguin data to csv

@@ -40,3 +40,13 @@ penguins_lda <- read_csv("data/penguins_lda.csv")
 ggscatmat(penguins_data, columns = 2:5, col="species")
 ggscatmat(penguins_rf, columns = 1:4, col="species", alpha=0.1)
 ggscatmat(penguins_lda, columns = 1:4, col="species", alpha=0.1)
+
+
+# saving the data to look at in mathematica
+penguins_data %>%
+  dplyr::select(c(2,3,4,5,1)) %>%
+  mutate(species_flag =
+           as.numeric(factor(species,
+                             levels = # making sure we have the same flags
+                               levels(as.factor(rf_samples$species))))) %>%
+  write_csv(file = "data/penguins.csv")

### 2022-07-07 paper/paper.Rmd
changed title

@@
 title: |
-  A new simplified manual tour, with examples in mathematica and R
+  A new simplified manual tour, with application to exploring classification boundaries in high dimensions
 type: Short Technical Note

### 2022-07-07 paper/paper.Rmd
revised intro and manual tour explanation

@@
 These are dynamic data visualizations methods and are collected under the term *tours*.
-Tours involve views of high-dimensional ($p$) data in low-dimensional ($d$) projections.
+Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
 In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
@@
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-It is impractical to impossible to steer through all possible projections, unlike Asimov's tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, as summarized in @lee2021.
+(A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
+After Asimov there have been many, many tour developments, which are summarized in @lee2021.
 
@@
 One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour.
-@cook_manual_1997 describes controls for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into or out of or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
-@spyrison_spinifex_2020 refines this algorithm and implements them to generate animations.
+@cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, or out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
+@spyrison_spinifex_2020 refined this algorithm and implements them to generate animation sequences.
 
@@
 
-## The new simpler definition
+## A new simpler and broadly applicable approach
 
@@
 
-### Algorithm 1
+### Algorithm
 
@@
 Changes in the other components happen as a result of the orthonormalisation, but are uncontrolled.
-
-### Algorithm 2
-
-The problem with Algorithm 1 is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
-This modification will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)).
-The algorithm is as follows:
-
-1. Provide $A$, and $m$.
-2. Change values in row $m$, giving $A^*$.
-3. Store row $m$ separately, and zero the values of row $m$ in $A^*$, giving $A^{*0}$. 
-4. Orthonormalise $A^{*0}$, using Gram-Schmidt.
-5. Replace row $m$ with the original values, giving $A^{**}$. 
-6. For $d=2$, adjust the values of ${\boldmath a}^{**}_{.2}$ using
-
-$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m$$.
-
-which ensures that
-
-$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0$$.
-
-If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
-If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
-
-<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
-
-Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
-
-### Algorithm 3
-
-For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
-Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
-Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
-Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
-One possible algorithm to achieve this is
-\begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
-\begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
-\end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
-\end{enumerate}
-The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
-
-\begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
-\CodeBefore
-       \columncolor{red!15}{1,2}
-              \rowcolor{blue!15}{4}
-     \Body
-       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
-e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
-e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
- e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
-e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
-\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
-\end{align}
 
@@
 
-### Potential Algorithm 3
+## Refinements to enforce exact position
+
+The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
+
+### Adjustment method 1
+
+A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)). It is as follows:
+
+1. Provide $A$, and $m$.
+2. Change values in row $m$, giving $A^*$.
+3. Store row $m$ separately, and zero the values of row $m$ in $A^*$, giving $A^{*0}$. 
+4. Orthonormalise $A^{*0}$, using Gram-Schmidt.
+5. Replace row $m$ with the original values, giving $A^{**}$. 
+6. For $d=2$, adjust the values of ${\boldmath a}^{**}_{.2}$ using
+
+$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m$$.
+
+which ensures that
+
+$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0$$.
+
+If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
+If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
+
+<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+
+Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
+
+### Adjustment method 2
+
+For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
+Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
+Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+One possible algorithm to achieve this is
+\begin{enumerate}
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
+\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
+\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\end{enumerate}
+\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\end{enumerate}
+The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
+
+\begin{align}
+A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
+e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
+e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
+\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
+\end{align}
+
+### Adjustment method 3
 

### 2022-07-07 paper/paper.tex
changed title
revised intro and manual tour explanation

@@
 
-\title{A new simplified manual tour, with examples in mathematica and R}
+\title{A new simplified manual tour, with application to exploring
+classification boundaries in high dimensions}
 
@@
 collected under the term \emph{tours}. Tours involve views of
-high-dimensional (\(p\)) data in low-dimensional (\(d\)) projections. In
-his original paper on the grand tour, \citet{As85} provided several
+high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
+In his original paper on the grand tour, \citet{As85} provided several
 algorithms for tour paths that could theoretically show the viewer the
@@
 user-controlled rotations on coordinate axes, allowing one to manually
-tour through low-dimensional projections. It is impractical to
-impossible to steer through all possible projections, unlike Asimov's
+tour through low-dimensional projections. (A video illustrating the
+capabilities is available through video library of \citet{ASA22}.)
+Steering through all possible projections is impossible, unlike Asimov's
 tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, as summarized
-in \citet{lee2021}.
+After Asimov there have been many, many tour developments, which are
+summarized in \citet{lee2021}.
 
 One such direction of work develops the ideas from PRIM-9, to provide
-manual control of a tour. \citet{cook_manual_1997} describes controls
-for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing
-the user to select any variable axis, and rotate it into or out of or
-around the projection through horizontal, vertical, oblique, radial or
-angular changes in value. \citet{spyrison_spinifex_2020} refines this
-algorithm and implements them to generate animations.
+manual control of a tour. \citet{cook_manual_1997} describe controls for
+1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space,
+allowing the user to select any variable axis, and rotate it into, or
+out of, or around the projection through horizontal, vertical, oblique,
+radial or angular changes in value. \citet{spyrison_spinifex_2020}
+refined this algorithm and implements them to generate animation
+sequences.
 
@@
 
-\hypertarget{the-new-simpler-definition}{%
-\subsection{The new simpler
-definition}\label{the-new-simpler-definition}}
+\hypertarget{a-new-simpler-and-broadly-applicable-approach}{%
+\subsection{A new simpler and broadly applicable
+approach}\label{a-new-simpler-and-broadly-applicable-approach}}
 
@@
 
-\hypertarget{algorithm-1}{%
-\subsubsection{Algorithm 1}\label{algorithm-1}}
+\hypertarget{algorithm}{%
+\subsubsection{Algorithm}\label{algorithm}}
 
@@
 
-\hypertarget{algorithm-2}{%
-\subsubsection{Algorithm 2}\label{algorithm-2}}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+The dot (orange) indicates the chosen values for the controlled variable.
+For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation.}\label{fig:manualsequence}
+\end{figure}
 
-The problem with Algorithm 1 is that the precise values for \(V_m\)
-cannot be specified because the orthonormalisation will change them.
-This modification will maintain the components of \(V_m\) precisely
-(Figure \ref{fig:manualsequence} (bottom row)). The algorithm is as
-follows:
+\hypertarget{refinements-to-enforce-exact-position}{%
+\subsection{Refinements to enforce exact
+position}\label{refinements-to-enforce-exact-position}}
+
+The problem with new simple method (Algorithm 1) is that the precise
+values for \(V_m\) cannot be specified because the orthonormalisation
+will change them.
+
+\hypertarget{adjustment-method-1}{%
+\subsubsection{Adjustment method 1}\label{adjustment-method-1}}
+
+A small modification to algorithm 1 will maintain the components of
+\(V_m\) precisely (Figure \ref{fig:manualsequence} (bottom row)). It is
+as follows:
 
@@
 
-\hypertarget{algorithm-3}{%
-\subsubsection{Algorithm 3}\label{algorithm-3}}
+\hypertarget{adjustment-method-2}{%
+\subsubsection{Adjustment method 2}\label{adjustment-method-2}}
 
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
-The dot (orange) indicates the chosen values for the controlled variable.
-For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation.}\label{fig:manualsequence}
-\end{figure}
-
-\hypertarget{potential-algorithm-3}{%
-\subsubsection{Potential Algorithm 3}\label{potential-algorithm-3}}
+\hypertarget{adjustment-method-3}{%
+\subsubsection{Adjustment method 3}\label{adjustment-method-3}}
 

### 2022-07-07 paper/paper.Rmd
start on penguins

@@
 
-# What is a manual tour {#sec:method}
+# What is a manual tour? {#sec:method}
 
@@
 
+To illustrate the usefulness of the manual tour use the penguins data [@penguins].
+We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
+
+```{r penguins-scatmat, out.width="100%", width=10, height=10, fig.cap="Scatterplot matrix of the penguins data."}
+library(tidyverse)
+library(palmerpenguins)
+library(GGally)
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x)))
+ggscatmat(penguins_data, columns = 2:5, col="species") +
+  theme(legend.position = "bottom")
+```
+
 # Extension added in the R package `tourr`
+
+Explain `radial_tour`
 

### 2022-07-07 paper/paper.tex
start on penguins

@@
 \hypertarget{sec:method}{%
-\section{What is a manual tour}\label{sec:method}}
+\section{What is a manual tour?}\label{sec:method}}
 
@@
 
+To illustrate the usefulness of the manual tour use the penguins data
+\citep{penguins}. We will show how classification boundaries can be
+explored and better understood using a manual tour, on projections and
+slices through 4D space.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/penguins-scatmat-1} \caption{Scatterplot matrix of the penguins data.}\label{fig:penguins-scatmat}
+\end{figure}
+
 \hypertarget{extension-added-in-the-r-package-tourr}{%
@@
 \texttt{tourr}}{Extension added in the R package tourr}}\label{extension-added-in-the-r-package-tourr}}
+
+Explain \texttt{radial\_tour}
 

### 2022-07-07 paper/paper.Rmd
penguins colour matching

--- 2022-07-07 paper/paper.Rmd
@@
 
+To illustrate the usefulness of the manual tour use the penguins data [@penguins].
+We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
+
+```{r penguins-scatmat, out.width="100%", width=10, height=10, fig.cap="Scatterplot matrix of the (standardised) penguins data."}
+library(tidyverse)
+library(palmerpenguins)
+library(GGally)
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x)))
+ggscatmat(penguins_data, columns = 2:5, col="species") +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  theme(legend.position = "bottom")
+```
+
 # Extension added in the R package `tourr`
+
+Explain `radial_tour`
 

### 2022-07-07 paper/paper.tex
penguins colour matching

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/penguins-scatmat-1} \caption{Scatterplot matrix of the penguins data.}\label{fig:penguins-scatmat}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/penguins-scatmat-1} \caption{Scatterplot matrix of the (standardised) penguins data.}\label{fig:penguins-scatmat}
 \end{figure}

### 2022-07-07 paper/paper.Rmd
shortened variable names of penguins

--- 2022-07-07 paper/paper.Rmd
@@
 
+To illustrate the usefulness of the manual tour use the penguins data [@penguins].
+We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
+
+```{r penguins-scatmat, out.width="100%", width=10, height=10, fig.cap="Scatterplot matrix of the (standardised) penguins data."}
+library(tidyverse)
+library(palmerpenguins)
+library(GGally)
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
+  rename("bl"="bill_length_mm",
+         "bd"="bill_depth_mm",
+         "fl"="flipper_length_mm",
+         "bm"="body_mass_g")
+ggscatmat(penguins_data, columns = 2:5, col="species") +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  theme(legend.position = "bottom")
+```
+
 # Extension added in the R package `tourr`
+
+Explain `radial_tour`
 

### 2022-07-07 paper/paper.Rmd
Some writing about the application

--- 2022-07-07 paper/paper.Rmd
@@
 
+We use the `classifly` package [@classifly] to generate predictions across the 4D space spanned by the data, with two classification models (linear discriminant analysis and random forest).
+
+The first projection we will start from is looking at bill depth vs flipper length (x2 vs x3).
+In a projection of the data we see that the Gentoo penguins are separate from the other two species which are on top of each other.
+If we instead look at a slice through the center of the distribution, we see that shifting the center point to large values of x1 (bill length), e.g.
+x1=2, results in Adelie penguins dropping out of the slice, while a slice with low values (x1=-2) in the center point only contains Adelie penguin observations.
+
+We can next explore how this is mapped in the two classification models.
+Both random forest and LDA can separate out the region for Gentoo penguins in a thin slice through the center (h = 0.5), while Adelie and Chinstrap predictions overlap (blue and yellow, an effect of the slice thickness).
+We also notice the different boundaries: while LDA produces linear boundaries by construction, random forest insead produces "boxes" from cutting up the data in the tree structure.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.4\textwidth]{figures/view1_lda.png}
+\includegraphics[width=0.4\textwidth]{figures/view1_rf.png}}
+\caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model.
+The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinugishing features of the two models in terms of the different boundaries.}
+\label{view1}
+\end{figure*}
+
+We can then start shifting the slice along x1 as we did when looking at the data, again moving to $\pm 1$ will separate the Adelie from the Chinstrap penguins, now in terms of the model predictions.
+(Could include screenshot here as well.)
+
+Thus, stepping through the different slices shows the relation between the species and the first variable (x1, bill length).
+Instead of moving the center point we could also keep the slice fixed as passing through the center of the distribution, and use the manual controls to rotate x1 into the projection.
+Here we started from the projected view of the data to identify a combination that can separate all three species, and then use the projection matrix as input to the slice displays for the model.
+Interestingly, the two models agree on how to classify the "empty" regions (where there are no data points) for this example.
+(this is applying A3 as the starting projection).
+
+Finally we use the manual tour on the random forest predictions first, to identify a centered slice with an interesting classification boundary.
+Here a combination of all four predictors is found to result in an interesting slice, where the predictions for the Gentoo penguins gets split into two regions of the slice, indicating the non-linearity of the random forest model.
+While x4 (body mass) is already contributing a lot in the projection, we can learn more by also moving the center point along that direction.
+Moving to high values (x4 = 2) results in a much larger part of the slice being predicted as Gentoo penguins, while now the yellow region (Chinstrap penguins) gets split.
+Moving in the opposite direction (x4 = -1) instead shows primarily Chinstrap penguins in the predictions, with Gentoo only appearing along the edge.
+
+Looking at the data, and shifting a fat slice (h = 1.5) as specified above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have observations from the Gentoo species, while x4 = 2 contains the Gentoo observations, but hardly any of the Chinstrap penguins.
+
+The last step is comparing to the LDA predictions in those same slices.
+With RF we could see the non-linear aspects in the boundaries, while LDA can only produce linear boundaries, so these look very different.
+On the other hand, the same dependence is observed when shifting the slice along the x4 direction.
+
 # Extension added in the R package `tourr`

### 2022-07-07 paper/paper.Rmd

@@
   drop_na() %>%
-  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x)))
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
+  rename("bl"="bill_length_mm",
+         "bd"="bill_depth_mm",
+         "fl"="flipper_length_mm",
+         "bm"="body_mass_g")
 ggscatmat(penguins_data, columns = 2:5, col="species") +

### 2022-07-07 paper/paper.tex
compiled tex file

@@
 
+We use the \texttt{classifly} package \citep{classifly} to generate
+predictions across the 4D space spanned by the data, with two
+classification models (linear discriminant analysis and random forest).
+
+The first projection we will start from is looking at bill depth vs
+flipper length (x2 vs x3). In a projection of the data we see that the
+Gentoo penguins are separate from the other two species which are on top
+of each other. If we instead look at a slice through the center of the
+distribution, we see that shifting the center point to large values of
+x1 (bill length), e.g.~x1=2, results in Adelie penguins dropping out of
+the slice, while a slice with low values (x1=-2) in the center point
+only contains Adelie penguin observations.
+
+We can next explore how this is mapped in the two classification models.
+Both random forest and LDA can separate out the region for Gentoo
+penguins in a thin slice through the center (h = 0.5), while Adelie and
+Chinstrap predictions overlap (blue and yellow, an effect of the slice
+thickness). We also notice the different boundaries: while LDA produces
+linear boundaries by construction, random forest insead produces
+``boxes'' from cutting up the data in the tree structure.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.4\textwidth]{figures/view1_lda.png}
+\includegraphics[width=0.4\textwidth]{figures/view1_rf.png}}
+\caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model.
+The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinugishing features of the two models in terms of the different boundaries.}
+\label{view1}
+\end{figure*}
+
+We can then start shifting the slice along x1 as we did when looking at
+the data, again moving to \(\pm 1\) will separate the Adelie from the
+Chinstrap penguins, now in terms of the model predictions. (Could
+include screenshot here as well.)
+
+Thus, stepping through the different slices shows the relation between
+the species and the first variable (x1, bill length). Instead of moving
+the center point we could also keep the slice fixed as passing through
+the center of the distribution, and use the manual controls to rotate x1
+into the projection. Here we started from the projected view of the data
+to identify a combination that can separate all three species, and then
+use the projection matrix as input to the slice displays for the model.
+Interestingly, the two models agree on how to classify the ``empty''
+regions (where there are no data points) for this example. (this is
+applying A3 as the starting projection).
+
+Finally we use the manual tour on the random forest predictions first,
+to identify a centered slice with an interesting classification
+boundary. Here a combination of all four predictors is found to result
+in an interesting slice, where the predictions for the Gentoo penguins
+gets split into two regions of the slice, indicating the non-linearity
+of the random forest model. While x4 (body mass) is already contributing
+a lot in the projection, we can learn more by also moving the center
+point along that direction. Moving to high values (x4 = 2) results in a
+much larger part of the slice being predicted as Gentoo penguins, while
+now the yellow region (Chinstrap penguins) gets split. Moving in the
+opposite direction (x4 = -1) instead shows primarily Chinstrap penguins
+in the predictions, with Gentoo only appearing along the edge.
+
+Looking at the data, and shifting a fat slice (h = 1.5) as specified
+above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have
+observations from the Gentoo species, while x4 = 2 contains the Gentoo
+observations, but hardly any of the Chinstrap penguins.
+
+The last step is comparing to the LDA predictions in those same slices.
+With RF we could see the non-linear aspects in the boundaries, while LDA
+can only produce linear boundaries, so these look very different. On the
+other hand, the same dependence is observed when shifting the slice
+along the x4 direction.
+
 \hypertarget{extension-added-in-the-r-package-tourr}{%

### 2022-07-08 paper/paper.Rmd
refininf manual tour explanation

@@
 
-# What is a manual tour? {#sec:method}
+# How to construct a manual tour {#sec:method}
 
-An orthonormal basis ($A_{p\times d}$) and a variable id ($m \in \{1, ..., p\}$) to control are provided to initialize a manual tour.
-A method to update the values of the component of the controlled variable $V_m$ is needed.
+A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d-D$ projection.
+The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed.
+A method to update the values of the component of the controlled variable $V_m$ is then needed.
 
@@
 
-# Applications {#sec:examples}
+# Application {#sec:examples}
 

### 2022-07-08 paper/paper.tex
refininf manual tour explanation

@@
 \hypertarget{sec:method}{%
-\section{What is a manual tour?}\label{sec:method}}
+\section{How to construct a manual tour}\label{sec:method}}
 
-An orthonormal basis (\(A_{p\times d}\)) and a variable id
-(\(m \in \{1, ..., p\}\)) to control are provided to initialize a manual
-tour. A method to update the values of the component of the controlled
-variable \(V_m\) is needed.
+A manual tour allows the user to alter the coefficients of one (or more)
+variables contributing to a \(d-D\) projection. The initial ingredients
+are an orthonormal basis (\(A_{p\times d}\)) defining the projection of
+the data, and a variable id (\(m \in \{1, ..., p\}\)) specifying which
+coefficient will be changed. A method to update the values of the
+component of the controlled variable \(V_m\) is then needed.
 
@@
 \hypertarget{sec:examples}{%
-\section{Applications}\label{sec:examples}}
+\section{Application}\label{sec:examples}}
 

### 2022-07-08 paper/paper.Rmd
application explanation of penguins

@@
 
-In the original work, the method for updating component values, for a 2D projection, was built trackball controls in 3D.
-A 3D manipulation space is created, as illustrated in Figure \ref{manipspace}, where the controlled variable has full range of motion from -1 to 1.
-Movements of a cursor are recorded and converted into changes in the values of $V_m$ to change it's values in the displayed 2D projection.
-Movement could also be constrained to be only in horizontal, vertical, radial or angular motions.
+The method for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
+The first step was to construct a 3D manipulation space from a 2D projection, as illustrated in Figure \ref{manipspace}.
+In this space the coefficient of the controlled variable ranges between -1 and 1.
+Movements of a cursor are recorded and converted into changes in the values of $V_m$ thus changing the displayed 2D projection.
+The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.
 
@@
 
-To illustrate the usefulness of the manual tour use the penguins data [@penguins].
+To illustrate the usefulness of the manual tour we use the 4D penguins data [@penguins].
 We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
+Figure \ref(fig:penguins-scatmat) shows a scatterplot matrix of this data.
+There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
+The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
 
-```{r penguins-scatmat, out.width="100%", width=10, height=10, fig.cap="Scatterplot matrix of the (standardised) penguins data."}
+```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
+The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
 library(tidyverse)

### 2022-07-08 paper/paper.tex
application explanation of penguins

@@
 
-In the original work, the method for updating component values, for a 2D
-projection, was built trackball controls in 3D. A 3D manipulation space
-is created, as illustrated in Figure \ref{manipspace}, where the
-controlled variable has full range of motion from -1 to 1. Movements of
-a cursor are recorded and converted into changes in the values of
-\(V_m\) to change it's values in the displayed 2D projection. Movement
-could also be constrained to be only in horizontal, vertical, radial or
-angular motions.
+The method for updating component values in \citet{cook_manual_1997}
+(and utilised in \citet{spyrison_spinifex_2020}) are prescribed
+primarily for a 2D projection, to take advantage of (then) newly
+developed 3D trackball controls made available for computer gaming. The
+first step was to construct a 3D manipulation space from a 2D
+projection, as illustrated in Figure \ref{manipspace}. In this space the
+coefficient of the controlled variable ranges between -1 and 1.
+Movements of a cursor are recorded and converted into changes in the
+values of \(V_m\) thus changing the displayed 2D projection. The
+algorithm also provided constraints to horizontal, vertical, radial or
+angular motions only.
 
@@
 
-To illustrate the usefulness of the manual tour use the penguins data
-\citep{penguins}. We will show how classification boundaries can be
+To illustrate the usefulness of the manual tour we use the 4D penguins
+data \citep{penguins}. We will show how classification boundaries can be
 explored and better understood using a manual tour, on projections and
-slices through 4D space.
+slices through 4D space. Figure \ref(fig:penguins-scatmat) shows a
+scatterplot matrix of this data. There are four variables
+(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
+measuring the size of the penguins from three species (Adelie, Chinstrap
+and Gentoo). The scatterplot matrix shows that the three species appear
+to be likely separable, and that at least the Gentoo can be
+distinguished from the other two species when \texttt{bd} is paired with
+\texttt{fl} or \texttt{bm}.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/penguins-scatmat-1} \caption{Scatterplot matrix of the (standardised) penguins data.}\label{fig:penguins-scatmat}
+
+{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
+
+}
+
+\caption{Scatterplot matrix of the (standardised) penguins data.
+The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
 \end{figure}

### 2022-07-08 paper/paper.Rmd
splitting out wlements of the application section

@@
 title: |
-  A new simplified manual tour, with application to exploring classification boundaries in high dimensions
+  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
 type: Short Technical Note
@@
 The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
+The steps for exploring boundaries are as follows:
+
+1. Build classification model, here we use a random forest and linear discriminant analysis.
+2. Predict the class for a dense grid of values covering the data space.
+3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
+4. Slice through the center, to explore where the boundaries will likely meet.
+5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
 
@@
 
+## Constructing the 4D prediction regions
+
 We use the `classifly` package [@classifly] to generate predictions across the 4D space spanned by the data, with two classification models (linear discriminant analysis and random forest).
+
+## Exploring projections manually
 
@@
 
+## Slicing through the center
+
 We can then start shifting the slice along x1 as we did when looking at the data, again moving to $\pm 1$ will separate the Adelie from the Chinstrap penguins, now in terms of the model predictions.
 (Could include screenshot here as well.)
+
+## Shiftinh the slice center
 
@@
 Looking at the data, and shifting a fat slice (h = 1.5) as specified above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have observations from the Gentoo species, while x4 = 2 contains the Gentoo observations, but hardly any of the Chinstrap penguins.
+
+## Comparison of boundaries produced by different classifiers
 

### 2022-07-08 paper/paper.tex
splitting out wlements of the application section

@@
 
-\title{A new simplified manual tour, with application to exploring
-classification boundaries in high dimensions}
+\title{New and simplified manual controls for projection and slice
+tours, with application to exploring classification boundaries in high
+dimensions}
 
@@
 distinguished from the other two species when \texttt{bd} is paired with
-\texttt{fl} or \texttt{bm}.
+\texttt{fl} or \texttt{bm}. The steps for exploring boundaries are as
+follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Build classification model, here we use a random forest and linear
+  discriminant analysis.
+\item
+  Predict the class for a dense grid of values covering the data space.
+\item
+  Examine projections, using a manual tour so that the contribution of
+  any variable is controlled.
+\item
+  Slice through the center, to explore where the boundaries will likely
+  meet.
+\item
+  Move the slice by changing the center in the direction of a single
+  variable to explore the extent of a boundary for a single group
+  relative to a variable.
+\end{enumerate}
 
@@
 
+\hypertarget{constructing-the-4d-prediction-regions}{%
+\subsection{Constructing the 4D prediction
+regions}\label{constructing-the-4d-prediction-regions}}
+
 We use the \texttt{classifly} package \citep{classifly} to generate
@@
 classification models (linear discriminant analysis and random forest).
+
+\hypertarget{exploring-projections-manually}{%
+\subsection{Exploring projections
+manually}\label{exploring-projections-manually}}
 
@@
 
+\hypertarget{slicing-through-the-center}{%
+\subsection{Slicing through the
+center}\label{slicing-through-the-center}}
+
 We can then start shifting the slice along x1 as we did when looking at
@@
 include screenshot here as well.)
+
+\hypertarget{shiftinh-the-slice-center}{%
+\subsection{Shiftinh the slice center}\label{shiftinh-the-slice-center}}
 
@@
 observations, but hardly any of the Chinstrap penguins.
+
+\hypertarget{comparison-of-boundaries-produced-by-different-classifiers}{%
+\subsection{Comparison of boundaries produced by different
+classifiers}\label{comparison-of-boundaries-produced-by-different-classifiers}}
 

### 2022-07-08 paper/paper.Rmd
figure refences, and fig 3 was not needed

@@
 
-```{r othermethod, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
+```{r othermethod, eval=FALSE, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
 source("../code/plot_basis.R")
@@
 We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
-Figure \ref(fig:penguins-scatmat) shows a scatterplot matrix of this data.
+Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
 There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).

### 2022-07-08 paper/paper.tex
figure refences, and fig 3 was not needed

@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/othermethod-1} \caption{Manual controls with algorithm 2. The precise location of the axis is maintained.}\label{fig:othermethod}
-\end{figure}
-
 \hypertarget{sec:implementation}{%
@@
 explored and better understood using a manual tour, on projections and
-slices through 4D space. Figure \ref(fig:penguins-scatmat) shows a
+slices through 4D space. Figure \ref{fig:penguins-scatmat} shows a
 scatterplot matrix of this data. There are four variables

### 2022-07-08 paper/paper.Rmd
typos

@@
 \caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model.
-The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinugishing features of the two models in terms of the different boundaries.}
+The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinguishing features of the two models in terms of the different boundaries.}
 \label{view1}
@@
 
-## Shiftinh the slice center
+## Shifting the slice center
 

### 2022-07-08 paper/paper.tex
typos

@@
 
-\hypertarget{shiftinh-the-slice-center}{%
-\subsection{Shiftinh the slice center}\label{shiftinh-the-slice-center}}
+\hypertarget{shifting-the-slice-center}{%
+\subsection{Shifting the slice center}\label{shifting-the-slice-center}}
 

### 2022-07-08 paper/paper.Rmd
Worked out penguins example

@@
 
-We use the `classifly` package [@classifly] to generate predictions across the 4D space spanned by the data, with two classification models (linear discriminant analysis and random forest).
+We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest(RF).
 
@@
 
-The first projection we will start from is looking at bill depth vs flipper length (x2 vs x3).
-In a projection of the data we see that the Gentoo penguins are separate from the other two species which are on top of each other.
-If we instead look at a slice through the center of the distribution, we see that shifting the center point to large values of x1 (bill length), e.g.
-x1=2, results in Adelie penguins dropping out of the slice, while a slice with low values (x1=-2) in the center point only contains Adelie penguin observations.
-
-We can next explore how this is mapped in the two classification models.
-Both random forest and LDA can separate out the region for Gentoo penguins in a thin slice through the center (h = 0.5), while Adelie and Chinstrap predictions overlap (blue and yellow, an effect of the slice thickness).
-We also notice the different boundaries: while LDA produces linear boundaries by construction, random forest insead produces "boxes" from cutting up the data in the tree structure.
+We start by first exploring the projections of the model prediction.
+Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
+To illustrate this difference we have rotated the projection view from the RF model to identify a projection that shows the non-linear, block-type structure introduced by the model, see Fig.
+\ref{proj1}.
+We then export the corresponding matrix $A_1$ to look at the same projection for the LDA model and the original data.
+Here we notice that while the overall picture is similar, the linearity of the LDA leads to a big difference in how two of the three species overlap across the projection (blue vs green, i.e.
+Adelie vs Gentoo penguins).
+Interestingly that same projection leads to almost perfect separation of the three species when looking at the data points, the overlap between the models is driven by what is happening in the orthogonal space to the plane.
 
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.4\textwidth]{figures/view1_lda.png}
-\includegraphics[width=0.4\textwidth]{figures/view1_rf.png}}
-\caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model.
-The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinguishing features of the two models in terms of the different boundaries.}
-\label{view1}
+\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
+We can clearly see the block structure, while the LDA model (middle) produces linear boundaries instead.
+The three groups are separated in this projection of the data (right)}.
+\label{proj1}
 \end{figure*}
@@
 
-We can then start shifting the slice along x1 as we did when looking at the data, again moving to $\pm 1$ will separate the Adelie from the Chinstrap penguins, now in terms of the model predictions.
-(Could include screenshot here as well.)
+XXXX currently in the notebook $A_1$ is called A5, and $A_2$ is called A6
+
+We continue the investigation by now slicing based on the projection $A_1$.
+For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0$.
+As a first option we explore how manually changing the projection away from $A_1$ can help with understanding the boundary better.
+For our example we notice that $A_1$ does not contain any contribution from the second variable (`bd`), and in our illustration we will first rotate this variable into the view.
+
+We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
+\ref{slice1}, top-left).
+This is similar for the LDA model (Fig.
+\ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
+
+By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
+\ref{slice1}).
+
+Finally we look at the data projection based on $A_2$ and find that the third group (green, Gentoo penguins) is further away from the other two groups (compared to $A_1$), see Fig. \ref{proj2}.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices based on $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With the starting projection $A_1$ we see one group overlapping (green, Gentoo), while $A_2$ results in clear boundaries inside the slice.
+The block vs linear structure of the two models is clearly visible in both slices.}
+\label{slice1}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
+\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
+\label{proj2}
+\end{figure*}
 
@@
 
-Thus, stepping through the different slices shows the relation between the species and the first variable (x1, bill length).
-Instead of moving the center point we could also keep the slice fixed as passing through the center of the distribution, and use the manual controls to rotate x1 into the projection.
-Here we started from the projected view of the data to identify a combination that can separate all three species, and then use the projection matrix as input to the slice displays for the model.
-Interestingly, the two models agree on how to classify the "empty" regions (where there are no data points) for this example.
-(this is applying A3 as the starting projection).
+We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
+Instead of rotating to a different projection, we might also change the view by moving the slice along one direction in the 4D space.
+Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$ to either large or small values of the bill depth ($\pm 1.5$ after centering and scaling).
+We will label these slices as $S_1^{\pm}$.
+Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
 
-Finally we use the manual tour on the random forest predictions first, to identify a centered slice with an interesting classification boundary.
-Here a combination of all four predictors is found to result in an interesting slice, where the predictions for the Gentoo penguins gets split into two regions of the slice, indicating the non-linearity of the random forest model.
-While x4 (body mass) is already contributing a lot in the projection, we can learn more by also moving the center point along that direction.
-Moving to high values (x4 = 2) results in a much larger part of the slice being predicted as Gentoo penguins, while now the yellow region (Chinstrap penguins) gets split.
-Moving in the opposite direction (x4 = -1) instead shows primarily Chinstrap penguins in the predictions, with Gentoo only appearing along the edge.
+We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Fig.
+\ref{slice1p}.
+We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
 
-Looking at the data, and shifting a fat slice (h = 1.5) as specified above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have observations from the Gentoo species, while x4 = 2 contains the Gentoo observations, but hardly any of the Chinstrap penguins.
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{...}.
+\label{slice1p}
+\end{figure*}
 
-## Comparison of boundaries produced by different classifiers
+A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Fig.
+\ref{slice1m}.
+The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
+On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
+Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
 
-The last step is comparing to the LDA predictions in those same slices.
-With RF we could see the non-linear aspects in the boundaries, while LDA can only produce linear boundaries, so these look very different.
-On the other hand, the same dependence is observed when shifting the slice along the x4 direction.
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
+\caption{...}.
+\label{slice1m}
+\end{figure*}
+
+Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
+\ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.
 

### 2022-07-08 paper/paper.tex
Worked out penguins example

@@
 We use the \texttt{classifly} package \citep{classifly} to generate
-predictions across the 4D space spanned by the data, with two
-classification models (linear discriminant analysis and random forest).
+predictions across the 4D cube spanned by the data, with two
+classification models: linear discriminant analysis (LDA) and random
+forest(RF).
 
@@
 
-The first projection we will start from is looking at bill depth vs
-flipper length (x2 vs x3). In a projection of the data we see that the
-Gentoo penguins are separate from the other two species which are on top
-of each other. If we instead look at a slice through the center of the
-distribution, we see that shifting the center point to large values of
-x1 (bill length), e.g.~x1=2, results in Adelie penguins dropping out of
-the slice, while a slice with low values (x1=-2) in the center point
-only contains Adelie penguin observations.
-
-We can next explore how this is mapped in the two classification models.
-Both random forest and LDA can separate out the region for Gentoo
-penguins in a thin slice through the center (h = 0.5), while Adelie and
-Chinstrap predictions overlap (blue and yellow, an effect of the slice
-thickness). We also notice the different boundaries: while LDA produces
-linear boundaries by construction, random forest insead produces
-``boxes'' from cutting up the data in the tree structure.
+We start by first exploring the projections of the model prediction.
+Through manual rotation of the view we can get a feeling for where in
+the space we primarily predict each of the three species, and we also
+get a sense of the difference between the two models. To illustrate this
+difference we have rotated the projection view from the RF model to
+identify a projection that shows the non-linear, block-type structure
+introduced by the model, see Fig. \ref{proj1}. We then export the
+corresponding matrix \(A_1\) to look at the same projection for the LDA
+model and the original data. Here we notice that while the overall
+picture is similar, the linearity of the LDA leads to a big difference
+in how two of the three species overlap across the projection (blue vs
+green, i.e.~Adelie vs Gentoo penguins). Interestingly that same
+projection leads to almost perfect separation of the three species when
+looking at the data points, the overlap between the models is driven by
+what is happening in the orthogonal space to the plane.
 
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.4\textwidth]{figures/view1_lda.png}
-\includegraphics[width=0.4\textwidth]{figures/view1_rf.png}}
-\caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model.
-The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinguishing features of the two models in terms of the different boundaries.}
-\label{view1}
+\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
+We can clearly see the block structure, while the LDA model (middle) produces linear boundaries instead.
+The three groups are separated in this projection of the data (right)}.
+\label{proj1}
 \end{figure*}
@@
 
-We can then start shifting the slice along x1 as we did when looking at
-the data, again moving to \(\pm 1\) will separate the Adelie from the
-Chinstrap penguins, now in terms of the model predictions. (Could
-include screenshot here as well.)
+XXXX currently in the notebook \(A_1\) is called A5, and \(A_2\) is
+called A6
+
+We continue the investigation by now slicing based on the projection
+\(A_1\). For both models we look at a thin (\(h=0.5\)) slice through the
+center, \(S_1^0\). As a first option we explore how manually changing
+the projection away from \(A_1\) can help with understanding the
+boundary better. For our example we notice that \(A_1\) does not contain
+any contribution from the second variable (\texttt{bd}), and in our
+illustration we will first rotate this variable into the view.
+
+We again start by looking at the RF model, the slice \(S_1^0\) shows the
+block structure of the model, with the third group (green, Gentoo
+penguins) overlapping with both of the other groups (Fig. \ref{slice1},
+top-left). This is similar for the LDA model (Fig. \ref{slice1},
+top-right), but again the linearity results in different boundaries and
+thus differences in how the classes overlap.
+
+By rotating in the second variable we can find a view that shows three
+neatly separated blocks for the RF model, and we export the
+corresponding projection matrix \(A_2\) to see the same views for the
+LDA model which also shows a neater separation of the classes along the
+boundary (see bottom row of Fig. \ref{slice1}).
+
+Finally we look at the data projection based on \(A_2\) and find that
+the third group (green, Gentoo penguins) is further away from the other
+two groups (compared to \(A_1\)), see Fig. \ref{proj2}.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices based on $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With the starting projection $A_1$ we see one group overlapping (green, Gentoo), while $A_2$ results in clear boundaries inside the slice.
+The block vs linear structure of the two models is clearly visible in both slices.}
+\label{slice1}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
+\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
+\label{proj2}
+\end{figure*}
 
@@
 
-Thus, stepping through the different slices shows the relation between
-the species and the first variable (x1, bill length). Instead of moving
-the center point we could also keep the slice fixed as passing through
-the center of the distribution, and use the manual controls to rotate x1
-into the projection. Here we started from the projected view of the data
-to identify a combination that can separate all three species, and then
-use the projection matrix as input to the slice displays for the model.
-Interestingly, the two models agree on how to classify the ``empty''
-regions (where there are no data points) for this example. (this is
-applying A3 as the starting projection).
+We have seen that starting from \(A_1\) using the manual controls to
+change the contribution of the second variable we could find a clear
+separation boundary indicating the relation between this variable (bill
+depth) and the Gentoo penguin species. Instead of rotating to a
+different projection, we might also change the view by moving the slice
+along one direction in the 4D space. Here we will continue our
+exploration of the dependence on \texttt{bd} and move the slice defined
+by \(A_1\) to either large or small values of the bill depth
+(\(\pm 1.5\) after centering and scaling). We will label these slices as
+\(S_1^{\pm}\). Here we will also look at slices of the observed data
+points, using a thicker slice (\(h=1.5\)) to capture enough points in a
+given view.
 
-Finally we use the manual tour on the random forest predictions first,
-to identify a centered slice with an interesting classification
-boundary. Here a combination of all four predictors is found to result
-in an interesting slice, where the predictions for the Gentoo penguins
-gets split into two regions of the slice, indicating the non-linearity
-of the random forest model. While x4 (body mass) is already contributing
-a lot in the projection, we can learn more by also moving the center
-point along that direction. Moving to high values (x4 = 2) results in a
-much larger part of the slice being predicted as Gentoo penguins, while
-now the yellow region (Chinstrap penguins) gets split. Moving in the
-opposite direction (x4 = -1) instead shows primarily Chinstrap penguins
-in the predictions, with Gentoo only appearing along the edge.
+We start by a comparison of the two models and the data distribution in
+\(S_1^{+}\), thus the slice is localized towards high values of bill
+depth in Fig. \ref{slice1p}. We can see that all three slices (the two
+models and the data) contain almost no points from the third class
+(green, Gentoo), and that the decision boundary between the two models
+is very similar.
 
-Looking at the data, and shifting a fat slice (h = 1.5) as specified
-above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have
-observations from the Gentoo species, while x4 = 2 contains the Gentoo
-observations, but hardly any of the Chinstrap penguins.
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{...}.
+\label{slice1p}
+\end{figure*}
 
-\hypertarget{comparison-of-boundaries-produced-by-different-classifiers}{%
-\subsection{Comparison of boundaries produced by different
-classifiers}\label{comparison-of-boundaries-produced-by-different-classifiers}}
+A more interesting comparison is found for \(S_1^{-}\), thus the slice
+localized towards low values of bill depth, shown in Fig. \ref{slice1m}.
+The RF model (left) predicts all three species within this slice, with
+an interesting boundary for the third class (green, Gentoo). On the
+other hand the LDA model (middle) predominantly predicts the third class
+within the slice, this appears to be enforced through the linear
+structure of the model. Looking finally at the thick slice through the
+data we see that there are primarily observations from this class within
+the slice we can conclude that the two models have filled in the
+``empty'' space (where we do not have any training observations) in very
+different ways and according to what we might expect given the model
+structure.
 
-The last step is comparing to the LDA predictions in those same slices.
-With RF we could see the non-linear aspects in the boundaries, while LDA
-can only produce linear boundaries, so these look very different. On the
-other hand, the same dependence is observed when shifting the slice
-along the x4 direction.
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
+\caption{...}.
+\label{slice1m}
+\end{figure*}
+
+Finally it is interesting to compare the slice views to the projection
+of the models seen in Fig. \ref{proj1} to better understand how the
+boundaries change along the \texttt{bd} direction and where the
+differences in the projections come from.
 

### 2022-07-11 paper/paper.Rmd
projections 4.2

@@
 
-# Application {#sec:examples}
-
-To illustrate the usefulness of the manual tour we use the 4D penguins data [@penguins].
-We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
-Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
-There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
-The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
-The steps for exploring boundaries are as follows:
-
-1. Build classification model, here we use a random forest and linear discriminant analysis.
-2. Predict the class for a dense grid of values covering the data space.
-3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
-4. Slice through the center, to explore where the boundaries will likely meet.
-5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
-
 ```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
@@
 
+# Application {#sec:examples}
+
+To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins].
+We will show how classification boundaries can be explored and better understood on projections and slices through 4D space.
+Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
+There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
+The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
+The steps for exploring boundaries in this example are as follows:
+
+1. Build your classification model. 
+2. Predict the class for a dense grid of values covering the data space.
+3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
+4. Slice through the center, to explore where the boundaries will likely meet.
+5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
+
 ## Constructing the 4D prediction regions
 
-We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest(RF).
+We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest (RF).
 
@@
 We start by first exploring the projections of the model prediction.
+Figure \ref{proj1} summarises the process.
 Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
-To illustrate this difference we have rotated the projection view from the RF model to identify a projection that shows the non-linear, block-type structure introduced by the model, see Fig.
-\ref{proj1}.
-We then export the corresponding matrix $A_1$ to look at the same projection for the LDA model and the original data.
-Here we notice that while the overall picture is similar, the linearity of the LDA leads to a big difference in how two of the three species overlap across the projection (blue vs green, i.e.
-Adelie vs Gentoo penguins).
-Interestingly that same projection leads to almost perfect separation of the three species when looking at the data points, the overlap between the models is driven by what is happening in the orthogonal space to the plane.
+To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a projection that shows the non-linear but block-type structure that is typical of this type of model.
+This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
+What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes.
+And, interestingly this particular projection of the original data shows very distinct clusters of the three species.
+That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
 

### 2022-07-11 paper/paper.tex
projections 4.2

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
+
+}
+
+\caption{Scatterplot matrix of the (standardised) penguins data.
+The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
+\end{figure}
+
 \hypertarget{sec:examples}{%
@@
 
-To illustrate the usefulness of the manual tour we use the 4D penguins
-data \citep{penguins}. We will show how classification boundaries can be
-explored and better understood using a manual tour, on projections and
+To illustrate the usefulness of the manual controls we use the 4D
+penguins data \citep{penguins}. We will show how classification
+boundaries can be explored and better understood on projections and
 slices through 4D space. Figure \ref{fig:penguins-scatmat} shows a
@@
 distinguished from the other two species when \texttt{bd} is paired with
-\texttt{fl} or \texttt{bm}. The steps for exploring boundaries are as
-follows:
+\texttt{fl} or \texttt{bm}. The steps for exploring boundaries in this
+example are as follows:
 
@@
 \item
-  Build classification model, here we use a random forest and linear
-  discriminant analysis.
+  Build your classification model.
 \item
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
-
-}
-
-\caption{Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
-\end{figure}
-
 \hypertarget{constructing-the-4d-prediction-regions}{%
@@
 classification models: linear discriminant analysis (LDA) and random
-forest(RF).
+forest (RF).
 
@@
 We start by first exploring the projections of the model prediction.
-Through manual rotation of the view we can get a feeling for where in
-the space we primarily predict each of the three species, and we also
-get a sense of the difference between the two models. To illustrate this
-difference we have rotated the projection view from the RF model to
-identify a projection that shows the non-linear, block-type structure
-introduced by the model, see Fig. \ref{proj1}. We then export the
-corresponding matrix \(A_1\) to look at the same projection for the LDA
-model and the original data. Here we notice that while the overall
-picture is similar, the linearity of the LDA leads to a big difference
-in how two of the three species overlap across the projection (blue vs
-green, i.e.~Adelie vs Gentoo penguins). Interestingly that same
-projection leads to almost perfect separation of the three species when
-looking at the data points, the overlap between the models is driven by
-what is happening in the orthogonal space to the plane.
+Figure \ref{proj1} summarises the process. Through manual rotation of
+the view we can get a feeling for where in the space we primarily
+predict each of the three species, and we also get a sense of the
+difference between the two models. To illustrate this difference we have
+manually rotated the projection for the RF model (left plot) to identify
+a projection that shows the non-linear but block-type structure that is
+typical of this type of model. This particular projection (\(A_1\)) is
+exported so that the same projection can be used to show the LDA model
+(middle plot) and the actual data (right plot). What can be seen is the
+linearity of the LDA model, where the boundaries are linear and oblique
+to the variable axes. And, interestingly this particular projection of
+the original data shows very distinct clusters of the three species.
+That means, the obscuring of the boundaries between groups for both of
+the models is driven by what is happening in the orthogonal space to the
+plane of the selected projection.
 

### 2022-07-11 paper/paper.Rmd
slice caption

@@
 \caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see the block structure, while the LDA model (middle) produces linear boundaries instead.
-The three groups are separated in this projection of the data (right)}.
+We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
+The three groups are nicely separated in this projection of the data (right)}.
 \label{proj1}
@@
 \includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With the starting projection $A_1$ we see one group overlapping (green, Gentoo), while $A_2$ results in clear boundaries inside the slice.
-The block vs linear structure of the two models is clearly visible in both slices.}
+\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
+The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
 \label{slice1}

### 2022-07-11 paper/appendix.Rmd
re-structured paper

--- 2022-07-08 paper/paper.Rmd
@@
 title: |
-  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
+  Appendix: New and simplified manual controls for projection and slice tours
 type: Short Technical Note
@@
 geometry: margin=2.5cm
-abstract: |
-  Something here
 keywords: |
@@
 <!-- \doublespacing-->
-# Introduction
 
-<!-- Background -->
-From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space.
-These are dynamic data visualizations methods and are collected under the term *tours*.
-Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
-In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
-Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
-PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-(A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, which are summarized in @lee2021.
-
-<!-- Manual tours -->
-One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour.
-@cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, or out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
-@spyrison_spinifex_2020 refined this algorithm and implements them to generate animation sequences.
-
-<!-- Purpose -->
-Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection.
-There are many places where it is useful.
-In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering.
-For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
-These controls can also be used to interactively generate faceted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
-Having the user interact with a projection is extremely valuable for understanding high-dimensional data.
-However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, (2) extending to higher dimensional control is difficult.
-
-Through experiments with the relatively new interactive graphics capabilities in mathematica(?), we have realized that there is a simpler approach, which is more direct, and extensible for generating user interaction.
-This paper explains this, and is organized as follows.
-The next section describes the new algorithm for manual control.
-This is followed by details on implementation.
-The applications section illustrate how these can be used.
-
-# How to construct a manual tour {#sec:method}
-
-A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d-D$ projection.
-The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed.
-A method to update the values of the component of the controlled variable $V_m$ is then needed.
-
-## Existing methods
-
-The method for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
-The first step was to construct a 3D manipulation space from a 2D projection, as illustrated in Figure \ref{manipspace}.
-In this space the coefficient of the controlled variable ranges between -1 and 1.
-Movements of a cursor are recorded and converted into changes in the values of $V_m$ thus changing the displayed 2D projection.
-The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
-\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
-(Figure 3 from Cook and Buja (1997).)}
-\label{manipspace}
-\end{figure*}
-
-## A new simpler and broadly applicable approach
-
-The new approach emerged from experiments in mathematica.
-The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
-The updated matrix is then orthonormalised.
-
-### Algorithm
-
-1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
-2.
-Change values in row $m$, giving $A^*$.
-A large change in these values would correspond to making a large jump from the current projection.
-Small changes would correspond to tracking a cursor, making small jumps from the current projection.
-3. Orthonormalise $A^*$, using Gram-Schmidt. For $d=2$, and $A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]$, the steps are:
-
-    i. Normalise ${\boldmath a}_{.1}$, and ${\boldmath a}_{.2}$.
-    ii. ${\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}$.
-    iii. Normalise ${\boldmath a}^*_{.2}$.
-
-This algorithm will produce the changes to a projection as illustrated in Figure \ref{fig:manualsequence} (top row).
-The controlled variable, $V_m$, corresponds to the black line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (orange dot).
-Changes in the other components happen as a result of the orthonormalisation, but are uncontrolled.
-
-```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
-The dot (orange) indicates the chosen values for the controlled variable.
-For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation."}
-source("../code/plot_basis.R")
-source("../code/linear_alg.R")
-
-# Base plot set up
-plot.new()
-par(pty="s", xaxt="n", yaxt="n", bty="n",
-      omi=c(0,0,0,0), mar=c(0,0,0,0),
-    mfrow=c(2,4))
-
-# Create manual tour example
-p <- 4
-d <- 2
-set.seed(24)
-A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
-colnames(A) <- c("P1", "P2")
-rownames(A) <- paste0("V", 1:p)
-A <- tourr::orthonormalise(A)
-
-# Now set a new position
-vchange <- 3
-eps <- 0.1
-plot_basis(A, vchange, vcol="black")
-
-# Iterate and save images
-# Unconstrained orthonormalisation
-Anew <- A
-for (i in 1:3) {
-  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-  exact <- Anew[vchange,]
-  Anew <- tourr::orthonormalise(Anew)
-  plot_basis(Anew, vchange, vcol="black")
-  points(exact[1], exact[2], pch=16, col="orange")
-}
-
-# Iterate and save images
-# Constrained orthonormalisation
-Anew <- A
-plot_basis(A, vchange, vcol="black")
-for (i in 1:3) {
-  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-  exact <- Anew[vchange,]
-  Anew <- orthonormalise_frozen(Anew, vchange)
-  plot_basis(Anew, vchange, vcol="black")
-  points(exact[1], exact[2], pch=16, col="orange")
-}
-```
-
-## Refinements to enforce exact position
+# Refinements to enforce exact position
 
@@
 
-### Adjustment method 1
+## Adjustment method 1
 
-A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)). It is as follows:
+A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:othermethod}. It is as follows:
 
@@
 
-### Adjustment method 2
-
-For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
-Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
-Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
-Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
-One possible algorithm to achieve this is
-\begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
-\begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
-\end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
-\end{enumerate}
-The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
-
-\begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
-\CodeBefore
-       \columncolor{red!15}{1,2}
-              \rowcolor{blue!15}{4}
-     \Body
-       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
-e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
-e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
- e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
-e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
-\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
-\end{align}
-
-### Adjustment method 3
-
-For now just sketching the idea:
-
-- first step is to click on the axis display to change the contribution of one variable $m$
-- capture that position and replace the corresponding row in the projection matrix
-- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
-- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
-- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
-- rotated back to the original xy basis to update the plots
-
-XXX need to check calculation of angle when $m_1 < 0$
-
-```{r othermethod, eval=FALSE, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
+```{r othermethod, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
 source("../code/plot_basis.R")
@@
 
-# Experimenting with new techniques using Mathematica {#sec:implementation}
+## Adjustment method 2
+
+For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
+Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
+Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+One possible algorithm to achieve this is
+\begin{enumerate}
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
+\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
+\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\end{enumerate}
+\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\end{enumerate}
+The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
+
+\begin{align}
+A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
+e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
+e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
+\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
+\end{align}
+
+## Adjustment method 3
+
+For now just sketching the idea:
+
+- first step is to click on the axis display to change the contribution of one variable $m$
+- capture that position and replace the corresponding row in the projection matrix
+- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
+- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
+- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
+- rotated back to the original xy basis to update the plots
+
+XXX need to check calculation of angle when $m_1 < 0$
+
+# Software details
 
@@
     - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
-
-# Application {#sec:examples}
-
-To illustrate the usefulness of the manual tour we use the 4D penguins data [@penguins].
-We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.
-Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
-There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
-The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
-The steps for exploring boundaries are as follows:
-
-1. Build classification model, here we use a random forest and linear discriminant analysis.
-2. Predict the class for a dense grid of values covering the data space.
-3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
-4. Slice through the center, to explore where the boundaries will likely meet.
-5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
-
-```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
-library(tidyverse)
-library(palmerpenguins)
-library(GGally)
-penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
-  drop_na() %>%
-  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
-  rename("bl"="bill_length_mm",
-         "bd"="bill_depth_mm",
-         "fl"="flipper_length_mm",
-         "bm"="body_mass_g")
-ggscatmat(penguins_data, columns = 2:5, col="species") +
-  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-  theme(legend.position = "bottom")
-```
-
-## Constructing the 4D prediction regions
-
-We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest(RF).
-
-## Exploring projections manually
-
-We start by first exploring the projections of the model prediction.
-Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
-To illustrate this difference we have rotated the projection view from the RF model to identify a projection that shows the non-linear, block-type structure introduced by the model, see Fig.
-\ref{proj1}.
-We then export the corresponding matrix $A_1$ to look at the same projection for the LDA model and the original data.
-Here we notice that while the overall picture is similar, the linearity of the LDA leads to a big difference in how two of the three species overlap across the projection (blue vs green, i.e.
-Adelie vs Gentoo penguins).
-Interestingly that same projection leads to almost perfect separation of the three species when looking at the data points, the overlap between the models is driven by what is happening in the orthogonal space to the plane.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see the block structure, while the LDA model (middle) produces linear boundaries instead.
-The three groups are separated in this projection of the data (right)}.
-\label{proj1}
-\end{figure*}
-
-## Slicing through the center
-
-XXXX currently in the notebook $A_1$ is called A5, and $A_2$ is called A6
-
-We continue the investigation by now slicing based on the projection $A_1$.
-For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0$.
-As a first option we explore how manually changing the projection away from $A_1$ can help with understanding the boundary better.
-For our example we notice that $A_1$ does not contain any contribution from the second variable (`bd`), and in our illustration we will first rotate this variable into the view.
-
-We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
-\ref{slice1}, top-left).
-This is similar for the LDA model (Fig.
-\ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
-
-By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
-\ref{slice1}).
-
-Finally we look at the data projection based on $A_2$ and find that the third group (green, Gentoo penguins) is further away from the other two groups (compared to $A_1$), see Fig. \ref{proj2}.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With the starting projection $A_1$ we see one group overlapping (green, Gentoo), while $A_2$ results in clear boundaries inside the slice.
-The block vs linear structure of the two models is clearly visible in both slices.}
-\label{slice1}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
-\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
-\label{proj2}
-\end{figure*}
-
-## Shifting the slice center
-
-We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
-Instead of rotating to a different projection, we might also change the view by moving the slice along one direction in the 4D space.
-Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$ to either large or small values of the bill depth ($\pm 1.5$ after centering and scaling).
-We will label these slices as $S_1^{\pm}$.
-Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
-
-We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Fig.
-\ref{slice1p}.
-We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{...}.
-\label{slice1p}
-\end{figure*}
-
-A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Fig.
-\ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
-On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
-Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{...}.
-\label{slice1m}
-\end{figure*}
-
-Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
-\ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.
-
-# Extension added in the R package `tourr`
-
-Explain `radial_tour`
-
-# What would be desirable for implementations in R?
-
-# Discussion {#sec:discussion}
-
-# Acknowledgements {-}
-
-The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
-
-# Supplementary material {-}
-
-The source material and animated gifs for this paper are available at

### 2022-07-11 paper/appendix.tex
re-structured paper

--- 2022-07-08 paper/paper.tex
@@
 
-\title{New and simplified manual controls for projection and slice
-tours, with application to exploring classification boundaries in high
-dimensions}
+\title{Appendix: New and simplified manual controls for projection and
+slice tours}
 
@@
 
-\begin{abstract}
-Something here
-\end{abstract}
-
 \begin{keywords}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-From a statistical perspective it is rare to have data that are strictly
-3D, and so unlike in most computer graphics applications, the more
-useful methods for data analysis show projections from an arbitrary
-dimensional space. These are dynamic data visualizations methods and are
-collected under the term \emph{tours}. Tours involve views of
-high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
-In his original paper on the grand tour, \citet{As85} provided several
-algorithms for tour paths that could theoretically show the viewer the
-data \emph{from all sides}. Prior to Asimov's work, there were numerous
-preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
-user-controlled rotations on coordinate axes, allowing one to manually
-tour through low-dimensional projections. (A video illustrating the
-capabilities is available through video library of \citet{ASA22}.)
-Steering through all possible projections is impossible, unlike Asimov's
-tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, which are
-summarized in \citet{lee2021}.
-
-One such direction of work develops the ideas from PRIM-9, to provide
-manual control of a tour. \citet{cook_manual_1997} describe controls for
-1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space,
-allowing the user to select any variable axis, and rotate it into, or
-out of, or around the projection through horizontal, vertical, oblique,
-radial or angular changes in value. \citet{spyrison_spinifex_2020}
-refined this algorithm and implements them to generate animation
-sequences.
-
-Manual controls are especially useful for assessing sensitivity of
-structure to particular elements of the projection. There are many
-places where it is useful. In exploratory data analysis, where one sees
-clusters in a projection, can some variables be removed from the
-projection without affecting the clustering. For interpreting models,
-one can reduce or increase a variable's contribution to examine the
-variable importance. These controls can also be used to interactively
-generate faceted plots \citep{XXX}, or spatiotemporal glyphmaps
-\citep{XXX}. Having the user interact with a projection is extremely
-valuable for understanding high-dimensional data. However, these
-algorithms have two problems: (1) the pre-processing of creating a
-manipulation space overly complicates the algorithm, (2) extending to
-higher dimensional control is difficult.
-
-Through experiments with the relatively new interactive graphics
-capabilities in mathematica(?), we have realized that there is a simpler
-approach, which is more direct, and extensible for generating user
-interaction. This paper explains this, and is organized as follows. The
-next section describes the new algorithm for manual control. This is
-followed by details on implementation. The applications section
-illustrate how these can be used.
-
-\hypertarget{sec:method}{%
-\section{How to construct a manual tour}\label{sec:method}}
-
-A manual tour allows the user to alter the coefficients of one (or more)
-variables contributing to a \(d-D\) projection. The initial ingredients
-are an orthonormal basis (\(A_{p\times d}\)) defining the projection of
-the data, and a variable id (\(m \in \{1, ..., p\}\)) specifying which
-coefficient will be changed. A method to update the values of the
-component of the controlled variable \(V_m\) is then needed.
-
-\hypertarget{existing-methods}{%
-\subsection{Existing methods}\label{existing-methods}}
-
-The method for updating component values in \citet{cook_manual_1997}
-(and utilised in \citet{spyrison_spinifex_2020}) are prescribed
-primarily for a 2D projection, to take advantage of (then) newly
-developed 3D trackball controls made available for computer gaming. The
-first step was to construct a 3D manipulation space from a 2D
-projection, as illustrated in Figure \ref{manipspace}. In this space the
-coefficient of the controlled variable ranges between -1 and 1.
-Movements of a cursor are recorded and converted into changes in the
-values of \(V_m\) thus changing the displayed 2D projection. The
-algorithm also provided constraints to horizontal, vertical, radial or
-angular motions only.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
-\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
-(Figure 3 from Cook and Buja (1997).)}
-\label{manipspace}
-\end{figure*}
-
-\hypertarget{a-new-simpler-and-broadly-applicable-approach}{%
-\subsection{A new simpler and broadly applicable
-approach}\label{a-new-simpler-and-broadly-applicable-approach}}
-
-The new approach emerged from experiments in mathematica. The components
-corresponding to \(V_m\) are directly controlled by cursor movement,
-which updates row \(m\) of \(A\). The updated matrix is then
-orthonormalised.
-
-\hypertarget{algorithm}{%
-\subsubsection{Algorithm}\label{algorithm}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Provide \(A\), and \(m\). (Note that \(m\) could also be automatically
-  chosen as the component that is closest to the cursor position.)
-\item
-  Change values in row \(m\), giving \(A^*\). A large change in these
-  values would correspond to making a large jump from the current
-  projection. Small changes would correspond to tracking a cursor,
-  making small jumps from the current projection.
-\item
-  Orthonormalise \(A^*\), using Gram-Schmidt. For \(d=2\), and
-  \(A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]\), the
-  steps are:
-
-  \begin{enumerate}
-  \def\labelenumii{\roman{enumii}.}
-  \tightlist
-  \item
-    Normalise \({\boldmath a}_{.1}\), and \({\boldmath a}_{.2}\).
-  \item
-    \({\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}\).
-  \item
-    Normalise \({\boldmath a}^*_{.2}\).
-  \end{enumerate}
-\end{enumerate}
-
-This algorithm will produce the changes to a projection as illustrated
-in Figure \ref{fig:manualsequence} (top row). The controlled variable,
-\(V_m\), corresponds to the black line, and sequential changes to row
-\(m\) of \(A\) can be seen to roughly follow a specified position
-(orange dot). Changes in the other components happen as a result of the
-orthonormalisation, but are uncontrolled.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
-The dot (orange) indicates the chosen values for the controlled variable.
-For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation.}\label{fig:manualsequence}
-\end{figure}
-
 \hypertarget{refinements-to-enforce-exact-position}{%
-\subsection{Refinements to enforce exact
+\section{Refinements to enforce exact
 position}\label{refinements-to-enforce-exact-position}}
@@
 \hypertarget{adjustment-method-1}{%
-\subsubsection{Adjustment method 1}\label{adjustment-method-1}}
+\subsection{Adjustment method 1}\label{adjustment-method-1}}
 
 A small modification to algorithm 1 will maintain the components of
-\(V_m\) precisely (Figure \ref{fig:manualsequence} (bottom row)). It is
-as follows:
+\(V_m\) precisely (Figure \ref{fig:othermethod}. It is as follows:
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{appendix_files/figure-latex/othermethod-1} \caption{Manual controls with algorithm 2. The precise location of the axis is maintained.}\label{fig:othermethod}
+\end{figure}
+
 \hypertarget{adjustment-method-2}{%
-\subsubsection{Adjustment method 2}\label{adjustment-method-2}}
+\subsection{Adjustment method 2}\label{adjustment-method-2}}
 
@@
 \hypertarget{adjustment-method-3}{%
-\subsubsection{Adjustment method 3}\label{adjustment-method-3}}
+\subsection{Adjustment method 3}\label{adjustment-method-3}}
 
@@
 
-\hypertarget{sec:implementation}{%
-\section{Experimenting with new techniques using
-Mathematica}\label{sec:implementation}}
+\hypertarget{software-details}{%
+\section{Software details}\label{software-details}}
 
@@
 
-\hypertarget{sec:examples}{%
-\section{Application}\label{sec:examples}}
-
-To illustrate the usefulness of the manual tour we use the 4D penguins
-data \citep{penguins}. We will show how classification boundaries can be
-explored and better understood using a manual tour, on projections and
-slices through 4D space. Figure \ref{fig:penguins-scatmat} shows a
-scatterplot matrix of this data. There are four variables
-(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
-measuring the size of the penguins from three species (Adelie, Chinstrap
-and Gentoo). The scatterplot matrix shows that the three species appear
-to be likely separable, and that at least the Gentoo can be
-distinguished from the other two species when \texttt{bd} is paired with
-\texttt{fl} or \texttt{bm}. The steps for exploring boundaries are as
-follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Build classification model, here we use a random forest and linear
-  discriminant analysis.
-\item
-  Predict the class for a dense grid of values covering the data space.
-\item
-  Examine projections, using a manual tour so that the contribution of
-  any variable is controlled.
-\item
-  Slice through the center, to explore where the boundaries will likely
-  meet.
-\item
-  Move the slice by changing the center in the direction of a single
-  variable to explore the extent of a boundary for a single group
-  relative to a variable.
-\end{enumerate}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
-
-}
-
-\caption{Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
-\end{figure}
-
-\hypertarget{constructing-the-4d-prediction-regions}{%
-\subsection{Constructing the 4D prediction
-regions}\label{constructing-the-4d-prediction-regions}}
-
-We use the \texttt{classifly} package \citep{classifly} to generate
-predictions across the 4D cube spanned by the data, with two
-classification models: linear discriminant analysis (LDA) and random
-forest(RF).
-
-\hypertarget{exploring-projections-manually}{%
-\subsection{Exploring projections
-manually}\label{exploring-projections-manually}}
-
-We start by first exploring the projections of the model prediction.
-Through manual rotation of the view we can get a feeling for where in
-the space we primarily predict each of the three species, and we also
-get a sense of the difference between the two models. To illustrate this
-difference we have rotated the projection view from the RF model to
-identify a projection that shows the non-linear, block-type structure
-introduced by the model, see Fig. \ref{proj1}. We then export the
-corresponding matrix \(A_1\) to look at the same projection for the LDA
-model and the original data. Here we notice that while the overall
-picture is similar, the linearity of the LDA leads to a big difference
-in how two of the three species overlap across the projection (blue vs
-green, i.e.~Adelie vs Gentoo penguins). Interestingly that same
-projection leads to almost perfect separation of the three species when
-looking at the data points, the overlap between the models is driven by
-what is happening in the orthogonal space to the plane.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see the block structure, while the LDA model (middle) produces linear boundaries instead.
-The three groups are separated in this projection of the data (right)}.
-\label{proj1}
-\end{figure*}
-
-\hypertarget{slicing-through-the-center}{%
-\subsection{Slicing through the
-center}\label{slicing-through-the-center}}
-
-XXXX currently in the notebook \(A_1\) is called A5, and \(A_2\) is
-called A6
-
-We continue the investigation by now slicing based on the projection
-\(A_1\). For both models we look at a thin (\(h=0.5\)) slice through the
-center, \(S_1^0\). As a first option we explore how manually changing
-the projection away from \(A_1\) can help with understanding the
-boundary better. For our example we notice that \(A_1\) does not contain
-any contribution from the second variable (\texttt{bd}), and in our
-illustration we will first rotate this variable into the view.
-
-We again start by looking at the RF model, the slice \(S_1^0\) shows the
-block structure of the model, with the third group (green, Gentoo
-penguins) overlapping with both of the other groups (Fig. \ref{slice1},
-top-left). This is similar for the LDA model (Fig. \ref{slice1},
-top-right), but again the linearity results in different boundaries and
-thus differences in how the classes overlap.
-
-By rotating in the second variable we can find a view that shows three
-neatly separated blocks for the RF model, and we export the
-corresponding projection matrix \(A_2\) to see the same views for the
-LDA model which also shows a neater separation of the classes along the
-boundary (see bottom row of Fig. \ref{slice1}).
-
-Finally we look at the data projection based on \(A_2\) and find that
-the third group (green, Gentoo penguins) is further away from the other
-two groups (compared to \(A_1\)), see Fig. \ref{proj2}.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With the starting projection $A_1$ we see one group overlapping (green, Gentoo), while $A_2$ results in clear boundaries inside the slice.
-The block vs linear structure of the two models is clearly visible in both slices.}
-\label{slice1}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
-\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
-\label{proj2}
-\end{figure*}
-
-\hypertarget{shifting-the-slice-center}{%
-\subsection{Shifting the slice center}\label{shifting-the-slice-center}}
-
-We have seen that starting from \(A_1\) using the manual controls to
-change the contribution of the second variable we could find a clear
-separation boundary indicating the relation between this variable (bill
-depth) and the Gentoo penguin species. Instead of rotating to a
-different projection, we might also change the view by moving the slice
-along one direction in the 4D space. Here we will continue our
-exploration of the dependence on \texttt{bd} and move the slice defined
-by \(A_1\) to either large or small values of the bill depth
-(\(\pm 1.5\) after centering and scaling). We will label these slices as
-\(S_1^{\pm}\). Here we will also look at slices of the observed data
-points, using a thicker slice (\(h=1.5\)) to capture enough points in a
-given view.
-
-We start by a comparison of the two models and the data distribution in
-\(S_1^{+}\), thus the slice is localized towards high values of bill
-depth in Fig. \ref{slice1p}. We can see that all three slices (the two
-models and the data) contain almost no points from the third class
-(green, Gentoo), and that the decision boundary between the two models
-is very similar.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{...}.
-\label{slice1p}
-\end{figure*}
-
-A more interesting comparison is found for \(S_1^{-}\), thus the slice
-localized towards low values of bill depth, shown in Fig. \ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with
-an interesting boundary for the third class (green, Gentoo). On the
-other hand the LDA model (middle) predominantly predicts the third class
-within the slice, this appears to be enforced through the linear
-structure of the model. Looking finally at the thick slice through the
-data we see that there are primarily observations from this class within
-the slice we can conclude that the two models have filled in the
-``empty'' space (where we do not have any training observations) in very
-different ways and according to what we might expect given the model
-structure.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{...}.
-\label{slice1m}
-\end{figure*}
-
-Finally it is interesting to compare the slice views to the projection
-of the models seen in Fig. \ref{proj1} to better understand how the
-boundaries change along the \texttt{bd} direction and where the
-differences in the projections come from.
-
-\hypertarget{extension-added-in-the-r-package-tourr}{%
-\section{\texorpdfstring{Extension added in the R package
-\texttt{tourr}}{Extension added in the R package tourr}}\label{extension-added-in-the-r-package-tourr}}
-
-Explain \texttt{radial\_tour}
-
-\hypertarget{what-would-be-desirable-for-implementations-in-r}{%
-\section{What would be desirable for implementations in
-R?}\label{what-would-be-desirable-for-implementations-in-r}}
-
-\hypertarget{sec:discussion}{%
-\section{Discussion}\label{sec:discussion}}
-
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
-
-The authors gratefully acknowledge the support of the Australian
-Research Council. The paper was written in \texttt{rmarkdown}
-\citep{rmarkdown} using \texttt{knitr} \citep{knitr}.
-
-\hypertarget{supplementary-material}{%
-\section*{Supplementary material}\label{supplementary-material}}
-\addcontentsline{toc}{section}{Supplementary material}
-
-The source material and animated gifs for this paper are available at
-
 \bibliographystyle{tfcad}

### 2022-07-11 paper/paper.Rmd
re-structured paper

@@
 The method for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
-The first step was to construct a 3D manipulation space from a 2D projection, as illustrated in Figure \ref{manipspace}.
+The first step was to construct a 3D manipulation space from a 2D projection.
 In this space the coefficient of the controlled variable ranges between -1 and 1.
@@
 The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.
+The construction of the manipulation space overly complicates the manual controls, especially when considering a technique that will generally apply to arbitrary $d$.
 
+<!--
 \begin{figure*}[ht]
@@
 \end{figure*}
+-->
 
@@
 
-```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
 The dot (orange) indicates the chosen values for the controlled variable.
-For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation."}
+It can be seen that the actual axis does not precisely match the chosen position, but it is close."}
 source("../code/plot_basis.R")
@@
 # Constrained orthonormalisation
-Anew <- A
-plot_basis(A, vchange, vcol="black")
-for (i in 1:3) {
-  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-  exact <- Anew[vchange,]
-  Anew <- orthonormalise_frozen(Anew, vchange)
-  plot_basis(Anew, vchange, vcol="black")
-  points(exact[1], exact[2], pch=16, col="orange")
-}
+#Anew <- A
+#plot_basis(A, vchange, vcol="black")
+#for (i in 1:3) {
+#  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+#  exact <- Anew[vchange,]
+#  Anew <- orthonormalise_frozen(Anew, vchange)
+#  plot_basis(Anew, vchange, vcol="black")
+#  points(exact[1], exact[2], pch=16, col="orange")
+#}
 ```
@@
 
-The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
+The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ are not followed because the orthonormalisation will change them.
+There are numerous ways that this can be enforced, and details are provided in the Appendix.
 
-### Adjustment method 1
+## Manual control for slice center
 
-A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)). It is as follows:
+How does it work?
 
-1. Provide $A$, and $m$.
-2. Change values in row $m$, giving $A^*$.
-3. Store row $m$ separately, and zero the values of row $m$ in $A^*$, giving $A^{*0}$. 
-4. Orthonormalise $A^{*0}$, using Gram-Schmidt.
-5. Replace row $m$ with the original values, giving $A^{**}$. 
-6. For $d=2$, adjust the values of ${\boldmath a}^{**}_{.2}$ using
-
-$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m$$.
-
-which ensures that
-
-$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0$$.
-
-If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
-If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
-
-<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
-
-Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
-
-### Adjustment method 2
-
-For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
-Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
-Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
-Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
-One possible algorithm to achieve this is
-\begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
-\begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
-\end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
-\end{enumerate}
-The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
-
-\begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
-\CodeBefore
-       \columncolor{red!15}{1,2}
-              \rowcolor{blue!15}{4}
-     \Body
-       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
-e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
-e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
- e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
-e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
-\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
-\end{align}
-
-### Adjustment method 3
-
-For now just sketching the idea:
-
-- first step is to click on the axis display to change the contribution of one variable $m$
-- capture that position and replace the corresponding row in the projection matrix
-- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
-- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
-- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
-- rotated back to the original xy basis to update the plots
-
-XXX need to check calculation of angle when $m_1 < 0$
-
-```{r othermethod, eval=FALSE, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
-source("../code/plot_basis.R")
-source("../code/linear_alg.R")
-
-# Base plot set up
-plot.new()
-par(pty="s", xaxt="n", yaxt="n", bty="n",
-      omi=c(0,0,0,0), mar=c(0,0,0,0),
-    mfrow=c(1,4))
-
-# Create manual tour example
-p <- 4
-d <- 2
-set.seed(24)
-A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
-colnames(A) <- c("P1", "P2")
-rownames(A) <- paste0("V", 1:p)
-A <- tourr::orthonormalise(A)
-
-# Now set a new position
-vchange <- 3
-eps <- 0.1
-plot_basis(A, vchange, vcol="black")
-
-# j is the component that should be kept fixed
-orthonormalize_rotated <- function(A, j){
-  m1 <- A[j, 1]
-  m2 <- A[j, 2]
-  c <- 1 / (sqrt(1 + (m2/m1)^2))
-  s <- (m2/m1) / (sqrt(1 + (m2/m1)^2))
-  rot <- matrix(c(c, -s, s, c), nrow = 2)
-  rot_inv <- matrix(c(c, s, -s, c), nrow = 2)
-  for(i in 1:nrow(A)){
-    A[i,] <- rot %*% A[i,]
-  }
-  A[,1] <- normalize_w_fixed(A[,1], j)
-  A <- tourr::orthonormalise(A)
-  for(i in 1:nrow(A)){
-    A[i,] <- rot_inv %*% A[i,]
-    }
-  A
-}
-
-# normalize a vector while keeping component j fixed
-normalize_w_fixed <- function(v, j){
-  c <- v[j]
-  k <- sqrt((1-c^2) / sum(v[-j]^2))
-  v[-j] <- k * v[-j]
-  v
-}
-
-# Iterate and save images
-# Unconstrained orthonormalisation
-Anew <- A
-for (i in 1:3) {
-  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-  exact <- Anew[vchange,]
-  Anew <- orthonormalize_rotated(Anew, vchange)
-  plot_basis(Anew, vchange, vcol="black")
-  points(exact[1], exact[2], pch=16, col="orange")
-}
-
-```
+Diagram of interface? Scaling
 
@@
 
+<!-- 
 The following functions are implemented in the Mathematica package `mmtour.wl`:
@@
     - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
+-->
 

### 2022-07-11 paper/paper.tex
re-structured paper

@@
 first step was to construct a 3D manipulation space from a 2D
-projection, as illustrated in Figure \ref{manipspace}. In this space the
-coefficient of the controlled variable ranges between -1 and 1.
-Movements of a cursor are recorded and converted into changes in the
-values of \(V_m\) thus changing the displayed 2D projection. The
-algorithm also provided constraints to horizontal, vertical, radial or
-angular motions only.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
-\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
-(Figure 3 from Cook and Buja (1997).)}
-\label{manipspace}
-\end{figure*}
+projection. In this space the coefficient of the controlled variable
+ranges between -1 and 1. Movements of a cursor are recorded and
+converted into changes in the values of \(V_m\) thus changing the
+displayed 2D projection. The algorithm also provided constraints to
+horizontal, vertical, radial or angular motions only. The construction
+of the manipulation space overly complicates the manual controls,
+especially when considering a technique that will generally apply to
+arbitrary \(d\).
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed: (top) unconstrained orthonormalisation, (bottom) constrained as specified.
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
 The dot (orange) indicates the chosen values for the controlled variable.
-For the constrained orthonormalisation it can be seen to precisely match the axis, but not so for the unconstrained orthonormalisation.}\label{fig:manualsequence}
+It can be seen that the actual axis does not precisely match the chosen position, but it is close.}\label{fig:manualsequence}
 \end{figure}
@@
 The problem with new simple method (Algorithm 1) is that the precise
-values for \(V_m\) cannot be specified because the orthonormalisation
-will change them.
+values for \(V_m\) are not followed because the orthonormalisation will
+change them. There are numerous ways that this can be enforced, and
+details are provided in the Appendix.
 
-\hypertarget{adjustment-method-1}{%
-\subsubsection{Adjustment method 1}\label{adjustment-method-1}}
+\hypertarget{manual-control-for-slice-center}{%
+\subsection{Manual control for slice
+center}\label{manual-control-for-slice-center}}
 
-A small modification to algorithm 1 will maintain the components of
-\(V_m\) precisely (Figure \ref{fig:manualsequence} (bottom row)). It is
-as follows:
+How does it work?
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Provide \(A\), and \(m\).
-\item
-  Change values in row \(m\), giving \(A^*\).
-\item
-  Store row \(m\) separately, and zero the values of row \(m\) in
-  \(A^*\), giving \(A^{*0}\).
-\item
-  Orthonormalise \(A^{*0}\), using Gram-Schmidt.
-\item
-  Replace row \(m\) with the original values, giving \(A^{**}\).
-\item
-  For \(d=2\), adjust the values of \({\boldmath a}^{**}_{.2}\) using
-\end{enumerate}
-
-\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m\].
-
-which ensures that
-
-\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0\].
-
-If \(d>2\) the process would be sequentially repeated in the same manner
-that Gram-Schmidt is applied sequentially to orthormalise the columns of
-a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
-vector would simply need to be normalized after each adjustment.
-
-\hypertarget{adjustment-method-2}{%
-\subsubsection{Adjustment method 2}\label{adjustment-method-2}}
-
-For \(d=2\) projections, the projection matrix is the sub-matrix of
-\(A\) formed by its first two columns. Whereas orthonormality of the
-basis for the p-dimensional space is given by
-\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}\), orthonormality of the
-projection matrix is expressed as
-\(P_i\cdot P_j=\delta_{ij}, {i,j=1,2}\). Movement of the cursor takes
-the two components \({x_{m1},x_{m2}}\) into a selected new value
-\({a,b}\). Although the motion is constrained by \(a^2+b^2\leq 1\), this
-is not sufficient to guarantee orthonormality of the new projection
-matrix. One possible algorithm to achieve this is
-
-\begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
-\begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
-\end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
-\end{enumerate}
-
-The random choice for the components of \(A\) not in the projection
-matrix allows the exploration of dimensions perpendicular to the
-projection plane.
-
-\begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
-\CodeBefore
-       \columncolor{red!15}{1,2}
-              \rowcolor{blue!15}{4}
-     \Body
-       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
-e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
-e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
- e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
-\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
-e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
-\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
-\end{align}
-
-\hypertarget{adjustment-method-3}{%
-\subsubsection{Adjustment method 3}\label{adjustment-method-3}}
-
-For now just sketching the idea:
-
-\begin{itemize}
-\tightlist
-\item
-  first step is to click on the axis display to change the contribution
-  of one variable \(m\)
-\item
-  capture that position and replace the corresponding row in the
-  projection matrix
-\item
-  for 2D projection this gives components \(m_1\) along ``x'' direction
-  and \(m_2\) along \(y\) direction
-\item
-  next step: rotate basis such that direction of row m now corresponds
-  to the first basis vector, this means we apply 2x2 rotation matrix to
-  each row of the projection matrix, where the rotation angle is
-  \(tan(\theta) = m_2/m_1\) (that matrix can be written just in terms of
-  that ratio, \(x= m_2/m_1\), when \(m_1<0\) I think we just need to
-  translate \(\theta \rightarrow \theta + \pi\))
-\item
-  take the rotated basis, apply Gram-Schmidt, now the direction of \(m\)
-  will not change (but the length will change during normalization)
-\item
-  rotated back to the original xy basis to update the plots
-\end{itemize}
-
-XXX need to check calculation of angle when \(m_1 < 0\)
+Diagram of interface? Scaling
 
@@
   \textbf{Explain the functionality available in the notebooks}
-\end{itemize}
-
-The following functions are implemented in the Mathematica package
-\texttt{mmtour.wl}:
-
-\begin{itemize}
-\item
-  \texttt{ProjectionPlot{[}data,\ projmat,\ xRange(=Automatic),\ yRange(=Automatic),\ \ colour(=Automatic){]}}
-  with arguments
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \texttt{data}: A list of data matrices (they must all have the same
-    dimensionality)
-  \item
-    `\texttt{projmat}: A projection matrix which describes the
-    projection plane. This is the initial projection and does not need
-    to be orthonormal (XXX is this right?).
-  \item
-    `\texttt{xRange}: A list which details the range of the horizontal
-    axis.
-  \item
-    `\texttt{yRange}: A list which details the range of the vertical
-    axis.
-  \item
-    `\texttt{colour}: A list of graphics directives to be applied to
-    each data matrix.
-  \end{itemize}
-\end{itemize}
-
-There are problems with the default values for x and y ranges and
-colour. It would be better to fix the ranges and colour based on the
-data as well as the origin of axes and the size of the points.
-
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data? Fix 1}
-
-The format for the printed projection matrix can be improved also. Maybe
-fix the number of digits and the position of the label.
-
-\begin{itemize}
-\item
-  \texttt{SlicePlot{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1(=0.005),\ ptSize2(=0.004),\ minDist(=0){]}}
-  with arguments (used to be called SliceDynamic in Alex's first
-  version, the name SliceDynamic was taken by something else below)
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \texttt{data}: A data matrix
-  \item
-    \texttt{projmat}: The inital projection matrix
-  \item
-    \texttt{centrepoint}: The initial position of the slice
-  \item
-    \texttt{height}: The initial height of the slice
-  \item
-    \texttt{heightRange}: The range of heights
-  \item
-    \texttt{ptSize1}: The size of the points that exist within the slice
-  \item
-    \texttt{ptSize2}: The size of the points that exist outside of the
-    slice
-  \end{itemize}
-\end{itemize}
-
-This function requires some auxiliary functions. It does not accept more
-than one data set.
-
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data?}
-
-It may be good to fix the size of the points, the center of the axes and
-the scales of the axes rather than allow them to change as the
-projection and slicing changes
-
-\begin{itemize}
-\item
-  \texttt{SliceDynamic{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1\ :\ 0.005,\ ptSize2\ :\ 0.004{]}}
-  with arguments
-\item
-  \texttt{data}: A data matrix with an extra grouping or clustering
-  column (numerical)
-\item
-  \texttt{projmat}: The initial projection matrix
-\item
-  \texttt{centrepoint}: The initial position of the slice
-\item
-  \texttt{height}: The initial height of the slice
-\item
-  \texttt{heightRange}: The range of heights
-\item
-  \texttt{ptSize1}: The size of the points that exist within the slice
-\item
-  \texttt{ptSize2}: The size of the points that exist outside of the
-  slice
-\end{itemize}
-
-\textbf{Does this need to be a separate function from SlicePlot? Fix 2}
-
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data? Fix 3}
-
-\textbf{It may be good to fix the size of the points, the center of the
-axes and the scales of the axes rather than allow them to change as the
-projection and slicing changes. Fix 4}
-
-\textbf{I have found it useful to input different data sets rather than
-have this function sort them automatically using the last column. One
-data set with a flag column at the end, is best! Is the code too
-complicated? Fix 5}
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{Projected2DSliderPlot{[}{]}}
-\end{itemize}
-
-This function uses a different display for the projection that I find
-very useful. However as it stands it is specific to one data set. Needs
-to be generalised to arbitrary data if we want to keep it.
-
-\textbf{For Alex to do}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{We would need a screenshot (and video) for each of these.}
-\item
-  \textbf{Fix 1, 2, 3, 4, 5}
-\item
-  \textbf{Code consistent for applications to work - see
-  slice\_tour\_compare\_sets.nb. See below}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{We need to make sure the drawing range is fixed (no resizing
-    of the axis when changing projection or slicing). It could be nice
-    if there was a manual zoom in/zoom out to change scale
-    occasionally.}
-  \item
-    \textbf{Should also make sure the size of the points is fixed (maybe
-    this will be automatic if the drawing range is fixed?), bigger
-    points would be preferred, currently the size is sometimes really
-    tiny and it becomes difficult to see anything.}
-  \item
-    \textbf{Can we have some heuristic to decide the range for the slice
-    thickness slider? In the second graph there is a tiny part of the
-    slider where the changes are interesting, difficult to navigate
-    since this makes it super sensitive to any small movement, while
-    most of the options are not interesting. My suggestion would be to
-    keep (as I understand is the case now) the maximum thickness (all
-    points in) as the upper bound, but instead of zero use a thickness
-    below which slices tend to be empty as the lower bound.}
-  \item
-    \textbf{For the olives example I think the relation between x2 and
-    the blue region could be interesting.}
-  \item
-    \textbf{pdfsense data: can we use random samples instead of the
-    first 10k rows? Then we should still be able to see the bigger
-    picture (with picking the first ones we get some artificial
-    structure from how the data was sampled initially).}
-  \item
-    \textbf{I was playing around with this today. Looks like the axes
-    are fixed when changing slice thickness, but not when rotating?}
-  \item
-    \textbf{Anyways, some observations that might be useful, but all of
-    this seems difficult for including in a paper:}
-  \item
-    \textbf{In the first example we can show how the yellow group is
-    associated with the first variable (x1, I guess this is C9?) - when
-    slicing through the center it only appears when x1 has a bigger
-    component on the projection, it is associated with low values of x1.
-    Since they are not near the average value of x1 the points do not
-    get captured in a slice where x1 is not important in the projection.
-    We can move the center point to -1 in the first component to then
-    explore how the yellow group relates to the other three variables.
-    Much less clear, maybe some indication of smaller x2 and larger x3
-    for this.}
-  \item
-    \textbf{The second example is too busy for manual exploration I
-    would say (too many axes to move around separately).}
-  \item
-    \textbf{Third example could be similar to the first: blue region
-    requires small values of x2, can be understood from the exploration,
-    then move center point to -0.5 in x2 to explore further. Now
-    interpretation becomes tricky. Sometimes blue and yellow seem just
-    randomly everywhere, sometimes blue more contained in an ellipsoid
-    around the center (indicating some correlation between variables).
-    For example try zeroing out components 2 and 3 to see this slice in
-    x1 vs x4.}
-  \item
-    \textbf{For the last example x4 seems to be the one to work with. I
-    feel like this is again very similar, maybe better to instead
-    explore other examples (maybe something without slicing?)}
-  \end{itemize}
 \end{itemize}

### 2022-07-11 paper/paper.Rmd
slicing 4.3

@@
 
-We continue the investigation by now slicing based on the projection $A_1$.
-For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0$.
-As a first option we explore how manually changing the projection away from $A_1$ can help with understanding the boundary better.
-For our example we notice that $A_1$ does not contain any contribution from the second variable (`bd`), and in our illustration we will first rotate this variable into the view.
+We continue the investigation by now slicing orthogonally to  the projection $A_1$.
+For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0 = (0,0,0,0)$.
+At  first we explore how changing the projection, and thus the slice) away from $A_1$ can help with understanding the boundary better.
+For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
 
-We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
+Figure \ref{slice1} shows snapshots of the exploration.
+The rop row is the initial projection ($A_1$ and $S_1^0$), and the bottom row is a later projection containing more of `bd` (call them $A_2$ and $S_2^0$).
+The columns correspond to the two models, with Rf on the left and LDA on the right.
+
+Both models have some overlap at the center, even for this thin slice ($S_1^0$).
+The second slice ($S_2^0$) mostly resolves this overlap and reveals the primary differences between the models.
+The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
+The boundary between Chinstrap and Gentoo (green) is where they differ.
+
+The RF is almost straight in this view, as something we might expect from a tree model where splitting occurs on single variables only.
+However, it is straight in a combination of the second and third variable (`bd` and `fl`).
+By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two speciesfor any tree, and likely only one is needed.
+Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
+
+In contrast, the LDA boundary has an oblique split between the two, and roughly divides the space into three similarly sized areas.
+It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary half-way between the group means.
+
+<!-- We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
 \ref{slice1}, top-left).
@@
 By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
-\ref{slice1}).
+\ref{slice1}).-->
 
-Finally we look at the data projection based on $A_2$ and find that the third group (green, Gentoo penguins) is further away from the other two groups (compared to $A_1$), see Fig. \ref{proj2}.
+Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
+It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
+The LDA boundary is better placed for all species.
 

### 2022-07-11 paper/paper.tex
slicing 4.3

@@
 
-We continue the investigation by now slicing based on the projection
-\(A_1\). For both models we look at a thin (\(h=0.5\)) slice through the
-center, \(S_1^0\). As a first option we explore how manually changing
-the projection away from \(A_1\) can help with understanding the
-boundary better. For our example we notice that \(A_1\) does not contain
-any contribution from the second variable (\texttt{bd}), and in our
-illustration we will first rotate this variable into the view.
+We continue the investigation by now slicing orthogonally to the
+projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
+through the center, \(S_1^0 = (0,0,0,0)\). At first we explore how
+changing the projection, and thus the slice) away from \(A_1\) can help
+with understanding the boundary better. For our example, notice that
+\(A_1\) does not contain any contribution from the second variable
+(\texttt{bd}), so we will first rotate this variable into the view.
 
-We again start by looking at the RF model, the slice \(S_1^0\) shows the
-block structure of the model, with the third group (green, Gentoo
-penguins) overlapping with both of the other groups (Fig. \ref{slice1},
-top-left). This is similar for the LDA model (Fig. \ref{slice1},
-top-right), but again the linearity results in different boundaries and
-thus differences in how the classes overlap.
+Figure \ref{slice1} shows snapshots of the exploration. The rop row is
+the initial projection (\(A_1\) and \(S_1^0\)), and the bottom row is a
+later projection containing more of \texttt{bd} (call them \(A_2\) and
+\(S_2^0\)). The columns correspond to the two models, with Rf on the
+left and LDA on the right.
 
-By rotating in the second variable we can find a view that shows three
-neatly separated blocks for the RF model, and we export the
-corresponding projection matrix \(A_2\) to see the same views for the
-LDA model which also shows a neater separation of the classes along the
-boundary (see bottom row of Fig. \ref{slice1}).
+Both models have some overlap at the center, even for this thin slice
+(\(S_1^0\)). The second slice (\(S_2^0\)) mostly resolves this overlap
+and reveals the primary differences between the models. The boundary
+between Adelie (blue) and Chinstrap (yellow) is very similar for both
+models. The boundary between Chinstrap and Gentoo (green) is where they
+differ.
 
-Finally we look at the data projection based on \(A_2\) and find that
-the third group (green, Gentoo penguins) is further away from the other
-two groups (compared to \(A_1\)), see Fig. \ref{proj2}.
+The RF is almost straight in this view, as something we might expect
+from a tree model where splitting occurs on single variables only.
+However, it is straight in a combination of the second and third
+variable (\texttt{bd} and \texttt{fl}). By examining the scatterplot
+matrix, we can understand how this boundary was built: on each of
+\texttt{bd} and \texttt{fl} it is possible to cut on a value where most
+of the Gentoo penguins are different from the other two speciesfor any
+tree, and likely only one is needed. Thus the forest construction is
+providing a sample of trees that use one variable or the other variable
+to split, producing the blocky boundary on the combination of the two.
+
+In contrast, the LDA boundary has an oblique split between the two, and
+roughly divides the space into three similarly sized areas. It is almost
+like a textbook illustration of how LDA works for two variables (2D)
+where by assuming the clusters have equal variance-covariance, it places
+a boundary half-way between the group means.
+
+Finally, to determine which model does the boundaries better, compare
+them with the \(A_2\) projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one
+can imagine that the trees in the forest has greedily grasped any one of
+many places to make a split to separate the group. It might be argued
+though the that RF boundary is cut too close to the Chinstrap species,
+and might lead to some unnecessary misclassification with new data. The
+LDA boundary is better placed for all species.
 

### 2022-07-11 paper/paper.Rmd
Some explanation of slice interactions (but slice tour not explained anywhere?)

--- 2022-07-11 paper/paper.Rmd
@@
 
-How does it work?
+To better explore the space we combine the manual controls for the projection with manual controls for slicing.
+A slice is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].
+In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points in steps along that direction to move the slice and fully cover the data space.
+This no longer works in higher-dimensional spaces, and we can think of picking one direction and shifting the slice along the component orthogonal to the projection plane.
+In practice we define the center point explicitly and calculate the slice display as explained in @slicetour.
+The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.
 
-Diagram of interface? Scaling
+Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here.
+An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
+
+In addition it is also useful to interactively change the slice thickness, in particular when selecting the preferred thickness $h$ (also called the slice radius) for exploring the input data.
+This was implemented as a slicer, and to allow the user to efficiently explore the interesting range we can provide the slider range as an input.
+For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with
+\begin{equation}
+N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
+\label{eq:count}
+\end{equation}
+
+The current implementation contains a switch to change between the slice view and the projected view of the data, while keeping all other settings fixed.
 

### 2022-07-11 paper/paper.tex
Some explanation of slice interactions (but slice tour not explained anywhere?)

--- 2022-07-11 paper/paper.tex
@@
 
-How does it work?
+To better explore the space we combine the manual controls for the
+projection with manual controls for slicing. A slice is defined by a
+projection, a center point that is anchoring it in the high-dimensional
+space and the slice thickness \(h\) \citep{slicetour}. In the case of a
+single orthogonal direction on the projection plane we can pick a
+sequence of center points in steps along that direction to move the
+slice and fully cover the data space. This no longer works in
+higher-dimensional spaces, and we can think of picking one direction and
+shifting the slice along the component orthogonal to the projection
+plane. In practice we define the center point explicitly and calculate
+the slice display as explained in \citet{slicetour}. The current
+implementation allows the user to individually enter the components of
+the vector defining the center point, and updates the display by jumping
+to the new slice.
 
-Diagram of interface? Scaling
+Potential adjustments for future work include a better visualization of
+the current center point, potentially with interactive graphics to move
+the slice similar to how the manual tour was implemented here. An
+alternative would be a step-wise shift of the slice towards the new
+center point to get a better understanding of the distribution along the
+selected direction.
+
+In addition it is also useful to interactively change the slice
+thickness, in particular when selecting the preferred thickness \(h\)
+(also called the slice radius) for exploring the input data. This was
+implemented as a slicer, and to allow the user to efficiently explore
+the interesting range we can provide the slider range as an input. For
+guidance the estimates of the number of points inside the slice as a
+function of the original sample size \(N\) and the number of dimensions
+\(p\) from \citet{sectionpursuit} can be used: in case of a uniform
+distribution inside a sphere of radius \(R\) a slice with thickness
+\(h\) will contain \(N_S\) points, with \begin{equation}
+N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
+\label{eq:count}
+\end{equation}
+
+The current implementation contains a switch to change between the slice
+view and the projected view of the data, while keeping all other
+settings fixed.
 

### 2022-07-11 paper/paper.Rmd
captions for fig 6 and 7

@@
 
-The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ are not followed because the orthonormalisation will change them.
+The problem with new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
 There are numerous ways that this can be enforced, and details are provided in the Appendix.
 
-## Manual control for slice center
+## Manual control for slices
 
@@
 A slice is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].
+
+### Shifting the center
+
 In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points in steps along that direction to move the slice and fully cover the data space.
@@
 An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
+
+### Changing the thickness
 
@@
 
-Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
-Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
-It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
-The LDA boundary is better placed for all species.
-
 \begin{figure*}[ht]
@@
 
+Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
+It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
+The LDA boundary is better placed for all species.
+
 ## Shifting the slice center
@@
 We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
-Instead of rotating to a different projection, we might also change the view by moving the slice along one direction in the 4D space.
-Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$ to either large or small values of the bill depth ($\pm 1.5$ after centering and scaling).
+Instead of rotating to a different projection, we might also change the view by moving the slice along one axis in the 4D space.
+Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$, $S_1^c$ to either large positive or negative values ($c_{bd} = \pm 1.5$ after centering and scaling).
 We will label these slices as $S_1^{\pm}$.
@@
 
-We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Fig.
-\ref{slice1p}.
+We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
 We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{...}.
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
+The two models have similar boundaries except that the RF is more of a step function.}
 \label{slice1p}
 \end{figure*}
-
-A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Fig.
-\ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
-On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
-Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
 
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{...}.
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
 \label{slice1m}
 \end{figure*}
+
+A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
+The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
+On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
+Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
 

### 2022-07-11 paper/paper.tex
captions for fig 6 and 7

@@
 
-The problem with new simple method (Algorithm 1) is that the precise
-values for \(V_m\) are not followed because the orthonormalisation will
-change them. There are numerous ways that this can be enforced, and
-details are provided in the Appendix.
+The problem with new simple method is that it is not faithful to the
+precise values for \(V_m\) because the orthonormalisation will change
+them. There are numerous ways that this can be enforced, and details are
+provided in the Appendix.
 
-\hypertarget{manual-control-for-slice-center}{%
-\subsection{Manual control for slice
-center}\label{manual-control-for-slice-center}}
+\hypertarget{manual-control-for-slices}{%
+\subsection{Manual control for slices}\label{manual-control-for-slices}}
 
@@
 projection, a center point that is anchoring it in the high-dimensional
-space and the slice thickness \(h\) \citep{slicetour}. In the case of a
-single orthogonal direction on the projection plane we can pick a
-sequence of center points in steps along that direction to move the
-slice and fully cover the data space. This no longer works in
+space and the slice thickness \(h\) \citep{slicetour}.
+
+\hypertarget{shifting-the-center}{%
+\subsubsection{Shifting the center}\label{shifting-the-center}}
+
+In the case of a single orthogonal direction on the projection plane we
+can pick a sequence of center points in steps along that direction to
+move the slice and fully cover the data space. This no longer works in
 higher-dimensional spaces, and we can think of picking one direction and
@@
 selected direction.
+
+\hypertarget{changing-the-thickness}{%
+\subsubsection{Changing the thickness}\label{changing-the-thickness}}
 
@@
 
-Finally, to determine which model does the boundaries better, compare
-them with the \(A_2\) projection of the data (Figure \ref{proj2}).
-Gentoo is more separated from the other two in this projection, and one
-can imagine that the trees in the forest has greedily grasped any one of
-many places to make a split to separate the group. It might be argued
-though the that RF boundary is cut too close to the Chinstrap species,
-and might lead to some unnecessary misclassification with new data. The
-LDA boundary is better placed for all species.
-
 \begin{figure*}[ht]
@@
 
+Finally, to determine which model does the boundaries better, compare
+them with the \(A_2\) projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one
+can imagine that the trees in the forest has greedily grasped any one of
+many places to make a split to separate the group. It might be argued
+though the that RF boundary is cut too close to the Chinstrap species,
+and might lead to some unnecessary misclassification with new data. The
+LDA boundary is better placed for all species.
+
 \hypertarget{shifting-the-slice-center}{%
@@
 different projection, we might also change the view by moving the slice
-along one direction in the 4D space. Here we will continue our
-exploration of the dependence on \texttt{bd} and move the slice defined
-by \(A_1\) to either large or small values of the bill depth
-(\(\pm 1.5\) after centering and scaling). We will label these slices as
-\(S_1^{\pm}\). Here we will also look at slices of the observed data
-points, using a thicker slice (\(h=1.5\)) to capture enough points in a
-given view.
+along one axis in the 4D space. Here we will continue our exploration of
+the dependence on \texttt{bd} and move the slice defined by \(A_1\),
+\(S_1^c\) to either large positive or negative values
+(\(c_{bd} = \pm 1.5\) after centering and scaling). We will label these
+slices as \(S_1^{\pm}\). Here we will also look at slices of the
+observed data points, using a thicker slice (\(h=1.5\)) to capture
+enough points in a given view.
 
@@
 \(S_1^{+}\), thus the slice is localized towards high values of bill
-depth in Fig. \ref{slice1p}. We can see that all three slices (the two
+depth in Figure \ref{slice1p}. We can see that all three slices (the two
 models and the data) contain almost no points from the third class
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{...}.
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
+The two models have similar boundaries except that the RF is more of a step function.}
 \label{slice1p}
 \end{figure*}
-
-A more interesting comparison is found for \(S_1^{-}\), thus the slice
-localized towards low values of bill depth, shown in Fig. \ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with
-an interesting boundary for the third class (green, Gentoo). On the
-other hand the LDA model (middle) predominantly predicts the third class
-within the slice, this appears to be enforced through the linear
-structure of the model. Looking finally at the thick slice through the
-data we see that there are primarily observations from this class within
-the slice we can conclude that the two models have filled in the
-``empty'' space (where we do not have any training observations) in very
-different ways and according to what we might expect given the model
-structure.
 
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{...}.
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
 \label{slice1m}
 \end{figure*}
+
+A more interesting comparison is found for \(S_1^{-}\), thus the slice
+localized towards low values of bill depth, shown in Figure
+\ref{slice1m}. The RF model (left) predicts all three species within
+this slice, with an interesting boundary for the third class (green,
+Gentoo). On the other hand the LDA model (middle) predominantly predicts
+the third class within the slice, this appears to be enforced through
+the linear structure of the model. Looking finally at the thick slice
+through the data we see that there are primarily observations from this
+class within the slice we can conclude that the two models have filled
+in the ``empty'' space (where we do not have any training observations)
+in very different ways and according to what we might expect given the
+model structure.
 

### 2022-07-12 paper/appendix.Rmd
Adjustment method 3 as algorithm

@@
 
-For now just sketching the idea:
+A very simple adjustment method works purely within the (hyper)plane, and is using the fact that the first direction will not be rotated when applying Gram-Schmidt. The following steps are needed:
 
-- first step is to click on the axis display to change the contribution of one variable $m$
-- capture that position and replace the corresponding row in the projection matrix
-- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
-- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
-- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
-- rotated back to the original xy basis to update the plots
+1. Provide $A$, and $m$.
+2. Change values in row $m$, giving $A^*$.
+3.
+Rotate the basis within the projection (hyper-)plane, such that the first direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$.
+In a 2D projection this means finding the angle $\theta$ by which the basis should be rotated.
+4. Normalise the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
+5. Orthonormalise $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
+6. Rotate $\hat A^{**}$ back into the original basis to obtain $A^{**}$.
 
-XXX need to check calculation of angle when $m_1 < 0$
+Diagram needed?
 

### 2022-07-12 paper/appendix.tex
Adjustment method 3 as algorithm

@@
 
-For now just sketching the idea:
+A very simple adjustment method works purely within the (hyper)plane,
+and is using the fact that the first direction will not be rotated when
+applying Gram-Schmidt. The following steps are needed:
 
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
 \tightlist
 \item
-  first step is to click on the axis display to change the contribution
-  of one variable \(m\)
+  Provide \(A\), and \(m\).
 \item
-  capture that position and replace the corresponding row in the
-  projection matrix
+  Change values in row \(m\), giving \(A^*\).
 \item
-  for 2D projection this gives components \(m_1\) along ``x'' direction
-  and \(m_2\) along \(y\) direction
+  Rotate the basis within the projection (hyper-)plane, such that the
+  first direction is aligned with the direction of \(m\) in \(A^*\),
+  giving \(\hat A^*\). In a 2D projection this means finding the angle
+  \(\theta\) by which the basis should be rotated.
 \item
-  next step: rotate basis such that direction of row m now corresponds
-  to the first basis vector, this means we apply 2x2 rotation matrix to
-  each row of the projection matrix, where the rotation angle is
-  \(tan(\theta) = m_2/m_1\) (that matrix can be written just in terms of
-  that ratio, \(x= m_2/m_1\), when \(m_1<0\) I think we just need to
-  translate \(\theta \rightarrow \theta + \pi\))
+  Normalise the first direction in \(\hat A^*\) while keeping the
+  contribution in row \(m\) fixed, giving \(\hat A^{*0}\).
 \item
-  take the rotated basis, apply Gram-Schmidt, now the direction of \(m\)
-  will not change (but the length will change during normalization)
+  Orthonormalise \(\hat A^{*0}\) using Gram-Schmidt, giving
+  \(\hat A^{**}\). By construction the contribution of \(m\) will remain
+  fixed.
 \item
-  rotated back to the original xy basis to update the plots
-\end{itemize}
+  Rotate \(\hat A^{**}\) back into the original basis to obtain
+  \(A^{**}\).
+\end{enumerate}
 
-XXX need to check calculation of angle when \(m_1 < 0\)
+Diagram needed?
 

### 2022-07-13 paper/paper.Rmd
Sentence about preprocessing and reference to remove the blindfold paper added.

@@
 
-To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins].
+To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins] and look at classification models following @sam11271.
 We will show how classification boundaries can be explored and better understood on projections and slices through 4D space.
@@
 We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest (RF).
+Both the data and the model points in the grid are centered and scaled (standard deviation $= 1$).
 

### 2022-07-13 paper/paper.tex
Sentence about preprocessing and reference to remove the blindfold paper added.

@@
 To illustrate the usefulness of the manual controls we use the 4D
-penguins data \citep{penguins}. We will show how classification
-boundaries can be explored and better understood on projections and
-slices through 4D space. Figure \ref{fig:penguins-scatmat} shows a
-scatterplot matrix of this data. There are four variables
+penguins data \citep{penguins} and look at classification models
+following \citet{sam11271}. We will show how classification boundaries
+can be explored and better understood on projections and slices through
+4D space. Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix
+of this data. There are four variables
 (\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
@@
 classification models: linear discriminant analysis (LDA) and random
-forest (RF).
+forest (RF). Both the data and the model points in the grid are centered
+and scaled (standard deviation \(= 1\)).
 

### 2022-07-14 paper/paper.Rmd
introduction revised

@@
 For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
-These controls can also be used to interactively generate faceted plots [@XXX], or spatiotemporal glyphmaps [@XXX].
 Having the user interact with a projection is extremely valuable for understanding high-dimensional data.
@@
 
-Through experiments with the relatively new interactive graphics capabilities in mathematica(?), we have realized that there is a simpler approach, which is more direct, and extensible for generating user interaction.
-This paper explains this, and is organized as follows.
-The next section describes the new algorithm for manual control.
-This is followed by details on implementation.
-The applications section illustrate how these can be used.
+This paper explains a simpler approach, which is more direct, and extensible for generating user interaction, and is organized as follows.
+The next section describes the new algorithm for manual control, for both projections and slices.
+The use of these methods are illustrated to compare and contrast boundaries constructed by different classifiers.
+The software section describes a mathematica notebook that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
+The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two mathematica notebooks, one focused on the application, and the other that could be used generically for new data.
 

### 2022-07-14 paper/paper.tex
introduction revised

@@
 one can reduce or increase a variable's contribution to examine the
-variable importance. These controls can also be used to interactively
-generate faceted plots \citep{XXX}, or spatiotemporal glyphmaps
-\citep{XXX}. Having the user interact with a projection is extremely
-valuable for understanding high-dimensional data. However, these
-algorithms have two problems: (1) the pre-processing of creating a
+variable importance. Having the user interact with a projection is
+extremely valuable for understanding high-dimensional data. However,
+these algorithms have two problems: (1) the pre-processing of creating a
 manipulation space overly complicates the algorithm, (2) extending to
@@
 
-Through experiments with the relatively new interactive graphics
-capabilities in mathematica(?), we have realized that there is a simpler
-approach, which is more direct, and extensible for generating user
-interaction. This paper explains this, and is organized as follows. The
-next section describes the new algorithm for manual control. This is
-followed by details on implementation. The applications section
-illustrate how these can be used.
+This paper explains a simpler approach, which is more direct, and
+extensible for generating user interaction, and is organized as follows.
+The next section describes the new algorithm for manual control, for
+both projections and slices. The use of these methods are illustrated to
+compare and contrast boundaries constructed by different classifiers.
+The software section describes a mathematica notebook that is used for
+the application, and describes the interactive environment that would be
+desirable within R as new technology becomes available. The paper is
+accompanied by an appendix with more details and adjustments to the
+manual controls, and two mathematica notebooks, one focused on the
+application, and the other that could be used generically for new data.
 

### 2022-07-14 paper/paper.tex
reference ggplot2 and ggally

@@
 
-\hypertarget{sec:implementation}{%
-\section{Experimenting with new techniques using
-Mathematica}\label{sec:implementation}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Why is this a good sandbox}
-\item
-  \textbf{Explain the functionality available in the notebooks}
-\end{itemize}
-
 \begin{figure}
@@
 
-\hypertarget{extension-added-in-the-r-package-tourr}{%
-\section{\texorpdfstring{Extension added in the R package
-\texttt{tourr}}{Extension added in the R package tourr}}\label{extension-added-in-the-r-package-tourr}}
+\hypertarget{sec:implementation}{%
+\section{Software}\label{sec:implementation}}
+
+\hypertarget{mathematica-notebook}{%
+\subsection{Mathematica notebook}\label{mathematica-notebook}}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Why is this a good sandbox}
+\item
+  \textbf{Explain the functionality available in the notebooks}
+\end{itemize}
+
+\hypertarget{extensions-to-the-r-package-tourr}{%
+\subsection{\texorpdfstring{Extensions to the R package
+\texttt{tourr}}{Extensions to the R package tourr}}\label{extensions-to-the-r-package-tourr}}
 
@@
 
-\hypertarget{what-would-be-desirable-for-implementations-in-r}{%
-\section{What would be desirable for implementations in
-R?}\label{what-would-be-desirable-for-implementations-in-r}}
+Changing slice center
+
+\hypertarget{desirable-interactivity}{%
+\subsection{Desirable interactivity}\label{desirable-interactivity}}
 
@@
 Research Council. The paper was written in \texttt{rmarkdown}
-\citep{rmarkdown} using \texttt{knitr} \citep{knitr}.
+\citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
+matrix of the penguins data was produced by the GGally \citep{GGally}
+package built on ggplot2 \citep{ggplot2} graphics.
 

### 2022-07-14 paper/paper.Rmd
reference ggplot2 and ggally
mapping out discussion

@@
 
+XXX Equation for arbitrary center for slice; and move the "the implementation ..." into the software section once that is constructed; and "potential adjustments for future" into the section on desirable controls
+
 Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here.
@@
 The current implementation contains a switch to change between the slice view and the projected view of the data, while keeping all other settings fixed.
-
-# Experimenting with new techniques using Mathematica {#sec:implementation}
-
-- **Why is this a good sandbox**
-- **Explain the functionality available in the notebooks**
-
-<!-- 
-The following functions are implemented in the Mathematica package `mmtour.wl`:
-
-- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
-colour(=Automatic)]` with arguments
-
-    - `data`: A list of data matrices (they must all have the same dimensionality)
-    - ``projmat`: A projection matrix which describes the projection plane. This is the 
-initial projection and does not need to be orthonormal (XXX is this right?).
-    - ``xRange`: A list which details the range of the horizontal axis.
-    - ``yRange`: A list which details the range of the vertical axis.
-    - ``colour`: A list of graphics directives to be applied to each data matrix.
-
-There are problems with the default values for x and y ranges and colour. It would be 
-better to fix the ranges and colour based on the data as well as the origin of axes 
-and the size of the points.
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**
-
-The format for the printed projection matrix can be improved also. Maybe fix the 
-number of digits and the position of the label.
-
-- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
-
-    - `data`: A data matrix
-    - `projmat`: The inital projection matrix
-    - `centrepoint`: The initial position of the slice
-    - `height`: The initial height of the slice
-    - `heightRange`: The range of heights
-    - `ptSize1`: The size of the points that exist within the slice
-    - `ptSize2`: The size of the points that exist outside of the slice
-
-This function requires some auxiliary functions. It does not accept more than one data set.
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
-
-It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
-
-- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments
-
-- `data`: A data matrix with an extra grouping or clustering column (numerical) 
-- `projmat`: The initial projection matrix
-- `centrepoint`: The initial position of the slice
-- `height`: The initial height of the slice
-- `heightRange`: The range of heights
-- `ptSize1`: The size of the points that exist within the slice
-- `ptSize2`: The size of the points that exist outside of the slice
-
-**Does this need to be a separate function from SlicePlot? Fix 2**
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**
-
-**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**
-
-**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
-One data set with a flag column at the end, is best!
-Is the code too complicated?
-Fix 5**
-
-- `Projected2DSliderPlot[]`
-
-This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
-
-**For Alex to do**
-
-- **We would need a screenshot (and video) for each of these.**
-- **Fix 1, 2, 3, 4, 5**
-- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
-    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
-It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
-    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
-    - **Can we have some heuristic to decide the range for the slice thickness slider?
-In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
-My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
-    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
-    - **pdfsense data: can we use random samples instead of the first 10k rows?
-Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
-    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
-    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
-    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
-Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
-We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
-Much less clear, maybe some indication of smaller x2 and larger x3 for this.**
-    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
-    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
-Now interpretation becomes tricky.
-Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
-For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
-    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
--->
 
@@
 
-# Extension added in the R package `tourr`
+# Software {#sec:implementation}
+
+## Mathematica notebook
+
+- **Why is this a good sandbox**
+- **Explain the functionality available in the notebooks**
+
+<!-- 
+The following functions are implemented in the Mathematica package `mmtour.wl`:
+
+- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
+colour(=Automatic)]` with arguments
+
+    - `data`: A list of data matrices (they must all have the same dimensionality)
+    - ``projmat`: A projection matrix which describes the projection plane. This is the 
+initial projection and does not need to be orthonormal (XXX is this right?).
+    - ``xRange`: A list which details the range of the horizontal axis.
+    - ``yRange`: A list which details the range of the vertical axis.
+    - ``colour`: A list of graphics directives to be applied to each data matrix.
+
+There are problems with the default values for x and y ranges and colour. It would be 
+better to fix the ranges and colour based on the data as well as the origin of axes 
+and the size of the points.
+
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**
+
+The format for the printed projection matrix can be improved also. Maybe fix the 
+number of digits and the position of the label.
+
+- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
+
+    - `data`: A data matrix
+    - `projmat`: The inital projection matrix
+    - `centrepoint`: The initial position of the slice
+    - `height`: The initial height of the slice
+    - `heightRange`: The range of heights
+    - `ptSize1`: The size of the points that exist within the slice
+    - `ptSize2`: The size of the points that exist outside of the slice
+
+This function requires some auxiliary functions. It does not accept more than one data set.
+
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
+
+It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+
+- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments
+
+- `data`: A data matrix with an extra grouping or clustering column (numerical) 
+- `projmat`: The initial projection matrix
+- `centrepoint`: The initial position of the slice
+- `height`: The initial height of the slice
+- `heightRange`: The range of heights
+- `ptSize1`: The size of the points that exist within the slice
+- `ptSize2`: The size of the points that exist outside of the slice
+
+**Does this need to be a separate function from SlicePlot? Fix 2**
+
+**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**
+
+**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**
+
+**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
+One data set with a flag column at the end, is best!
+Is the code too complicated?
+Fix 5**
+
+- `Projected2DSliderPlot[]`
+
+This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
+
+**For Alex to do**
+
+- **We would need a screenshot (and video) for each of these.**
+- **Fix 1, 2, 3, 4, 5**
+- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
+    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
+It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
+    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
+    - **Can we have some heuristic to decide the range for the slice thickness slider?
+In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
+My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
+    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
+    - **pdfsense data: can we use random samples instead of the first 10k rows?
+Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
+    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
+    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
+    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
+Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
+We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
+Much less clear, maybe some indication of smaller x2 and larger x3 for this.**
+    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
+    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
+Now interpretation becomes tricky.
+Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
+For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
+    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
+-->
+
+## Extensions to the R package `tourr`
 
@@
 
-# What would be desirable for implementations in R?
+Changing slice center
+
+## Desirable interactivity
 
@@
 
+Recap contribution
+
+Where else might it be useful to use mathematica
+
 # Acknowledgements {-}
 
-The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
+The authors gratefully acknowledge the support of the Australian Research Council.
+The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
+The scatterplot matrix of the penguins data was produced by the GGally [@GGally] package built on ggplot2 [@ggplot2] graphics.
 
@@
 The source material and animated gifs for this paper are available at
+
+Mathematica notebook
+
+Appendix

### 2022-07-21 paper/paper.Rmd
discussion points
Acknowledgements updated according to German

@@
 
+Other applications
+
 Where else might it be useful to use mathematica
+
+Maybe extensions for slicing
 
@@
 
-The authors gratefully acknowledge the support of the Australian Research Council.
+The authors gratefully acknowledge the support of the Australian Research Council and the ResearchFirst undergraduate research program at Monash University.
 The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
 The scatterplot matrix of the penguins data was produced by the GGally [@GGally] package built on ggplot2 [@ggplot2] graphics.
+We thank the Institute of Statistics, BOKU, for their hospitality while part of this work was conducted.
 

### 2022-07-31 paper/paper.Rmd
started my homework

@@
 abstract: | 
-  Something here
+  We implement a new version of a manual tour that takes advantage of existing features in Mathematica to construct functions that interact with the position of the cursor.
+We demonstrate the usefulness of this new tool with an example exploring region boundaries in classification problems.
 keywords: |
@@
 
+Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
+Most of this inbuilt functionality is quite user-friendly and can be found in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
+Also, it is relatively simple  to create dynamic objects via the inbuilt commands Manipulate or DynamicModule.
+Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that dynamic variable will be updated.
+This is the essential ingredient for this implementation of the manual tour.
+However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different metric(?)  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example) Also, Mathematica can be difficult to debug depending on the size of the project as there are no line numbers and there isn’t a way to backtrack evaluations to see how the state of local variables changes throughout the evaluation.
+(PyCharm has this functionality) \textcolor{red}{needs rewording here}
+
+The examples presented in the penguins’ notebooks all utilise our primary new function, SliceDynamic.
+This function typically accepts data in the form of a matrix where the second last column details the name of the group and the last column details the group index, although this can be changed.
+Additionally, the initial slice height and the slice range need to be specified as function arguments.
+The user is then presented with an interactive(?) in which the control objects appear on the left and the slice appears on the right.
+The user can change the slice via the locator pane, which changes the projection matrix;  a slider, which changes the slice height; and the input field, which changes the slice center.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
+Also, the current projection can be displayed instead of the slice and the current projection matrix can be displayed via checkboxes.
+The Slice Tour* notebooks display some other functions such as ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic.
+ProjectionPlot is very similar to SliceDynamic, except it only displays projections and VisualiseSliceDynamic allows you to discern which points exist in the slice and which don’t.
+However, ProjectedLocatorPlot** displays the interactive slightly different.
+Several locator panes are displayed above the plot and each one corresponds to a row in the projection matrix but behaves the same as ProjectionPlot.
+(this may be easier to implement in another language (?) )
+
 <!-- 

### 2022-08-19 paper/appendix.Rmd
first pass at my changes

@@
 author:
+  - name: Ursula Laa
+    affil: a
+    email: ursula.laa@boku.ac.at
   - name: Alex Aumann
-    affil: a
+    affil: b
     email: aaum0002@student.monash.edu
-  - name: German Valencia
-    affil: a
-    email: german.valencia@monash.edu
-  - name: Ursula Laa
-    affil: b
-    email: ursula.laa@boku.ac.at
   - name: Dianne Cook
@@
     email: dicook@monash.edu
+  - name: German Valencia
+    affil: b
+    email: german.valencia@monash.edu
 affiliation:
@@
     address: |
-      School of Physics and Astronomy, Monash University
+      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
   - num: b
     address: |
-      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
-  - num: b
+      School of Physics and Astronomy, Monash University
+  - num: c
     address: |
@@
 \end{enumerate}
-The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.
+The random completion of $A$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane.
+However, it renders projections in a way that is not continuouns and may be distracting.
+This can be alleviated by replacing the random completion with a rule restricting the size of the jumps.
 
@@
 
+{\textcolor{blue} Do we need three methods? is this one implemented somewhere?}
+
 # Software details
@@
 
-The following functions are implemented in the Mathematica package `mmtour.wl`:
+Here we describe the functions implemented in the Mathematica package `mmtour.wl`. The main function is
 
-- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
-colour(=Automatic)]` with arguments
+- `SliceDynamic[data, projmat, height, heightRange, legendQ(=1), flagQ(=1), colorFunc(=ColorData[97])]` with arguments
 
-    - `data`: A list of data matrices (they must all have the same dimensionality)
-    - ``projmat`: A projection matrix which describes the projection plane. This is the 
-initial projection and does not need to be orthonormal (XXX is this right?).
-    - ``xRange`: A list which details the range of the horizontal axis.
-    - ``yRange`: A list which details the range of the vertical axis.
-    - ``colour`: A list of graphics directives to be applied to each data matrix.
+- `data`: A data matrix with an extra grouping that has a label (string) and a flag  (numerical) 
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `height`: The initial height of the slice.
+- `heightRange`: The range of slice heights to be explored.
+- `legendQ`: A flag signaling whether the data matrix includes a column with group labels (1) or not (0).
+- `flagQ`: A flag signaling whether the data matrix includes a numerical flag for the groups.
+- `colorFunc`: specifies the colour mapping for the different groups.
 
-There are problems with the default values for x and y ranges and colour. It would be 
-better to fix the ranges and colour based on the data as well as the origin of axes 
-and the size of the points.
+The function renders a manual tour with controls that allow the user to navigate through different projections.
+For a given projection, a slider permits the user to change the slide width within its range and a separate box switches the view to a projection.
+The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders.
+An additional box displays the coordinates of the projection matrix explicitly.
 
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**
+This function can be used for ungrouped data setting  legendQ and flagQ to 0.
 
-The format for the printed projection matrix can be improved also. Maybe fix the 
-number of digits and the position of the label.
+To remove a group from the visual display, the user can choose specific colours for the entry colorFunc.
+For example, {Red,White,Green} would display groups 1 (red) and 3 (green) while making 2 invisible.
 
-- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
+Other functions are:
 
-    - `data`: A data matrix
-    - `projmat`: The inital projection matrix
-    - `centrepoint`: The initial position of the slice
-    - `height`: The initial height of the slice
-    - `heightRange`: The range of heights
-    - `ptSize1`: The size of the points that exist within the slice
-    - `ptSize2`: The size of the points that exist outside of the slice
+- `ProjectionPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments
 
-This function requires some auxiliary functions. It does not accept more than one data set.
+- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `legendNames`: an optional list of labels for the groups.
+- `colorFunc`: specifies the colour mapping for the different groups.
 
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
+This function is similar to SliceDynamic but can only display projections.
+Using it instead of SliceDynamic simplifies the display and is more efficient.
+The data file should not contain labels for groups, if these are desired they are passed through the optional argument legendNames as {"group 1", "group 2",...}, for example.
 
-It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
+The display allows the user to navigate through different projections, choose the point size and zoom the factor, and explicitly display the coordinates of the projection.
 
-- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments
+- `ProjectedLocatorPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments
 
-- `data`: A data matrix with an extra grouping or clustering column (numerical) 
-- `projmat`: The initial projection matrix
-- `centrepoint`: The initial position of the slice
-- `height`: The initial height of the slice
-- `heightRange`: The range of heights
-- `ptSize1`: The size of the points that exist within the slice
-- `ptSize2`: The size of the points that exist outside of the slice
+- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `legendNames`: an optional list of labels for the groups.
+- `colorFunc`: specifies the colour mapping for the different groups.
 
-**Does this need to be a separate function from SlicePlot? Fix 2**
+This function is very similar to ProjectionPlot but uses a different display for navigation of  projections in which each coordinate is represented on a separate dial.
 
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**
+- `VisualiseSliceDynamic[data, projmat, height, heightRange]` with arguments
 
-**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**
+- `data`: A data matrix containg only one set.In this case the data matrix should not contain column labels.
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `height`: The initial height of the slice.
+- `heightRange`: The range of slice heights to be explored.
 
-**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
-One data set with a flag column at the end, is best!
-Is the code too complicated?
-Fix 5**
-
-- `Projected2DSliderPlot[]`
-
-This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
-
-**For Alex to do**
-
-- **We would need a screenshot (and video) for each of these.**
-- **Fix 1, 2, 3, 4, 5**
-- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
-    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
-It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
-    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
-    - **Can we have some heuristic to decide the range for the slice thickness slider?
-In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
-My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
-    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
-    - **pdfsense data: can we use random samples instead of the first 10k rows?
-Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
-    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
-    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
-    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
-Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
-We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
-Much less clear, maybe some indication of smaller x2 and larger x3 for this.**
-    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
-    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
-Now interpretation becomes tricky.
-Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
-For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
-    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
+This function differs from SliceDynamic in that it looks at only one data set but displays both the points inside and outside the slice.
+The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, centre point and zoom level.

### 2022-08-19 paper/appendix.tex
first pass at my changes

@@
 
-\author{\name{Alex Aumann$^{a}$, German Valencia$^{a}$, Ursula
-Laa$^{b}$, Dianne Cook$^{c}$}
-\affil{$^{a}$School of Physics and Astronomy, Monash
-University; $^{b}$Institute of Statistics, University of Natural
-Resources and Life Sciences, Vienna; $^{b}$Department of Econometrics
-and Business Statistics, Monash University}
+\author{\name{Ursula Laa$^{a}$, Alex Aumann$^{b}$, Dianne
+Cook$^{c}$, German Valencia$^{b}$}
+\affil{$^{a}$Institute of Statistics, University of Natural Resources
+and Life Sciences, Vienna; $^{b}$School of Physics and Astronomy, Monash
+University; $^{c}$Department of Econometrics and Business Statistics,
+Monash University}
 }
 
-\thanks{CONTACT Alex
-Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, German
-Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}, Ursula
-Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Dianne
-Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}}
+\thanks{CONTACT Ursula
+Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Alex
+Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, Dianne
+Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, German
+Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}}
 
@@
 
-The random choice for the components of \(A\) not in the projection
-matrix allows the exploration of dimensions perpendicular to the
-projection plane.
+The random completion of \(A\) outside the projection matrix provides an
+exploration of dimensions orthogonal to the projection plane. However,
+it renders projections in a way that is not continuouns and may be
+distracting. This can be alleviated by replacing the random completion
+with a rule restricting the size of the jumps.
 
@@
 
+\{\textcolor{blue} Do we need three methods? is this one implemented
+somewhere?\}
+
 \hypertarget{software-details}{%
@@
 
-The following functions are implemented in the Mathematica package
-\texttt{mmtour.wl}:
+Here we describe the functions implemented in the Mathematica package
+\texttt{mmtour.wl}. The main function is
 
@@
 \item
-  \texttt{ProjectionPlot{[}data,\ projmat,\ xRange(=Automatic),\ yRange(=Automatic),\ \ colour(=Automatic){]}}
+  \texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
   with arguments
-
-  \begin{itemize}
-  \tightlist
 \item
-    \texttt{data}: A list of data matrices (they must all have the same
-    dimensionality)
+  \texttt{data}: A data matrix with an extra grouping that has a label
+  (string) and a flag (numerical)
 \item
-    `\texttt{projmat}: A projection matrix which describes the
-    projection plane. This is the initial projection and does not need
-    to be orthonormal (XXX is this right?).
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
 \item
-    `\texttt{xRange}: A list which details the range of the horizontal
-    axis.
+  \texttt{height}: The initial height of the slice.
 \item
-    `\texttt{yRange}: A list which details the range of the vertical
-    axis.
+  \texttt{heightRange}: The range of slice heights to be explored.
 \item
-    `\texttt{colour}: A list of graphics directives to be applied to
-    each data matrix.
-  \end{itemize}
+  \texttt{legendQ}: A flag signaling whether the data matrix includes a
+  column with group labels (1) or not (0).
+\item
+  \texttt{flagQ}: A flag signaling whether the data matrix includes a
+  numerical flag for the groups.
+\item
+  \texttt{colorFunc}: specifies the colour mapping for the different
+  groups.
 \end{itemize}
 
-There are problems with the default values for x and y ranges and
-colour. It would be better to fix the ranges and colour based on the
-data as well as the origin of axes and the size of the points.
+The function renders a manual tour with controls that allow the user to
+navigate through different projections. For a given projection, a slider
+permits the user to change the slide width within its range and a
+separate box switches the view to a projection. The user can also change
+the centre point of slices; the size of the points rendered and the
+scale of the plot through sliders. An additional box displays the
+coordinates of the projection matrix explicitly.
 
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data? Fix 1}
+This function can be used for ungrouped data setting legendQ and flagQ
+to 0.
 
-The format for the printed projection matrix can be improved also. Maybe
-fix the number of digits and the position of the label.
+To remove a group from the visual display, the user can choose specific
+colours for the entry colorFunc. For example, \{Red,White,Green\} would
+display groups 1 (red) and 3 (green) while making 2 invisible.
+
+Other functions are:
 
@@
 \item
-  \texttt{SlicePlot{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1(=0.005),\ ptSize2(=0.004),\ minDist(=0){]}}
-  with arguments (used to be called SliceDynamic in Alex's first
-  version, the name SliceDynamic was taken by something else below)
-
-  \begin{itemize}
-  \tightlist
+  \texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
+  with arguments
 \item
-    \texttt{data}: A data matrix
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical)
 \item
-    \texttt{projmat}: The inital projection matrix
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
 \item
-    \texttt{centrepoint}: The initial position of the slice
+  \texttt{legendNames}: an optional list of labels for the groups.
 \item
-    \texttt{height}: The initial height of the slice
-  \item
-    \texttt{heightRange}: The range of heights
-  \item
-    \texttt{ptSize1}: The size of the points that exist within the slice
-  \item
-    \texttt{ptSize2}: The size of the points that exist outside of the
-    slice
-  \end{itemize}
+  \texttt{colorFunc}: specifies the colour mapping for the different
+  groups.
 \end{itemize}
 
-This function requires some auxiliary functions. It does not accept more
-than one data set.
+This function is similar to SliceDynamic but can only display
+projections. Using it instead of SliceDynamic simplifies the display and
+is more efficient. The data file should not contain labels for groups,
+if these are desired they are passed through the optional argument
+legendNames as \{``group 1'', ``group 2'',\ldots\}, for example.
 
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data?}
-
-It may be good to fix the size of the points, the center of the axes and
-the scales of the axes rather than allow them to change as the
-projection and slicing changes
+The display allows the user to navigate through different projections,
+choose the point size and zoom the factor, and explicitly display the
+coordinates of the projection.
 
@@
 \item
-  \texttt{SliceDynamic{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1\ :\ 0.005,\ ptSize2\ :\ 0.004{]}}
+  \texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
   with arguments
 \item
-  \texttt{data}: A data matrix with an extra grouping or clustering
-  column (numerical)
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical)
 \item
-  \texttt{projmat}: The initial projection matrix
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
 \item
-  \texttt{centrepoint}: The initial position of the slice
+  \texttt{legendNames}: an optional list of labels for the groups.
 \item
-  \texttt{height}: The initial height of the slice
-\item
-  \texttt{heightRange}: The range of heights
-\item
-  \texttt{ptSize1}: The size of the points that exist within the slice
-\item
-  \texttt{ptSize2}: The size of the points that exist outside of the
-  slice
+  \texttt{colorFunc}: specifies the colour mapping for the different
+  groups.
 \end{itemize}
 
-\textbf{Does this need to be a separate function from SlicePlot? Fix 2}
-
-\textbf{It would be good if the initial projection matrix (and maybe the
-center) had default values. Perhaps a random matrix with the correct
-dimensions read from the data? Fix 3}
-
-\textbf{It may be good to fix the size of the points, the center of the
-axes and the scales of the axes rather than allow them to change as the
-projection and slicing changes. Fix 4}
-
-\textbf{I have found it useful to input different data sets rather than
-have this function sort them automatically using the last column. One
-data set with a flag column at the end, is best! Is the code too
-complicated? Fix 5}
+This function is very similar to ProjectionPlot but uses a different
+display for navigation of projections in which each coordinate is
+represented on a separate dial.
 
 \begin{itemize}
-\tightlist
 \item
-  \texttt{Projected2DSliderPlot{[}{]}}
+  \texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
+  with arguments
+\item
+  \texttt{data}: A data matrix containg only one set.In this case the
+  data matrix should not contain column labels.
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{height}: The initial height of the slice.
+\item
+  \texttt{heightRange}: The range of slice heights to be explored.
 \end{itemize}
 
-This function uses a different display for the projection that I find
-very useful. However as it stands it is specific to one data set. Needs
-to be generalised to arbitrary data if we want to keep it.
-
-\textbf{For Alex to do}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{We would need a screenshot (and video) for each of these.}
-\item
-  \textbf{Fix 1, 2, 3, 4, 5}
-\item
-  \textbf{Code consistent for applications to work - see
-  slice\_tour\_compare\_sets.nb. See below}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{We need to make sure the drawing range is fixed (no resizing
-    of the axis when changing projection or slicing). It could be nice
-    if there was a manual zoom in/zoom out to change scale
-    occasionally.}
-  \item
-    \textbf{Should also make sure the size of the points is fixed (maybe
-    this will be automatic if the drawing range is fixed?), bigger
-    points would be preferred, currently the size is sometimes really
-    tiny and it becomes difficult to see anything.}
-  \item
-    \textbf{Can we have some heuristic to decide the range for the slice
-    thickness slider? In the second graph there is a tiny part of the
-    slider where the changes are interesting, difficult to navigate
-    since this makes it super sensitive to any small movement, while
-    most of the options are not interesting. My suggestion would be to
-    keep (as I understand is the case now) the maximum thickness (all
-    points in) as the upper bound, but instead of zero use a thickness
-    below which slices tend to be empty as the lower bound.}
-  \item
-    \textbf{For the olives example I think the relation between x2 and
-    the blue region could be interesting.}
-  \item
-    \textbf{pdfsense data: can we use random samples instead of the
-    first 10k rows? Then we should still be able to see the bigger
-    picture (with picking the first ones we get some artificial
-    structure from how the data was sampled initially).}
-  \item
-    \textbf{I was playing around with this today. Looks like the axes
-    are fixed when changing slice thickness, but not when rotating?}
-  \item
-    \textbf{Anyways, some observations that might be useful, but all of
-    this seems difficult for including in a paper:}
-  \item
-    \textbf{In the first example we can show how the yellow group is
-    associated with the first variable (x1, I guess this is C9?) - when
-    slicing through the center it only appears when x1 has a bigger
-    component on the projection, it is associated with low values of x1.
-    Since they are not near the average value of x1 the points do not
-    get captured in a slice where x1 is not important in the projection.
-    We can move the center point to -1 in the first component to then
-    explore how the yellow group relates to the other three variables.
-    Much less clear, maybe some indication of smaller x2 and larger x3
-    for this.}
-  \item
-    \textbf{The second example is too busy for manual exploration I
-    would say (too many axes to move around separately).}
-  \item
-    \textbf{Third example could be similar to the first: blue region
-    requires small values of x2, can be understood from the exploration,
-    then move center point to -0.5 in x2 to explore further. Now
-    interpretation becomes tricky. Sometimes blue and yellow seem just
-    randomly everywhere, sometimes blue more contained in an ellipsoid
-    around the center (indicating some correlation between variables).
-    For example try zeroing out components 2 and 3 to see this slice in
-    x1 vs x4.}
-  \item
-    \textbf{For the last example x4 seems to be the one to work with. I
-    feel like this is again very similar, maybe better to instead
-    explore other examples (maybe something without slicing?)}
-  \end{itemize}
-\end{itemize}
+This function differs from SliceDynamic in that it looks at only one
+data set but displays both the points inside and outside the slice. The
+display panel allows the user to specify the size of both sets of points
+(inside and outside) along with the slice height, centre point and zoom
+level.
 

### 2022-08-19 paper/paper.tex
mapping out discussion
discussion points
first pass at my changes

@@
 \begin{abstract}
-Something here
+We implement a new version of a manual tour that takes advantage of
+existing features in Mathematica to construct functions that interact
+with the position of the cursor. We demonstrate the usefulness of this
+new tool with an example exploring region boundaries in classification
+problems.
 \end{abstract}
@@
 higher dimensional control is difficult.
+
+\{\textcolor{blue}Another very useful feature is that the user can keep
+the projection fixed while slicing across the orthogonal space and
+changing the centre point used for slicing. This can serve to illustrate
+how the orthogonal space affects correlations in a slice and how they
+may change or disappear in a projection.\}
 
@@
 to the new slice.
+
+XXX Equation for arbitrary center for slice; and move the ``the
+implementation \ldots{}'' into the software section once that is
+constructed; and ``potential adjustments for future'' into the section
+on desirable controls
 
@@
 
+Mathematica provides much utility and versatility, such as inbuilt data
+visualisation, data manipulation and analysis, dynamic functionality,
+and symbolic and numeric computation. Most of this inbuilt functionality
+is quite user-friendly and can be found in the Mathematica
+documentation; typically, numerous examples are provided within the
+documentation, and some possible issues are outlined there. Also, it is
+relatively simple to create dynamic objects via the inbuilt commands
+Manipulate or DynamicModule. Control objects, such as sliders, locator
+panes, and input fields, can then be used on dynamic variables, and when
+there is a change in the dynamic variable the dynamic objects which
+contain that dynamic variable will be updated. This is the essential
+ingredient for this implementation of the manual tour. However, it can
+be hard to alter the behaviour of the control objects such as setting
+the 1D slider on a different scale (Slider increments add 10 points into
+the slice rather than increasing the slice height by 0.1 for example).
+\{\textcolor{red}Alex, is this still true with the controls you are
+using now?\}
+
+The examples presented in the penguins' notebooks all utilise our
+primary new function, SliceDynamic. This function typically accepts
+grouped data in the form of a matrix where the second last column
+details the name of the group and the last column details the group
+index. After specifying the initial slice height and the slice range,
+the user is then presented with an interactive display in which the
+control objects appear on the left and the slice visulisation appears on
+the right. The user can change the orientation of the slice via the
+locator pane, which changes the projection matrix; a slider controls the
+slice height; and there is an input field, which changes the slice
+centre. The user can also change the appearance of the plot by zooming
+into the center or changing the point size with sliders provided. The
+current projection can be displayed to contrast it with the slice by
+ticking a box. The explicit projection matrix can be displayed via
+another checkbox. The Slice Tour* notebooks include the additional
+functions: ProjectionPlot, Projected2DSliderPlot, and
+VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic,
+except it only displays projections and VisualiseSliceDynamic allows you
+to discern which points exist inside the slice and outside it. However,
+ProjectedLocatorPlot displays the interactive slightly differently.
+Several locator panes are displayed above the plot and each one
+corresponds to a row in the projection matrix but behaves the same as
+ProjectionPlot. \{\textcolor{blue}(this may be easier to implement in
+another language (?) )\}
+
 \hypertarget{extensions-to-the-r-package-tourr}{%
@@
 
+Recap contribution
+
+Other applications
+
+Where else might it be useful to use mathematica
+
+Maybe extensions for slicing
+
 \hypertarget{acknowledgements}{%
@@
 The authors gratefully acknowledge the support of the Australian
-Research Council. The paper was written in \texttt{rmarkdown}
+Research Council and the ResearchFirst undergraduate research program at
+Monash University. The paper was written in \texttt{rmarkdown}
 \citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
 matrix of the penguins data was produced by the GGally \citep{GGally}
-package built on ggplot2 \citep{ggplot2} graphics.
+package built on ggplot2 \citep{ggplot2} graphics. We thank the
+Institute of Statistics, BOKU, for their hospitality while part of this
+work was conducted.
 
@@
 
+Mathematica notebook
+
+Appendix
+
 \bibliographystyle{tfcad}

### 2022-08-19 paper/paper.Rmd

--- 2022-07-21 paper/paper.Rmd
@@
 However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, (2) extending to higher dimensional control is difficult.
+
+{\textcolor{blue}Another very useful feature is that the user can keep the projection fixed while slicing across the orthogonal space and changing the centre point used for slicing.
+This can serve to illustrate how the orthogonal space affects correlations in a slice and how they may change or disappear in a projection.}
 
@@
 
+Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
+Most of this inbuilt functionality is quite user-friendly and can be found in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
+Also, it is relatively simple  to create dynamic objects via the inbuilt commands Manipulate or DynamicModule.
+Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that dynamic variable will be updated.
+This is the essential ingredient for this implementation of the manual tour.
+However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different scale  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example).
+{\textcolor{red}Alex, is this still true with the controls you are using now?}
+
+The examples presented in the penguins’ notebooks all utilise our primary new function, SliceDynamic.
+This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index.
+After specifying the initial slice height and the slice range, the user is then presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
+The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice height; and there is an input field, which changes the slice centre.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
+The current projection can be displayed to contrast it with the slice by ticking a box.
+The explicit projection matrix can be displayed via another checkbox.
+The Slice Tour* notebooks include the additional functions: ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic.
+ProjectionPlot is very similar to SliceDynamic, except it only displays projections and VisualiseSliceDynamic allows you to discern which points exist inside the slice and outside it.
+However, ProjectedLocatorPlot displays the interactive slightly differently.
+Several locator panes are displayed above the plot and each one corresponds to a row in the projection matrix but behaves the same as ProjectionPlot.
+{\textcolor{blue}(this may be easier to implement in another language (?) )}
+
 <!-- 

### 2022-08-24 paper/paper.Rmd
revised abstract

@@
 abstract: | 
-  We implement a new version of a manual tour that takes advantage of existing features in Mathematica to construct functions that interact with the position of the cursor.
+  This paper describes new user controls for examining high-dimensional data using low-dimensional linear projections and slices.
+A user can interactively change the coefficients of a variable in a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables.
+The user can also interactively shift the slice center, which is useful for exploring how structure changes in local subspaces.
+These controls are implemented in Mathematica to take advantage of it's mathematics functions, and interactivity tools that allow construction of functions that interact with the position of the cursor.
 We demonstrate the usefulness of this new tool with an example exploring region boundaries in classification problems.

### 2022-08-24 paper/paper.tex
revised abstract

@@
 \begin{abstract}
-We implement a new version of a manual tour that takes advantage of
-existing features in Mathematica to construct functions that interact
-with the position of the cursor. We demonstrate the usefulness of this
-new tool with an example exploring region boundaries in classification
-problems.
+This paper describes new user controls for examining high-dimensional
+data using low-dimensional linear projections and slices. A user can
+interactively change the coefficients of a variable in a low-dimensional
+projection, which is useful for exploring the sensitivity of structure
+to particular variables. The user can also interactively shift the slice
+center, which is useful for exploring how structure changes in local
+subspaces. These controls are implemented in Mathematica to take
+advantage of it's mathematics functions, and interactivity tools that
+allow construction of functions that interact with the position of the
+cursor. We demonstrate the usefulness of this new tool with an example
+exploring region boundaries in classification problems.
 \end{abstract}

### 2022-09-05 paper/paper.Rmd
algorithm fixed

@@
 <!-- Background -->
-From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space.
-These are dynamic data visualizations methods and are collected under the term *tours*.
+From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, the more useful methods for data analysis need to work for arbitrary dimension.
+A good approah is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
 Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
@@
 @cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, or out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
-@spyrison_spinifex_2020 refined this algorithm and implements them to generate animation sequences.
+@spyrison_spinifex_2020 refined this algorithm and implemented it to generate radial tour animation sequences.
 
@@
 
-{\textcolor{blue}Another very useful feature is that the user can keep the projection fixed while slicing across the orthogonal space and changing the centre point used for slicing.
-This can serve to illustrate how the orthogonal space affects correlations in a slice and how they may change or disappear in a projection.}
+Another potentially useful manual control, is to allow the user to control the position of the center of a slice.
+The slice tour was introduced in @slicetour.
+It operates by converting the projection plane into a slice, by removing or de-emphasizing points that are further than a fixed orthogonal distance from the plane.
+The projection plane is usually thought of as passing through the center of the data.
+Manual control would allow the user to change the value of center point, by shifting it along a coordinate axis, while keeping the orientation of the projection plane fixed.
+The purpose would be to explore how or if the shape of the data, in the space orthogonal to the projection, changes as one gets away from the center.
 
-This paper explains a simpler approach, which is more direct, and extensible for generating user interaction, and is organized as follows.
+This paper explains the new manual controls for projection and slice tours.
 The next section describes the new algorithm for manual control, for both projections and slices.
@@
 The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed.
-A method to update the values of the component of the controlled variable $V_m$ is then needed.
+A method to update the values of the component ($m^{th}$ row of $A_{p\times d}$) of the controlled variable $V_m$ is then needed.
 
@@
 The first step was to construct a 3D manipulation space from a 2D projection.
-In this space the coefficient of the controlled variable ranges between -1 and 1.
+In this space, the coefficient of the controlled variable ranges between -1 and 1.
 Movements of a cursor are recorded and converted into changes in the values of $V_m$ thus changing the displayed 2D projection.
 The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.
-The construction of the manipulation space overly complicates the manual controls, especially when considering a technique that will generally apply to arbitrary $d$.
+The construction of the manipulation space overly complicates the manual controls, especially when considering possible techniques that will apply to arbitrary $d$.
 
@@
 
-The new approach emerged from experiments in mathematica.
+The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in @Mathematica.
 The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
@@
 1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
-2.
-Change values in row $m$, giving $A^*$.
-A large change in these values would correspond to making a large jump from the current projection.
+2. Change values in row $m$, for example, if $d=2$ gives 
+$$
+A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
+                              \vdots & \vdots \\
+                              a^*_{m1} & a^*_{m2}\\
+                              \vdots & \vdots \\
+                              a_{p1} & a_{p2} 
+        \end{array}\right].
+$$ 
+\noindent A large change in these values would correspond to making a large jump from the current projection.
 Small changes would correspond to tracking a cursor, making small jumps from the current projection.
-3. Orthonormalise $A^*$, using Gram-Schmidt. For $d=2$, and $A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]$, the steps are:
+3. Orthonormalise $A^*$, using Gram-Schmidt. 
+    i. Normalise $\boldsymbol{a}^*_1$ and $\boldsymbol{a}^*_2$.
+    ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - <\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2>\boldsymbol{a}^*_1$.
 
-    i. Normalise ${\boldmath a}_{.1}$, and ${\boldmath a}_{.2}$.
-    ii. ${\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}$.
-    iii. Normalise ${\boldmath a}^*_{.2}$.
-
-This algorithm will produce the changes to a projection as illustrated in Figure \ref{fig:manualsequence} (top row).
+This algorithm will produce the changes to a projection as illustrated in Figure \ref{fig:manualsequence}.
 The controlled variable, $V_m$, corresponds to the black line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (orange dot).
@@
 The problem with new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
-There are numerous ways that this can be enforced, and details are provided in the Appendix.
+There are numerous ways that this can be enforced, and details are in the Appendix.
 

### 2022-09-05 paper/paper.tex
algorithm fixed

@@
 
-From a statistical perspective it is rare to have data that are strictly
-3D, and so unlike in most computer graphics applications, the more
-useful methods for data analysis show projections from an arbitrary
-dimensional space. These are dynamic data visualizations methods and are
-collected under the term \emph{tours}. Tours involve views of
+From a statistical perspective 3D is a rare data dimension, so unlike in
+most 3D rotation computer graphics applications, the more useful methods
+for data analysis need to work for arbitrary dimension. A good approah
+is to show projections from an arbitrary dimensional space to create
+dynamic data visualizations called \emph{tours}. Tours involve views of
 high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
@@
 radial or angular changes in value. \citet{spyrison_spinifex_2020}
-refined this algorithm and implements them to generate animation
-sequences.
+refined this algorithm and implemented it to generate radial tour
+animation sequences.
 
@@
 
-\{\textcolor{blue}Another very useful feature is that the user can keep
-the projection fixed while slicing across the orthogonal space and
-changing the centre point used for slicing. This can serve to illustrate
-how the orthogonal space affects correlations in a slice and how they
-may change or disappear in a projection.\}
+Another potentially useful manual control, is to allow the user to
+control the position of the center of a slice. The slice tour was
+introduced in \citet{slicetour}. It operates by converting the
+projection plane into a slice, by removing or de-emphasizing points that
+are further than a fixed orthogonal distance from the plane. The
+projection plane is usually thought of as passing through the center of
+the data. Manual control would allow the user to change the value of
+center point, by shifting it along a coordinate axis, while keeping the
+orientation of the projection plane fixed. The purpose would be to
+explore how or if the shape of the data, in the space orthogonal to the
+projection, changes as one gets away from the center.
 
-This paper explains a simpler approach, which is more direct, and
-extensible for generating user interaction, and is organized as follows.
-The next section describes the new algorithm for manual control, for
-both projections and slices. The use of these methods are illustrated to
-compare and contrast boundaries constructed by different classifiers.
-The software section describes a mathematica notebook that is used for
-the application, and describes the interactive environment that would be
-desirable within R as new technology becomes available. The paper is
-accompanied by an appendix with more details and adjustments to the
-manual controls, and two mathematica notebooks, one focused on the
-application, and the other that could be used generically for new data.
+This paper explains the new manual controls for projection and slice
+tours. The next section describes the new algorithm for manual control,
+for both projections and slices. The use of these methods are
+illustrated to compare and contrast boundaries constructed by different
+classifiers. The software section describes a mathematica notebook that
+is used for the application, and describes the interactive environment
+that would be desirable within R as new technology becomes available.
+The paper is accompanied by an appendix with more details and
+adjustments to the manual controls, and two mathematica notebooks, one
+focused on the application, and the other that could be used generically
+for new data.
 
@@
 coefficient will be changed. A method to update the values of the
-component of the controlled variable \(V_m\) is then needed.
+component (\(m^{th}\) row of \(A_{p\times d}\)) of the controlled
+variable \(V_m\) is then needed.
 
@@
 first step was to construct a 3D manipulation space from a 2D
-projection. In this space the coefficient of the controlled variable
+projection. In this space, the coefficient of the controlled variable
 ranges between -1 and 1. Movements of a cursor are recorded and
@@
 of the manipulation space overly complicates the manual controls,
-especially when considering a technique that will generally apply to
+especially when considering possible techniques that will apply to
 arbitrary \(d\).
@@
 
-The new approach emerged from experiments in mathematica. The components
-corresponding to \(V_m\) are directly controlled by cursor movement,
-which updates row \(m\) of \(A\). The updated matrix is then
-orthonormalised.
+The new approach emerged from experiments on the tour using the linear
+algebra capabilities, and relatively new interactive graphics interface,
+available in \citet{Mathematica}. The components corresponding to
+\(V_m\) are directly controlled by cursor movement, which updates row
+\(m\) of \(A\). The updated matrix is then orthonormalised.
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\tightlist
 \item
@@
 \item
-  Change values in row \(m\), giving \(A^*\). A large change in these
-  values would correspond to making a large jump from the current
-  projection. Small changes would correspond to tracking a cursor,
-  making small jumps from the current projection.
+  Change values in row \(m\), for example, if \(d=2\) gives \[
+  A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
+                             \vdots & \vdots \\
+                             a^*_{m1} & a^*_{m2}\\
+                             \vdots & \vdots \\
+                             a_{p1} & a_{p2} 
+       \end{array}\right].
+  \] \noindent A large change in these values would correspond to making
+  a large jump from the current projection. Small changes would
+  correspond to tracking a cursor, making small jumps from the current
+  projection.
 \item
-  Orthonormalise \(A^*\), using Gram-Schmidt. For \(d=2\), and
-  \(A^* = \left[ {\boldmath a}_{.1}~{\boldmath a}_{.2}\right]\), the
-  steps are:
+  Orthonormalise \(A^*\), using Gram-Schmidt.
 
@@
   \item
-    Normalise \({\boldmath a}_{.1}\), and \({\boldmath a}_{.2}\).
+    Normalise \(\boldsymbol{a}^*_1\) and \(\boldsymbol{a}^*_2\).
   \item
-    \({\boldmath a}^*_{.2} = {\boldmath a}_{.2} - {\boldmath a}_{.1}^T{\boldmath a}_{.2}{\boldmath a}_{.1}\).
-  \item
-    Normalise \({\boldmath a}^*_{.2}\).
+    \(\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - <\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2>\boldsymbol{a}^*_1\).
   \end{enumerate}
@@
 This algorithm will produce the changes to a projection as illustrated
-in Figure \ref{fig:manualsequence} (top row). The controlled variable,
-\(V_m\), corresponds to the black line, and sequential changes to row
-\(m\) of \(A\) can be seen to roughly follow a specified position
-(orange dot). Changes in the other components happen as a result of the
+in Figure \ref{fig:manualsequence}. The controlled variable, \(V_m\),
+corresponds to the black line, and sequential changes to row \(m\) of
+\(A\) can be seen to roughly follow a specified position (orange dot).
+Changes in the other components happen as a result of the
 orthonormalisation, but are uncontrolled.
@@
 them. There are numerous ways that this can be enforced, and details are
-provided in the Appendix.
+in the Appendix.
 

### 2022-09-05 paper/paper.tex
software section

@@
 
-Explain \texttt{radial\_tour}
+The \citet{tourr} package provides numerous types of tours. An
+additional function has been constructed that will generate a sequence
+of projections that will decrease the coordinates for \(V_m\) to
+\(\boldsymbol{0}\) and back to the original values. It is applicable for
+any projection dimension, \(d\).
 

### 2022-09-05 paper/paper.Rmd
software section
Reworking the slicing section

@@
 To better explore the space we combine the manual controls for the projection with manual controls for slicing.
-A slice is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].
+A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].
+A data point is inside the slice if its orthogonal distance from the projection plane (passing through the center point) is below the thickness $h$.
+This orthogonal distance is computed in terms of the component that is normal on the projection plane.
+For $\mathbf{x}_i$ a $p$ dimensional data point and $\mathbf{c}$ the center point (in the same $p$ dimensional space) we compute the orthogonal distance as
+\begin{equation}
+v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
+\label{eq:slice}
+\end{equation}
+with $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$, $\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
 
@@
 
-In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points in steps along that direction to move the slice and fully cover the data space.
+In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points $\mathbf{c}$ in steps along that direction to move the slice and fully cover the data space.
 This no longer works in higher-dimensional spaces, and we can think of picking one direction and shifting the slice along the component orthogonal to the projection plane.
-In practice we define the center point explicitly and calculate the slice display as explained in @slicetour.
-The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.
-
-XXX Equation for arbitrary center for slice; and move the "the implementation ..." into the software section once that is constructed; and "potential adjustments for future" into the section on desirable controls
-
-Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here.
-An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
+In practice we define the center point $\mathbf{c}$ explicitly in terms of all $p$ components and calculate the slice display following Eq.
+\ref{eq:slice}.
 
@@
 
-In addition it is also useful to interactively change the slice thickness, in particular when selecting the preferred thickness $h$ (also called the slice radius) for exploring the input data.
-This was implemented as a slicer, and to allow the user to efficiently explore the interesting range we can provide the slider range as an input.
+In addition it is also useful to interactively change the slice thickness $h$ (also called the slice radius), in particular to find the preferred value for exploring the input data.
+This was implemented as a slider, and to allow the user to efficiently explore the interesting range we provide the slider range as an input.
 For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with
@@
 \end{equation}
-
-The current implementation contains a switch to change between the slice view and the projected view of the data, while keeping all other settings fixed.
 
@@
 
+XXX copied from the slicing part to be weaved in below:
+
+The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.
+
+The current implementation contains a switch to change between the slice view and the projected view of the data, while keeping all other settings fixed.
+
 ## Mathematica notebook
@@
 
-Explain `radial_tour`
+The @tourr package provides numerous types of tours.
+An additional function has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.
+It is applicable for any projection dimension, $d$.
 
@@
 ## Desirable interactivity
+
+Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here.
+An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
 

### 2022-09-05 paper/paper.tex
Reworking the slicing section

@@
 To better explore the space we combine the manual controls for the
-projection with manual controls for slicing. A slice is defined by a
-projection, a center point that is anchoring it in the high-dimensional
-space and the slice thickness \(h\) \citep{slicetour}.
+projection with manual controls for slicing. A slice is a section of the
+data that is defined by a projection, a center point that is anchoring
+it in the high-dimensional space and the slice thickness \(h\)
+\citep{slicetour}. A data point is inside the slice if its orthogonal
+distance from the projection plane (passing through the center point) is
+below the thickness \(h\). This orthogonal distance is computed in terms
+of the component that is normal on the projection plane. For
+\(\mathbf{x}_i\) a \(p\) dimensional data point and \(\mathbf{c}\) the
+center point (in the same \(p\) dimensional space) we compute the
+orthogonal distance as \begin{equation}
+v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
+\label{eq:slice}
+\end{equation} with
+\(\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2\),
+\(\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2\)
+and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
+matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\).
 
@@
 In the case of a single orthogonal direction on the projection plane we
-can pick a sequence of center points in steps along that direction to
-move the slice and fully cover the data space. This no longer works in
-higher-dimensional spaces, and we can think of picking one direction and
-shifting the slice along the component orthogonal to the projection
-plane. In practice we define the center point explicitly and calculate
-the slice display as explained in \citet{slicetour}. The current
-implementation allows the user to individually enter the components of
-the vector defining the center point, and updates the display by jumping
-to the new slice.
-
-XXX Equation for arbitrary center for slice; and move the ``the
-implementation \ldots{}'' into the software section once that is
-constructed; and ``potential adjustments for future'' into the section
-on desirable controls
-
-Potential adjustments for future work include a better visualization of
-the current center point, potentially with interactive graphics to move
-the slice similar to how the manual tour was implemented here. An
-alternative would be a step-wise shift of the slice towards the new
-center point to get a better understanding of the distribution along the
-selected direction.
+can pick a sequence of center points \(\mathbf{c}\) in steps along that
+direction to move the slice and fully cover the data space. This no
+longer works in higher-dimensional spaces, and we can think of picking
+one direction and shifting the slice along the component orthogonal to
+the projection plane. In practice we define the center point
+\(\mathbf{c}\) explicitly in terms of all \(p\) components and calculate
+the slice display following Eq. \ref{eq:slice}.
 
@@
 In addition it is also useful to interactively change the slice
-thickness, in particular when selecting the preferred thickness \(h\)
-(also called the slice radius) for exploring the input data. This was
-implemented as a slicer, and to allow the user to efficiently explore
-the interesting range we can provide the slider range as an input. For
-guidance the estimates of the number of points inside the slice as a
-function of the original sample size \(N\) and the number of dimensions
-\(p\) from \citet{sectionpursuit} can be used: in case of a uniform
-distribution inside a sphere of radius \(R\) a slice with thickness
-\(h\) will contain \(N_S\) points, with \begin{equation}
+thickness \(h\) (also called the slice radius), in particular to find
+the preferred value for exploring the input data. This was implemented
+as a slider, and to allow the user to efficiently explore the
+interesting range we provide the slider range as an input. For guidance
+the estimates of the number of points inside the slice as a function of
+the original sample size \(N\) and the number of dimensions \(p\) from
+\citet{sectionpursuit} can be used: in case of a uniform distribution
+inside a sphere of radius \(R\) a slice with thickness \(h\) will
+contain \(N_S\) points, with \begin{equation}
 N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
@@
 \end{equation}
-
-The current implementation contains a switch to change between the slice
-view and the projected view of the data, while keeping all other
-settings fixed.
 
@@
 
+XXX copied from the slicing part to be weaved in below:
+
+The current implementation allows the user to individually enter the
+components of the vector defining the center point, and updates the
+display by jumping to the new slice.
+
+The current implementation contains a switch to change between the slice
+view and the projected view of the data, while keeping all other
+settings fixed.
+
 \hypertarget{mathematica-notebook}{%
@@
 
+Potential adjustments for future work include a better visualization of
+the current center point, potentially with interactive graphics to move
+the slice similar to how the manual tour was implemented here. An
+alternative would be a step-wise shift of the slice towards the new
+center point to get a better understanding of the distribution along the
+selected direction.
+
 \hypertarget{sec:discussion}{%

### 2022-09-07 paper/paper.Rmd
conclusions mapped out

@@
 
-The @tourr package provides numerous types of tours.
-An additional function has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.
+The R package tourr [@tourr] provides numerous types of tours.
+An additional function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.
 It is applicable for any projection dimension, $d$.
@@
 
-Recap contribution
+This short note has described new technology for manually interacting with tours.
+The manual control of projection coefficients is most useful for assessing variable importance to perceived structure, but can be generally used for steering a viewer through high-dimensional space.
+Changing the center of a slice manually enables exploring the space orthogonal to a projection, specifically in the direction of a single variable.
+This is a new tool, and as shown in the example, when used with the slice tour can very useful for understanding boundaries of classifiers.
+This is also likely a useful method for examining low-dimensional non-linear manifolds, and also functions of multiple parameters.
+
+Mathematica provided a useful sandbox to experiment with the ideas presented in this paper.
+Most of the data, though, is first constructed using R, especially the classification boundaries.
+If new technology for interactive graphics becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.
+
+<!-- Recap contribution
 
@@
 
-Maybe extensions for slicing
+Maybe extensions for slicing-->
 

### 2022-09-07 paper/paper.tex
conclusions mapped out

@@
 
-The \citet{tourr} package provides numerous types of tours. An
-additional function has been constructed that will generate a sequence
-of projections that will decrease the coordinates for \(V_m\) to
-\(\boldsymbol{0}\) and back to the original values. It is applicable for
-any projection dimension, \(d\).
+The R package tourr \citep{tourr} provides numerous types of tours. An
+additional function, \texttt{radial\_tour()} has been constructed that
+will generate a sequence of projections that will decrease the
+coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
+values. It is applicable for any projection dimension, \(d\).
 
@@
 
-Recap contribution
+This short note has described new technology for manually interacting
+with tours. The manual control of projection coefficients is most useful
+for assessing variable importance to perceived structure, but can be
+generally used for steering a viewer through high-dimensional space.
+Changing the center of a slice manually enables exploring the space
+orthogonal to a projection, specifically in the direction of a single
+variable. This is a new tool, and as shown in the example, when used
+with the slice tour can very useful for understanding boundaries of
+classifiers. This is also likely a useful method for examining
+low-dimensional non-linear manifolds, and also functions of multiple
+parameters.
 
-Other applications
-
-Where else might it be useful to use mathematica
-
-Maybe extensions for slicing
+Mathematica provided a useful sandbox to experiment with the ideas
+presented in this paper. Most of the data, though, is first constructed
+using R, especially the classification boundaries. If new technology for
+interactive graphics becomes available in R it would be useful to create
+a tighter coupling of models and visualization to allow exploring and
+comparing fits.
 

### 2022-09-07 paper/paper.Rmd
Software before application.
Re-doing mathematica plots from the new notebooks, to match videos

@@
 
-# Application {#sec:examples}
-
-To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins] and look at classification models following @sam11271.
-We will show how classification boundaries can be explored and better understood on projections and slices through 4D space.
-Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
-There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
-The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
-The steps for exploring boundaries in this example are as follows:
-
-1. Build your classification model. 
-2. Predict the class for a dense grid of values covering the data space.
-3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
-4. Slice through the center, to explore where the boundaries will likely meet.
-5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
-
-## Constructing the 4D prediction regions
-
-We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest (RF).
-Both the data and the model points in the grid are centered and scaled (standard deviation $= 1$).
-
-## Exploring projections manually
-
-We start by first exploring the projections of the model prediction.
-Figure \ref{proj1} summarises the process.
-Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
-To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a projection that shows the non-linear but block-type structure that is typical of this type of model.
-This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
-What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes.
-And, interestingly this particular projection of the original data shows very distinct clusters of the three species.
-That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
-The three groups are nicely separated in this projection of the data (right)}.
-\label{proj1}
-\end{figure*}
-
-## Slicing through the center
-
-XXXX currently in the notebook $A_1$ is called A5, and $A_2$ is called A6
-
-We continue the investigation by now slicing orthogonally to  the projection $A_1$.
-For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0 = (0,0,0,0)$.
-At  first we explore how changing the projection, and thus the slice) away from $A_1$ can help with understanding the boundary better.
-For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
-
-Figure \ref{slice1} shows snapshots of the exploration.
-The rop row is the initial projection ($A_1$ and $S_1^0$), and the bottom row is a later projection containing more of `bd` (call them $A_2$ and $S_2^0$).
-The columns correspond to the two models, with Rf on the left and LDA on the right.
-
-Both models have some overlap at the center, even for this thin slice ($S_1^0$).
-The second slice ($S_2^0$) mostly resolves this overlap and reveals the primary differences between the models.
-The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
-The boundary between Chinstrap and Gentoo (green) is where they differ.
-
-The RF is almost straight in this view, as something we might expect from a tree model where splitting occurs on single variables only.
-However, it is straight in a combination of the second and third variable (`bd` and `fl`).
-By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two speciesfor any tree, and likely only one is needed.
-Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
-
-In contrast, the LDA boundary has an oblique split between the two, and roughly divides the space into three similarly sized areas.
-It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary half-way between the group means.
-
-<!-- We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
-\ref{slice1}, top-left).
-This is similar for the LDA model (Fig.
-\ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
-
-By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
-\ref{slice1}).-->
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
-The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
-\label{slice1}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
-\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
-\label{proj2}
-\end{figure*}
-
-Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
-Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
-It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
-The LDA boundary is better placed for all species.
-
-## Shifting the slice center
-
-We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
-Instead of rotating to a different projection, we might also change the view by moving the slice along one axis in the 4D space.
-Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$, $S_1^c$ to either large positive or negative values ($c_{bd} = \pm 1.5$ after centering and scaling).
-We will label these slices as $S_1^{\pm}$.
-Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
-
-We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
-We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
-The two models have similar boundaries except that the RF is more of a step function.}
-\label{slice1p}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
-The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
-\label{slice1m}
-\end{figure*}
-
-A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
-On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
-Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
-
-Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
-\ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.
-
 # Software {#sec:implementation}
@@
 
+# Application {#sec:examples}
+
+To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins] and look at classification models following @sam11271.
+We will show how classification boundaries can be explored and better understood on projections and slices through 4D space.
+Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
+There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
+The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
+The steps for exploring boundaries in this example are as follows:
+
+1. Build your classification model. 
+2. Predict the class for a dense grid of values covering the data space.
+3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
+4. Slice through the center, to explore where the boundaries will likely meet.
+5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
+
+## Constructing the 4D prediction regions
+
+We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest (RF).
+Both the data and the model points in the grid are centered and scaled (standard deviation $= 1$).
+
+## Exploring projections manually
+
+We start by first exploring the projections of the model prediction.
+Figure \ref{proj1} summarises the process.
+Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
+To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a projection that shows the non-linear but block-type structure that is typical of this type of model.
+This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
+What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes.
+And, interestingly this particular projection of the original data shows very distinct clusters of the three species.
+That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
+We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
+The three groups are nicely separated in this projection of the data (right).}
+\label{proj1}
+\end{figure*}
+
+This example can be reproduced with the code in penguins_exploring_manually.nb and a run through is shown in the video at (post to vimeo, same name as the nb).
+
+## Slicing through the center
+
+XXXX currently in the notebook $A_1$ is called A5, and $A_2$ is called A6
+
+We continue the investigation by now slicing orthogonally to  the projection $A_1$.
+For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0 = (0,0,0,0)$.
+At  first we explore how changing the projection, and thus the slice) away from $A_1$ can help with understanding the boundary better.
+For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
+
+Figure \ref{slice1} shows snapshots of the exploration.
+The rop row is the initial projection ($A_1$ and $S_1^0$), and the bottom row is a later projection containing more of `bd` (call them $A_2$ and $S_2^0$).
+The columns correspond to the two models, with Rf on the left and LDA on the right.
+
+Both models have some overlap at the center, even for this thin slice ($S_1^0$).
+The second slice ($S_2^0$) mostly resolves this overlap and reveals the primary differences between the models.
+The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
+The boundary between Chinstrap and Gentoo (green) is where they differ.
+
+The RF is almost straight in this view, as something we might expect from a tree model where splitting occurs on single variables only.
+However, it is straight in a combination of the second and third variable (`bd` and `fl`).
+By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two speciesfor any tree, and likely only one is needed.
+Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
+
+In contrast, the LDA boundary has an oblique split between the two, and roughly divides the space into three similarly sized areas.
+It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary half-way between the group means.
+
+<!-- We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
+\ref{slice1}, top-left).
+This is similar for the LDA model (Fig.
+\ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
+
+By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
+\ref{slice1}).-->
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
+The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
+\label{slice1}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
+\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
+\label{proj2}
+\end{figure*}
+
+Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
+It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
+The LDA boundary is better placed for all species.
+
+This example can be reproduced with the code in penguins_slicing_through_center.nb and a run through is shown in the video at (post to vimeo, same name as the nb).
+
+## Shifting the slice center
+
+We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
+Instead of rotating to a different projection, we might also change the view by moving the slice along one axis in the 4D space.
+Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$, $S_1^c$ to either large positive or negative values ($c_{bd} = \pm 1.5$ after centering and scaling).
+We will label these slices as $S_1^{\pm}$.
+Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
+
+We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
+We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
+The two models have similar boundaries except that the RF is more of a step function.}
+\label{slice1p}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
+\label{slice1m}
+\end{figure*}
+
+A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
+The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
+On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
+Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
+
+Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
+\ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.
+
+This example can be reproduced with the code in penguins_shifting_slice_center.nb and a run through is shown in the video at (post to vimeo, same name as the nb but one for positive and one for negative shift).
+
 # Discussion {#sec:discussion}

### 2022-09-07 paper/paper.tex
Re-doing mathematica plots from the new notebooks, to match videos

@@
 
-\hypertarget{sec:examples}{%
-\section{Application}\label{sec:examples}}
-
-To illustrate the usefulness of the manual controls we use the 4D
-penguins data \citep{penguins} and look at classification models
-following \citet{sam11271}. We will show how classification boundaries
-can be explored and better understood on projections and slices through
-4D space. Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix
-of this data. There are four variables
-(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
-measuring the size of the penguins from three species (Adelie, Chinstrap
-and Gentoo). The scatterplot matrix shows that the three species appear
-to be likely separable, and that at least the Gentoo can be
-distinguished from the other two species when \texttt{bd} is paired with
-\texttt{fl} or \texttt{bm}. The steps for exploring boundaries in this
-example are as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Build your classification model.
-\item
-  Predict the class for a dense grid of values covering the data space.
-\item
-  Examine projections, using a manual tour so that the contribution of
-  any variable is controlled.
-\item
-  Slice through the center, to explore where the boundaries will likely
-  meet.
-\item
-  Move the slice by changing the center in the direction of a single
-  variable to explore the extent of a boundary for a single group
-  relative to a variable.
-\end{enumerate}
-
-\hypertarget{constructing-the-4d-prediction-regions}{%
-\subsection{Constructing the 4D prediction
-regions}\label{constructing-the-4d-prediction-regions}}
-
-We use the \texttt{classifly} package \citep{classifly} to generate
-predictions across the 4D cube spanned by the data, with two
-classification models: linear discriminant analysis (LDA) and random
-forest (RF). Both the data and the model points in the grid are centered
-and scaled (standard deviation \(= 1\)).
-
-\hypertarget{exploring-projections-manually}{%
-\subsection{Exploring projections
-manually}\label{exploring-projections-manually}}
-
-We start by first exploring the projections of the model prediction.
-Figure \ref{proj1} summarises the process. Through manual rotation of
-the view we can get a feeling for where in the space we primarily
-predict each of the three species, and we also get a sense of the
-difference between the two models. To illustrate this difference we have
-manually rotated the projection for the RF model (left plot) to identify
-a projection that shows the non-linear but block-type structure that is
-typical of this type of model. This particular projection (\(A_1\)) is
-exported so that the same projection can be used to show the LDA model
-(middle plot) and the actual data (right plot). What can be seen is the
-linearity of the LDA model, where the boundaries are linear and oblique
-to the variable axes. And, interestingly this particular projection of
-the original data shows very distinct clusters of the three species.
-That means, the obscuring of the boundaries between groups for both of
-the models is driven by what is happening in the orthogonal space to the
-plane of the selected projection.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
-The three groups are nicely separated in this projection of the data (right)}.
-\label{proj1}
-\end{figure*}
-
-\hypertarget{slicing-through-the-center}{%
-\subsection{Slicing through the
-center}\label{slicing-through-the-center}}
-
-XXXX currently in the notebook \(A_1\) is called A5, and \(A_2\) is
-called A6
-
-We continue the investigation by now slicing orthogonally to the
-projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
-through the center, \(S_1^0 = (0,0,0,0)\). At first we explore how
-changing the projection, and thus the slice) away from \(A_1\) can help
-with understanding the boundary better. For our example, notice that
-\(A_1\) does not contain any contribution from the second variable
-(\texttt{bd}), so we will first rotate this variable into the view.
-
-Figure \ref{slice1} shows snapshots of the exploration. The rop row is
-the initial projection (\(A_1\) and \(S_1^0\)), and the bottom row is a
-later projection containing more of \texttt{bd} (call them \(A_2\) and
-\(S_2^0\)). The columns correspond to the two models, with Rf on the
-left and LDA on the right.
-
-Both models have some overlap at the center, even for this thin slice
-(\(S_1^0\)). The second slice (\(S_2^0\)) mostly resolves this overlap
-and reveals the primary differences between the models. The boundary
-between Adelie (blue) and Chinstrap (yellow) is very similar for both
-models. The boundary between Chinstrap and Gentoo (green) is where they
-differ.
-
-The RF is almost straight in this view, as something we might expect
-from a tree model where splitting occurs on single variables only.
-However, it is straight in a combination of the second and third
-variable (\texttt{bd} and \texttt{fl}). By examining the scatterplot
-matrix, we can understand how this boundary was built: on each of
-\texttt{bd} and \texttt{fl} it is possible to cut on a value where most
-of the Gentoo penguins are different from the other two speciesfor any
-tree, and likely only one is needed. Thus the forest construction is
-providing a sample of trees that use one variable or the other variable
-to split, producing the blocky boundary on the combination of the two.
-
-In contrast, the LDA boundary has an oblique split between the two, and
-roughly divides the space into three similarly sized areas. It is almost
-like a textbook illustration of how LDA works for two variables (2D)
-where by assuming the clusters have equal variance-covariance, it places
-a boundary half-way between the group means.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
-The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
-\label{slice1}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
-\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
-\label{proj2}
-\end{figure*}
-
-Finally, to determine which model does the boundaries better, compare
-them with the \(A_2\) projection of the data (Figure \ref{proj2}).
-Gentoo is more separated from the other two in this projection, and one
-can imagine that the trees in the forest has greedily grasped any one of
-many places to make a split to separate the group. It might be argued
-though the that RF boundary is cut too close to the Chinstrap species,
-and might lead to some unnecessary misclassification with new data. The
-LDA boundary is better placed for all species.
-
-\hypertarget{shifting-the-slice-center}{%
-\subsection{Shifting the slice center}\label{shifting-the-slice-center}}
-
-We have seen that starting from \(A_1\) using the manual controls to
-change the contribution of the second variable we could find a clear
-separation boundary indicating the relation between this variable (bill
-depth) and the Gentoo penguin species. Instead of rotating to a
-different projection, we might also change the view by moving the slice
-along one axis in the 4D space. Here we will continue our exploration of
-the dependence on \texttt{bd} and move the slice defined by \(A_1\),
-\(S_1^c\) to either large positive or negative values
-(\(c_{bd} = \pm 1.5\) after centering and scaling). We will label these
-slices as \(S_1^{\pm}\). Here we will also look at slices of the
-observed data points, using a thicker slice (\(h=1.5\)) to capture
-enough points in a given view.
-
-We start by a comparison of the two models and the data distribution in
-\(S_1^{+}\), thus the slice is localized towards high values of bill
-depth in Figure \ref{slice1p}. We can see that all three slices (the two
-models and the data) contain almost no points from the third class
-(green, Gentoo), and that the decision boundary between the two models
-is very similar.
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
-The two models have similar boundaries except that the RF is more of a step function.}
-\label{slice1p}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
-The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
-\label{slice1m}
-\end{figure*}
-
-A more interesting comparison is found for \(S_1^{-}\), thus the slice
-localized towards low values of bill depth, shown in Figure
-\ref{slice1m}. The RF model (left) predicts all three species within
-this slice, with an interesting boundary for the third class (green,
-Gentoo). On the other hand the LDA model (middle) predominantly predicts
-the third class within the slice, this appears to be enforced through
-the linear structure of the model. Looking finally at the thick slice
-through the data we see that there are primarily observations from this
-class within the slice we can conclude that the two models have filled
-in the ``empty'' space (where we do not have any training observations)
-in very different ways and according to what we might expect given the
-model structure.
-
-Finally it is interesting to compare the slice views to the projection
-of the models seen in Fig. \ref{proj1} to better understand how the
-boundaries change along the \texttt{bd} direction and where the
-differences in the projections come from.
-
 \hypertarget{sec:implementation}{%
@@
 
+\hypertarget{sec:examples}{%
+\section{Application}\label{sec:examples}}
+
+To illustrate the usefulness of the manual controls we use the 4D
+penguins data \citep{penguins} and look at classification models
+following \citet{sam11271}. We will show how classification boundaries
+can be explored and better understood on projections and slices through
+4D space. Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix
+of this data. There are four variables
+(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
+measuring the size of the penguins from three species (Adelie, Chinstrap
+and Gentoo). The scatterplot matrix shows that the three species appear
+to be likely separable, and that at least the Gentoo can be
+distinguished from the other two species when \texttt{bd} is paired with
+\texttt{fl} or \texttt{bm}. The steps for exploring boundaries in this
+example are as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Build your classification model.
+\item
+  Predict the class for a dense grid of values covering the data space.
+\item
+  Examine projections, using a manual tour so that the contribution of
+  any variable is controlled.
+\item
+  Slice through the center, to explore where the boundaries will likely
+  meet.
+\item
+  Move the slice by changing the center in the direction of a single
+  variable to explore the extent of a boundary for a single group
+  relative to a variable.
+\end{enumerate}
+
+\hypertarget{constructing-the-4d-prediction-regions}{%
+\subsection{Constructing the 4D prediction
+regions}\label{constructing-the-4d-prediction-regions}}
+
+We use the \texttt{classifly} package \citep{classifly} to generate
+predictions across the 4D cube spanned by the data, with two
+classification models: linear discriminant analysis (LDA) and random
+forest (RF). Both the data and the model points in the grid are centered
+and scaled (standard deviation \(= 1\)).
+
+\hypertarget{exploring-projections-manually}{%
+\subsection{Exploring projections
+manually}\label{exploring-projections-manually}}
+
+We start by first exploring the projections of the model prediction.
+Figure \ref{proj1} summarises the process. Through manual rotation of
+the view we can get a feeling for where in the space we primarily
+predict each of the three species, and we also get a sense of the
+difference between the two models. To illustrate this difference we have
+manually rotated the projection for the RF model (left plot) to identify
+a projection that shows the non-linear but block-type structure that is
+typical of this type of model. This particular projection (\(A_1\)) is
+exported so that the same projection can be used to show the LDA model
+(middle plot) and the actual data (right plot). What can be seen is the
+linearity of the LDA model, where the boundaries are linear and oblique
+to the variable axes. And, interestingly this particular projection of
+the original data shows very distinct clusters of the three species.
+That means, the obscuring of the boundaries between groups for both of
+the models is driven by what is happening in the orthogonal space to the
+plane of the selected projection.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
+We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
+The three groups are nicely separated in this projection of the data (right).}
+\label{proj1}
+\end{figure*}
+
+This example can be reproduced with the code in
+penguins\_exploring\_manually.nb and a run through is shown in the video
+at (post to vimeo, same name as the nb).
+
+\hypertarget{slicing-through-the-center}{%
+\subsection{Slicing through the
+center}\label{slicing-through-the-center}}
+
+XXXX currently in the notebook \(A_1\) is called A5, and \(A_2\) is
+called A6
+
+We continue the investigation by now slicing orthogonally to the
+projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
+through the center, \(S_1^0 = (0,0,0,0)\). At first we explore how
+changing the projection, and thus the slice) away from \(A_1\) can help
+with understanding the boundary better. For our example, notice that
+\(A_1\) does not contain any contribution from the second variable
+(\texttt{bd}), so we will first rotate this variable into the view.
+
+Figure \ref{slice1} shows snapshots of the exploration. The rop row is
+the initial projection (\(A_1\) and \(S_1^0\)), and the bottom row is a
+later projection containing more of \texttt{bd} (call them \(A_2\) and
+\(S_2^0\)). The columns correspond to the two models, with Rf on the
+left and LDA on the right.
+
+Both models have some overlap at the center, even for this thin slice
+(\(S_1^0\)). The second slice (\(S_2^0\)) mostly resolves this overlap
+and reveals the primary differences between the models. The boundary
+between Adelie (blue) and Chinstrap (yellow) is very similar for both
+models. The boundary between Chinstrap and Gentoo (green) is where they
+differ.
+
+The RF is almost straight in this view, as something we might expect
+from a tree model where splitting occurs on single variables only.
+However, it is straight in a combination of the second and third
+variable (\texttt{bd} and \texttt{fl}). By examining the scatterplot
+matrix, we can understand how this boundary was built: on each of
+\texttt{bd} and \texttt{fl} it is possible to cut on a value where most
+of the Gentoo penguins are different from the other two speciesfor any
+tree, and likely only one is needed. Thus the forest construction is
+providing a sample of trees that use one variable or the other variable
+to split, producing the blocky boundary on the combination of the two.
+
+In contrast, the LDA boundary has an oblique split between the two, and
+roughly divides the space into three similarly sized areas. It is almost
+like a textbook illustration of how LDA works for two variables (2D)
+where by assuming the clusters have equal variance-covariance, it places
+a boundary half-way between the group means.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
+The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
+\label{slice1}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
+\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
+\label{proj2}
+\end{figure*}
+
+Finally, to determine which model does the boundaries better, compare
+them with the \(A_2\) projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one
+can imagine that the trees in the forest has greedily grasped any one of
+many places to make a split to separate the group. It might be argued
+though the that RF boundary is cut too close to the Chinstrap species,
+and might lead to some unnecessary misclassification with new data. The
+LDA boundary is better placed for all species.
+
+This example can be reproduced with the code in
+penguins\_slicing\_through\_center.nb and a run through is shown in the
+video at (post to vimeo, same name as the nb).
+
+\hypertarget{shifting-the-slice-center}{%
+\subsection{Shifting the slice center}\label{shifting-the-slice-center}}
+
+We have seen that starting from \(A_1\) using the manual controls to
+change the contribution of the second variable we could find a clear
+separation boundary indicating the relation between this variable (bill
+depth) and the Gentoo penguin species. Instead of rotating to a
+different projection, we might also change the view by moving the slice
+along one axis in the 4D space. Here we will continue our exploration of
+the dependence on \texttt{bd} and move the slice defined by \(A_1\),
+\(S_1^c\) to either large positive or negative values
+(\(c_{bd} = \pm 1.5\) after centering and scaling). We will label these
+slices as \(S_1^{\pm}\). Here we will also look at slices of the
+observed data points, using a thicker slice (\(h=1.5\)) to capture
+enough points in a given view.
+
+We start by a comparison of the two models and the data distribution in
+\(S_1^{+}\), thus the slice is localized towards high values of bill
+depth in Figure \ref{slice1p}. We can see that all three slices (the two
+models and the data) contain almost no points from the third class
+(green, Gentoo), and that the decision boundary between the two models
+is very similar.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
+The two models have similar boundaries except that the RF is more of a step function.}
+\label{slice1p}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
+\label{slice1m}
+\end{figure*}
+
+A more interesting comparison is found for \(S_1^{-}\), thus the slice
+localized towards low values of bill depth, shown in Figure
+\ref{slice1m}. The RF model (left) predicts all three species within
+this slice, with an interesting boundary for the third class (green,
+Gentoo). On the other hand the LDA model (middle) predominantly predicts
+the third class within the slice, this appears to be enforced through
+the linear structure of the model. Looking finally at the thick slice
+through the data we see that there are primarily observations from this
+class within the slice we can conclude that the two models have filled
+in the ``empty'' space (where we do not have any training observations)
+in very different ways and according to what we might expect given the
+model structure.
+
+Finally it is interesting to compare the slice views to the projection
+of the models seen in Fig. \ref{proj1} to better understand how the
+boundaries change along the \texttt{bd} direction and where the
+differences in the projections come from.
+
+This example can be reproduced with the code in
+penguins\_shifting\_slice\_center.nb and a run through is shown in the
+video at (post to vimeo, same name as the nb but one for positive and
+one for negative shift).
+
 \hypertarget{sec:discussion}{%

### 2022-09-08 paper/paper.Rmd
Video links added.

--- 2022-09-07 paper/paper.Rmd
@@
 
+# Application {#sec:examples}
+
+To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins] and look at classification models following @sam11271.
+We will show how classification boundaries can be explored and better understood on projections and slices through 4D space.
+Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
+There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
+The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
+The steps for exploring boundaries in this example are as follows:
+
+1. Build your classification model. 
+2. Predict the class for a dense grid of values covering the data space.
+3. Examine projections, using a manual tour so that the contribution of any variable is controlled.
+4. Slice through the center, to explore where the boundaries will likely meet.
+5. Move the slice by changing the center in the direction of a single variable to explore the extent of a boundary for a single group relative to a variable.
+
+## Constructing the 4D prediction regions
+
+We use the `classifly` package [@classifly] to generate predictions across the 4D cube spanned by the data, with two classification models: linear discriminant analysis (LDA) and random forest (RF).
+Both the data and the model points in the grid are centered and scaled (standard deviation $= 1$).
+
+## Exploring projections manually
+
+We start by first exploring the projections of the model prediction.
+Figure \ref{proj1} summarises the process.
+Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
+To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a projection that shows the non-linear but block-type structure that is typical of this type of model.
+This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
+What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes.
+And, interestingly this particular projection of the original data shows very distinct clusters of the three species.
+That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
+We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
+The three groups are nicely separated in this projection of the data (right).}
+\label{proj1}
+\end{figure*}
+
+This example can be reproduced with the code in penguins_exploring_manually.nb and a run through is shown in the video at https://vimeo.com/747585410.
+
+## Slicing through the center
+
+We continue the investigation by now slicing orthogonally to  the projection $A_1$.
+For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0 = (0,0,0,0)$.
+At  first we explore how changing the projection, and thus the slice) away from $A_1$ can help with understanding the boundary better.
+For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
+
+Figure \ref{slice1} shows snapshots of the exploration.
+The rop row is the initial projection ($A_1$ and $S_1^0$), and the bottom row is a later projection containing more of `bd` (call them $A_2$ and $S_2^0$).
+The columns correspond to the two models, with Rf on the left and LDA on the right.
+
+Both models have some overlap at the center, even for this thin slice ($S_1^0$).
+The second slice ($S_2^0$) mostly resolves this overlap and reveals the primary differences between the models.
+The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
+The boundary between Chinstrap and Gentoo (green) is where they differ.
+
+The RF is almost straight in this view, as something we might expect from a tree model where splitting occurs on single variables only.
+However, it is straight in a combination of the second and third variable (`bd` and `fl`).
+By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two speciesfor any tree, and likely only one is needed.
+Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
+
+In contrast, the LDA boundary has an oblique split between the two, and roughly divides the space into three similarly sized areas.
+It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary half-way between the group means.
+
+<!-- We again start by looking at the RF model, the slice $S_1^0$ shows the block structure of the model, with the third group (green, Gentoo penguins) overlapping with both of the other groups (Fig.
+\ref{slice1}, top-left).
+This is similar for the LDA model (Fig.
+\ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
+
+By rotating in the second variable we can find a view that shows three neatly separated blocks for the RF model, and we export the corresponding projection matrix $A_2$ to see the same views for the LDA model which also shows a neater separation of the classes along the boundary (see bottom row of Fig.
+\ref{slice1}).-->
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
+With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
+The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
+\label{slice1}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
+\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
+\label{proj2}
+\end{figure*}
+
+Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
+It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
+The LDA boundary is better placed for all species.
+
+This example can be reproduced with the code in penguins_slicing_through_center.nb and a run through is shown in the first part of the video at https://vimeo.com/747590472.
+
+## Shifting the slice center
+
+We have seen that starting from $A_1$ using the manual controls to change the contribution of the second variable we could find a clear separation boundary indicating the relation between this variable (bill depth) and the Gentoo penguin species.
+Instead of rotating to a different projection, we might also change the view by moving the slice along one axis in the 4D space.
+Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$, $S_1^c$ to either large positive or negative values ($c_{bd} = \pm 1.5$ after centering and scaling).
+We will label these slices as $S_1^{\pm}$.
+Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
+
+We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
+We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
+The two models have similar boundaries except that the RF is more of a step function.}
+\label{slice1p}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
+\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
+\label{slice1m}
+\end{figure*}
+
+A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
+The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
+On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
+Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
+
+Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
+\ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.
+
+This example can be reproduced with the code in penguins_shifting_slice_center.nb and a run through is shown in the second part of the video at https://vimeo.com/747590472.
+
 # Discussion {#sec:discussion}

### 2022-09-08 paper/paper.Rmd
Went over sections 1, 2, 4, updated notation for the application section.

@@
 From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, the more useful methods for data analysis need to work for arbitrary dimension.
-A good approah is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
+A good approch is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
 Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
@@
 
-A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d-D$ projection.
+A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d$ dimensional projection.
 The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed.
@@
     i. Normalise $\boldsymbol{a}^*_1$ and $\boldsymbol{a}^*_2$.
-    ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - <\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2>\boldsymbol{a}^*_1$.
+    ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1$.
+
+<!--
+Note: I removed the angular bracets since cdot denotes scalar product already (also used like this below for the slice definitions)
+-->
 
@@
 There are numerous ways that this can be enforced, and details are in the Appendix.
+\textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
 
@@
 
+A natural starting point is to place $\mathbf{c}$ in the center of the data distribution, but shifting it away from the mean can provide additional insights.
 In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points $\mathbf{c}$ in steps along that direction to move the slice and fully cover the data space.
 This no longer works in higher-dimensional spaces, and we can think of picking one direction and shifting the slice along the component orthogonal to the projection plane.
-In practice we define the center point $\mathbf{c}$ explicitly in terms of all $p$ components and calculate the slice display following Eq.
-\ref{eq:slice}.
 
@@
 In addition it is also useful to interactively change the slice thickness $h$ (also called the slice radius), in particular to find the preferred value for exploring the input data.
-This was implemented as a slider, and to allow the user to efficiently explore the interesting range we provide the slider range as an input.
 For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with
 \begin{equation}
-N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
+N_S(h, p, R, N) = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
 \label{eq:count}
 \end{equation}
-
-```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
-library(tidyverse)
-library(palmerpenguins)
-library(GGally)
-penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
-  drop_na() %>%
-  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
-  rename("bl"="bill_length_mm",
-         "bd"="bill_depth_mm",
-         "fl"="flipper_length_mm",
-         "bm"="body_mass_g")
-ggscatmat(penguins_data, columns = 2:5, col="species") +
-  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-  theme(legend.position = "bottom")
-```
 
@@
 
+In practice we define the center point $\mathbf{c}$ explicitly in terms of all $p$ components and calculate the slice display following Eq. \ref{eq:slice}.
+
+This was implemented as a slider, and to allow the user to efficiently explore the interesting range we provide the slider range as an input.
+
 ## Mathematica notebook
@@
 Most of this inbuilt functionality is quite user-friendly and can be found in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
-Also, it is relatively simple  to create dynamic objects via the inbuilt commands Manipulate or DynamicModule.
+Also, it is relatively simple  to create dynamic objects via the inbuilt commands `Manipulate` or `DynamicModule`.
 Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that dynamic variable will be updated.
@@
 
-The examples presented in the penguins’ notebooks all utilise our primary new function, SliceDynamic.
+The examples presented in the penguins’ notebooks all utilise our primary new function, `SliceDynamic`.
 This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index.
@@
 
+```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
+The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
+library(tidyverse)
+library(palmerpenguins)
+library(GGally)
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
+  rename("bl"="bill_length_mm",
+         "bd"="bill_depth_mm",
+         "fl"="flipper_length_mm",
+         "bm"="body_mass_g")
+ggscatmat(penguins_data, columns = 2:5, col="species") +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  theme(legend.position = "bottom")
+```
+
 To illustrate the usefulness of the manual controls we use the 4D penguins data [@penguins] and look at classification models following @sam11271.
@@
 Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
-To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a projection that shows the non-linear but block-type structure that is typical of this type of model.
+To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model.
 This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
@@
 We continue the investigation by now slicing orthogonally to  the projection $A_1$.
-For both models we look at a thin ($h=0.5$) slice through the center, $S_1^0 = (0,0,0,0)$.
-At  first we explore how changing the projection, and thus the slice) away from $A_1$ can help with understanding the boundary better.
+For both models we look at a thin ($h=0.5$) slice through the center, $\mathbf{c}^0 = (0,0,0,0)$.
+At  first we explore how changing the projection away from $A_1$ can help with understanding the boundary better.
 For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
@@
 Figure \ref{slice1} shows snapshots of the exploration.
-The rop row is the initial projection ($A_1$ and $S_1^0$), and the bottom row is a later projection containing more of `bd` (call them $A_2$ and $S_2^0$).
-The columns correspond to the two models, with Rf on the left and LDA on the right.
+The top row is the initial projection ($A_1$ and $\mathbf{c}^0$), and the bottom row is a later projection containing more of `bd` (new projection $A_2$ while keeping the center fixed at $\mathbf{c}^0$).
+The columns correspond to the two models, with RF on the left and LDA on the right.
 
-Both models have some overlap at the center, even for this thin slice ($S_1^0$).
-The second slice ($S_2^0$) mostly resolves this overlap and reveals the primary differences between the models.
+Both models have some overlap at the center for the thin slice based on $A_1$ (despite the small value for $h$).
+The second slice (based on $A_2$) mostly resolves this overlap and reveals the primary differences between the models.
 The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
@@
 However, it is straight in a combination of the second and third variable (`bd` and `fl`).
-By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two speciesfor any tree, and likely only one is needed.
+By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two species for any tree, and likely only one is needed.
 Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
@@
 Instead of rotating to a different projection, we might also change the view by moving the slice along one axis in the 4D space.
-Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$, $S_1^c$ to either large positive or negative values ($c_{bd} = \pm 1.5$ after centering and scaling).
-We will label these slices as $S_1^{\pm}$.
+Here we will continue our exploration of the dependence on `bd` and move the slice defined by $A_1$ to either large positive or negative values ($c^{\pm}_{bd} = \pm 1.5$ after centering and scaling).
+Thus we define $\mathbf{c}^{\pm} = (0, c^{\pm}_{bd}, 0, 0)$.
 Here we will also look at slices of the observed data points, using a thicker slice ($h=1.5$) to capture enough points in a given view.
 
-We start by a comparison of the two models and the data distribution in $S_1^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
+We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
 We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
@@
 
-A more interesting comparison is found for $S_1^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
+A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
 The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).

### 2022-09-08 paper/paper.tex
Went over sections 1, 2, 4, updated notation for the application section.

--- 2022-09-07 paper/paper.tex
@@
 most 3D rotation computer graphics applications, the more useful methods
-for data analysis need to work for arbitrary dimension. A good approah
+for data analysis need to work for arbitrary dimension. A good approch
 is to show projections from an arbitrary dimensional space to create
@@
 A manual tour allows the user to alter the coefficients of one (or more)
-variables contributing to a \(d-D\) projection. The initial ingredients
-are an orthonormal basis (\(A_{p\times d}\)) defining the projection of
-the data, and a variable id (\(m \in \{1, ..., p\}\)) specifying which
-coefficient will be changed. A method to update the values of the
-component (\(m^{th}\) row of \(A_{p\times d}\)) of the controlled
-variable \(V_m\) is then needed.
+variables contributing to a \(d\) dimensional projection. The initial
+ingredients are an orthonormal basis (\(A_{p\times d}\)) defining the
+projection of the data, and a variable id (\(m \in \{1, ..., p\}\))
+specifying which coefficient will be changed. A method to update the
+values of the component (\(m^{th}\) row of \(A_{p\times d}\)) of the
+controlled variable \(V_m\) is then needed.
 
@@
   \item
-    \(\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - <\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2>\boldsymbol{a}^*_1\).
+    \(\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1\).
   \end{enumerate}
@@
 in the Appendix.
+\textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
 
@@
 
-In the case of a single orthogonal direction on the projection plane we
-can pick a sequence of center points \(\mathbf{c}\) in steps along that
-direction to move the slice and fully cover the data space. This no
-longer works in higher-dimensional spaces, and we can think of picking
-one direction and shifting the slice along the component orthogonal to
-the projection plane. In practice we define the center point
-\(\mathbf{c}\) explicitly in terms of all \(p\) components and calculate
-the slice display following Eq. \ref{eq:slice}.
+A natural starting point is to place \(\mathbf{c}\) in the center of the
+data distribution, but shifting it away from the mean can provide
+additional insights. In the case of a single orthogonal direction on the
+projection plane we can pick a sequence of center points \(\mathbf{c}\)
+in steps along that direction to move the slice and fully cover the data
+space. This no longer works in higher-dimensional spaces, and we can
+think of picking one direction and shifting the slice along the
+component orthogonal to the projection plane.
 
@@
 thickness \(h\) (also called the slice radius), in particular to find
-the preferred value for exploring the input data. This was implemented
-as a slider, and to allow the user to efficiently explore the
-interesting range we provide the slider range as an input. For guidance
-the estimates of the number of points inside the slice as a function of
-the original sample size \(N\) and the number of dimensions \(p\) from
+the preferred value for exploring the input data. For guidance the
+estimates of the number of points inside the slice as a function of the
+original sample size \(N\) and the number of dimensions \(p\) from
 \citet{sectionpursuit} can be used: in case of a uniform distribution
@@
 contain \(N_S\) points, with \begin{equation}
-N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
+N_S(h, p, R, N) = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
 \label{eq:count}
 \end{equation}
+
+\hypertarget{sec:implementation}{%
+\section{Software}\label{sec:implementation}}
+
+XXX copied from the slicing part to be weaved in below:
+
+The current implementation allows the user to individually enter the
+components of the vector defining the center point, and updates the
+display by jumping to the new slice.
+
+The current implementation contains a switch to change between the slice
+view and the projected view of the data, while keeping all other
+settings fixed.
+
+In practice we define the center point \(\mathbf{c}\) explicitly in
+terms of all \(p\) components and calculate the slice display following
+Eq. \ref{eq:slice}.
+
+This was implemented as a slider, and to allow the user to efficiently
+explore the interesting range we provide the slider range as an input.
+
+\hypertarget{mathematica-notebook}{%
+\subsection{Mathematica notebook}\label{mathematica-notebook}}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Why is this a good sandbox}
+\item
+  \textbf{Explain the functionality available in the notebooks}
+\end{itemize}
+
+Mathematica provides much utility and versatility, such as inbuilt data
+visualisation, data manipulation and analysis, dynamic functionality,
+and symbolic and numeric computation. Most of this inbuilt functionality
+is quite user-friendly and can be found in the Mathematica
+documentation; typically, numerous examples are provided within the
+documentation, and some possible issues are outlined there. Also, it is
+relatively simple to create dynamic objects via the inbuilt commands
+\texttt{Manipulate} or \texttt{DynamicModule}. Control objects, such as
+sliders, locator panes, and input fields, can then be used on dynamic
+variables, and when there is a change in the dynamic variable the
+dynamic objects which contain that dynamic variable will be updated.
+This is the essential ingredient for this implementation of the manual
+tour. However, it can be hard to alter the behaviour of the control
+objects such as setting the 1D slider on a different scale (Slider
+increments add 10 points into the slice rather than increasing the slice
+height by 0.1 for example). \{\textcolor{red}Alex, is this still true
+with the controls you are using now?\}
+
+The examples presented in the penguins' notebooks all utilise our
+primary new function, \texttt{SliceDynamic}. This function typically
+accepts grouped data in the form of a matrix where the second last
+column details the name of the group and the last column details the
+group index. After specifying the initial slice height and the slice
+range, the user is then presented with an interactive display in which
+the control objects appear on the left and the slice visulisation
+appears on the right. The user can change the orientation of the slice
+via the locator pane, which changes the projection matrix; a slider
+controls the slice height; and there is an input field, which changes
+the slice centre. The user can also change the appearance of the plot by
+zooming into the center or changing the point size with sliders
+provided. The current projection can be displayed to contrast it with
+the slice by ticking a box. The explicit projection matrix can be
+displayed via another checkbox. The Slice Tour* notebooks include the
+additional functions: ProjectionPlot, Projected2DSliderPlot, and
+VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic,
+except it only displays projections and VisualiseSliceDynamic allows you
+to discern which points exist inside the slice and outside it. However,
+ProjectedLocatorPlot displays the interactive slightly differently.
+Several locator panes are displayed above the plot and each one
+corresponds to a row in the projection matrix but behaves the same as
+ProjectionPlot. \{\textcolor{blue}(this may be easier to implement in
+another language (?) )\}
+
+\hypertarget{extensions-to-the-r-package-tourr}{%
+\subsection{\texorpdfstring{Extensions to the R package
+\texttt{tourr}}{Extensions to the R package tourr}}\label{extensions-to-the-r-package-tourr}}
+
+The R package tourr \citep{tourr} provides numerous types of tours. An
+additional function, \texttt{radial\_tour()} has been constructed that
+will generate a sequence of projections that will decrease the
+coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
+values. It is applicable for any projection dimension, \(d\).
+
+Changing slice center
+
+\hypertarget{desirable-interactivity}{%
+\subsection{Desirable interactivity}\label{desirable-interactivity}}
+
+Potential adjustments for future work include a better visualization of
+the current center point, potentially with interactive graphics to move
+the slice similar to how the manual tour was implemented here. An
+alternative would be a step-wise shift of the slice towards the new
+center point to get a better understanding of the distribution along the
+selected direction.
+
+\hypertarget{sec:examples}{%
+\section{Application}\label{sec:examples}}
 
@@
 \end{figure}
-
-\hypertarget{sec:examples}{%
-\section{Application}\label{sec:examples}}
 
@@
 manually rotated the projection for the RF model (left plot) to identify
-a projection that shows the non-linear but block-type structure that is
+a view that shows the non-linear but block-type structure that is
 typical of this type of model. This particular projection (\(A_1\)) is
@@
 We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
-The three groups are nicely separated in this projection of the data (right)}.
+The three groups are nicely separated in this projection of the data (right).}
 \label{proj1}
 \end{figure*}
+
+This example can be reproduced with the code in
+penguins\_exploring\_manually.nb and a run through is shown in the video
+at \url{https://vimeo.com/747585410}.
 
@@
 
-XXXX currently in the notebook \(A_1\) is called A5, and \(A_2\) is
-called A6
-
 We continue the investigation by now slicing orthogonally to the
 projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
-through the center, \(S_1^0 = (0,0,0,0)\). At first we explore how
-changing the projection, and thus the slice) away from \(A_1\) can help
-with understanding the boundary better. For our example, notice that
-\(A_1\) does not contain any contribution from the second variable
+through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first we explore
+how changing the projection away from \(A_1\) can help with
+understanding the boundary better. For our example, notice that \(A_1\)
+does not contain any contribution from the second variable
 (\texttt{bd}), so we will first rotate this variable into the view.
 
-Figure \ref{slice1} shows snapshots of the exploration. The rop row is
-the initial projection (\(A_1\) and \(S_1^0\)), and the bottom row is a
-later projection containing more of \texttt{bd} (call them \(A_2\) and
-\(S_2^0\)). The columns correspond to the two models, with Rf on the
-left and LDA on the right.
+Figure \ref{slice1} shows snapshots of the exploration. The top row is
+the initial projection (\(A_1\) and \(\mathbf{c}^0\)), and the bottom
+row is a later projection containing more of \texttt{bd} (new projection
+\(A_2\) while keeping the center fixed at \(\mathbf{c}^0\)). The columns
+correspond to the two models, with RF on the left and LDA on the right.
 
-Both models have some overlap at the center, even for this thin slice
-(\(S_1^0\)). The second slice (\(S_2^0\)) mostly resolves this overlap
-and reveals the primary differences between the models. The boundary
-between Adelie (blue) and Chinstrap (yellow) is very similar for both
-models. The boundary between Chinstrap and Gentoo (green) is where they
-differ.
+Both models have some overlap at the center for the thin slice based on
+\(A_1\) (despite the small value for \(h\)). The second slice (based on
+\(A_2\)) mostly resolves this overlap and reveals the primary
+differences between the models. The boundary between Adelie (blue) and
+Chinstrap (yellow) is very similar for both models. The boundary between
+Chinstrap and Gentoo (green) is where they differ.
 
@@
 \texttt{bd} and \texttt{fl} it is possible to cut on a value where most
-of the Gentoo penguins are different from the other two speciesfor any
+of the Gentoo penguins are different from the other two species for any
 tree, and likely only one is needed. Thus the forest construction is
@@
 
+This example can be reproduced with the code in
+penguins\_slicing\_through\_center.nb and a run through is shown in the
+first part of the video at \url{https://vimeo.com/747590472}.
+
 \hypertarget{shifting-the-slice-center}{%
@@
 along one axis in the 4D space. Here we will continue our exploration of
-the dependence on \texttt{bd} and move the slice defined by \(A_1\),
-\(S_1^c\) to either large positive or negative values
-(\(c_{bd} = \pm 1.5\) after centering and scaling). We will label these
-slices as \(S_1^{\pm}\). Here we will also look at slices of the
-observed data points, using a thicker slice (\(h=1.5\)) to capture
-enough points in a given view.
+the dependence on \texttt{bd} and move the slice defined by \(A_1\) to
+either large positive or negative values (\(c^{\pm}_{bd} = \pm 1.5\)
+after centering and scaling). Thus we define
+\(\mathbf{c}^{\pm} = (0, c^{\pm}_{bd}, 0, 0)\). Here we will also look
+at slices of the observed data points, using a thicker slice (\(h=1.5\))
+to capture enough points in a given view.
 
-We start by a comparison of the two models and the data distribution in
-\(S_1^{+}\), thus the slice is localized towards high values of bill
-depth in Figure \ref{slice1p}. We can see that all three slices (the two
-models and the data) contain almost no points from the third class
-(green, Gentoo), and that the decision boundary between the two models
-is very similar.
+We start by a comparison of the two models and the data distribution
+shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
+values of bill depth in Figure \ref{slice1p}. We can see that all three
+slices (the two models and the data) contain almost no points from the
+third class (green, Gentoo), and that the decision boundary between the
+two models is very similar.
 
@@
 
-A more interesting comparison is found for \(S_1^{-}\), thus the slice
-localized towards low values of bill depth, shown in Figure
+A more interesting comparison is found for \(\mathbf{c}^{-}\), thus the
+slice localized towards low values of bill depth, shown in Figure
 \ref{slice1m}. The RF model (left) predicts all three species within
@@
 
-\hypertarget{sec:implementation}{%
-\section{Software}\label{sec:implementation}}
-
-XXX copied from the slicing part to be weaved in below:
-
-The current implementation allows the user to individually enter the
-components of the vector defining the center point, and updates the
-display by jumping to the new slice.
-
-The current implementation contains a switch to change between the slice
-view and the projected view of the data, while keeping all other
-settings fixed.
-
-\hypertarget{mathematica-notebook}{%
-\subsection{Mathematica notebook}\label{mathematica-notebook}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Why is this a good sandbox}
-\item
-  \textbf{Explain the functionality available in the notebooks}
-\end{itemize}
-
-Mathematica provides much utility and versatility, such as inbuilt data
-visualisation, data manipulation and analysis, dynamic functionality,
-and symbolic and numeric computation. Most of this inbuilt functionality
-is quite user-friendly and can be found in the Mathematica
-documentation; typically, numerous examples are provided within the
-documentation, and some possible issues are outlined there. Also, it is
-relatively simple to create dynamic objects via the inbuilt commands
-Manipulate or DynamicModule. Control objects, such as sliders, locator
-panes, and input fields, can then be used on dynamic variables, and when
-there is a change in the dynamic variable the dynamic objects which
-contain that dynamic variable will be updated. This is the essential
-ingredient for this implementation of the manual tour. However, it can
-be hard to alter the behaviour of the control objects such as setting
-the 1D slider on a different scale (Slider increments add 10 points into
-the slice rather than increasing the slice height by 0.1 for example).
-\{\textcolor{red}Alex, is this still true with the controls you are
-using now?\}
-
-The examples presented in the penguins' notebooks all utilise our
-primary new function, SliceDynamic. This function typically accepts
-grouped data in the form of a matrix where the second last column
-details the name of the group and the last column details the group
-index. After specifying the initial slice height and the slice range,
-the user is then presented with an interactive display in which the
-control objects appear on the left and the slice visulisation appears on
-the right. The user can change the orientation of the slice via the
-locator pane, which changes the projection matrix; a slider controls the
-slice height; and there is an input field, which changes the slice
-centre. The user can also change the appearance of the plot by zooming
-into the center or changing the point size with sliders provided. The
-current projection can be displayed to contrast it with the slice by
-ticking a box. The explicit projection matrix can be displayed via
-another checkbox. The Slice Tour* notebooks include the additional
-functions: ProjectionPlot, Projected2DSliderPlot, and
-VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic,
-except it only displays projections and VisualiseSliceDynamic allows you
-to discern which points exist inside the slice and outside it. However,
-ProjectedLocatorPlot displays the interactive slightly differently.
-Several locator panes are displayed above the plot and each one
-corresponds to a row in the projection matrix but behaves the same as
-ProjectionPlot. \{\textcolor{blue}(this may be easier to implement in
-another language (?) )\}
-
-\hypertarget{extensions-to-the-r-package-tourr}{%
-\subsection{\texorpdfstring{Extensions to the R package
-\texttt{tourr}}{Extensions to the R package tourr}}\label{extensions-to-the-r-package-tourr}}
-
-The R package tourr \citep{tourr} provides numerous types of tours. An
-additional function, \texttt{radial\_tour()} has been constructed that
-will generate a sequence of projections that will decrease the
-coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
-values. It is applicable for any projection dimension, \(d\).
-
-Changing slice center
-
-\hypertarget{desirable-interactivity}{%
-\subsection{Desirable interactivity}\label{desirable-interactivity}}
-
-Potential adjustments for future work include a better visualization of
-the current center point, potentially with interactive graphics to move
-the slice similar to how the manual tour was implemented here. An
-alternative would be a step-wise shift of the slice towards the new
-center point to get a better understanding of the distribution along the
-selected direction.
+This example can be reproduced with the code in
+penguins\_shifting\_slice\_center.nb and a run through is shown in the
+second part of the video at \url{https://vimeo.com/747590472}.
 

### 2022-09-12 paper/appendix.Rmd
went through sections 1 and 2 and moved a paragraph of mathematica to appendix as requested by Ursula

@@
 
-Other functions are:
+The Slice Tour* notebooks include the additional functions: ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic.
+ProjectionPlot is very similar to SliceDynamic, except it only displays projections and VisualiseSliceDynamic allows you to discern which points exist inside the slice and outside it.
+However, ProjectedLocatorPlot displays the interactive slightly differently.
+Several locator panes are displayed above the plot and each one corresponds to a row in the projection matrix but behaves the same as ProjectionPlot.
 

### 2022-09-12 paper/appendix.tex
went through sections 1 and 2 and moved a paragraph of mathematica to appendix as requested by Ursula

@@
 
-Other functions are:
+The Slice Tour* notebooks include the additional functions:
+ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic.
+ProjectionPlot is very similar to SliceDynamic, except it only displays
+projections and VisualiseSliceDynamic allows you to discern which points
+exist inside the slice and outside it. However, ProjectedLocatorPlot
+displays the interactive slightly differently. Several locator panes are
+displayed above the plot and each one corresponds to a row in the
+projection matrix but behaves the same as ProjectionPlot.
 

### 2022-09-12 paper/paper.Rmd
went through sections 1 and 2 and moved a paragraph of mathematica to appendix as requested by Ursula

@@
   This paper describes new user controls for examining high-dimensional data using low-dimensional linear projections and slices.
-A user can interactively change the coefficients of a variable in a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables.
+A user can interactively change the \textcolor{blue} {coefficients of a variable in} \textcolor{red}{contribution of a given variable to} a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables.
 The user can also interactively shift the slice center, which is useful for exploring how structure changes in local subspaces.
-These controls are implemented in Mathematica to take advantage of it's mathematics functions, and interactivity tools that allow construction of functions that interact with the position of the cursor.
+These controls are implemented in Mathematica to take advantage of \textcolor{blue}{it's} \textcolor{red}{its} \textcolor{blue}{remove this: mathematics functions, and interactivity tools} \textcolor{red}{location controls} that allow construction of functions that interact with the position of the cursor.
 We demonstrate the usefulness of this new tool with an example exploring region boundaries in classification problems.
@@
 (A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, which are summarized in @lee2021.
+After Asimov there have been many \textcolor{blue}{remove: }, many} tour developments, which are summarized in @lee2021.
 
@@
 There are many places where it is useful.
-In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering.
+In exploratory data analysis, where one sees clusters in a projection, \textcolor{red}{} one may ask whether some} variables can be removed from the projection without affecting the clustering.
 For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
@@
 
-Another potentially useful manual control, is to allow the user to control the position of the center of a slice.
+Another potentially useful manual control, is to allow the user to \textcolor{blue}{control}\textcolor{red}{choose} the position of the center of a slice.
 The slice tour was introduced in @slicetour.
@@
 The purpose would be to explore how or if the shape of the data, in the space orthogonal to the projection, changes as one gets away from the center.
+\textcolor{red}{It would also allow the user to interactively decide on the thickness of the slice.}
 
@@
 The next section describes the new algorithm for manual control, for both projections and slices.
-The use of these methods are illustrated to compare and contrast boundaries constructed by different classifiers.
+The use of these methods \textcolor{blue}{are}\textcolor{red}{ is} illustrated to compare and contrast boundaries constructed by different classifiers.
 The software section describes a mathematica notebook that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
-The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two mathematica notebooks, one focused on the application, and the other that could be used generically for new data.
+The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two mathematica notebooks, one focused on the application, and another that could be used generically for new data.
 
@@
 
-The method for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
+The methods for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
 The first step was to construct a 3D manipulation space from a 2D projection.
@@
 
-The problem with new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
-There are numerous ways that this can be enforced, and details are in the Appendix.
+The problem with \textcolor{red}{the} new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
+\textcolor{red}{Even though these changes are for the most part imperceptible, one may wits to remove them and} there are numerous ways that this can be enforced, and details are given in the Appendix.
 \textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
@@
 However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different scale  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example).
-{\textcolor{red}Alex, is this still true with the controls you are using now?}
+\textcolor{red}{but this is no longer true if you replace it with animate??}
+\textcolor{red}{Alex, is this still true with the controls you are using now?}
 
 The examples presented in the penguins’ notebooks all utilise our primary new function, `SliceDynamic`.
-This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index.
-After specifying the initial slice height and the slice range, the user is then presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
-The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice height; and there is an input field, which changes the slice centre.
-The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
-The current projection can be displayed to contrast it with the slice by ticking a box.
-The explicit projection matrix can be displayed via another checkbox.
-The Slice Tour* notebooks include the additional functions: ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic.
-ProjectionPlot is very similar to SliceDynamic, except it only displays projections and VisualiseSliceDynamic allows you to discern which points exist inside the slice and outside it.
-However, ProjectedLocatorPlot displays the interactive slightly differently.
-Several locator panes are displayed above the plot and each one corresponds to a row in the projection matrix but behaves the same as ProjectionPlot.
-{\textcolor{blue}(this may be easier to implement in another language (?) )}
+This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index \textcolor{red}{(which can be one if there is only one group)}.
+After specifying the initial slice thickness and the slice range, the user is then presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
+The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.\textcolor{red}{this zoom doesn't work very well} The current projection can be displayed to contrast it with the slice by ticking a box.
+The explicit projection matrix can also be displayed via another checkbox in order to store it for further study.
 

### 2022-09-12 paper/paper.tex
went through sections 1 and 2 and moved a paragraph of mathematica to appendix as requested by Ursula

@@
 data using low-dimensional linear projections and slices. A user can
-interactively change the coefficients of a variable in a low-dimensional
+interactively change the
+\textcolor{blue} {coefficients of a variable in}
+\textcolor{red}{contribution of a given variable to} a low-dimensional
 projection, which is useful for exploring the sensitivity of structure
@@
 subspaces. These controls are implemented in Mathematica to take
-advantage of it's mathematics functions, and interactivity tools that
-allow construction of functions that interact with the position of the
-cursor. We demonstrate the usefulness of this new tool with an example
-exploring region boundaries in classification problems.
+advantage of \textcolor{blue}{it's} \textcolor{red}{its}
+\textcolor{blue}{remove this: mathematics functions, and interactivity tools}
+\textcolor{red}{location controls} that allow construction of functions
+that interact with the position of the cursor. We demonstrate the
+usefulness of this new tool with an example exploring region boundaries
+in classification problems.
 \end{abstract}
@@
 tours which allows one to quickly see many, many different projections.
-After Asimov there have been many, many tour developments, which are
-summarized in \citet{lee2021}.
+After Asimov there have been many \textcolor{blue}{remove: }, many\}
+tour developments, which are summarized in \citet{lee2021}.
 
@@
 places where it is useful. In exploratory data analysis, where one sees
-clusters in a projection, can some variables be removed from the
-projection without affecting the clustering. For interpreting models,
-one can reduce or increase a variable's contribution to examine the
-variable importance. Having the user interact with a projection is
-extremely valuable for understanding high-dimensional data. However,
-these algorithms have two problems: (1) the pre-processing of creating a
-manipulation space overly complicates the algorithm, (2) extending to
-higher dimensional control is difficult.
+clusters in a projection, \textcolor{red}{} one may ask whether some\}
+variables can be removed from the projection without affecting the
+clustering. For interpreting models, one can reduce or increase a
+variable's contribution to examine the variable importance. Having the
+user interact with a projection is extremely valuable for understanding
+high-dimensional data. However, these algorithms have two problems: (1)
+the pre-processing of creating a manipulation space overly complicates
+the algorithm, (2) extending to higher dimensional control is difficult.
 
 Another potentially useful manual control, is to allow the user to
-control the position of the center of a slice. The slice tour was
-introduced in \citet{slicetour}. It operates by converting the
-projection plane into a slice, by removing or de-emphasizing points that
-are further than a fixed orthogonal distance from the plane. The
-projection plane is usually thought of as passing through the center of
-the data. Manual control would allow the user to change the value of
-center point, by shifting it along a coordinate axis, while keeping the
-orientation of the projection plane fixed. The purpose would be to
-explore how or if the shape of the data, in the space orthogonal to the
-projection, changes as one gets away from the center.
+\textcolor{blue}{control}\textcolor{red}{choose} the position of the
+center of a slice. The slice tour was introduced in \citet{slicetour}.
+It operates by converting the projection plane into a slice, by removing
+or de-emphasizing points that are further than a fixed orthogonal
+distance from the plane. The projection plane is usually thought of as
+passing through the center of the data. Manual control would allow the
+user to change the value of center point, by shifting it along a
+coordinate axis, while keeping the orientation of the projection plane
+fixed. The purpose would be to explore how or if the shape of the data,
+in the space orthogonal to the projection, changes as one gets away from
+the center.
+\textcolor{red}{It would also allow the user to interactively decide on the thickness of the slice.}
 
@@
 tours. The next section describes the new algorithm for manual control,
-for both projections and slices. The use of these methods are
-illustrated to compare and contrast boundaries constructed by different
-classifiers. The software section describes a mathematica notebook that
-is used for the application, and describes the interactive environment
-that would be desirable within R as new technology becomes available.
-The paper is accompanied by an appendix with more details and
-adjustments to the manual controls, and two mathematica notebooks, one
-focused on the application, and the other that could be used generically
-for new data.
+for both projections and slices. The use of these methods
+\textcolor{blue}{are}\textcolor{red}{ is} illustrated to compare and
+contrast boundaries constructed by different classifiers. The software
+section describes a mathematica notebook that is used for the
+application, and describes the interactive environment that would be
+desirable within R as new technology becomes available. The paper is
+accompanied by an appendix with more details and adjustments to the
+manual controls, and two mathematica notebooks, one focused on the
+application, and another that could be used generically for new data.
 
@@
 
-The method for updating component values in \citet{cook_manual_1997}
+The methods for updating component values in \citet{cook_manual_1997}
 (and utilised in \citet{spyrison_spinifex_2020}) are prescribed
@@
 
-The problem with new simple method is that it is not faithful to the
-precise values for \(V_m\) because the orthonormalisation will change
-them. There are numerous ways that this can be enforced, and details are
+The problem with \textcolor{red}{the} new simple method is that it is
+not faithful to the precise values for \(V_m\) because the
+orthonormalisation will change them.
+\textcolor{red}{Even though these changes are for the most part imperceptible, one may wits to remove them and}
+there are numerous ways that this can be enforced, and details are given
 in the Appendix.

### 2022-09-12 paper/paper.Rmd
Some work on the software section

@@
 
-XXX copied from the slicing part to be weaved in below:
+## Structure
 
-The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.
+The implementation of the manual tour with slice control needs a visualization of the current projection in terms of an axis display.
+This display should be able to track the mouse position and adjust the projection based on the user selection.
+The user can select one axis (corresponding to one variable) via clicking on it, and then adjust the position of that axis by dragging it during the click event.
+In practice the closest axis will be selected, and the dragging results in a tracking of the mouse position that will update the projection in small steps, rendering interpolation unnecessary.
 
-The current implementation contains a switch to change between the slice view and the projected view of the data, while keeping all other settings fixed.
+The interaction in the axis display need to be mapped back onto the projection matrix, which in turn needs to be linked to both the orthonormalization procedure and to a second display that is showing the projected data.
 
-In practice we define the center point $\mathbf{c}$ explicitly in terms of all $p$ components and calculate the slice display following Eq. \ref{eq:slice}.
+In addition, we also want to be able to look at slices of the data and select slicing parameters interactively.
+Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice height $h$ and a numeric input to specify the center point $c$ explicitly.
+The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).
 
-This was implemented as a slider, and to allow the user to efficiently explore the interesting range we provide the slider range as an input.
+One might implement such an interactive interface via R Shiny, but in particular the tracking of the mouse position in small increments might pose a challenge.
+Here we found the dynamic graphics interface from Mathematica to be especially useful.
+In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approach might be implemented in the `tourr` package.
 
@@
 
-- **Why is this a good sandbox**
-- **Explain the functionality available in the notebooks**
+Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
+Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
 
-Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
-Most of this inbuilt functionality is quite user-friendly and can be found in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
-Also, it is relatively simple  to create dynamic objects via the inbuilt commands `Manipulate` or `DynamicModule`.
-Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that dynamic variable will be updated.
+Importantly for our work, it is relatively simple  to create dynamic objects via the inbuilt commands `Manipulate` or `DynamicModule`.
+Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.
 This is the essential ingredient for this implementation of the manual tour.
+
 However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different scale  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example).
@@
 
-The examples presented in the penguins’ notebooks all utilise our primary new function, `SliceDynamic`.
+\textcolor{red}{Ursula: this comment about slider scale seems to specific, can we have a more general comment about difficulties, e.g.
+I remember some issues with the color mapping?
+Also there seem to be a lot of loops needed?
+Can you explain what the locator pane is?}
+
+The examples presented in the penguins’ notebooks all use our primary new function, `SliceDynamic`.
 This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index \textcolor{red}{(which can be one if there is only one group)}.
-After specifying the initial slice thickness and the slice range, the user is then presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
+After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
 The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre.
-The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.\textcolor{red}{this zoom doesn't work very well} The current projection can be displayed to contrast it with the slice by ticking a box.
-The explicit projection matrix can also be displayed via another checkbox in order to store it for further study.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.\textcolor{red}{this zoom doesn't work very well} The projected data can be displayed to contrast it with the slice by ticking a box.
+The explicit projection matrix can also be displayed via another checkbox.
+This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration later, as shown in the applications below.
+
+\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}
 

### 2022-09-12 paper/paper.tex
Some work on the software section

@@
 
-XXX copied from the slicing part to be weaved in below:
+\hypertarget{structure}{%
+\subsection{Structure}\label{structure}}
 
-The current implementation allows the user to individually enter the
-components of the vector defining the center point, and updates the
-display by jumping to the new slice.
+The implementation of the manual tour with slice control needs a
+visualization of the current projection in terms of an axis display.
+This display should be able to track the mouse position and adjust the
+projection based on the user selection. The user can select one axis
+(corresponding to one variable) via clicking on it, and then adjust the
+position of that axis by dragging it during the click event. In practice
+the closest axis will be selected, and the dragging results in a
+tracking of the mouse position that will update the projection in small
+steps, rendering interpolation unnecessary.
 
-The current implementation contains a switch to change between the slice
-view and the projected view of the data, while keeping all other
-settings fixed.
+The interaction in the axis display need to be mapped back onto the
+projection matrix, which in turn needs to be linked to both the
+orthonormalization procedure and to a second display that is showing the
+projected data.
 
-In practice we define the center point \(\mathbf{c}\) explicitly in
-terms of all \(p\) components and calculate the slice display following
-Eq. \ref{eq:slice}.
+In addition, we also want to be able to look at slices of the data and
+select slicing parameters interactively. Here we have implemented a
+switch to change between projection and slice view, a slider that
+adjusts the slice height \(h\) and a numeric input to specify the center
+point \(c\) explicitly. The display will update based on these inputs,
+in particular the view will jump to a new slice when \(c\) is changed
+(though an interpolation might also be useful).
 
-This was implemented as a slider, and to allow the user to efficiently
-explore the interesting range we provide the slider range as an input.
+One might implement such an interactive interface via R Shiny, but in
+particular the tracking of the mouse position in small increments might
+pose a challenge. Here we found the dynamic graphics interface from
+Mathematica to be especially useful. In the following we first describe
+the relevant Mathematica functionalities and how they were used in the
+implementation, followed by a brief sketch of how the new approach might
+be implemented in the \texttt{tourr} package.
 
@@
 \subsection{Mathematica notebook}\label{mathematica-notebook}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Why is this a good sandbox}
-\item
-  \textbf{Explain the functionality available in the notebooks}
-\end{itemize}
 
@@
 increments add 10 points into the slice rather than increasing the slice
-height by 0.1 for example). \{\textcolor{red}Alex, is this still true
-with the controls you are using now?\}
+height by 0.1 for example).
+\textcolor{red}{but this is no longer true if you replace it with animate??}
+\textcolor{red}{Alex, is this still true with the controls you are using now?}
 
@@
 column details the name of the group and the last column details the
-group index. After specifying the initial slice height and the slice
-range, the user is then presented with an interactive display in which
-the control objects appear on the left and the slice visulisation
-appears on the right. The user can change the orientation of the slice
-via the locator pane, which changes the projection matrix; a slider
-controls the slice height; and there is an input field, which changes
-the slice centre. The user can also change the appearance of the plot by
-zooming into the center or changing the point size with sliders
-provided. The current projection can be displayed to contrast it with
-the slice by ticking a box. The explicit projection matrix can be
-displayed via another checkbox. The Slice Tour* notebooks include the
-additional functions: ProjectionPlot, Projected2DSliderPlot, and
-VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic,
-except it only displays projections and VisualiseSliceDynamic allows you
-to discern which points exist inside the slice and outside it. However,
-ProjectedLocatorPlot displays the interactive slightly differently.
-Several locator panes are displayed above the plot and each one
-corresponds to a row in the projection matrix but behaves the same as
-ProjectionPlot. \{\textcolor{blue}(this may be easier to implement in
-another language (?) )\}
+group index
+\textcolor{red}{(which can be one if there is only one group)}. After
+specifying the initial slice thickness and the slice range, the user is
+then presented with an interactive display in which the control objects
+appear on the left and the slice visulisation appears on the right. The
+user can change the orientation of the slice via the locator pane, which
+changes the projection matrix; a slider controls the slice thickness;
+and there is an input field, which changes the slice centre. The user
+can also change the appearance of the plot by zooming into the center or
+changing the point size with sliders
+provided.\textcolor{red}{this zoom doesn't work very well} The current
+projection can be displayed to contrast it with the slice by ticking a
+box. The explicit projection matrix can also be displayed via another
+checkbox in order to store it for further study.
 

### 2022-09-15 paper/paper.Rmd
typos

@@
 The problem with \textcolor{red}{the} new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
-\textcolor{red}{Even though these changes are for the most part imperceptible, one may wits to remove them and} there are numerous ways that this can be enforced, and details are given in the Appendix.
+\textcolor{red}{Even though these changes are for the most part imperceptible, one may wish to avoid them and} there are numerous ways that this can be enforced, a few are detailed in the Appendix.
 \textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}

### 2022-09-15 paper/paper.tex
typos

@@
 and symbolic and numeric computation. Most of this inbuilt functionality
-is quite user-friendly and can be found in the Mathematica
-documentation; typically, numerous examples are provided within the
-documentation, and some possible issues are outlined there. Also, it is
-relatively simple to create dynamic objects via the inbuilt commands
-\texttt{Manipulate} or \texttt{DynamicModule}. Control objects, such as
-sliders, locator panes, and input fields, can then be used on dynamic
-variables, and when there is a change in the dynamic variable the
-dynamic objects which contain that dynamic variable will be updated.
-This is the essential ingredient for this implementation of the manual
-tour. However, it can be hard to alter the behaviour of the control
-objects such as setting the 1D slider on a different scale (Slider
-increments add 10 points into the slice rather than increasing the slice
-height by 0.1 for example).
+is user-friendly and described in the Mathematica documentation;
+typically, numerous examples are provided within the documentation, and
+some possible issues are outlined there.
+
+Importantly for our work, it is relatively simple to create dynamic
+objects via the inbuilt commands \texttt{Manipulate} or
+\texttt{DynamicModule}. Control objects, such as sliders, locator panes,
+and input fields, can then be used on dynamic variables, and when there
+is a change in the dynamic variable the dynamic objects which contain
+that variable will be updated. This is the essential ingredient for this
+implementation of the manual tour.
+
+However, it can be hard to alter the behaviour of the control objects
+such as setting the 1D slider on a different scale (Slider increments
+add 10 points into the slice rather than increasing the slice height by
+0.1 for example).
 \textcolor{red}{but this is no longer true if you replace it with animate??}
@@
 
-The examples presented in the penguins' notebooks all utilise our
-primary new function, \texttt{SliceDynamic}. This function typically
-accepts grouped data in the form of a matrix where the second last
-column details the name of the group and the last column details the
-group index
-\textcolor{red}{(which can be one if there is only one group)}. After
-specifying the initial slice thickness and the slice range, the user is
-then presented with an interactive display in which the control objects
-appear on the left and the slice visulisation appears on the right. The
-user can change the orientation of the slice via the locator pane, which
-changes the projection matrix; a slider controls the slice thickness;
-and there is an input field, which changes the slice centre. The user
-can also change the appearance of the plot by zooming into the center or
-changing the point size with sliders
-provided.\textcolor{red}{this zoom doesn't work very well} The current
-projection can be displayed to contrast it with the slice by ticking a
-box. The explicit projection matrix can also be displayed via another
-checkbox in order to store it for further study.
+\textcolor{red}{Ursula: this comment about slider scale seems to specific, can we have a more general comment about difficulties, e.g.
+I remember some issues with the color mapping?
+Also there seem to be a lot of loops needed?
+Can you explain what the locator pane is?}
+
+The examples presented in the penguins' notebooks all use our primary
+new function, \texttt{SliceDynamic}. This function typically accepts
+grouped data in the form of a matrix where the second last column
+details the name of the group and the last column details the group
+index \textcolor{red}{(which can be one if there is only one group)}.
+After specifying the initial slice thickness and the slice range, the
+user is presented with an interactive display in which the control
+objects appear on the left and the slice visulisation appears on the
+right. The user can change the orientation of the slice via the locator
+pane, which changes the projection matrix; a slider controls the slice
+thickness; and there is an input field, which changes the slice centre.
+The user can also change the appearance of the plot by zooming into the
+center or changing the point size with sliders
+provided.\textcolor{red}{this zoom doesn't work very well} The projected
+data can be displayed to contrast it with the slice by ticking a box.
+The explicit projection matrix can also be displayed via another
+checkbox. This is especially useful when we wish to import a projection
+that was identified as interesting in the manual exploration later, as
+shown in the applications below.
+
+\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}
 

### 2022-09-21 paper/paper.Rmd
fixed few sentences

@@
 
-However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different scale  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example).
-\textcolor{red}{but this is no longer true if you replace it with animate??}
-\textcolor{red}{Alex, is this still true with the controls you are using now?}
+The most important inbuilt Mathematica function for our purpose is the locator pane.
+This creates a region on the screen where the position of the mouse is converted to input for the graphics functions, enabling the manual navigation of the tour.
 
-\textcolor{red}{Ursula: this comment about slider scale seems to specific, can we have a more general comment about difficulties, e.g.
-I remember some issues with the color mapping?
-Also there seem to be a lot of loops needed?
-Can you explain what the locator pane is?}
+\textcolor{red}{To Ursula:I removed the comment on the slider.
+There are no problems with color mapping.
+What do you mean by:" Also there seem to be a lot of loops needed?"?
+as far as I can tell the code does not use any loop.
+Is my explanantion of the locator pane enough?
+Please modify as you see fit and finish up this section.}
 
@@
 The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre.
-The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.\textcolor{red}{this zoom doesn't work very well} The projected data can be displayed to contrast it with the slice by ticking a box.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
+\textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.} The projected data can be displayed to contrast it with the slice by ticking a box.
 The explicit projection matrix can also be displayed via another checkbox.

### 2022-09-21 paper/paper.tex
fixed few sentences

@@
 orthonormalisation will change them.
-\textcolor{red}{Even though these changes are for the most part imperceptible, one may wits to remove them and}
-there are numerous ways that this can be enforced, and details are given
-in the Appendix.
+\textcolor{red}{Even though these changes are for the most part imperceptible, one may wish to avoid them and}
+there are numerous ways that this can be enforced, a few are detailed in
+the Appendix.
 \textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
@@
 
-However, it can be hard to alter the behaviour of the control objects
-such as setting the 1D slider on a different scale (Slider increments
-add 10 points into the slice rather than increasing the slice height by
-0.1 for example).
-\textcolor{red}{but this is no longer true if you replace it with animate??}
-\textcolor{red}{Alex, is this still true with the controls you are using now?}
+The most important inbuilt Mathematica function for our purpose is the
+locator pane. This creates a region on the screen where the position of
+the mouse is converted to input for the graphics functions, enabling the
+manual navigation of the tour.
 
-\textcolor{red}{Ursula: this comment about slider scale seems to specific, can we have a more general comment about difficulties, e.g.
-I remember some issues with the color mapping?
-Also there seem to be a lot of loops needed?
-Can you explain what the locator pane is?}
+\textcolor{red}{To Ursula:I removed the comment on the slider.
+There are no problems with color mapping.
+What do you mean by:" Also there seem to be a lot of loops needed?"?
+as far as I can tell the code does not use any loop.
+Is my explanantion of the locator pane enough?
+Please modify as you see fit and finish up this section.}
 
@@
 The user can also change the appearance of the plot by zooming into the
-center or changing the point size with sliders
-provided.\textcolor{red}{this zoom doesn't work very well} The projected
-data can be displayed to contrast it with the slice by ticking a box.
-The explicit projection matrix can also be displayed via another
-checkbox. This is especially useful when we wish to import a projection
-that was identified as interesting in the manual exploration later, as
-shown in the applications below.
+center or changing the point size with sliders provided.
+\textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.}
+The projected data can be displayed to contrast it with the slice by
+ticking a box. The explicit projection matrix can also be displayed via
+another checkbox. This is especially useful when we wish to import a
+projection that was identified as interesting in the manual exploration
+later, as shown in the applications below.
 

### 2022-09-22 paper/paper.Rmd
Iteration on the software section

@@
 
-The implementation of the manual tour with slice control needs a visualization of the current projection in terms of an axis display.
+The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example).
 This display should be able to track the mouse position and adjust the projection based on the user selection.
-The user can select one axis (corresponding to one variable) via clicking on it, and then adjust the position of that axis by dragging it during the click event.
-In practice the closest axis will be selected, and the dragging results in a tracking of the mouse position that will update the projection in small steps, rendering interpolation unnecessary.
+The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it during the click event.
+In practice the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.
 
-The interaction in the axis display need to be mapped back onto the projection matrix, which in turn needs to be linked to both the orthonormalization procedure and to a second display that is showing the projected data.
+The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized before feeding back into the axis display.
+A second, linked display shows the projected data in sync with the updates from the axis display.
 
@@
 
-The most important inbuilt Mathematica function for our purpose is the locator pane.
-This creates a region on the screen where the position of the mouse is converted to input for the graphics functions, enabling the manual navigation of the tour.
+The most relevant inbuilt Mathematica function for our purpose is the locator pane.
+This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
 
-\textcolor{red}{To Ursula:I removed the comment on the slider.
-There are no problems with color mapping.
-What do you mean by:" Also there seem to be a lot of loops needed?"?
-as far as I can tell the code does not use any loop.
-Is my explanantion of the locator pane enough?
-Please modify as you see fit and finish up this section.}
-
-The examples presented in the penguins’ notebooks all use our primary new function, `SliceDynamic`.
-This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index \textcolor{red}{(which can be one if there is only one group)}.
+The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`.
+This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group).
 After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
@@
 The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
-\textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.} The projected data can be displayed to contrast it with the slice by ticking a box.
+It is worth noting that this zoom works best with data that has been centred and scaled.
+The projected data can be displayed to contrast it with the slice by ticking a box.
 The explicit projection matrix can also be displayed via another checkbox.
@@
 
-\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}
+Details about the implementation and usage instructions are given in the Appendix.
 
@@
 Changing the center of a slice manually enables exploring the space orthogonal to a projection, specifically in the direction of a single variable.
-This is a new tool, and as shown in the example, when used with the slice tour can very useful for understanding boundaries of classifiers.
+This is a new tool, and as shown in the example, when used with the slice tour can be very useful for understanding boundaries of classifiers.
 This is also likely a useful method for examining low-dimensional non-linear manifolds, and also functions of multiple parameters.

### 2022-09-22 paper/paper.tex
Iteration on the software section

@@
 
-The implementation of the manual tour with slice control needs a
-visualization of the current projection in terms of an axis display.
-This display should be able to track the mouse position and adjust the
-projection based on the user selection. The user can select one axis
-(corresponding to one variable) via clicking on it, and then adjust the
-position of that axis by dragging it during the click event. In practice
-the closest axis will be selected, and the dragging results in a
-tracking of the mouse position that will update the projection in small
-steps, rendering interpolation unnecessary.
+The implementation of the manual tour as suggested here requires the
+visualization of the current projection in terms of an axis display (see
+Figure \ref{fig:manualsequence} for an example). This display should be
+able to track the mouse position and adjust the projection based on the
+user selection. The user can select one axis (corresponding to one
+variable) by clicking on it, and then adjust the position of that axis
+by dragging it during the click event. In practice the closest axis will
+be selected, and the dragging results in tracking of the mouse position
+which will update the projection in small steps, such that no further
+interpolation is required.
 
-The interaction in the axis display need to be mapped back onto the
-projection matrix, which in turn needs to be linked to both the
-orthonormalization procedure and to a second display that is showing the
-projected data.
+The interactions in the axis display need to be mapped back onto the
+current projection matrix, which will then be orthonormalized before
+feeding back into the axis display. A second, linked display shows the
+projected data in sync with the updates from the axis display.
 
@@
 
-The most important inbuilt Mathematica function for our purpose is the
+The most relevant inbuilt Mathematica function for our purpose is the
 locator pane. This creates a region on the screen where the position of
-the mouse is converted to input for the graphics functions, enabling the
-manual navigation of the tour.
+the mouse is captured and then converted to input that updates the
+graphics functions, enabling the manual navigation of the tour.
 
-\textcolor{red}{To Ursula:I removed the comment on the slider.
-There are no problems with color mapping.
-What do you mean by:" Also there seem to be a lot of loops needed?"?
-as far as I can tell the code does not use any loop.
-Is my explanantion of the locator pane enough?
-Please modify as you see fit and finish up this section.}
+The examples presented in the application section below (and attached in
+notebook format in the supplementary material) all use our primary new
+function, \texttt{SliceDynamic}. This function typically accepts grouped
+data in the form of a matrix where the second last column details the
+name of the group and the last column details the group index (which can
+be one if there is only one group). After specifying the initial slice
+thickness and the slice range, the user is presented with an interactive
+display in which the control objects appear on the left and the slice
+visulisation appears on the right. The user can change the orientation
+of the slice via the locator pane, which changes the projection matrix;
+a slider controls the slice thickness; and there is an input field,
+which changes the slice centre. The user can also change the appearance
+of the plot by zooming into the center or changing the point size with
+sliders provided. It is worth noting that this zoom works best with data
+that has been centred and scaled. The projected data can be displayed to
+contrast it with the slice by ticking a box. The explicit projection
+matrix can also be displayed via another checkbox. This is especially
+useful when we wish to import a projection that was identified as
+interesting in the manual exploration later, as shown in the
+applications below.
 
-The examples presented in the penguins' notebooks all use our primary
-new function, \texttt{SliceDynamic}. This function typically accepts
-grouped data in the form of a matrix where the second last column
-details the name of the group and the last column details the group
-index \textcolor{red}{(which can be one if there is only one group)}.
-After specifying the initial slice thickness and the slice range, the
-user is presented with an interactive display in which the control
-objects appear on the left and the slice visulisation appears on the
-right. The user can change the orientation of the slice via the locator
-pane, which changes the projection matrix; a slider controls the slice
-thickness; and there is an input field, which changes the slice centre.
-The user can also change the appearance of the plot by zooming into the
-center or changing the point size with sliders provided.
-\textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.}
-The projected data can be displayed to contrast it with the slice by
-ticking a box. The explicit projection matrix can also be displayed via
-another checkbox. This is especially useful when we wish to import a
-projection that was identified as interesting in the manual exploration
-later, as shown in the applications below.
-
-\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}
+Details about the implementation and usage instructions are given in the
+Appendix.
 
@@
 variable. This is a new tool, and as shown in the example, when used
-with the slice tour can very useful for understanding boundaries of
+with the slice tour can be very useful for understanding boundaries of
 classifiers. This is also likely a useful method for examining

### 2022-09-22 paper/appendix.Rmd
Some work on the appendix

--- 2022-08-19 paper/appendix.Rmd
@@
 
-A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:othermethod}. It is as follows:
+A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:othermethod}). It is as follows:
 
@@
 
-$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m$$.
+$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m.$$
 
@@
 
-$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0$$.
+$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0.$$
 
@@
 
+\textcolor{red}{Notation does not match, everywhere else $A$ is the projection matrix onto 2d, so just the two columns, this is called $P$ here...maybe switch?
+$A$ can be considered as first two directions of the complete rotation matrix $P$?}
+
 For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
-Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
 Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
@@
 
-A very simple adjustment method works purely within the (hyper)plane, and is using the fact that the first direction will not be rotated when applying Gram-Schmidt. The following steps are needed:
+Finally we may also consider an adjustment that works within the projection plane.
+The idea is that when applying Gram-Schmidt, the direction of the first basis vector is not changed.
+We thus rotate the projection matrix within the plane such that the first basis is aligned with the direction of the selected variable $m$ before applying orthonormalization.
+In addition we need to keep the contribution of variable $m$ along that direction fixed when normalizing the vector.
+The following steps are needed:
 
@@
 3.
-Rotate the basis within the projection (hyper-)plane, such that the first direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$.
+Rotate the basis within the projection plane, such that the first basis direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$.
 In a 2D projection this means finding the angle $\theta$ by which the basis should be rotated.
-4. Normalise the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
-5. Orthonormalise $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
+4. Normalize the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
+5. Orthonormalize $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
 6. Rotate $\hat A^{**}$ back into the original basis to obtain $A^{**}$.
@@
 
-- **Why is this a good sandbox**
-- **Explain the functionality available in the notebooks**
+Here we describe the functions implemented in the Mathematica package `mmtour.wl`. The main function and its arguments are given below:
 
-Here we describe the functions implemented in the Mathematica package `mmtour.wl`. The main function is
+`SliceDynamic[data, projmat, height, heightRange, legendQ(=1), flagQ(=1), colorFunc(=ColorData[97])]`
 
-- `SliceDynamic[data, projmat, height, heightRange, legendQ(=1), flagQ(=1), colorFunc(=ColorData[97])]` with arguments
-
-- `data`: A data matrix with an extra grouping that has a label (string) and a flag  (numerical) 
+- `data`: A data matrix, potentially with grouping that has a label (string) and a flag  (numerical) in the last two columns.
 - `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
@@
 
+\textcolor{red}{The first four arguments are required.
+As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.}
+
 The function renders a manual tour with controls that allow the user to navigate through different projections.
@@
 
-This function can be used for ungrouped data setting  legendQ and flagQ to 0.
-
+This function can be used for ungrouped data by setting legendQ and flagQ to 0.
 To remove a group from the visual display, the user can choose specific colours for the entry colorFunc.
@@
 
-Other functions are:
+The Slice Tour* notebooks include the additional functions: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
 
-- `ProjectionPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments
+`ProjectionPlot` is very similar to `SliceDynamic`, except it only displays projections.
+Using it instead of `SliceDynamic` simplifies the display and is more efficient.
+The interface is slightly different, the data file should not contain labels for groups, if these are desired they are passed through the optional argument legendNames as {"group 1", "group 2",...}, for example.
+The function and its arguments are given below:
 
-- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
+`ProjectionPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]`
+
+- `data`: A data matrix with one or more groups labeled by a flag (numerical, last column).
 - `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
@@
 
-This function is similar to SliceDynamic but can only display projections.
-Using it instead of SliceDynamic simplifies the display and is more efficient.
-The data file should not contain labels for groups, if these are desired they are passed through the optional argument legendNames as {"group 1", "group 2",...}, for example.
+`ProjectedLocatorPlot` displays the interactive controls slightly differently.
+Several locator panes are displayed above the display of the projected data and each one corresponds to a row in the projection matrix.
+The updating behaviour of the projection is the same as in `ProjectionPlot`.
+The function and its arguments are given below:
 
-The display allows the user to navigate through different projections, choose the point size and zoom the factor, and explicitly display the coordinates of the projection.
+`ProjectedLocatorPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]`
 
-- `ProjectedLocatorPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments
-
-- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
+- `data`: A data matrix with one or more groups labeled by a flag  (numerical, last column). 
 - `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
@@
 
-This function is very similar to ProjectionPlot but uses a different display for navigation of  projections in which each coordinate is represented on a separate dial.
+Finally `VisualiseSliceDynamic` allows to discern which points exist inside the slice and outside it.
+This function differs from `SliceDynamic` in that it looks at only one data set but displays both the points inside and outside the slice.
+The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, centre point and zoom level.
+The function and its arguments are given below:
 
-- `VisualiseSliceDynamic[data, projmat, height, heightRange]` with arguments
+`VisualiseSliceDynamic[data, projmat, height, heightRange]`
 
-- `data`: A data matrix containg only one set.In this case the data matrix should not contain column labels.
+- `data`: A data matrix containing only one group. In this case the data matrix should not contain column labels.
 - `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
@@
 
-This function differs from SliceDynamic in that it looks at only one data set but displays both the points inside and outside the slice.
-The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, centre point and zoom level.
+\textcolor{red}{How are they shown, different colors? Fading points that are outside?}

### 2022-09-22 paper/appendix.tex
Some work on the appendix

--- 2022-08-19 paper/appendix.tex
@@
 A small modification to algorithm 1 will maintain the components of
-\(V_m\) precisely (Figure \ref{fig:othermethod}. It is as follows:
+\(V_m\) precisely (Figure \ref{fig:othermethod}). It is as follows:
 
@@
 
-\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m\].
+\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m.\]
 
@@
 
-\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0\].
+\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0.\]
 
@@
 
+\textcolor{red}{Notation does not match, everywhere else $A$ is the projection matrix onto 2d, so just the two columns, this is called $P$ here...maybe switch?
+$A$ can be considered as first two directions of the complete rotation matrix $P$?}
+
 For \(d=2\) projections, the projection matrix is the sub-matrix of
@@
 basis for the p-dimensional space is given by
-\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}\), orthonormality of the
+\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
 projection matrix is expressed as
@@
 
-A very simple adjustment method works purely within the (hyper)plane,
-and is using the fact that the first direction will not be rotated when
-applying Gram-Schmidt. The following steps are needed:
+Finally we may also consider an adjustment that works within the
+projection plane. The idea is that when applying Gram-Schmidt, the
+direction of the first basis vector is not changed. We thus rotate the
+projection matrix within the plane such that the first basis is aligned
+with the direction of the selected variable \(m\) before applying
+orthonormalization. In addition we need to keep the contribution of
+variable \(m\) along that direction fixed when normalizing the vector.
+The following steps are needed:
 
@@
 \item
-  Rotate the basis within the projection (hyper-)plane, such that the
-  first direction is aligned with the direction of \(m\) in \(A^*\),
+  Rotate the basis within the projection plane, such that the first
+  basis direction is aligned with the direction of \(m\) in \(A^*\),
   giving \(\hat A^*\). In a 2D projection this means finding the angle
@@
 \item
-  Normalise the first direction in \(\hat A^*\) while keeping the
+  Normalize the first direction in \(\hat A^*\) while keeping the
   contribution in row \(m\) fixed, giving \(\hat A^{*0}\).
 \item
-  Orthonormalise \(\hat A^{*0}\) using Gram-Schmidt, giving
+  Orthonormalize \(\hat A^{*0}\) using Gram-Schmidt, giving
   \(\hat A^{**}\). By construction the contribution of \(m\) will remain
@@
 
+Here we describe the functions implemented in the Mathematica package
+\texttt{mmtour.wl}. The main function and its arguments are given below:
+
+\texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
+
 \begin{itemize}
@@
 \item
-  \textbf{Why is this a good sandbox}
-\item
-  \textbf{Explain the functionality available in the notebooks}
-\end{itemize}
-
-Here we describe the functions implemented in the Mathematica package
-\texttt{mmtour.wl}. The main function is
-
-\begin{itemize}
-\item
-  \texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
-  with arguments
-\item
-  \texttt{data}: A data matrix with an extra grouping that has a label
-  (string) and a flag (numerical)
+  \texttt{data}: A data matrix, potentially with grouping that has a
+  label (string) and a flag (numerical) in the last two columns.
 \item
@@
 
+\textcolor{red}{The first four arguments are required.
+As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.}
+
 The function renders a manual tour with controls that allow the user to
@@
 
-This function can be used for ungrouped data setting legendQ and flagQ
-to 0.
+This function can be used for ungrouped data by setting legendQ and
+flagQ to 0. To remove a group from the visual display, the user can
+choose specific colours for the entry colorFunc. For example,
+\{Red,White,Green\} would display groups 1 (red) and 3 (green) while
+making 2 invisible.
 
-To remove a group from the visual display, the user can choose specific
-colours for the entry colorFunc. For example, \{Red,White,Green\} would
-display groups 1 (red) and 3 (green) while making 2 invisible.
+The Slice Tour* notebooks include the additional functions:
+\texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot}, and
+\texttt{VisualiseSliceDyanmic}.
 
-Other functions are:
+\texttt{ProjectionPlot} is very similar to \texttt{SliceDynamic}, except
+it only displays projections. Using it instead of \texttt{SliceDynamic}
+simplifies the display and is more efficient. The interface is slightly
+different, the data file should not contain labels for groups, if these
+are desired they are passed through the optional argument legendNames as
+\{``group 1'', ``group 2'',\ldots\}, for example. The function and its
+arguments are given below:
+
+\texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
 
 \begin{itemize}
-\item
-  \texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
-  with arguments
+\tightlist
 \item
   \texttt{data}: A data matrix with one or more groups labeled by a flag
-  (numerical)
+  (numerical, last column).
 \item
@@
 
-This function is similar to SliceDynamic but can only display
-projections. Using it instead of SliceDynamic simplifies the display and
-is more efficient. The data file should not contain labels for groups,
-if these are desired they are passed through the optional argument
-legendNames as \{``group 1'', ``group 2'',\ldots\}, for example.
+\texttt{ProjectedLocatorPlot} displays the interactive controls slightly
+differently. Several locator panes are displayed above the display of
+the projected data and each one corresponds to a row in the projection
+matrix. The updating behaviour of the projection is the same as in
+\texttt{ProjectionPlot}. The function and its arguments are given below:
 
-The display allows the user to navigate through different projections,
-choose the point size and zoom the factor, and explicitly display the
-coordinates of the projection.
+\texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
 
 \begin{itemize}
-\item
-  \texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
-  with arguments
+\tightlist
 \item
   \texttt{data}: A data matrix with one or more groups labeled by a flag
-  (numerical)
+  (numerical, last column).
 \item
@@
 
-This function is very similar to ProjectionPlot but uses a different
-display for navigation of projections in which each coordinate is
-represented on a separate dial.
+Finally \texttt{VisualiseSliceDynamic} allows to discern which points
+exist inside the slice and outside it. This function differs from
+\texttt{SliceDynamic} in that it looks at only one data set but displays
+both the points inside and outside the slice. The display panel allows
+the user to specify the size of both sets of points (inside and outside)
+along with the slice height, centre point and zoom level. The function
+and its arguments are given below:
+
+\texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
 
 \begin{itemize}
+\tightlist
 \item
-  \texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
-  with arguments
-\item
-  \texttt{data}: A data matrix containg only one set.In this case the
-  data matrix should not contain column labels.
+  \texttt{data}: A data matrix containing only one group. In this case
+  the data matrix should not contain column labels.
 \item
@@
 
-This function differs from SliceDynamic in that it looks at only one
-data set but displays both the points inside and outside the slice. The
-display panel allows the user to specify the size of both sets of points
-(inside and outside) along with the slice height, centre point and zoom
-level.
+\textcolor{red}{How are they shown, different colors? Fading points that are outside?}
 

### 2022-09-22 paper/appendix.Rmd
fixed details appendix

@@
 
-\textcolor{red}{Notation does not match, everywhere else $A$ is the projection matrix onto 2d, so just the two columns, this is called $P$ here...maybe switch?
-$A$ can be considered as first two directions of the complete rotation matrix $P$?}
-
-For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns.
-Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$.
-Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$.
-Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$, formed by its first two columns.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$.
+Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$.
+Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
 One possible algorithm to achieve this is
 \begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a^*_{m1},a^*_{m2}}$
+\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
 \begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
+\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
 \item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
 \end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
 \end{enumerate}
-The random completion of $A$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane.
+The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane.
 However, it renders projections in a way that is not continuouns and may be distracting.
-This can be alleviated by replacing the random completion with a rule restricting the size of the jumps.
+This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.
 
 \begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
 \CodeBefore
@@
      \Body
-       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
-e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
-e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
+       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
+e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
+e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
 \vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
- e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
+ e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
 \vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
-e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
-\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
+e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
+\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
 \end{align}
@@
 
-\textcolor{red}{How are they shown, different colors? Fading points that are outside?}
+The points inside the slice are shown in black whereas those outside the slice are shown in light blue, with labels indicating this.

### 2022-09-22 paper/appendix.tex
fixed details appendix

@@
 
-\textcolor{red}{Notation does not match, everywhere else $A$ is the projection matrix onto 2d, so just the two columns, this is called $P$ here...maybe switch?
-$A$ can be considered as first two directions of the complete rotation matrix $P$?}
-
-For \(d=2\) projections, the projection matrix is the sub-matrix of
-\(A\) formed by its first two columns. Whereas orthonormality of the
+For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
+of \(O\), formed by its first two columns. Whereas orthonormality of the
 basis for the p-dimensional space is given by
@@
 projection matrix is expressed as
-\(P_i\cdot P_j=\delta_{ij}, {i,j=1,2}\). Movement of the cursor takes
-the two components \({x_{m1},x_{m2}}\) into a selected new value
-\({a,b}\). Although the motion is constrained by \(a^2+b^2\leq 1\), this
-is not sufficient to guarantee orthonormality of the new projection
-matrix. One possible algorithm to achieve this is
+\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
+Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
+a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
+constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
+to guarantee orthonormality of the new projection matrix. One possible
+algorithm to achieve this is
 
 \begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
-\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
+\item Cursor movement takes ${x_{m1},x_{m2}}\to {a^*_{m1},a^*_{m2}}$
+\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
 \begin{enumerate}
-\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
-\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
+\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
 \item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
 \end{enumerate}
-\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
+\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
 \end{enumerate}
 
-The random completion of \(A\) outside the projection matrix provides an
+The random completion of \(O\) outside the projection matrix provides an
 exploration of dimensions orthogonal to the projection plane. However,
@@
 distracting. This can be alleviated by replacing the random completion
-with a rule restricting the size of the jumps.
+with a rule restricting the size of the jumps, for example in step (b)
+the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
+orthogonalisation.
 
 \begin{align}
-A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
 \CodeBefore
@@
 
-\textcolor{red}{How are they shown, different colors? Fading points that are outside?}
+The points inside the slice are shown in black whereas those outside the
+slice are shown in light blue, with labels indicating this.
 

### 2022-09-23 paper/paper.Rmd
handled all the red/blue text

@@
   This paper describes new user controls for examining high-dimensional data using low-dimensional linear projections and slices.
-A user can interactively change the \textcolor{blue} {coefficients of a variable in} \textcolor{red}{contribution of a given variable to} a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables.
-The user can also interactively shift the slice center, which is useful for exploring how structure changes in local subspaces.
-These controls are implemented in Mathematica to take advantage of \textcolor{blue}{it's} \textcolor{red}{its} \textcolor{blue}{remove this: mathematics functions, and interactivity tools} \textcolor{red}{location controls} that allow construction of functions that interact with the position of the cursor.
-We demonstrate the usefulness of this new tool with an example exploring region boundaries in classification problems.
+A user can interactively change the contribution of a given variable to a low-dimensional projection, which is useful for exploring the sensitivity of structure to particular variables.
+The user can also interactively shift the center of a slice, for example, to explore how structure changes in local subspaces.
+Two Mathematica notebooks are provided, which contains functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
+The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls.
+Some limited implementation has also been made available in the R package \tt{tourr}.
 keywords: |
@@
 (A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
-After Asimov there have been many \textcolor{blue}{remove: }, many} tour developments, which are summarized in @lee2021.
+After Asimov there have been many tour developments, which are summarized in @lee2021.
 
@@
 There are many places where it is useful.
-In exploratory data analysis, where one sees clusters in a projection, \textcolor{red}{} one may ask whether some} variables can be removed from the projection without affecting the clustering.
+In exploratory data analysis, where one sees clusters in a projection, one may ask whether some variables can be removed from the projection without affecting the clustering.
 For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
@@
 
-Another potentially useful manual control, is to allow the user to \textcolor{blue}{control}\textcolor{red}{choose} the position of the center of a slice.
+Another potentially useful manual control, is to allow the user to choose the position of the center of a slice.
 The slice tour was introduced in @slicetour.
@@
 The purpose would be to explore how or if the shape of the data, in the space orthogonal to the projection, changes as one gets away from the center.
-\textcolor{red}{It would also allow the user to interactively decide on the thickness of the slice.}
+It would also allow the user to interactively decide on the thickness of the slice.
 
@@
 The next section describes the new algorithm for manual control, for both projections and slices.
-The use of these methods \textcolor{blue}{are}\textcolor{red}{ is} illustrated to compare and contrast boundaries constructed by different classifiers.
+The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers.
 The software section describes a mathematica notebook that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
-The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two mathematica notebooks, one focused on the application, and another that could be used generically for new data.
+The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.
 
@@
 
-The problem with \textcolor{red}{the} new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
-\textcolor{red}{Even though these changes are for the most part imperceptible, one may wish to avoid them and} there are numerous ways that this can be enforced, a few are detailed in the Appendix.
-\textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
+The problem with the new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
+Even though these changes are for the most part imperceptible, one may wish to avoid them and there are numerous ways that this can be enforced, a few are detailed in the Appendix.
+These primarily differ in how the remaining variables are adjusted during orthonormalisation.
 
@@
 # Software {#sec:implementation}
-
-## Structure
 

### 2022-09-23 paper/paper.tex
handled all the red/blue text

@@
 data using low-dimensional linear projections and slices. A user can
-interactively change the
-\textcolor{blue} {coefficients of a variable in}
-\textcolor{red}{contribution of a given variable to} a low-dimensional
-projection, which is useful for exploring the sensitivity of structure
-to particular variables. The user can also interactively shift the slice
-center, which is useful for exploring how structure changes in local
-subspaces. These controls are implemented in Mathematica to take
-advantage of \textcolor{blue}{it's} \textcolor{red}{its}
-\textcolor{blue}{remove this: mathematics functions, and interactivity tools}
-\textcolor{red}{location controls} that allow construction of functions
-that interact with the position of the cursor. We demonstrate the
-usefulness of this new tool with an example exploring region boundaries
-in classification problems.
+interactively change the contribution of a given variable to a
+low-dimensional projection, which is useful for exploring the
+sensitivity of structure to particular variables. The user can also
+interactively shift the center of a slice, for example, to explore how
+structure changes in local subspaces. Two Mathematica notebooks are
+provided, which contains functions enabling the user to experiment with
+these new manual controls, with one specifically for exploring regions
+and boundaries produced by classification models. The advantage of
+Mathematica is its linear algebra capabilities, and interactive cursor
+location controls. Some limited implementation has also been made
+available in the R package \tt{tourr}. 
 \end{abstract}
@@
 tours which allows one to quickly see many, many different projections.
-After Asimov there have been many \textcolor{blue}{remove: }, many\}
-tour developments, which are summarized in \citet{lee2021}.
+After Asimov there have been many tour developments, which are
+summarized in \citet{lee2021}.
 
@@
 places where it is useful. In exploratory data analysis, where one sees
-clusters in a projection, \textcolor{red}{} one may ask whether some\}
-variables can be removed from the projection without affecting the
-clustering. For interpreting models, one can reduce or increase a
-variable's contribution to examine the variable importance. Having the
-user interact with a projection is extremely valuable for understanding
+clusters in a projection, one may ask whether some variables can be
+removed from the projection without affecting the clustering. For
+interpreting models, one can reduce or increase a variable's
+contribution to examine the variable importance. Having the user
+interact with a projection is extremely valuable for understanding
 high-dimensional data. However, these algorithms have two problems: (1)
@@
 Another potentially useful manual control, is to allow the user to
-\textcolor{blue}{control}\textcolor{red}{choose} the position of the
-center of a slice. The slice tour was introduced in \citet{slicetour}.
-It operates by converting the projection plane into a slice, by removing
-or de-emphasizing points that are further than a fixed orthogonal
-distance from the plane. The projection plane is usually thought of as
-passing through the center of the data. Manual control would allow the
-user to change the value of center point, by shifting it along a
-coordinate axis, while keeping the orientation of the projection plane
-fixed. The purpose would be to explore how or if the shape of the data,
-in the space orthogonal to the projection, changes as one gets away from
-the center.
-\textcolor{red}{It would also allow the user to interactively decide on the thickness of the slice.}
+choose the position of the center of a slice. The slice tour was
+introduced in \citet{slicetour}. It operates by converting the
+projection plane into a slice, by removing or de-emphasizing points that
+are further than a fixed orthogonal distance from the plane. The
+projection plane is usually thought of as passing through the center of
+the data. Manual control would allow the user to change the value of
+center point, by shifting it along a coordinate axis, while keeping the
+orientation of the projection plane fixed. The purpose would be to
+explore how or if the shape of the data, in the space orthogonal to the
+projection, changes as one gets away from the center. It would also
+allow the user to interactively decide on the thickness of the slice.
 
@@
 tours. The next section describes the new algorithm for manual control,
-for both projections and slices. The use of these methods
-\textcolor{blue}{are}\textcolor{red}{ is} illustrated to compare and
-contrast boundaries constructed by different classifiers. The software
-section describes a mathematica notebook that is used for the
-application, and describes the interactive environment that would be
+for both projections and slices. The use of these methods is illustrated
+to compare and contrast boundaries constructed by different classifiers.
+The software section describes a mathematica notebook that is used for
+the application, and describes the interactive environment that would be
 desirable within R as new technology becomes available. The paper is
 accompanied by an appendix with more details and adjustments to the
-manual controls, and two mathematica notebooks, one focused on the
+manual controls, and two Mathematica notebooks, one focused on the
 application, and another that could be used generically for new data.
@@
 
-The problem with \textcolor{red}{the} new simple method is that it is
-not faithful to the precise values for \(V_m\) because the
-orthonormalisation will change them.
-\textcolor{red}{Even though these changes are for the most part imperceptible, one may wish to avoid them and}
-there are numerous ways that this can be enforced, a few are detailed in
-the Appendix.
-\textcolor{red}{While they all enforce the same position for the selected variable, they differ in how the remaining variables are adjusted during orthonormalisation.}
+The problem with the new simple method is that it is not faithful to the
+precise values for \(V_m\) because the orthonormalisation will change
+them. Even though these changes are for the most part imperceptible, one
+may wish to avoid them and there are numerous ways that this can be
+enforced, a few are detailed in the Appendix. These primarily differ in
+how the remaining variables are adjusted during orthonormalisation.
 
@@
 \section{Software}\label{sec:implementation}}
-
-\hypertarget{structure}{%
-\subsection{Structure}\label{structure}}
 

### 2022-09-23 paper/paper.Rmd
tinkered with scatterplot matrix

@@
 
-```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
+```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
 The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
@@
          "bm"="body_mass_g")
-ggscatmat(penguins_data, columns = 2:5, col="species") +
+#ggscatmat(penguins_data, columns = 2:5, col="species") +
+penguins_diag <- tibble(x=1, y=c(3,2,1),
+     species=levels(penguins_data$species))
+pscat <- ggpairs(penguins_data, columns = 2:5, 
+          mapping = ggplot2::aes(color = species, 
+                                 fill = species),
+          upper = list(continuous = "points"),
+          diag = list(continuous = "densityDiag")) +
   scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-  theme(legend.position = "bottom")
+  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  theme(aspect.ratio=1, 
+        #legend.position = "bottom", 
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
+pscat[1,1] <- ggplot(penguins_diag, 
+                     aes(x=x, 
+                         y=y, 
+                         label=species, 
+                         colour=species)) + 
+  geom_text() +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  ylim(c(0,4))
+pscat[2,2] <- ggplot(penguins_diag, 
+                     aes(x=x,
+                         y=y, 
+                         label=species, 
+                         colour=species)) + 
+  geom_text() +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  ylim(c(0,4))
+pscat[3,3] <- ggplot(penguins_diag, 
+                     aes(x=x,
+                         y=y, 
+                         label=species, 
+                         colour=species)) + 
+  geom_text() +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  ylim(c(0,4))
+pscat[4,4] <- ggplot(penguins_diag, 
+                     aes(x=x,
+                         y=y, 
+                         label=species, 
+                         colour=species)) + 
+  geom_text() +
+  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  ylim(c(0,4))
+pscat
 ```

### 2022-09-23 paper/paper.Rmd
added some sentences about manual slice

@@
 Here we found the dynamic graphics interface from Mathematica to be especially useful.
-In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approach might be implemented in the `tourr` package.
+In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approaches are implemented in the `tourr` package.
 
@@
 It is applicable for any projection dimension, $d$.
-
-Changing slice center
+Changing slice center manually can be accomplished with the new function, `manual_slice()`.
+This changes the value of the center point ($\mathbf{c}$) of the slice, along a selected variable axis.
+It slices from the center of the data out to one side, back to the center, out to the opposite side, and then back to the center.
 
@@
 
-The source material and animated gifs for this paper are available at
+The source material and animated gifs for this paper are available at [https://github.com/uschiLaa/mmtour](https://github.com/uschiLaa/mmtour).
 
-Mathematica notebook
+The supplementary materials include:
 
-Appendix
+- Two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.
+- Appendix with additional details about manual controls, and Mathematica functions.
+- Animated gifs illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2022-09-23 paper/paper.tex
tinkered with scatterplot matrix
added some sentences about manual slice

@@
 the relevant Mathematica functionalities and how they were used in the
-implementation, followed by a brief sketch of how the new approach might
-be implemented in the \texttt{tourr} package.
+implementation, followed by a brief sketch of how the new approaches are
+implemented in the \texttt{tourr} package.
 
@@
 coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
-values. It is applicable for any projection dimension, \(d\).
-
-Changing slice center
+values. It is applicable for any projection dimension, \(d\). Changing
+slice center manually can be accomplished with the new function,
+\texttt{manual\_slice()}. This changes the value of the center point
+(\(\mathbf{c}\)) of the slice, along a selected variable axis. It slices
+from the center of the data out to one side, back to the center, out to
+the opposite side, and then back to the center.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
+{\centering \includegraphics[width=0.6\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
 
@@
 The source material and animated gifs for this paper are available at
+\url{https://github.com/uschiLaa/mmtour}.
 
-Mathematica notebook
+The supplementary materials include:
 
-Appendix
+\begin{itemize}
+\tightlist
+\item
+  Two Mathematica notebooks, one focused on the application, and another
+  that could be used generically for new data.
+\item
+  Appendix with additional details about manual controls, and
+  Mathematica functions.
+\item
+  Animated gifs illustrating the manual tour and slicing, matching the
+  static figures in the paper. These are also available at
+  \url{https://vimeo.com/747585410} and
+  \url{https://vimeo.com/747590472}.
+\end{itemize}
 

### 2022-09-24 paper/appendix.Rmd
I fixed some typos

@@
 
-The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
+The problem with the new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
 
@@
 
-For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$, formed by its first two columns.
+For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}.
 Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$.
@@
 \begin{enumerate}
-\item Cursor movement takes ${x_{m1},x_{m2}}\to {a^*_{m1},a^*_{m2}}$
+\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
 \item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
@@
 \item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
 \end{enumerate}
@@
 The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane.
-However, it renders projections in a way that is not continuouns and may be distracting.
+However, it renders projections in a way that is not continuous  and may be distracting.
 This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.
@@
 \end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
+\label{matrices}
 \end{align}
@@
 
-{\textcolor{blue} Do we need three methods? is this one implemented somewhere?}
+{\textcolor{blue} Ursula, I am more convinced now that this is exactly method 2, the only difference is that you want to rotate to get to the "first" direction, whereas I simply relabel the rows of the matrix}
 
@@
 The function renders a manual tour with controls that allow the user to navigate through different projections.
-For a given projection, a slider permits the user to change the slide width within its range and a separate box switches the view to a projection.
+For a given projection, a slider permits the user to change the slice width within its range and a separate box switches the view to a projection.
 The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders.
@@
 The Slice Tour* notebooks include the additional functions: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
+
+\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}
 

### 2022-09-24 paper/appendix.tex
I fixed some typos

@@
 
-The problem with new simple method (Algorithm 1) is that the precise
+The problem with the new simple method (Algorithm 1) is that the precise
 values for \(V_m\) cannot be specified because the orthonormalisation
@@
 For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
-of \(O\), formed by its first two columns. Whereas orthonormality of the
-basis for the p-dimensional space is given by
+of \(O\) (an orthonormal basis for the p-dimensional space), formed by
+its first two columns as illustrated in
+Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
+for the p-dimensional space is given by
 \(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
@@
 exploration of dimensions orthogonal to the projection plane. However,
-it renders projections in a way that is not continuouns and may be
+it renders projections in a way that is not continuous and may be
 distracting. This can be alleviated by replacing the random completion
@@
 
-\{\textcolor{blue} Do we need three methods? is this one implemented
-somewhere?\}
+\{\textcolor{blue} Ursula, I am more convinced now that this is exactly
+method 2, the only difference is that you want to rotate to get to the
+``first'' direction, whereas I simply relabel the rows of the matrix\}
 
@@
 navigate through different projections. For a given projection, a slider
-permits the user to change the slide width within its range and a
+permits the user to change the slice width within its range and a
 separate box switches the view to a projection. The user can also change
@@
 \texttt{VisualiseSliceDyanmic}.
+
+\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}
 

### 2022-09-24 paper/paper.Rmd
I fixed some typos

@@
 In addition, we also want to be able to look at slices of the data and select slicing parameters interactively.
-Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice height $h$ and a numeric input to specify the center point $c$ explicitly.
+Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness $h$ (dubbed `height` in the `Mathematica` package) and a numeric input to specify the center point $c$ explicitly.
 The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).
@@
 
-The most relevant inbuilt Mathematica function for our purpose is the locator pane.
+The most relevant inbuilt Mathematica function for our purpose is the `LocatorPane`.
 This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
@@
 This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group).
-After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right.
+After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualisation appears on the right.
 The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre.
-The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided.
+The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided.
 It is worth noting that this zoom works best with data that has been centred and scaled.
@@
 The explicit projection matrix can also be displayed via another checkbox.
-This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration later, as shown in the applications below.
+This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration into a later study, as shown in the applications below.
 
@@
 
-Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here.
+Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice, similarly to how the manual tour was implemented here.
 An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
@@
 
-We start by first exploring the projections of the model prediction.
+We start by exploring the projections of the model prediction.
 Figure \ref{proj1} summarises the process.
-Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models.
-To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model.
-This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot).
+Rotating manually the view, we can visualise the location of predictions for each of the three species, and also get a sense of the difference between the two models.
+To illustrate this difference, we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model.
+This particular projection ($A_1$) is exported so that it can later be used to show the LDA model (middle plot) and the actual data (right plot).
 What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes.
-And, interestingly this particular projection of the original data shows very distinct clusters of the three species.
+And, interestingly, this particular projection of the original data shows very distinct clusters of the three species.
 That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
@@
 
-We continue the investigation by now slicing orthogonally to  the projection $A_1$.
+We now continue the investigation by slicing orthogonally to  the projection $A_1$.
 For both models we look at a thin ($h=0.5$) slice through the center, $\mathbf{c}^0 = (0,0,0,0)$.
@@
 
-Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Finally, to determine which model describes the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
 Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.

### 2022-09-24 paper/paper.tex
I fixed some typos

@@
 switch to change between projection and slice view, a slider that
-adjusts the slice height \(h\) and a numeric input to specify the center
+adjusts the slice thickness \(h\) (dubbed \texttt{height} in the
+\texttt{Mathematica} package) and a numeric input to specify the center
 point \(c\) explicitly. The display will update based on these inputs,
@@
 The most relevant inbuilt Mathematica function for our purpose is the
-locator pane. This creates a region on the screen where the position of
-the mouse is captured and then converted to input that updates the
-graphics functions, enabling the manual navigation of the tour.
+\texttt{LocatorPane}. This creates a region on the screen where the
+position of the mouse is captured and then converted to input that
+updates the graphics functions, enabling the manual navigation of the
+tour.
 
@@
 display in which the control objects appear on the left and the slice
-visulisation appears on the right. The user can change the orientation
+visualisation appears on the right. The user can change the orientation
 of the slice via the locator pane, which changes the projection matrix;
@@
 of the plot by zooming into the center or changing the point size with
-sliders provided. It is worth noting that this zoom works best with data
-that has been centred and scaled. The projected data can be displayed to
-contrast it with the slice by ticking a box. The explicit projection
-matrix can also be displayed via another checkbox. This is especially
-useful when we wish to import a projection that was identified as
-interesting in the manual exploration later, as shown in the
-applications below.
+the sliders provided. It is worth noting that this zoom works best with
+data that has been centred and scaled. The projected data can be
+displayed to contrast it with the slice by ticking a box. The explicit
+projection matrix can also be displayed via another checkbox. This is
+especially useful when we wish to import a projection that was
+identified as interesting in the manual exploration into a later study,
+as shown in the applications below.
 
@@
 the current center point, potentially with interactive graphics to move
-the slice similar to how the manual tour was implemented here. An
+the slice, similarly to how the manual tour was implemented here. An
 alternative would be a step-wise shift of the slice towards the new
@@
 
-We start by first exploring the projections of the model prediction.
-Figure \ref{proj1} summarises the process. Through manual rotation of
-the view we can get a feeling for where in the space we primarily
-predict each of the three species, and we also get a sense of the
-difference between the two models. To illustrate this difference we have
-manually rotated the projection for the RF model (left plot) to identify
-a view that shows the non-linear but block-type structure that is
-typical of this type of model. This particular projection (\(A_1\)) is
-exported so that the same projection can be used to show the LDA model
-(middle plot) and the actual data (right plot). What can be seen is the
-linearity of the LDA model, where the boundaries are linear and oblique
-to the variable axes. And, interestingly this particular projection of
-the original data shows very distinct clusters of the three species.
-That means, the obscuring of the boundaries between groups for both of
-the models is driven by what is happening in the orthogonal space to the
-plane of the selected projection.
+We start by exploring the projections of the model prediction. Figure
+\ref{proj1} summarises the process. Rotating manually the view, we can
+visualise the location of predictions for each of the three species, and
+also get a sense of the difference between the two models. To illustrate
+this difference, we have manually rotated the projection for the RF
+model (left plot) to identify a view that shows the non-linear but
+block-type structure that is typical of this type of model. This
+particular projection (\(A_1\)) is exported so that it can later be used
+to show the LDA model (middle plot) and the actual data (right plot).
+What can be seen is the linearity of the LDA model, where the boundaries
+are linear and oblique to the variable axes. And, interestingly, this
+particular projection of the original data shows very distinct clusters
+of the three species. That means, the obscuring of the boundaries
+between groups for both of the models is driven by what is happening in
+the orthogonal space to the plane of the selected projection.
 
@@
 
-We continue the investigation by now slicing orthogonally to the
+We now continue the investigation by slicing orthogonally to the
 projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
@@
 
-Finally, to determine which model does the boundaries better, compare
-them with the \(A_2\) projection of the data (Figure \ref{proj2}).
-Gentoo is more separated from the other two in this projection, and one
-can imagine that the trees in the forest has greedily grasped any one of
-many places to make a split to separate the group. It might be argued
-though the that RF boundary is cut too close to the Chinstrap species,
-and might lead to some unnecessary misclassification with new data. The
-LDA boundary is better placed for all species.
+Finally, to determine which model describes the boundaries better,
+compare them with the \(A_2\) projection of the data (Figure
+\ref{proj2}). Gentoo is more separated from the other two in this
+projection, and one can imagine that the trees in the forest has
+greedily grasped any one of many places to make a split to separate the
+group. It might be argued though the that RF boundary is cut too close
+to the Chinstrap species, and might lead to some unnecessary
+misclassification with new data. The LDA boundary is better placed for
+all species.
 

### 2022-09-30 paper/appendix.Rmd
navigation for slice center in appendix and removed small section from main paper

@@
 
+# Slice center guide
+
+The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space.
+By default it is placed in the center of the data (plot A).
+Moving this along a variable axis moves the slice outwards from the center of the data (plots B, C).
+
+```{r anchornav, fig.width=10, fig.height=3, out.width="100%", fig.cap="Visual guide for the slice center.
+Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+The position of the center corresponds to the dark polygon.
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis."}
+par(pty="s", mfrow=c(1,3))
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="A. data center")
+anchor = data.frame(matrix(rep(0.5, 6), ncol=6))
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="B. center moved towards max value of X3")
+anchor[3] <- 0.75
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="C. center moved towards min value of X3")
+anchor[3] <- 0.25
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+```
+
 # Software details

### 2022-09-30 paper/appendix.tex
navigation for slice center in appendix and removed small section from main paper

@@
 
+\hypertarget{slice-center-guide}{%
+\section{Slice center guide}\label{slice-center-guide}}
+
+The slice center, \(c\), from Equation 1 in the main paper, is a point
+in the \(p\)-dimensional space. By default it is placed in the center of
+the data (plot A). Moving this along a variable axis moves the slice
+outwards from the center of the data (plots B, C).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{appendix_files/figure-latex/anchornav-1} \caption{Visual guide for the slice center.
+Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+The position of the center corresponds to the dark polygon.
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis.}\label{fig:anchornav}
+\end{figure}
+
 \hypertarget{software-details}{%

### 2022-09-30 paper/paper.Rmd
navigation for slice center in appendix and removed small section from main paper

@@
 In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approaches are implemented in the `tourr` package.
+There is hope that better interactivity will be available for R soon, which might allow for implementation of the techniques developed in Mathematica.
 
@@
 It slices from the center of the data out to one side, back to the center, out to the opposite side, and then back to the center.
+The appendix contains a discussion of a possible visual guide for the slice center location.
 
+<!--
+I don't think the paper gains anything by having this section.
 ## Desirable interactivity
@@
 An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
+-->
 

### 2022-09-30 paper/paper.tex
navigation for slice center in appendix and removed small section from main paper

@@
 implementation, followed by a brief sketch of how the new approaches are
-implemented in the \texttt{tourr} package.
+implemented in the \texttt{tourr} package. There is hope that better
+interactivity will be available for R soon, which might allow for
+implementation of the techniques developed in Mathematica.
 
@@
 from the center of the data out to one side, back to the center, out to
-the opposite side, and then back to the center.
-
-\hypertarget{desirable-interactivity}{%
-\subsection{Desirable interactivity}\label{desirable-interactivity}}
-
-Potential adjustments for future work include a better visualization of
-the current center point, potentially with interactive graphics to move
-the slice, similarly to how the manual tour was implemented here. An
-alternative would be a step-wise shift of the slice towards the new
-center point to get a better understanding of the distribution along the
-selected direction.
+the opposite side, and then back to the center. The appendix contains a
+discussion of a possible visual guide for the slice center location.
 

### 2022-09-30 paper/appendix.Rmd
figure link made

@@
 The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space.
-By default it is placed in the center of the data (plot A).
-Moving this along a variable axis moves the slice outwards from the center of the data (plots B, C).
+By default it is placed in the center of the data (Figure \ref{fig:anchornav} A).
+Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C).
 

### 2022-09-30 paper/appendix.tex
figure link made

@@
 in the \(p\)-dimensional space. By default it is placed in the center of
-the data (plot A). Moving this along a variable axis moves the slice
-outwards from the center of the data (plots B, C).
+the data (Figure \ref{fig:anchornav} A). Moving this along a variable
+axis moves the slice outwards from the center of the data (Figure
+\ref{fig:anchornav} B, C).
 

### 2022-10-01 paper/appendix.Rmd
finished paragraph on visual display of slice position

@@
 The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space.
-By default it is placed in the center of the data (Figure \ref{fig:anchornav} A).
+By default, it is placed in the center of the data (Figure \ref{fig:anchornav} A).
 Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C).
+The visual display of the center position is a form of star plot as suggested useful in \textcolor{red}{REF to Catherine's paper}, where the position of the point is marked by a polygon on radial axes indicating its position relative to the variable minimum to maximum values.
 

### 2022-10-01 paper/appendix.tex
finished paragraph on visual display of slice position

@@
 The slice center, \(c\), from Equation 1 in the main paper, is a point
-in the \(p\)-dimensional space. By default it is placed in the center of
-the data (Figure \ref{fig:anchornav} A). Moving this along a variable
+in the \(p\)-dimensional space. By default, it is placed in the center
+of the data (Figure \ref{fig:anchornav} A). Moving this along a variable
 axis moves the slice outwards from the center of the data (Figure
-\ref{fig:anchornav} B, C).
+\ref{fig:anchornav} B, C). The visual display of the center position is
+a form of star plot as suggested useful in
+\textcolor{red}{REF to Catherine's paper}, where the position of the
+point is marked by a polygon on radial axes indicating its position
+relative to the variable minimum to maximum values.
 

### 2022-10-04 paper/appendix.Rmd
Fixing the appendix

--- 2022-09-30 paper/appendix.Rmd
@@
 
-## Adjustment method 3
-
-Finally we may also consider an adjustment that works within the projection plane.
-The idea is that when applying Gram-Schmidt, the direction of the first basis vector is not changed.
-We thus rotate the projection matrix within the plane such that the first basis is aligned with the direction of the selected variable $m$ before applying orthonormalization.
-In addition we need to keep the contribution of variable $m$ along that direction fixed when normalizing the vector.
-The following steps are needed:
-
-1. Provide $A$, and $m$.
-2. Change values in row $m$, giving $A^*$.
-3.
-Rotate the basis within the projection plane, such that the first basis direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$.
-In a 2D projection this means finding the angle $\theta$ by which the basis should be rotated.
-4. Normalize the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
-5. Orthonormalize $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
-6. Rotate $\hat A^{**}$ back into the original basis to obtain $A^{**}$.
-
-Diagram needed?
-
-{\textcolor{blue} Ursula, I am more convinced now that this is exactly method 2, the only difference is that you want to rotate to get to the "first" direction, whereas I simply relabel the rows of the matrix}
+\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
+In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
+Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
+The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}
 
@@
 The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space.
-By default it is placed in the center of the data (Figure \ref{fig:anchornav} A).
+By default, it is placed in the center of the data (Figure \ref{fig:anchornav} A).
 Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C).
+The visual display of the center position is a form of star plot as suggested useful in @condviz2, where the position of the point is marked by a polygon on radial axes indicating its position relative to the variable minimum to maximum values.
 
@@
 
-The Slice Tour* notebooks include the additional functions: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
-
-\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}
+The package includes the additional functions not used in the example application: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
 

### 2022-10-04 paper/appendix.tex
Fixing the appendix

--- 2022-09-30 paper/appendix.tex
@@
 
-\hypertarget{adjustment-method-3}{%
-\subsection{Adjustment method 3}\label{adjustment-method-3}}
-
-Finally we may also consider an adjustment that works within the
-projection plane. The idea is that when applying Gram-Schmidt, the
-direction of the first basis vector is not changed. We thus rotate the
-projection matrix within the plane such that the first basis is aligned
-with the direction of the selected variable \(m\) before applying
-orthonormalization. In addition we need to keep the contribution of
-variable \(m\) along that direction fixed when normalizing the vector.
-The following steps are needed:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Provide \(A\), and \(m\).
-\item
-  Change values in row \(m\), giving \(A^*\).
-\item
-  Rotate the basis within the projection plane, such that the first
-  basis direction is aligned with the direction of \(m\) in \(A^*\),
-  giving \(\hat A^*\). In a 2D projection this means finding the angle
-  \(\theta\) by which the basis should be rotated.
-\item
-  Normalize the first direction in \(\hat A^*\) while keeping the
-  contribution in row \(m\) fixed, giving \(\hat A^{*0}\).
-\item
-  Orthonormalize \(\hat A^{*0}\) using Gram-Schmidt, giving
-  \(\hat A^{**}\). By construction the contribution of \(m\) will remain
-  fixed.
-\item
-  Rotate \(\hat A^{**}\) back into the original basis to obtain
-  \(A^{**}\).
-\end{enumerate}
-
-Diagram needed?
-
-\{\textcolor{blue} Ursula, I am more convinced now that this is exactly
-method 2, the only difference is that you want to rotate to get to the
-``first'' direction, whereas I simply relabel the rows of the matrix\}
+\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
+In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
+Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
+The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}
 
@@
 The slice center, \(c\), from Equation 1 in the main paper, is a point
-in the \(p\)-dimensional space. By default it is placed in the center of
-the data (Figure \ref{fig:anchornav} A). Moving this along a variable
+in the \(p\)-dimensional space. By default, it is placed in the center
+of the data (Figure \ref{fig:anchornav} A). Moving this along a variable
 axis moves the slice outwards from the center of the data (Figure
-\ref{fig:anchornav} B, C).
+\ref{fig:anchornav} B, C). The visual display of the center position is
+a form of star plot as suggested useful in \citet{condviz2}, where the
+position of the point is marked by a polygon on radial axes indicating
+its position relative to the variable minimum to maximum values.
 
@@
 
-The Slice Tour* notebooks include the additional functions:
-\texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot}, and
-\texttt{VisualiseSliceDyanmic}.
-
-\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}
+The package includes the additional functions not used in the example
+application: \texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot},
+and \texttt{VisualiseSliceDyanmic}.
 

### 2022-10-04 paper/appendix.Rmd
Spellcheck and typos

@@
 - `flagQ`: A flag signaling whether the data matrix includes a numerical flag for the groups.
-- `colorFunc`: specifies the colour mapping for the different groups.
+- `colorFunc`: specifies the color mapping for the different groups.
 
@@
 For a given projection, a slider permits the user to change the slice width within its range and a separate box switches the view to a projection.
-The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders.
+The user can also change the center point of slices; the size of the points rendered and the scale of the plot through sliders.
 An additional box displays the coordinates of the projection matrix explicitly.
@@
 This function can be used for ungrouped data by setting legendQ and flagQ to 0.
-To remove a group from the visual display, the user can choose specific colours for the entry colorFunc.
+To remove a group from the visual display, the user can choose specific colors for the entry colorFunc.
 For example, {Red,White,Green} would display groups 1 (red) and 3 (green) while making 2 invisible.
@@
 - `legendNames`: an optional list of labels for the groups.
-- `colorFunc`: specifies the colour mapping for the different groups.
+- `colorFunc`: specifies the color mapping for the different groups.
 
@@
 Several locator panes are displayed above the display of the projected data and each one corresponds to a row in the projection matrix.
-The updating behaviour of the projection is the same as in `ProjectionPlot`.
+The updating behavior of the projection is the same as in `ProjectionPlot`.
 The function and its arguments are given below:
@@
 This function differs from `SliceDynamic` in that it looks at only one data set but displays both the points inside and outside the slice.
-The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, centre point and zoom level.
+The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, center point and zoom level.
 The function and its arguments are given below:

### 2022-10-04 paper/appendix.tex
Spellcheck and typos

@@
 \item
-  \texttt{colorFunc}: specifies the colour mapping for the different
+  \texttt{colorFunc}: specifies the color mapping for the different
   groups.
@@
 separate box switches the view to a projection. The user can also change
-the centre point of slices; the size of the points rendered and the
+the center point of slices; the size of the points rendered and the
 scale of the plot through sliders. An additional box displays the
@@
 flagQ to 0. To remove a group from the visual display, the user can
-choose specific colours for the entry colorFunc. For example,
+choose specific colors for the entry colorFunc. For example,
 \{Red,White,Green\} would display groups 1 (red) and 3 (green) while
@@
 the projected data and each one corresponds to a row in the projection
-matrix. The updating behaviour of the projection is the same as in
+matrix. The updating behavior of the projection is the same as in
 \texttt{ProjectionPlot}. The function and its arguments are given below:
@@
 the user to specify the size of both sets of points (inside and outside)
-along with the slice height, centre point and zoom level. The function
+along with the slice height, center point and zoom level. The function
 and its arguments are given below:

### 2022-10-04 paper/paper.Rmd
Spellcheck and typos

@@
 The user can also interactively shift the center of a slice, for example, to explore how structure changes in local subspaces.
-Two Mathematica notebooks are provided, which contains functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
+Two Mathematica notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
 The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls.
@@
 From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, the more useful methods for data analysis need to work for arbitrary dimension.
-A good approch is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
+A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
 Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
@@
 The projection plane is usually thought of as passing through the center of the data.
-Manual control would allow the user to change the value of center point, by shifting it along a coordinate axis, while keeping the orientation of the projection plane fixed.
+Manual control would allow the user to change the position of the center point, by shifting it along a coordinate axis, while keeping the orientation of the projection plane fixed.
 The purpose would be to explore how or if the shape of the data, in the space orthogonal to the projection, changes as one gets away from the center.
@@
 
-The methods for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
+The methods for updating component values in @cook_manual_1997 (and utilized in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming.
 The first step was to construct a 3D manipulation space from a 2D projection.
@@
 
-The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in @Mathematica.
+The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in Mathematica [@Mathematica].
 The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
@@
 
-Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
+Mathematica provides much utility and versatility, such as inbuilt data visualization, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
 Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
@@
 This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group).
-After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualisation appears on the right.
-The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre.
+After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualization appears on the right.
+The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice center.
 The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided.
-It is worth noting that this zoom works best with data that has been centred and scaled.
+It is worth noting that this zoom works best with data that has been centered and scaled.
 The projected data can be displayed to contrast it with the slice by ticking a box.
@@
     - `data`: A data matrix
-    - `projmat`: The inital projection matrix
+    - `projmat`: The initial projection matrix
     - `centrepoint`: The initial position of the slice
@@
 
-This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.
+This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalized to arbitrary data if we want to keep it.
 
@@
 We start by exploring the projections of the model prediction.
-Figure \ref{proj1} summarises the process.
-Rotating manually the view, we can visualise the location of predictions for each of the three species, and also get a sense of the difference between the two models.
+Figure \ref{proj1} summarizes the process.
+Rotating manually the view, we can visualize the location of predictions for each of the three species, and also get a sense of the difference between the two models.
 To illustrate this difference, we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model.

### 2022-10-04 paper/paper.tex
Spellcheck and typos

@@
 structure changes in local subspaces. Two Mathematica notebooks are
-provided, which contains functions enabling the user to experiment with
+provided, which contain functions enabling the user to experiment with
 these new manual controls, with one specifically for exploring regions
@@
 most 3D rotation computer graphics applications, the more useful methods
-for data analysis need to work for arbitrary dimension. A good approch
+for data analysis need to work for arbitrary dimension. A good approach
 is to show projections from an arbitrary dimensional space to create
@@
 projection plane is usually thought of as passing through the center of
-the data. Manual control would allow the user to change the value of
-center point, by shifting it along a coordinate axis, while keeping the
-orientation of the projection plane fixed. The purpose would be to
+the data. Manual control would allow the user to change the position of
+the center point, by shifting it along a coordinate axis, while keeping
+the orientation of the projection plane fixed. The purpose would be to
 explore how or if the shape of the data, in the space orthogonal to the
@@
 The methods for updating component values in \citet{cook_manual_1997}
-(and utilised in \citet{spyrison_spinifex_2020}) are prescribed
+(and utilized in \citet{spyrison_spinifex_2020}) are prescribed
 primarily for a 2D projection, to take advantage of (then) newly
@@
 algebra capabilities, and relatively new interactive graphics interface,
-available in \citet{Mathematica}. The components corresponding to
-\(V_m\) are directly controlled by cursor movement, which updates row
-\(m\) of \(A\). The updated matrix is then orthonormalised.
+available in Mathematica \citep{Mathematica}. The components
+corresponding to \(V_m\) are directly controlled by cursor movement,
+which updates row \(m\) of \(A\). The updated matrix is then
+orthonormalised.
 
@@
 Mathematica provides much utility and versatility, such as inbuilt data
-visualisation, data manipulation and analysis, dynamic functionality,
+visualization, data manipulation and analysis, dynamic functionality,
 and symbolic and numeric computation. Most of this inbuilt functionality
@@
 display in which the control objects appear on the left and the slice
-visualisation appears on the right. The user can change the orientation
+visualization appears on the right. The user can change the orientation
 of the slice via the locator pane, which changes the projection matrix;
 a slider controls the slice thickness; and there is an input field,
-which changes the slice centre. The user can also change the appearance
+which changes the slice center. The user can also change the appearance
 of the plot by zooming into the center or changing the point size with
 the sliders provided. It is worth noting that this zoom works best with
-data that has been centred and scaled. The projected data can be
+data that has been centered and scaled. The projected data can be
 displayed to contrast it with the slice by ticking a box. The explicit
@@
 We start by exploring the projections of the model prediction. Figure
-\ref{proj1} summarises the process. Rotating manually the view, we can
-visualise the location of predictions for each of the three species, and
+\ref{proj1} summarizes the process. Rotating manually the view, we can
+visualize the location of predictions for each of the three species, and
 also get a sense of the difference between the two models. To illustrate

### 2022-10-04 paper/appendix.Rmd
removed red in appendix

@@
 
-\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
+We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
 In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
@@
 By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
-The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}
 
@@
 
-\textcolor{red}{The first four arguments are required.
-As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.}
+The first four arguments are required.
+As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.
 

### 2022-10-04 paper/appendix.tex
removed red in appendix

@@
 
-\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
-In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
-Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor.
-By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
-The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}
+We could also take a similar approach without completing the basis,
+through rotation within the \(d\)-dimensional hyperplane defined by the
+new projection. In that case we would again start from \(A^*\) as above,
+and then rotate the basis such that the first direction is aligned with
+the selected variable \(m\). Applying Gram-Schmid within that basis
+definition will not alter the direction for the first vector, thus the
+direction of \(m\) remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the
+lenght remains as selcted manually by the user.
 
@@
 
-\textcolor{red}{The first four arguments are required.
-As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.}
+The first four arguments are required. As possible starting values we
+suggest using a random input projection, and selecting slice height
+parameters using the estimate given in Section 2.4.2 of the paper.
 

### 2022-10-07 paper/paper.Rmd
Mathematica package vs notebook, doublespacing

@@
 The user can also interactively shift the center of a slice, for example, to explore how structure changes in local subspaces.
-Two Mathematica notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
+The Mathematica package as well as example notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
 The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls.
@@
   \def\tightlist{}
-  \usepackage{setspace}
+  \usepackage{setspace}\doublespacing
   \usepackage{graphicx}
@@
 The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers.
-The software section describes a mathematica notebook that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
-The paper is accompanied by an appendix with more details and adjustments to the manual controls, and two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.
+The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
+The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the application.
 
@@
 
-## Mathematica notebook
+## Mathematica package
 
@@
 
-- Two Mathematica notebooks, one focused on the application, and another that could be used generically for new data.
+- The Mathematica source code defining the new functions in mmtour.wl.
+- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
 - Appendix with additional details about manual controls, and Mathematica functions.
-- Animated gifs illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.
+- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2022-10-07 paper/paper.tex
Mathematica package vs notebook, doublespacing

@@
 \def\tightlist{}
-\usepackage{setspace}
+\usepackage{setspace}\doublespacing
 \usepackage{graphicx}
@@
 interactively shift the center of a slice, for example, to explore how
-structure changes in local subspaces. Two Mathematica notebooks are
-provided, which contain functions enabling the user to experiment with
-these new manual controls, with one specifically for exploring regions
-and boundaries produced by classification models. The advantage of
-Mathematica is its linear algebra capabilities, and interactive cursor
-location controls. Some limited implementation has also been made
-available in the R package \tt{tourr}. 
+structure changes in local subspaces. The Mathematica package as well as
+example notebooks are provided, which contain functions enabling the
+user to experiment with these new manual controls, with one specifically
+for exploring regions and boundaries produced by classification models.
+The advantage of Mathematica is its linear algebra capabilities, and
+interactive cursor location controls. Some limited implementation has
+also been made available in the R package \tt{tourr}. 
 \end{abstract}
@@
 to compare and contrast boundaries constructed by different classifiers.
-The software section describes a mathematica notebook that is used for
+The software section describes a mathematica package that is used for
 the application, and describes the interactive environment that would be
@@
 accompanied by an appendix with more details and adjustments to the
-manual controls, and two Mathematica notebooks, one focused on the
-application, and another that could be used generically for new data.
+manual controls, and three Mathematica notebooks that can be used to
+reproduce the application.
 
@@
 
-\hypertarget{mathematica-notebook}{%
-\subsection{Mathematica notebook}\label{mathematica-notebook}}
+\hypertarget{mathematica-package}{%
+\subsection{Mathematica package}\label{mathematica-package}}
 
@@
 \item
-  Two Mathematica notebooks, one focused on the application, and another
-  that could be used generically for new data.
+  The Mathematica source code defining the new functions in mmtour.wl.
+\item
+  Three Mathematica notebooks with the code used in the application
+  (corresponding to the three subsections).
 \item
@@
 \item
-  Animated gifs illustrating the manual tour and slicing, matching the
+  Animations illustrating the manual tour and slicing, matching the
   static figures in the paper. These are also available at

### 2022-10-11 arxiv/paper.tex
arxiv version

--- 2022-10-07 paper/paper.tex
@@
-% interactcadsample.tex
-% v1.03 - April 2017
+\documentclass{article}
 
-\documentclass[]{interact}
+\usepackage{arxiv}
 
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
-\usepackage{subfigure}% Support for small, `sub' figures and tables
-%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{graphicx}
 
-\usepackage{natbib}% Citation support using natbib.sty
-\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
-\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
+\title{New and simplified manual controls for projection and slice
+tours, with application to exploring classification boundaries in high
+dimensions}
 
-\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
-\newtheorem{notation}{Notation}
+\author{
+  }
 
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \usepackage{nicematrix}
-\NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
 
 \begin{document}
-
-\articletype{Short Technical Note}
-
-\title{New and simplified manual controls for projection and slice
-tours, with application to exploring classification boundaries in high
-dimensions}
-
-\author{\name{Ursula Laa$^{a}$, Alex Aumann$^{b}$, Dianne
-Cook$^{c}$, German Valencia$^{b}$}
-\affil{$^{a}$Institute of Statistics, University of Natural Resources
-and Life Sciences, Vienna; $^{b}$School of Physics and Astronomy, Monash
-University; $^{c}$Department of Econometrics and Business Statistics,
-Monash University}
-}
-
-\thanks{CONTACT Ursula
-Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Alex
-Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, Dianne
-Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, German
-Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}}
-
 \maketitle
@@
 
-\begin{keywords}
+\keywords{
     data visualisation; grand tour; statistical computing; statistical
     graphics; multivariate data; dynamic graphics
-\end{keywords}
+  }
 
@@
 high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
-In his original paper on the grand tour, \citet{As85} provided several
+In his original paper on the grand tour, Asimov (1985) provided several
 algorithms for tour paths that could theoretically show the viewer the
 data \emph{from all sides}. Prior to Asimov's work, there were numerous
-preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
-user-controlled rotations on coordinate axes, allowing one to manually
-tour through low-dimensional projections. (A video illustrating the
-capabilities is available through video library of \citet{ASA22}.)
-Steering through all possible projections is impossible, unlike Asimov's
-tours which allows one to quickly see many, many different projections.
-After Asimov there have been many tour developments, which are
-summarized in \citet{lee2021}.
+preparatory developments including Fisherkeller, Friedman, and Tukey
+(1974)'s PRIM-9. PRIM-9 had user-controlled rotations on coordinate
+axes, allowing one to manually tour through low-dimensional projections.
+(A video illustrating the capabilities is available through video
+library of ASA Statistical Graphics Section (2022).) Steering through
+all possible projections is impossible, unlike Asimov's tours which
+allows one to quickly see many, many different projections. After Asimov
+there have been many tour developments, which are summarized in Lee et
+al. (2021).
 
 One such direction of work develops the ideas from PRIM-9, to provide
-manual control of a tour. \citet{cook_manual_1997} describe controls for
-1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space,
+manual control of a tour. Cook and Buja (1997) describe controls for 1D
+(or 2D) projections, respectively in a 2D (or 3D) manipulation space,
 allowing the user to select any variable axis, and rotate it into, or
 out of, or around the projection through horizontal, vertical, oblique,
-radial or angular changes in value. \citet{spyrison_spinifex_2020}
-refined this algorithm and implemented it to generate radial tour
-animation sequences.
+radial or angular changes in value. Spyrison and Cook (2020) refined
+this algorithm and implemented it to generate radial tour animation
+sequences.
 
@@
 choose the position of the center of a slice. The slice tour was
-introduced in \citet{slicetour}. It operates by converting the
-projection plane into a slice, by removing or de-emphasizing points that
-are further than a fixed orthogonal distance from the plane. The
+introduced in Laa, Cook, and Valencia (2020). It operates by converting
+the projection plane into a slice, by removing or de-emphasizing points
+that are further than a fixed orthogonal distance from the plane. The
 projection plane is usually thought of as passing through the center of
@@
 
-The methods for updating component values in \citet{cook_manual_1997}
-(and utilized in \citet{spyrison_spinifex_2020}) are prescribed
-primarily for a 2D projection, to take advantage of (then) newly
-developed 3D trackball controls made available for computer gaming. The
-first step was to construct a 3D manipulation space from a 2D
-projection. In this space, the coefficient of the controlled variable
-ranges between -1 and 1. Movements of a cursor are recorded and
-converted into changes in the values of \(V_m\) thus changing the
-displayed 2D projection. The algorithm also provided constraints to
-horizontal, vertical, radial or angular motions only. The construction
-of the manipulation space overly complicates the manual controls,
-especially when considering possible techniques that will apply to
-arbitrary \(d\).
+The methods for updating component values in Cook and Buja (1997) (and
+utilized in Spyrison and Cook (2020)) are prescribed primarily for a 2D
+projection, to take advantage of (then) newly developed 3D trackball
+controls made available for computer gaming. The first step was to
+construct a 3D manipulation space from a 2D projection. In this space,
+the coefficient of the controlled variable ranges between -1 and 1.
+Movements of a cursor are recorded and converted into changes in the
+values of \(V_m\) thus changing the displayed 2D projection. The
+algorithm also provided constraints to horizontal, vertical, radial or
+angular motions only. The construction of the manipulation space overly
+complicates the manual controls, especially when considering possible
+techniques that will apply to arbitrary \(d\).
 
@@
 algebra capabilities, and relatively new interactive graphics interface,
-available in Mathematica \citep{Mathematica}. The components
+available in Mathematica (Wolfram Research, Inc. 2022). The components
 corresponding to \(V_m\) are directly controlled by cursor movement,
@@
 data that is defined by a projection, a center point that is anchoring
-it in the high-dimensional space and the slice thickness \(h\)
-\citep{slicetour}. A data point is inside the slice if its orthogonal
-distance from the projection plane (passing through the center point) is
-below the thickness \(h\). This orthogonal distance is computed in terms
-of the component that is normal on the projection plane. For
-\(\mathbf{x}_i\) a \(p\) dimensional data point and \(\mathbf{c}\) the
-center point (in the same \(p\) dimensional space) we compute the
-orthogonal distance as \begin{equation}
+it in the high-dimensional space and the slice thickness \(h\) (Laa,
+Cook, and Valencia 2020). A data point is inside the slice if its
+orthogonal distance from the projection plane (passing through the
+center point) is below the thickness \(h\). This orthogonal distance is
+computed in terms of the component that is normal on the projection
+plane. For \(\mathbf{x}_i\) a \(p\) dimensional data point and
+\(\mathbf{c}\) the center point (in the same \(p\) dimensional space) we
+compute the orthogonal distance as \begin{equation}
 v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
@@
 estimates of the number of points inside the slice as a function of the
-original sample size \(N\) and the number of dimensions \(p\) from
-\citet{sectionpursuit} can be used: in case of a uniform distribution
-inside a sphere of radius \(R\) a slice with thickness \(h\) will
-contain \(N_S\) points, with \begin{equation}
+original sample size \(N\) and the number of dimensions \(p\) from Laa
+et al. (2022) can be used: in case of a uniform distribution inside a
+sphere of radius \(R\) a slice with thickness \(h\) will contain \(N_S\)
+points, with \begin{equation}
 N_S(h, p, R, N) = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
@@
 
-The R package tourr \citep{tourr} provides numerous types of tours. An
-additional function, \texttt{radial\_tour()} has been constructed that
-will generate a sequence of projections that will decrease the
-coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
-values. It is applicable for any projection dimension, \(d\). Changing
-slice center manually can be accomplished with the new function,
-\texttt{manual\_slice()}. This changes the value of the center point
-(\(\mathbf{c}\)) of the slice, along a selected variable axis. It slices
-from the center of the data out to one side, back to the center, out to
-the opposite side, and then back to the center. The appendix contains a
-discussion of a possible visual guide for the slice center location.
+The R package tourr (Wickham et al. 2011) provides numerous types of
+tours. An additional function, \texttt{radial\_tour()} has been
+constructed that will generate a sequence of projections that will
+decrease the coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to
+the original values. It is applicable for any projection dimension,
+\(d\). Changing slice center manually can be accomplished with the new
+function, \texttt{manual\_slice()}. This changes the value of the center
+point (\(\mathbf{c}\)) of the slice, along a selected variable axis. It
+slices from the center of the data out to one side, back to the center,
+out to the opposite side, and then back to the center. The appendix
+contains a discussion of a possible visual guide for the slice center
+location.
 
@@
 To illustrate the usefulness of the manual controls we use the 4D
-penguins data \citep{penguins} and look at classification models
-following \citet{sam11271}. We will show how classification boundaries
-can be explored and better understood on projections and slices through
-4D space. Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix
-of this data. There are four variables
+penguins data (Horst, Hill, and Gorman 2020) and look at classification
+models following Wickham, Cook, and Hofmann (2015). We will show how
+classification boundaries can be explored and better understood on
+projections and slices through 4D space. Figure
+\ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
+There are four variables
 (\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
@@
 
-We use the \texttt{classifly} package \citep{classifly} to generate
+We use the \texttt{classifly} package (Wickham 2022) to generate
 predictions across the 4D cube spanned by the data, with two
@@
 Research Council and the ResearchFirst undergraduate research program at
-Monash University. The paper was written in \texttt{rmarkdown}
-\citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
-matrix of the penguins data was produced by the GGally \citep{GGally}
-package built on ggplot2 \citep{ggplot2} graphics. We thank the
-Institute of Statistics, BOKU, for their hospitality while part of this
-work was conducted.
+Monash University. The paper was written in \texttt{rmarkdown} (Xie,
+Allaire, and Grolemund 2018) using \texttt{knitr} (Xie 2015). The
+scatterplot matrix of the penguins data was produced by the GGally
+(Emerson et al. 2013) package built on ggplot2 (Wickham 2016) graphics.
+We thank the Institute of Statistics, BOKU, for their hospitality while
+part of this work was conducted.
 
@@
 
-\bibliographystyle{tfcad}
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-ASA22}{}}%
+ASA Statistical Graphics Section. 2022. {``Video {L}ibrary.''}
+\url{https://community.amstat.org/jointscsg-section/media/videos}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-As85}{}}%
+Asimov, D. 1985. {``{T}he {G}rand {T}our: {A} {T}ool for {V}iewing
+{M}ultidimensional {D}ata.''} \emph{SIAM Journal of Scientific and
+Statistical Computing} 6 (1): 128--43.
+https://doi.org/\url{https://doi.org/10.1137/0906011}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
+Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
+{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
+Computational and Graphical Statistics} 6 (4): 464--80.
+\url{https://doi.org/10.2307/1390747}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-GGally}{}}%
+Emerson, John W., Walton A. Green, Barret Schloerke, Jason Crowley,
+Dianne Cook, Heike Hofmann, and Hadley Wickham. 2013. {``The Generalized
+Pairs Plot.''} \emph{Journal of Computational and Graphical Statistics}
+22 (1): 79--91. \url{http://www.jstor.org/stable/43304816}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tukey}{}}%
+Fisherkeller, M. A., J. H. Friedman, and J. W. Tukey. 1974. {``{PRIM-9},
+an Interactive Multidimensional Data Display and Analysis System.''} In
+\emph{The Collected Works of John w. Tukey: Graphics 1965-1985, Volume
+v}, edited by William S. Cleveland, 340--46.
+
+\leavevmode\vadjust pre{\hypertarget{ref-penguins}{}}%
+Horst, Allison Marie, Alison Presmanes Hill, and Kristen B Gorman. 2020.
+\emph{Palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data}.
+\url{https://allisonhorst.github.io/palmerpenguins/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sectionpursuit}{}}%
+Laa, Ursula, Dianne Cook, Andreas Buja, and German Valencia. 2022.
+{``Hole or Grain? A Section Pursuit Index for Finding Hidden Structure
+in Multiple Dimensions.''} \emph{Journal of Computational and Graphical
+Statistics} 0 (0): 1--14.
+\url{https://doi.org/10.1080/10618600.2022.2035230}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-slicetour}{}}%
+Laa, Ursula, Dianne Cook, and German Valencia. 2020. {``A Slice Tour for
+Finding Hollowness in High-Dimensional Data.''} \emph{Journal of
+Computational and Graphical Statistics} 29 (3): 681--87.
+\url{https://doi.org/10.1080/10618600.2020.1777140}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee2021}{}}%
+Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Earo Wang, Nick
+Spyrison, and H. Sherry Zhang. 2021. {``Advanced {Review}: {The}
+{State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of
+{High}-Dimensional {Data}.''} \emph{arXiv:2104.08016 {[}Cs, Stat{]}},
+April. \url{http://arxiv.org/abs/2104.08016}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
+Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
+for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
+{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
+\url{https://doi.org/10.32614/RJ-2020-027}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
+Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
+Analysis}. Springer-Verlag New York.
+\url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-classifly}{}}%
+---------. 2022. \emph{Classifly: Explore Classification Models in High
+Dimensions}. \url{https://CRAN.R-project.org/package=classifly}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sam11271}{}}%
+Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
+Statistical Models: Removing the Blindfold.''} \emph{Statistical
+Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
+\url{https://doi.org/10.1002/sam.11271}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
+Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja. 2011.
+{``{tourr}: An {R} Package for Exploring Multivariate Data with
+Projections.''} \emph{Journal of Statistical Software} 40 (2): 1--18.
+https://doi.org/\url{http://dx.doi.org/10.18637/jss.v040.i02}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mathematica}{}}%
+Wolfram Research, Inc. 2022. \emph{Mathematica, {V}ersion 13.1}.
+Champaign, IL. \url{https://www.wolfram.com/mathematica}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed.
+Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
+Xie, Yihui, Joseph J. Allaire, and Garrett Grolemund. 2018. \emph{R
+Markdown: The Definitive Guide}. Chapman; Hall/CRC.
+\url{https://bookdown.org/yihui/rmarkdown}.
+
+\end{CSLReferences}
+
+\bibliographystyle{unsrt}
 \bibliography{biblio.bib}

### 2022-10-11 arxiv/paper.tex
version arxiv that almost works

@@
 dimensions}
-
-\author{
-  }
 
@@
 
+\NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
+
 \begin{document}
 \maketitle
+
+\author{\name{Ursula Laa$^{a}$, Alex Aumann$^{b}$, Dianne
+Cook$^{c}$, German Valencia$^{b}$}
+\affil{$^{a}$Institute of Statistics, University of Natural Resources
+and Life Sciences, Vienna; $^{b}$School of Physics and Astronomy, Monash
+University; $^{c}$Department of Econometrics and Business Statistics,
+Monash University}
+}
+
+\thanks{CONTACT Ursula
+Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Alex
+Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, Dianne
+Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, German
+Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}}
 
@@
 
+\appendix
+
+\hypertarget{refinements-to-enforce-exact-position}{%
+\section{Refinements to enforce exact
+position}\label{refinements-to-enforce-exact-position}}
+
+The problem with the new simple method (Algorithm 1) is that the precise
+values for \(V_m\) cannot be specified because the orthonormalisation
+will change them.
+
+\hypertarget{adjustment-method-1}{%
+\subsection{Adjustment method 1}\label{adjustment-method-1}}
+
+A small modification to algorithm 1 will maintain the components of
+\(V_m\) precisely (Figure \ref{fig:othermethod}). It is as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Provide \(A\), and \(m\).
+\item
+  Change values in row \(m\), giving \(A^*\).
+\item
+  Store row \(m\) separately, and zero the values of row \(m\) in
+  \(A^*\), giving \(A^{*0}\).
+\item
+  Orthonormalise \(A^{*0}\), using Gram-Schmidt.
+\item
+  Replace row \(m\) with the original values, giving \(A^{**}\).
+\item
+  For \(d=2\), adjust the values of \({\boldmath a}^{**}_{.2}\) using
+\end{enumerate}
+
+\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m.\]
+
+which ensures that
+
+\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0.\]
+
+If \(d>2\) the process would be sequentially repeated in the same manner
+that Gram-Schmidt is applied sequentially to orthormalise the columns of
+a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
+vector would simply need to be normalized after each adjustment.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{appendix_files/figure-latex/othermethod-1} \caption{Manual controls with algorithm 2. The precise location of the axis is maintained.}\label{fig:othermethod}
+\end{figure}
+
+\hypertarget{adjustment-method-2}{%
+\subsection{Adjustment method 2}\label{adjustment-method-2}}
+
+For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
+of \(O\) (an orthonormal basis for the p-dimensional space), formed by
+its first two columns as illustrated in
+Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
+for the p-dimensional space is given by
+\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
+projection matrix is expressed as
+\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
+Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
+a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
+constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
+to guarantee orthonormality of the new projection matrix. One possible
+algorithm to achieve this is
+
+\begin{enumerate}
+\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
+\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
+\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
+\end{enumerate}
+\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
+\end{enumerate}
+
+The random completion of \(O\) outside the projection matrix provides an
+exploration of dimensions orthogonal to the projection plane. However,
+it renders projections in a way that is not continuous and may be
+distracting. This can be alleviated by replacing the random completion
+with a rule restricting the size of the jumps, for example in step (b)
+the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
+orthogonalisation.
+
+\begin{align}
+O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
+e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
+e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
+\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
+\label{matrices}
+\end{align}
+
+We could also take a similar approach without completing the basis,
+through rotation within the \(d\)-dimensional hyperplane defined by the
+new projection. In that case we would again start from \(A^*\) as above,
+and then rotate the basis such that the first direction is aligned with
+the selected variable \(m\). Applying Gram-Schmid within that basis
+definition will not alter the direction for the first vector, thus the
+direction of \(m\) remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the
+lenght remains as selcted manually by the user.
+
+\hypertarget{slice-center-guide}{%
+\section{Slice center guide}\label{slice-center-guide}}
+
+The slice center, \(c\), from Equation 1 in the main paper, is a point
+in the \(p\)-dimensional space. By default, it is placed in the center
+of the data (Figure \ref{fig:anchornav} A). Moving this along a variable
+axis moves the slice outwards from the center of the data (Figure
+\ref{fig:anchornav} B, C). The visual display of the center position is
+a form of star plot as suggested useful in \cite{condviz2}, where the
+position of the point is marked by a polygon on radial axes indicating
+its position relative to the variable minimum to maximum values.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{appendix_files/figure-latex/anchornav-1} \caption{Visual guide for the slice center.
+Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+The position of the center corresponds to the dark polygon.
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis.}\label{fig:anchornav}
+\end{figure}
+
+\hypertarget{software-details}{%
+\section{Software details}\label{software-details}}
+
+Here we describe the functions implemented in the Mathematica package
+\texttt{mmtour.wl}. The main function and its arguments are given below:
+
+\texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix, potentially with grouping that has a
+  label (string) and a flag (numerical) in the last two columns.
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{height}: The initial height of the slice.
+\item
+  \texttt{heightRange}: The range of slice heights to be explored.
+\item
+  \texttt{legendQ}: A flag signaling whether the data matrix includes a
+  column with group labels (1) or not (0).
+\item
+  \texttt{flagQ}: A flag signaling whether the data matrix includes a
+  numerical flag for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+The first four arguments are required. As possible starting values we
+suggest using a random input projection, and selecting slice height
+parameters using the estimate given in Section 2.4.2 of the paper.
+
+The function renders a manual tour with controls that allow the user to
+navigate through different projections. For a given projection, a slider
+permits the user to change the slice width within its range and a
+separate box switches the view to a projection. The user can also change
+the center point of slices; the size of the points rendered and the
+scale of the plot through sliders. An additional box displays the
+coordinates of the projection matrix explicitly.
+
+This function can be used for ungrouped data by setting legendQ and
+flagQ to 0. To remove a group from the visual display, the user can
+choose specific colors for the entry colorFunc. For example,
+\{Red,White,Green\} would display groups 1 (red) and 3 (green) while
+making 2 invisible.
+
+The package includes the additional functions not used in the example
+application: \texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot},
+and \texttt{VisualiseSliceDyanmic}.
+
+\texttt{ProjectionPlot} is very similar to \texttt{SliceDynamic}, except
+it only displays projections. Using it instead of \texttt{SliceDynamic}
+simplifies the display and is more efficient. The interface is slightly
+different, the data file should not contain labels for groups, if these
+are desired they are passed through the optional argument legendNames as
+\{``group 1'', ``group 2'',\ldots\}, for example. The function and its
+arguments are given below:
+
+\texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical, last column).
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{legendNames}: an optional list of labels for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+\texttt{ProjectedLocatorPlot} displays the interactive controls slightly
+differently. Several locator panes are displayed above the display of
+the projected data and each one corresponds to a row in the projection
+matrix. The updating behavior of the projection is the same as in
+\texttt{ProjectionPlot}. The function and its arguments are given below:
+
+\texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical, last column).
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{legendNames}: an optional list of labels for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+Finally \texttt{VisualiseSliceDynamic} allows to discern which points
+exist inside the slice and outside it. This function differs from
+\texttt{SliceDynamic} in that it looks at only one data set but displays
+both the points inside and outside the slice. The display panel allows
+the user to specify the size of both sets of points (inside and outside)
+along with the slice height, center point and zoom level. The function
+and its arguments are given below:
+
+\texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix containing only one group. In this case
+  the data matrix should not contain column labels.
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{height}: The initial height of the slice.
+\item
+  \texttt{heightRange}: The range of slice heights to be explored.
+\end{itemize}
+
+The points inside the slice are shown in black whereas those outside the
+slice are shown in light blue, with labels indicating this.
+
 \hypertarget{refs}{}

### 2022-10-11 paper/arxiv.Rmd
arxiv version with both paper and appendix in one file

--- 2022-10-07 paper/paper.Rmd
@@
   New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
-type: Short Technical Note
 author:
@@
   \def\tightlist{}
-  \usepackage{setspace}\doublespacing
+  \usepackage{setspace}
   \usepackage{graphicx}
@@
 - Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.
+
+\appendix
+
+# Refinements to enforce exact position
+
+The problem with the new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
+
+## Adjustment method 1
+
+A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:othermethod}). It is as follows:
+
+1. Provide $A$, and $m$.
+2. Change values in row $m$, giving $A^*$.
+3. Store row $m$ separately, and zero the values of row $m$ in $A^*$, giving $A^{*0}$. 
+4. Orthonormalise $A^{*0}$, using Gram-Schmidt.
+5. Replace row $m$ with the original values, giving $A^{**}$. 
+6. For $d=2$, adjust the values of ${\boldmath a}^{**}_{.2}$ using
+
+$$a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m.$$
+
+which ensures that
+
+$$\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0.$$
+
+If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
+If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
+
+<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+
+Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
+
+```{r othermethod, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."}
+source("../code/plot_basis.R")
+source("../code/linear_alg.R")
+
+# Base plot set up
+plot.new()
+par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0),
+    mfrow=c(1,4))
+
+# Create manual tour example
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+
+# Now set a new position
+vchange <- 3
+eps <- 0.1
+plot_basis(A, vchange, vcol="black")
+
+# j is the component that should be kept fixed
+orthonormalize_rotated <- function(A, j){
+  m1 <- A[j, 1]
+  m2 <- A[j, 2]
+  c <- 1 / (sqrt(1 + (m2/m1)^2))
+  s <- (m2/m1) / (sqrt(1 + (m2/m1)^2))
+  rot <- matrix(c(c, -s, s, c), nrow = 2)
+  rot_inv <- matrix(c(c, s, -s, c), nrow = 2)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot %*% A[i,]
+  }
+  A[,1] <- normalize_w_fixed(A[,1], j)
+  A <- tourr::orthonormalise(A)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot_inv %*% A[i,]
+    }
+  A
+}
+
+# normalize a vector while keeping component j fixed
+normalize_w_fixed <- function(v, j){
+  c <- v[j]
+  k <- sqrt((1-c^2) / sum(v[-j]^2))
+  v[-j] <- k * v[-j]
+  v
+}
+
+# Iterate and save images
+# Unconstrained orthonormalisation
+Anew <- A
+for (i in 1:3) {
+  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+  exact <- Anew[vchange,]
+  Anew <- orthonormalize_rotated(Anew, vchange)
+  plot_basis(Anew, vchange, vcol="black")
+  points(exact[1], exact[2], pch=16, col="orange")
+}
+
+```
+
+## Adjustment method 2
+
+For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}.
+Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$.
+Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$.
+Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+One possible algorithm to achieve this is
+\begin{enumerate}
+\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
+\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
+\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
+\end{enumerate}
+\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
+\end{enumerate}
+The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane.
+However, it renders projections in a way that is not continuous  and may be distracting.
+This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.
+
+\begin{align}
+O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
+e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
+e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
+\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
+\label{matrices}
+\end{align}
+
+We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
+In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
+Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
+
+# Slice center guide
+
+The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space.
+By default, it is placed in the center of the data (Figure \ref{fig:anchornav} A).
+Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C).
+The visual display of the center position is a form of star plot as suggested useful in @condviz2, where the position of the point is marked by a polygon on radial axes indicating its position relative to the variable minimum to maximum values.
+
+```{r anchornav, fig.width=10, fig.height=3, out.width="100%", fig.cap="Visual guide for the slice center.
+Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+The position of the center corresponds to the dark polygon.
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis."}
+par(pty="s", mfrow=c(1,3))
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="A. data center")
+anchor = data.frame(matrix(rep(0.5, 6), ncol=6))
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="B. center moved towards max value of X3")
+anchor[3] <- 0.75
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+plot(c(-1, 1), c(-1, 1), type="n", axes=FALSE, xlab = "", ylab = "", main="C. center moved towards min value of X3")
+anchor[3] <- 0.25
+tourr:::draw_slice_center(
+   anchor, 
+   rng = matrix(rep(c(0,1), 6), ncol=6, byrow=FALSE),
+   limits = 1, anchor_nav = "center")
+```
+
+# Software details
+
+Here we describe the functions implemented in the Mathematica package `mmtour.wl`. The main function and its arguments are given below:
+
+`SliceDynamic[data, projmat, height, heightRange, legendQ(=1), flagQ(=1), colorFunc(=ColorData[97])]`
+
+- `data`: A data matrix, potentially with grouping that has a label (string) and a flag  (numerical) in the last two columns.
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `height`: The initial height of the slice.
+- `heightRange`: The range of slice heights to be explored.
+- `legendQ`: A flag signaling whether the data matrix includes a column with group labels (1) or not (0).
+- `flagQ`: A flag signaling whether the data matrix includes a numerical flag for the groups.
+- `colorFunc`: specifies the color mapping for the different groups.
+
+The first four arguments are required.
+As possible starting values we suggest using a random input projection, and selecting slice height parameters using the estimate given in Section 2.4.2 of the paper.
+
+The function renders a manual tour with controls that allow the user to navigate through different projections.
+For a given projection, a slider permits the user to change the slice width within its range and a separate box switches the view to a projection.
+The user can also change the center point of slices; the size of the points rendered and the scale of the plot through sliders.
+An additional box displays the coordinates of the projection matrix explicitly.
+
+This function can be used for ungrouped data by setting legendQ and flagQ to 0.
+To remove a group from the visual display, the user can choose specific colors for the entry colorFunc.
+For example, {Red,White,Green} would display groups 1 (red) and 3 (green) while making 2 invisible.
+
+The package includes the additional functions not used in the example application: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
+
+`ProjectionPlot` is very similar to `SliceDynamic`, except it only displays projections.
+Using it instead of `SliceDynamic` simplifies the display and is more efficient.
+The interface is slightly different, the data file should not contain labels for groups, if these are desired they are passed through the optional argument legendNames as {"group 1", "group 2",...}, for example.
+The function and its arguments are given below:
+
+`ProjectionPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]`
+
+- `data`: A data matrix with one or more groups labeled by a flag (numerical, last column).
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `legendNames`: an optional list of labels for the groups.
+- `colorFunc`: specifies the color mapping for the different groups.
+
+`ProjectedLocatorPlot` displays the interactive controls slightly differently.
+Several locator panes are displayed above the display of the projected data and each one corresponds to a row in the projection matrix.
+The updating behavior of the projection is the same as in `ProjectionPlot`.
+The function and its arguments are given below:
+
+`ProjectedLocatorPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]`
+
+- `data`: A data matrix with one or more groups labeled by a flag  (numerical, last column). 
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `legendNames`: an optional list of labels for the groups.
+- `colorFunc`: specifies the color mapping for the different groups.
+
+Finally `VisualiseSliceDynamic` allows to discern which points exist inside the slice and outside it.
+This function differs from `SliceDynamic` in that it looks at only one data set but displays both the points inside and outside the slice.
+The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, center point and zoom level.
+The function and its arguments are given below:
+
+`VisualiseSliceDynamic[data, projmat, height, heightRange]`
+
+- `data`: A data matrix containing only one group. In this case the data matrix should not contain column labels.
+- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
+- `height`: The initial height of the slice.
+- `heightRange`: The range of slice heights to be explored.
+
+The points inside the slice are shown in black whereas those outside the slice are shown in light blue, with labels indicating this.

### 2022-10-11 paper/arxiv.Rmd
Version working with arxiv

@@
 
-For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}.
+For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.
+(A1) .
 Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$.
@@
 
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.9\textwidth]{figures/matrix.png}}
+\end{figure*}
+
+<!-- 
 \begin{align}
@@
 \end{align}
+-->
 

### 2022-10-16 paper/arxiv.tex
removed old arxiv

--- 2022-10-07 paper/paper.tex
@@
 \def\tightlist{}
-\usepackage{setspace}\doublespacing
+\usepackage{setspace}
 \usepackage{graphicx}
@@
 
-\articletype{Short Technical Note}
+\articletype{}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
+\includegraphics[width=1\linewidth]{arxiv_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
 The dot (orange) indicates the chosen values for the controlled variable.
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{paper_files/figure-latex/penguins-scatmat-1}
+{\centering \includegraphics[width=0.6\linewidth]{arxiv_files/figure-latex/penguins-scatmat-1}
 
@@
 
+\appendix
+
+\hypertarget{refinements-to-enforce-exact-position-1}{%
+\section{Refinements to enforce exact
+position}\label{refinements-to-enforce-exact-position-1}}
+
+The problem with the new simple method (Algorithm 1) is that the precise
+values for \(V_m\) cannot be specified because the orthonormalisation
+will change them.
+
+\hypertarget{adjustment-method-1}{%
+\subsection{Adjustment method 1}\label{adjustment-method-1}}
+
+A small modification to algorithm 1 will maintain the components of
+\(V_m\) precisely (Figure \ref{fig:othermethod}). It is as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Provide \(A\), and \(m\).
+\item
+  Change values in row \(m\), giving \(A^*\).
+\item
+  Store row \(m\) separately, and zero the values of row \(m\) in
+  \(A^*\), giving \(A^{*0}\).
+\item
+  Orthonormalise \(A^{*0}\), using Gram-Schmidt.
+\item
+  Replace row \(m\) with the original values, giving \(A^{**}\).
+\item
+  For \(d=2\), adjust the values of \({\boldmath a}^{**}_{.2}\) using
+\end{enumerate}
+
+\[a^{**}_{j2}+\frac{a_{m1}a_{m2}}{p-1}, j=1, ..., p, j\neq m.\]
+
+which ensures that
+
+\[\sum_{j=1, j\neq m}^p a^{**}_{j1}a^{**}_{j2} + a_{m1}a_{m2} = 0.\]
+
+If \(d>2\) the process would be sequentially repeated in the same manner
+that Gram-Schmidt is applied sequentially to orthormalise the columns of
+a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
+vector would simply need to be normalized after each adjustment.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{arxiv_files/figure-latex/othermethod-1} \caption{Manual controls with algorithm 2. The precise location of the axis is maintained.}\label{fig:othermethod}
+\end{figure}
+
+\hypertarget{adjustment-method-2}{%
+\subsection{Adjustment method 2}\label{adjustment-method-2}}
+
+For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
+of \(O\) (an orthonormal basis for the p-dimensional space), formed by
+its first two columns as illustrated in
+Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
+for the p-dimensional space is given by
+\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
+projection matrix is expressed as
+\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
+Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
+a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
+constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
+to guarantee orthonormality of the new projection matrix. One possible
+algorithm to achieve this is
+
+\begin{enumerate}
+\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
+\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
+\begin{enumerate}
+\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
+\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
+\end{enumerate}
+\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
+\end{enumerate}
+
+The random completion of \(O\) outside the projection matrix provides an
+exploration of dimensions orthogonal to the projection plane. However,
+it renders projections in a way that is not continuous and may be
+distracting. This can be alleviated by replacing the random completion
+with a rule restricting the size of the jumps, for example in step (b)
+the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
+orthogonalisation.
+
+\begin{align}
+O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
+\CodeBefore
+       \columncolor{red!15}{1,2}
+              \rowcolor{blue!15}{4}
+     \Body
+       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
+e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
+e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
+ e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
+\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
+e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
+\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
+\label{matrices}
+\end{align}
+
+We could also take a similar approach without completing the basis,
+through rotation within the \(d\)-dimensional hyperplane defined by the
+new projection. In that case we would again start from \(A^*\) as above,
+and then rotate the basis such that the first direction is aligned with
+the selected variable \(m\). Applying Gram-Schmid within that basis
+definition will not alter the direction for the first vector, thus the
+direction of \(m\) remains fixed to what was selected with the cursor.
+By adjusting the normalization procedure we can also ensure that the
+lenght remains as selcted manually by the user.
+
+\hypertarget{slice-center-guide}{%
+\section{Slice center guide}\label{slice-center-guide}}
+
+The slice center, \(c\), from Equation 1 in the main paper, is a point
+in the \(p\)-dimensional space. By default, it is placed in the center
+of the data (Figure \ref{fig:anchornav} A). Moving this along a variable
+axis moves the slice outwards from the center of the data (Figure
+\ref{fig:anchornav} B, C). The visual display of the center position is
+a form of star plot as suggested useful in \citet{condviz2}, where the
+position of the point is marked by a polygon on radial axes indicating
+its position relative to the variable minimum to maximum values.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{arxiv_files/figure-latex/anchornav-1} \caption{Visual guide for the slice center.
+Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+The position of the center corresponds to the dark polygon.
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis.}\label{fig:anchornav}
+\end{figure}
+
+\hypertarget{software-details}{%
+\section{Software details}\label{software-details}}
+
+Here we describe the functions implemented in the Mathematica package
+\texttt{mmtour.wl}. The main function and its arguments are given below:
+
+\texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix, potentially with grouping that has a
+  label (string) and a flag (numerical) in the last two columns.
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{height}: The initial height of the slice.
+\item
+  \texttt{heightRange}: The range of slice heights to be explored.
+\item
+  \texttt{legendQ}: A flag signaling whether the data matrix includes a
+  column with group labels (1) or not (0).
+\item
+  \texttt{flagQ}: A flag signaling whether the data matrix includes a
+  numerical flag for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+The first four arguments are required. As possible starting values we
+suggest using a random input projection, and selecting slice height
+parameters using the estimate given in Section 2.4.2 of the paper.
+
+The function renders a manual tour with controls that allow the user to
+navigate through different projections. For a given projection, a slider
+permits the user to change the slice width within its range and a
+separate box switches the view to a projection. The user can also change
+the center point of slices; the size of the points rendered and the
+scale of the plot through sliders. An additional box displays the
+coordinates of the projection matrix explicitly.
+
+This function can be used for ungrouped data by setting legendQ and
+flagQ to 0. To remove a group from the visual display, the user can
+choose specific colors for the entry colorFunc. For example,
+\{Red,White,Green\} would display groups 1 (red) and 3 (green) while
+making 2 invisible.
+
+The package includes the additional functions not used in the example
+application: \texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot},
+and \texttt{VisualiseSliceDyanmic}.
+
+\texttt{ProjectionPlot} is very similar to \texttt{SliceDynamic}, except
+it only displays projections. Using it instead of \texttt{SliceDynamic}
+simplifies the display and is more efficient. The interface is slightly
+different, the data file should not contain labels for groups, if these
+are desired they are passed through the optional argument legendNames as
+\{``group 1'', ``group 2'',\ldots\}, for example. The function and its
+arguments are given below:
+
+\texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical, last column).
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{legendNames}: an optional list of labels for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+\texttt{ProjectedLocatorPlot} displays the interactive controls slightly
+differently. Several locator panes are displayed above the display of
+the projected data and each one corresponds to a row in the projection
+matrix. The updating behavior of the projection is the same as in
+\texttt{ProjectionPlot}. The function and its arguments are given below:
+
+\texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix with one or more groups labeled by a flag
+  (numerical, last column).
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{legendNames}: an optional list of labels for the groups.
+\item
+  \texttt{colorFunc}: specifies the color mapping for the different
+  groups.
+\end{itemize}
+
+Finally \texttt{VisualiseSliceDynamic} allows to discern which points
+exist inside the slice and outside it. This function differs from
+\texttt{SliceDynamic} in that it looks at only one data set but displays
+both the points inside and outside the slice. The display panel allows
+the user to specify the size of both sets of points (inside and outside)
+along with the slice height, center point and zoom level. The function
+and its arguments are given below:
+
+\texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{data}: A data matrix containing only one group. In this case
+  the data matrix should not contain column labels.
+\item
+  \texttt{projmat}: The initial projection matrix. It is possible to
+  start with a random projection with the input ``random'' (including
+  the quotes) in this entry field.
+\item
+  \texttt{height}: The initial height of the slice.
+\item
+  \texttt{heightRange}: The range of slice heights to be explored.
+\end{itemize}
+
+The points inside the slice are shown in black whereas those outside the
+slice are shown in light blue, with labels indicating this.
+
 \bibliographystyle{tfcad}

### 2023-01-06 review/response.Rmd
reviews and response document started

---
title: "Response to review: New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions"
author: "Ursula Laa, Alex Aumann, Dianne Cook, German Valencia"
date: "2023-01-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Editor**

A referee, an Associate Editor (AE), and I have reviewed your paper. There's enthusiasm for your paper but a number of issues that came up in the reviews require additional attention.  As such, I'd like to offer the opportunity to submit a revised version that addresses the issues brought up by the AE and referee.

1- Your revision should not increase the length of the manuscript.

2- I will download and run your code as part of the next round of reviews.  I suggest, but don't require, the tidyverse style for your code and, in any case, ask that you submit code that is documented and as user-friendly as possible.  Before resubmitting please run your code in a "clean" environment as issues such as required packages that aren't loaded or calls to files that don't exist are common and can delay publication (I'm not saying these exist in your case, just asking you to be aware!).

3- Your paper will be printed in black and white.  The online version will be in color but to accommodate printing without color please be sure you don't refer to color in either the figure captions or text and that all figures are still interpretable in black and white.

4- Please be sure all figures have font size of labels and any text in the figure of at least 10pt.

Articles published in JCGS are ordinarily accompanied by online publication of the data and computer code used to illustrate the proposed methods. The goal is to enable readers to replicate much of the analysis and simulations described in a paper and to apply proposed methods in their own data without having to write code from scratch.  When it is not practical to include a whole dataset in the manuscript, the manuscript should state how the complete dataset can be obtained. Please let me know if you would not be able to publish your data for reasons of security or confidentiality. Please provide the computer code and/or data that you hope to publish along with your revision. I have attached the instructions for including online materials here.

**Associate Editor**

Summary
This paper describes a method for interacting with multidimensional data using slice tours in Mathematica (as well as a reduced-feature implementation in R).

Major Issues
- Reproducibility:
Code should be reproducible and not depend on user paths.
Data files should be uploaded along with code files (they don’t seem to be part of supplemental material?)
Not having the ability to actually run the code (because the data wasn’t included) makes it somewhat diﬀicult to assess the software side of this paper.
- Grammar: There are a lot of minor grammatical issues with this paper. I suggest utilizing a tool such as Grammarly to spot and fix them. The ones that I noticed as I read through the first section are detailed in minor issues below, but I am trying to focus more on the content than the style, so I will try to not nit-pick through the rest of the paper and trust that the authors can use the online tools.
- Cite your software packages appropriately - they're tools you depend on to do your work and should be cited in the same way papers are cited.
– Page 3, line 28: Capitalize Mathematica and cite it appropriately. – Page 3, line 32: Cite R
– Page 7, line 28: Cite Shiny
- Notation
– Define your notation better: $V_m$ is used without being defined, and it would proba- bly be helpful to have a clearer mathematical definition of the relationship between your basis $A$ and $V_m$.

- What developments are specific to Mathematica and are not yet implemented in R? You allude several times to R’s deficiencies in this space but don’t ever explicitly spell out what is missing, beyond a nod to mouse tracking capabilities in Shiny (which would likely depend on the user’s screen resolution, but should be fairly precise once you condition on that - why would Mathematica have better resolution for mouse controls?)
Below is a set of comments relating to this issue, written as I read through the paper (so hopefully the chronology is preserved and you can see why this isn’t as direct as it
should be).

Page 7, line 46: "Mathematica provides much utility and versatility, such as inbuilt data visualization, data manipulation and analysis, dynamic functionality, and sym- bolic and numeric computation. Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there."
The same could be said of R, so I’m not sure why this is relevant. It seems mostly like Mathematica propaganda.
– Page 7-8: "Importantly for our work, it is relatively simple to create dynamic objects via the inbuilt commands Manipulate or DynamicModule. Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated. This is the essential ingredient for this implementation of the manual tour."
Again, the same dynamic control creation exists in Shiny, along with the basic reactivity to user inputs. It's hard to see why Mathematica is different here.
– Page 8, line 11: "The most relevant inbuilt Mathematica function for our purpose is the LocatorPane. This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour."
Finally, after all of this time, this is the one input that might be a bit harder to implement in R/Shiny, though I think via WebGL it should be workable. This is where the focus should be, rather than the previous two paragraphs, which are a distraction.

Minor issues and suggestions

- Pg 2, line 8: "the more useful methods for data analysis need to work for arbitrary dimension."
Perhaps instead: methods which work for arbitrary dimensions are more useful/have more general utility? Either way, dimension should be plural.
- Pg 2, line 17: "could theoretically show the viewer the data from all sides."
It might be helpful to have a small figure here to demonstrate what this might mean, for those who aren't familiar with tours and/or who can't visualize n-dimensional space projections into 3D out of thin air. Alternately, link to e.g. a YouTube video showing what you mean by "from all sides"
- Pg 2, line 23: "A video illustrating the capabilities is available through video library of ASA Statistical Graphics Section (2022)."
Add "the" before video

**Reviewer 1**

In this paper a new manual controls for projections and slice tours are presented. The controls are designed to explore high-dimensional data based on linear projections and slices.  In the paper a Mathematica package is described to explain the application.

Some comments:

In the Abstract you have mentioned “Some limited implementations has also been made available in the R package tourr” but in Sub-Section 3.2 you can mention in more detail the limitations in tourr implementation

Section 1 line 29 “The software section describes mathematica package”  mathematica should be capitalized as you did  through the paper in most cases.  Also since is the first time you have mentioned Mathematica you should include the reference here (Wolfram reference appears in page 4 line 26 for first time)

Section 2: “manual tour allows the user to alter the coefficients of one (or more) variables” but in the explanation it is not clear what is  V_m  and how to alter the coefficients for more than one variable.
In page 4 line 2.2 “The components corresponding to V_m are directly controlled by cursor movement which updates row m of A ''. Maybe a video will be illustrative in this sub-section.

Section 3: will be useful to include a schematic diagram illustrating the interactivity 
In the manual tour. implementation .
This part is difficult to follow, it is more clear after reading the application Section because you have included some videos but I think you need to include  something here to help the reader.

In the introduction “The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available ”
I think you need to be more specific about the R limitations.
Iin Section 3. Page 7, line 20 “One might implement such an interactive interface via R Shiny, but in particular the tracking of the mouse position in small increments might pose a challenge” This  is the only possible R limitation you have mentioned.

Section 4: Page 10 line 36: Figure 3summarizes (space)

Section 5: page 15 line 26 “If new technology for interactive graphics became available”  be more specific about the needed technology in R to implement the manual controls as you did with Mathematica

### 2023-01-06 review/response.Rmd
tidying of format of response

@@
 As such, I'd like to offer the opportunity to submit a revised version that addresses the issues brought up by the AE and referee.
+
+\textcolor{blue}{Thank you for the careful review of our paper. We have done our best to make changes to address each comment, as detailed in blue below.}
 
@@
 Summary
+
 This paper describes a method for interacting with multidimensional data using slice tours in Mathematica (as well as a reduced-feature implementation in R).
@@
 Major Issues
+
 - Reproducibility:
-Code should be reproducible and not depend on user paths.
-Data files should be uploaded along with code files (they don’t seem to be part of supplemental material?)
-Not having the ability to actually run the code (because the data wasn’t included) makes it somewhat diﬀicult to assess the software side of this paper.
+    - Code should be reproducible and not depend on user paths.
+    - Data files should be uploaded along with code files (they don't seem to be part of supplemental material?)
+    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat diﬀicult to assess the software side of this paper.
 - Grammar: There are a lot of minor grammatical issues with this paper.
@@
 - Cite your software packages appropriately - they're tools you depend on to do your work and should be cited in the same way papers are cited.
-– Page 3, line 28: Capitalize Mathematica and cite it appropriately. – Page 3, line 32: Cite R
+
+    – Page 3, line 28: Capitalize Mathematica and cite it appropriately.
+
+    – Page 3, line 32: Cite R
+
     – Page 7, line 28: Cite Shiny
+
 - Notation
-– Define your notation better: $V_m$ is used without being defined, and it would proba- bly be helpful to have a clearer mathematical definition of the relationship between your basis $A$ and $V_m$.
+
+    – Define your notation better: $V_m$ is used without being defined, and it would probably be helpful to have a clearer mathematical definition of the relationship between your basis $A$ and $V_m$.
 
 - What developments are specific to Mathematica and are not yet implemented in R?
-You allude several times to R’s deficiencies in this space but don’t ever explicitly spell out what is missing, beyond a nod to mouse tracking capabilities in Shiny (which would likely depend on the user’s screen resolution, but should be fairly precise once you condition on that - why would Mathematica have better resolution for mouse controls?)
+You allude several times to R's deficiencies in this space but don't ever explicitly spell out what is missing, beyond a nod to mouse tracking capabilities in Shiny (which would likely depend on the user's screen resolution, but should be fairly precise once you condition on that - why would Mathematica have better resolution for mouse controls?)
 Below is a set of comments relating to this issue, written as I read through the paper (so hopefully the chronology is preserved and you can see why this isn’t as direct as it
@@
 
-Page 7, line 46: "Mathematica provides much utility and versatility, such as inbuilt data visualization, data manipulation and analysis, dynamic functionality, and sym- bolic and numeric computation.
-Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there."
-The same could be said of R, so I’m not sure why this is relevant. It seems mostly like Mathematica propaganda.
+    - Page 7, line 46: "Mathematica provides much utility and versatility, such as inbuilt data visualization, data manipulation and analysis, dynamic functionality, and sym- bolic and numeric computation.
+Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there." The same could be said of R, so I'm not sure why this is relevant.
+It seems mostly like Mathematica propaganda.
+
         – Page 7-8: "Importantly for our work, it is relatively simple to create dynamic objects via the inbuilt commands Manipulate or DynamicModule.
 Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.
-This is the essential ingredient for this implementation of the manual tour."
-Again, the same dynamic control creation exists in Shiny, along with the basic reactivity to user inputs. It's hard to see why Mathematica is different here.
+This is the essential ingredient for this implementation of the manual tour." Again, the same dynamic control creation exists in Shiny, along with the basic reactivity to user inputs.
+It's hard to see why Mathematica is different here.
+
         – Page 8, line 11: "The most relevant inbuilt Mathematica function for our purpose is the LocatorPane.
 This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour."
-Finally, after all of this time, this is the one input that might be a bit harder to implement in R/Shiny, though I think via WebGL it should be workable.
+
+    - Finally, after all of this time, this is the one input that might be a bit harder to implement in R/Shiny, though I think via WebGL it should be workable.
 This is where the focus should be, rather than the previous two paragraphs, which are a distraction.
@@
 
-In the Abstract you have mentioned “Some limited implementations has also been made available in the R package tourr” but in Sub-Section 3.2 you can mention in more detail the limitations in tourr implementation
+In the Abstract you have mentioned "Some limited implementations has also been made available in the R package tourr" but in Sub-Section 3.2 you can mention in more detail the limitations in tourr implementation
 
-Section 1 line 29 “The software section describes mathematica package”  mathematica should be capitalized as you did  through the paper in most cases.
+Section 1 line 29 "The software section describes mathematica package"  mathematica should be capitalized as you did  through the paper in most cases.
 Also since is the first time you have mentioned Mathematica you should include the reference here (Wolfram reference appears in page 4 line 26 for first time)
 
-Section 2: “manual tour allows the user to alter the coefficients of one (or more) variables” but in the explanation it is not clear what is  V_m  and how to alter the coefficients for more than one variable.
-In page 4 line 2.2 “The components corresponding to V_m are directly controlled by cursor movement which updates row m of A ''. Maybe a video will be illustrative in this sub-section.
+Section 2: "manual tour allows the user to alter the coefficients of one (or more) variables" but in the explanation it is not clear what is  $V_m$  and how to alter the coefficients for more than one variable.
+
+In page 4 line 2.2 "The components corresponding to $V_m$ are directly controlled by cursor movement which updates row $m$ of $A$". Maybe a video will be illustrative in this sub-section.
 
 Section 3: will be useful to include a schematic diagram illustrating the interactivity
-In the manual tour. implementation .
+
+In the manual tour implementation.
 This part is difficult to follow, it is more clear after reading the application Section because you have included some videos but I think you need to include something here to help the reader.
 
-In the introduction “The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available ”
+In the introduction "The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available"
 I think you need to be more specific about the R limitations.
 Iin Section 3.
-Page 7, line 20 “One might implement such an interactive interface via R Shiny, but in particular the tracking of the mouse position in small increments might pose a challenge” This  is the only possible R limitation you have mentioned.
+Page 7, line 20 "One might implement such an interactive interface via R Shiny, but in particular the tracking of the mouse position in small increments might pose a challenge" This  is the only possible R limitation you have mentioned.
 
-Section 4: Page 10 line 36: Figure 3summarizes (space)
+Section 4: Page 10 line 36: Figure 3 summarizes (space)
 
-Section 5: page 15 line 26 “If new technology for interactive graphics became available”  be more specific about the needed technology in R to implement the manual controls as you did with Mathematica
+Section 5: page 15 line 26 "If new technology for interactive graphics became available"  be more specific about the needed technology in R to implement the manual controls as you did with Mathematica

### 2023-01-11 review/response.Rmd
start fixing errors

@@
     – Page 7, line 28: Cite Shiny
-
+\textcolor{blue}{done}    
 - Notation
@@
     – Define your notation better: $V_m$ is used without being defined, and it would probably be helpful to have a clearer mathematical definition of the relationship between your basis $A$ and $V_m$.
+
+\textcolor{blue}{see text in blue page 3, is this adequate?}
 
@@
 
+\textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny}
+
 Minor issues and suggestions
@@
 Perhaps instead: methods which work for arbitrary dimensions are more useful/have more general utility? Either way, dimension should be plural.
+
+\textcolor{blue}{done}
+
 - Pg 2, line 17: "could theoretically show the viewer the data from all sides."
@@
 a YouTube video showing what you mean by "from all sides"
+
 - Pg 2, line 23: "A video illustrating the capabilities is available through video library of ASA Statistical Graphics Section (2022)."
 Add "the" before video
+\textcolor{blue}{done}
 
@@
 Also since is the first time you have mentioned Mathematica you should include the reference here (Wolfram reference appears in page 4 line 26 for first time)
+\textcolor{blue}{done}
 
 Section 2: "manual tour allows the user to alter the coefficients of one (or more) variables" but in the explanation it is not clear what is  $V_m$  and how to alter the coefficients for more than one variable.
+
+\textcolor{blue}{definition of $V_m$ added but have not addressed the second part of the comment}
 

### 2023-01-11 paper/paper.Rmd
start fixing errors

@@
 <!-- Background -->
-From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, the more useful methods for data analysis need to work for arbitrary dimension.
+From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, \textcolor{blue}{methods which work for arbitrary dimensions have more general utility}.
 A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
@@
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-(A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
+(A video illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
 After Asimov there have been many tour developments, which are summarized in @lee2021.
@@
 The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers.
-The software section describes a mathematica package that is used for the application, and describes the interactive environment that would be desirable within R as new technology becomes available.
+The software section describes a Mathematica @Mathematica package that is used for the application, and describes the interactive environment that would be desirable within R @rref as new technology becomes available.
 The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the application.
@@
 A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d$ dimensional projection.
-The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed.
-A method to update the values of the component ($m^{th}$ row of $A_{p\times d}$) of the controlled variable $V_m$ is then needed.
+The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be \textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.} A method to update the values of the component of the controlled variable $V_m$ is then needed.
 
@@
 
-One might implement such an interactive interface via R Shiny, but in particular the tracking of the mouse position in small increments might pose a challenge.
+One might implement such an interactive interface via R Shiny @shinym, but in particular the tracking of the mouse position in small increments might pose a challenge.
 Here we found the dynamic graphics interface from Mathematica to be especially useful.
@@
 
-Mathematica provides much utility and versatility, such as inbuilt data visualization, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation.
-Most of this inbuilt functionality is user-friendly and described in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there.
+\textcolor{blue}{The Mathematica implementation relies on the inbuilt Mathematica function  `LocatorPane`.
+This function creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
+We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
+Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.
 
-Importantly for our work, it is relatively simple  to create dynamic objects via the inbuilt commands `Manipulate` or `DynamicModule`.
-Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.
-This is the essential ingredient for this implementation of the manual tour.
-
-The most relevant inbuilt Mathematica function for our purpose is the `LocatorPane`.
-This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
+Di will explain what part and why is difficult in R, also check grammar}
 

### 2023-01-11 paper/paper.tex
compiled version of G changes

@@
 From a statistical perspective 3D is a rare data dimension, so unlike in
-most 3D rotation computer graphics applications, the more useful methods
-for data analysis need to work for arbitrary dimension. A good approach
-is to show projections from an arbitrary dimensional space to create
-dynamic data visualizations called \emph{tours}. Tours involve views of
-high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
-In his original paper on the grand tour, \citet{As85} provided several
-algorithms for tour paths that could theoretically show the viewer the
-data \emph{from all sides}. Prior to Asimov's work, there were numerous
-preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
-user-controlled rotations on coordinate axes, allowing one to manually
-tour through low-dimensional projections. (A video illustrating the
-capabilities is available through video library of \citet{ASA22}.)
-Steering through all possible projections is impossible, unlike Asimov's
-tours which allows one to quickly see many, many different projections.
-After Asimov there have been many tour developments, which are
-summarized in \citet{lee2021}.
+most 3D rotation computer graphics applications,
+\textcolor{blue}{methods which work for arbitrary dimensions have more general utility}.
+A good approach is to show projections from an arbitrary dimensional
+space to create dynamic data visualizations called \emph{tours}. Tours
+involve views of high-dimensional (\(p\)) data with low-dimensional
+(\(d\)) projections. In his original paper on the grand tour,
+\citet{As85} provided several algorithms for tour paths that could
+theoretically show the viewer the data \emph{from all sides}. Prior to
+Asimov's work, there were numerous preparatory developments including
+\citet{tukey}'s PRIM-9. PRIM-9 had user-controlled rotations on
+coordinate axes, allowing one to manually tour through low-dimensional
+projections. (A video illustrating the capabilities is available through
+\textcolor{blue}{the} video library of \citet{ASA22}.) Steering through
+all possible projections is impossible, unlike Asimov's tours which
+allows one to quickly see many, many different projections. After Asimov
+there have been many tour developments, which are summarized in
+\citet{lee2021}.
 
@@
 to compare and contrast boundaries constructed by different classifiers.
-The software section describes a mathematica package that is used for
-the application, and describes the interactive environment that would be
-desirable within R as new technology becomes available. The paper is
-accompanied by an appendix with more details and adjustments to the
-manual controls, and three Mathematica notebooks that can be used to
-reproduce the application.
+The software section describes a Mathematica \citet{Mathematica} package
+that is used for the application, and describes the interactive
+environment that would be desirable within R \citet{rref} as new
+technology becomes available. The paper is accompanied by an appendix
+with more details and adjustments to the manual controls, and three
+Mathematica notebooks that can be used to reproduce the application.
 
@@
 projection of the data, and a variable id (\(m \in \{1, ..., p\}\))
-specifying which coefficient will be changed. A method to update the
-values of the component (\(m^{th}\) row of \(A_{p\times d}\)) of the
-controlled variable \(V_m\) is then needed.
+specifying which coefficient will be
+\textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.}
+A method to update the values of the component of the controlled
+variable \(V_m\) is then needed.
 
@@
 
-One might implement such an interactive interface via R Shiny, but in
-particular the tracking of the mouse position in small increments might
-pose a challenge. Here we found the dynamic graphics interface from
-Mathematica to be especially useful. In the following we first describe
-the relevant Mathematica functionalities and how they were used in the
-implementation, followed by a brief sketch of how the new approaches are
-implemented in the \texttt{tourr} package. There is hope that better
-interactivity will be available for R soon, which might allow for
-implementation of the techniques developed in Mathematica.
+One might implement such an interactive interface via R Shiny
+\citet{shinym}, but in particular the tracking of the mouse position in
+small increments might pose a challenge. Here we found the dynamic
+graphics interface from Mathematica to be especially useful. In the
+following we first describe the relevant Mathematica functionalities and
+how they were used in the implementation, followed by a brief sketch of
+how the new approaches are implemented in the \texttt{tourr} package.
+There is hope that better interactivity will be available for R soon,
+which might allow for implementation of the techniques developed in
+Mathematica.
 
@@
 
-Mathematica provides much utility and versatility, such as inbuilt data
-visualization, data manipulation and analysis, dynamic functionality,
-and symbolic and numeric computation. Most of this inbuilt functionality
-is user-friendly and described in the Mathematica documentation;
-typically, numerous examples are provided within the documentation, and
-some possible issues are outlined there.
+\textcolor{blue}{The Mathematica} implementation relies on the inbuilt
+Mathematica function \texttt{LocatorPane}. This function creates a
+region on the screen where the position of the mouse is captured and
+then converted to input that updates the graphics functions, enabling
+the manual navigation of the tour. We then use the graphics function
+\texttt{ListPlot} embedded in dynamic objects via \texttt{DynamicModule}
+to produce the display. Control objects, such as sliders, locator panes,
+and input fields, are used on dynamic variables, and when there is a
+change in the dynamic variable the dynamic objects which contain that
+variable will be updated.
 
-Importantly for our work, it is relatively simple to create dynamic
-objects via the inbuilt commands \texttt{Manipulate} or
-\texttt{DynamicModule}. Control objects, such as sliders, locator panes,
-and input fields, can then be used on dynamic variables, and when there
-is a change in the dynamic variable the dynamic objects which contain
-that variable will be updated. This is the essential ingredient for this
-implementation of the manual tour.
-
-The most relevant inbuilt Mathematica function for our purpose is the
-\texttt{LocatorPane}. This creates a region on the screen where the
-position of the mouse is captured and then converted to input that
-updates the graphics functions, enabling the manual navigation of the
-tour.
+Di will explain what part and why is difficult in R, also check
+grammar\}
 

### 2023-01-11 paper/paper.Rmd
fixed textcolor error

@@
 We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
-Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.
+Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.}
 
-Di will explain what part and why is difficult in R, also check grammar}
+\textcolor{blue}{Di will explain what part and why is difficult in R, also check grammar}
 

### 2023-01-11 paper/paper.tex
fixed textcolor error

@@
 
-\textcolor{blue}{The Mathematica} implementation relies on the inbuilt
-Mathematica function \texttt{LocatorPane}. This function creates a
-region on the screen where the position of the mouse is captured and
-then converted to input that updates the graphics functions, enabling
-the manual navigation of the tour. We then use the graphics function
-\texttt{ListPlot} embedded in dynamic objects via \texttt{DynamicModule}
-to produce the display. Control objects, such as sliders, locator panes,
-and input fields, are used on dynamic variables, and when there is a
-change in the dynamic variable the dynamic objects which contain that
-variable will be updated.
+\textcolor{blue}{The Mathematica implementation relies on the inbuilt Mathematica function  `LocatorPane`.
+This function creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
+We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
+Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.}
 
-Di will explain what part and why is difficult in R, also check
-grammar\}
+\textcolor{blue}{Di will explain what part and why is difficult in R, also check grammar}
 

### 2023-01-16 paper/paper.Rmd
experiments with colour

@@
 #ggscatmat(penguins_data, columns = 2:5, col="species") +
+# library(paletteer)
+clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
+clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
+clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
+clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
 penguins_diag <- tibble(x=1, y=c(3,2,1),
@@
           diag = list(continuous = "densityDiag")) +
-  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-  theme(aspect.ratio=1, 
-        #legend.position = "bottom", 
-        axis.text = element_blank(),
-        axis.ticks = element_blank())
+#  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+#  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
+  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
 pscat[1,1] <- ggplot(penguins_diag, 
@@
   geom_text() +
-  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
   ylim(c(0,4))
@@
   geom_text() +
-  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
+  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
   ylim(c(0,4))
-pscat
+pscat + theme_bw() +
+  theme(aspect.ratio=1, 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
 ```

### 2023-01-18 paper/paper.Rmd
fig 3 revised

--- 2023-01-11 paper/paper.Rmd
@@
 #ggscatmat(penguins_data, columns = 2:5, col="species") +
+# library(paletteer) 
+# rcartocolor::ag_Sunset
+# rcartocolor::ag_GrnYl
+# RColorBrewer::Spectral
+# clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
+# clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
+# clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
+clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
 penguins_diag <- tibble(x=1, y=c(3,2,1),
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\centerline{\includegraphics[width=0.95\textwidth]{figures/proj1.pdf}}
+%\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+%\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+%\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
 \caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).

### 2023-01-18 paper/paper.Rmd
fig 4 revised

@@
 %\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
-The three groups are nicely separated in this projection of the data (right).}
+\caption{A projection identified using the manual tour, because it shows differences in classification boundaries induced by different models.
+A RF model produces blocky, rectangular, boundaries, while the LDA model produces oblique linear boundaries.
+For comparison, in this projection the data shows the three groups are almost perfectly separable.}
 \label{proj1}
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
-\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
-\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
-\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right).
-With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice.
-The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
+\centerline{\includegraphics[width=0.95\textwidth]{figures/proj2.pdf}}
+%\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
+%\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
+%\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
+%\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
+\caption{Comparing slices from two different projections $A_1$ (top row) and $A_2$ (bottom row), to compare and contrast classification boundaries induced by two models RF and LDA.
+A thin slice through the center of the data is shown.
+The plot of the data shows that the Gentoo group (light grey/green) is less separated in $A_1$, and the rotation to $A_2$ produces a bigger gap.
+The boundaries between groups are clear in $A_2$ also.
+The LDA boundary is more oblique between Gentoo and Chinstrap than the RF, which is too close to the Chinstrap group.
+On the other hand between Adelie and Gentoo, the RF is better situated about half-way between the two groups.
+Interestingly, the RF boundaries, which can be highly nonlinear, albeit blocky, are relatively linear for this problem.
+}
 \label{slice1}
@@
 
+<!--
 \begin{figure*}[ht]
@@
 \end{figure*}
+-->
 
-Finally, to determine which model describes the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}).
+Finally, to determine which model describes the boundaries better, compare them with the $A_2$ projection of the data (bottom row of Figure \ref{slice1}).
 Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
@@
 
-This example can be reproduced with the code in penguins_slicing_through_center.nb and a run through is shown in the first part of the video at https://vimeo.com/747590472.
+This example can be reproduced with the code in `penguins_slicing_through_center.nb` and a run through is shown in the first part of the video at https://vimeo.com/747590472.
 

### 2023-01-18 paper/paper.tex
fig 3 revised
fig 4 revised

@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left).
-We can clearly see a block structure, while the LDA model (middle) produces linear boundaries.
-The three groups are nicely separated in this projection of the data (right).}
+\centerline{\includegraphics[width=0.95\textwidth]{figures/proj1.pdf}}
+%\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
+%\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
+%\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
+\caption{A projection identified using the manual tour, because it shows differences in classification boundaries induced by different models.
+A RF model produces blocky, rectangular, boundaries, while the LDA model produces oblique linear boundaries.
+For comparison, in this projection the data shows the three groups are almost perfectly separable.}
 \label{proj1}
@@
 
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
-\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
-\label{proj2}
-\end{figure*}
-
 Finally, to determine which model describes the boundaries better,
-compare them with the \(A_2\) projection of the data (Figure
-\ref{proj2}). Gentoo is more separated from the other two in this
-projection, and one can imagine that the trees in the forest has
+compare them with the \(A_2\) projection of the data (bottom row of
+Figure \ref{slice1}). Gentoo is more separated from the other two in
+this projection, and one can imagine that the trees in the forest has
 greedily grasped any one of many places to make a split to separate the
@@
 This example can be reproduced with the code in
-penguins\_slicing\_through\_center.nb and a run through is shown in the
-first part of the video at \url{https://vimeo.com/747590472}.
+\texttt{penguins\_slicing\_through\_center.nb} and a run through is
+shown in the first part of the video at
+\url{https://vimeo.com/747590472}.
 

### 2023-01-18 paper/paper.Rmd
removed references to color

@@
 The second slice (based on $A_2$) mostly resolves this overlap and reveals the primary differences between the models.
-The boundary between Adelie (blue) and Chinstrap (yellow) is very similar for both models.
-The boundary between Chinstrap and Gentoo (green) is where they differ.
+The boundary between Adelie and Chinstrap  is very similar for both models.
+The boundary between Chinstrap and Gentoo is where they differ.
 
@@
 We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
-We can see that all three slices (the two models and the data) contain almost no points from the third class (green, Gentoo), and that the decision boundary between the two models is very similar.
+We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
 
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green).
-The two models have similar boundaries except that the RF is more of a step function.}
+\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo. The two models have similar boundaries except that the RF is more of a step function.}
 \label{slice1p}
@@
 \includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green).
+\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo.
 The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
@@
 A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
-The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (green, Gentoo).
+The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
 On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.

### 2023-01-18 paper/paper.tex
removed references to color

@@
 \(A_2\)) mostly resolves this overlap and reveals the primary
-differences between the models. The boundary between Adelie (blue) and
-Chinstrap (yellow) is very similar for both models. The boundary between
-Chinstrap and Gentoo (green) is where they differ.
+differences between the models. The boundary between Adelie and
+Chinstrap is very similar for both models. The boundary between
+Chinstrap and Gentoo is where they differ.
 
@@
 slices (the two models and the data) contain almost no points from the
-third class (green, Gentoo), and that the decision boundary between the
-two models is very similar.
+third class (Gentoo), and that the decision boundary between the two
+models is very similar.
 
@@
 \ref{slice1m}. The RF model (left) predicts all three species within
-this slice, with an interesting boundary for the third class (green,
-Gentoo). On the other hand the LDA model (middle) predominantly predicts
-the third class within the slice, this appears to be enforced through
-the linear structure of the model. Looking finally at the thick slice
+this slice, with an interesting boundary for the third class (Gentoo).
+On the other hand the LDA model (middle) predominantly predicts the
+third class within the slice, this appears to be enforced through the
+linear structure of the model. Looking finally at the thick slice
 through the data we see that there are primarily observations from this

### 2023-01-18 review/response.Rmd
fig 3 revised
fig 4 revised
updated response

@@
 ```
+
+ XXX If use new slice center control for fig 7, need to make a note of it here, updated the mathematica notebook and check text and captions correct.
 
@@
 
+\textcolor{blue}{Colour in figures has been changed to work in black and white as well as colour. Captions have been adjusted accordingly. TODO}
+
 4- Please be sure all figures have font size of labels and any text in the figure of at least 10pt.
+
+\textcolor{blue}{Done}
 
@@
     – Page 7, line 28: Cite Shiny
-\textcolor{blue}{done}    
+
+\textcolor{blue}{Done - the citations were there, but not placed at the first mention.}
+
 - Notation
@@
 
+\textcolor{blue}{Modified}
+
       – Page 7-8: "Importantly for our work, it is relatively simple to create dynamic objects via the inbuilt commands Manipulate or DynamicModule.
@@
 It's hard to see why Mathematica is different here.
+
+\textcolor{blue}{This is better explained}
 

### 2023-01-18 paper/paper.Rmd
updated response

@@
 The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers.
-The software section describes a Mathematica @Mathematica package that is used for the application, and describes the interactive environment that would be desirable within R @rref as new technology becomes available.
-The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the application.
+The software section describes a Mathematica [@Mathematica] package that is used for the application, and describes the interactive environment that would be desirable within R [@rref] as new technology becomes available.
+The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the applications.
 
@@
 The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example).
-This display should be able to track the mouse position and adjust the projection based on the user selection.
-The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it during the click event.
-In practice the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.
+This display should be able to track the mouse position and adjust the projection based on the \textcolor{blue}{user's cursor position}.
+The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it.
+In practice, the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.
 
-The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized before feeding back into the axis display.
+The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized and \textcolor{blue}{re-displayed}.
 A second, linked display shows the projected data in sync with the updates from the axis display.
@@
 In addition, we also want to be able to look at slices of the data and select slicing parameters interactively.
-Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness $h$ (dubbed `height` in the `Mathematica` package) and a numeric input to specify the center point $c$ explicitly.
-The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).
+Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness, $h$ (named `height` in the Mathematica interface), and a numeric input to specify the center point $c$ explicitly.
+The display will update based on these input values, in particular the view will jump to a new slice when $c$ is changed.<!-- (though an interpolation might also be useful).-->
 
-One might implement such an interactive interface via R Shiny @shinym, but in particular the tracking of the mouse position in small increments might pose a challenge.
-Here we found the dynamic graphics interface from Mathematica to be especially useful.
-In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approaches are implemented in the `tourr` package.
-There is hope that better interactivity will be available for R soon, which might allow for implementation of the techniques developed in Mathematica.
+\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing.
+It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours.
+The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R.
+The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
+It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today.
+What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package.
+There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}
 
@@
 
-The R package tourr [@tourr] provides numerous types of tours.
+The R package `tourr` provides numerous types of tours.
 An additional function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.

### 2023-01-18 paper/paper.tex
updated response

@@
 to compare and contrast boundaries constructed by different classifiers.
-The software section describes a Mathematica \citet{Mathematica} package
+The software section describes a Mathematica \citep{Mathematica} package
 that is used for the application, and describes the interactive
-environment that would be desirable within R \citet{rref} as new
+environment that would be desirable within R \citep{rref} as new
 technology becomes available. The paper is accompanied by an appendix
 with more details and adjustments to the manual controls, and three
-Mathematica notebooks that can be used to reproduce the application.
+Mathematica notebooks that can be used to reproduce the applications.
 
@@
 able to track the mouse position and adjust the projection based on the
-user selection. The user can select one axis (corresponding to one
-variable) by clicking on it, and then adjust the position of that axis
-by dragging it during the click event. In practice the closest axis will
+\textcolor{blue}{user's cursor position}. The user can select one axis
+(corresponding to one variable) by clicking on it, and then adjust the
+position of that axis by dragging it. In practice, the closest axis will
 be selected, and the dragging results in tracking of the mouse position
@@
 The interactions in the axis display need to be mapped back onto the
-current projection matrix, which will then be orthonormalized before
-feeding back into the axis display. A second, linked display shows the
+current projection matrix, which will then be orthonormalized and
+\textcolor{blue}{re-displayed}. A second, linked display shows the
 projected data in sync with the updates from the axis display.
@@
 switch to change between projection and slice view, a slider that
-adjusts the slice thickness \(h\) (dubbed \texttt{height} in the
-\texttt{Mathematica} package) and a numeric input to specify the center
-point \(c\) explicitly. The display will update based on these inputs,
-in particular the view will jump to a new slice when \(c\) is changed
-(though an interpolation might also be useful).
+adjusts the slice thickness, \(h\) (named \texttt{height} in the
+Mathematica interface), and a numeric input to specify the center point
+\(c\) explicitly. The display will update based on these input values,
+in particular the view will jump to a new slice when \(c\) is changed.
 
-One might implement such an interactive interface via R Shiny
-\citet{shinym}, but in particular the tracking of the mouse position in
-small increments might pose a challenge. Here we found the dynamic
-graphics interface from Mathematica to be especially useful. In the
-following we first describe the relevant Mathematica functionalities and
-how they were used in the implementation, followed by a brief sketch of
-how the new approaches are implemented in the \texttt{tourr} package.
-There is hope that better interactivity will be available for R soon,
-which might allow for implementation of the techniques developed in
-Mathematica.
+\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing.
+It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours.
+The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R.
+The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
+It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today.
+What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package.
+There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}
 
@@
 
-The R package tourr \citep{tourr} provides numerous types of tours. An
+The R package \texttt{tourr} provides numerous types of tours. An
 additional function, \texttt{radial\_tour()} has been constructed that

### 2023-01-19 paper/paper.Rmd
why not R expanded

@@
 
-\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing.
-It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours.
-The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R.
-The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
-It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today.
+\textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing.
+However, the `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
+It lacks capability to implement the primary tool needed for a manual tour, that is the tracking of the cursor position in small increments.
+In theory, the historical `locator()` function of base R should provide the necessary capability, but it's behavior is not consistent between graphical devices.
+The cursor position control is possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary dimension of tours.
+The level of interactivity would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that currently operates across platforms.
 What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package.
-There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}
+There is hope that better interactivity will be available for R soon, perhaps in javascript [e.g.
+@langevitour, @detourr], which might allow for the  experimentation to develop new techniques.
 
@@
 Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.}
-
-\textcolor{blue}{Di will explain what part and why is difficult in R, also check grammar}
 

### 2023-01-19 paper/paper.tex
why not R expanded

@@
 
-\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing.
-It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours.
-The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R.
-The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
-It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today.
-What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package.
-There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}
+\textcolor{blue}{Within R, the natural implementation for an interactive interface would use}
+\texttt{shiny} \citep{shinym} with \texttt{plotly} \citep{plotly}
+drawing. However, the \texttt{plotly} package is designed to add on to
+\texttt{ggplot2} graphics, not to be a programmable interactive graphics
+system. It lacks capability to implement the primary tool needed for a
+manual tour, that is the tracking of the cursor position in small
+increments. In theory, the historical \texttt{locator()} function of
+base R should provide the necessary capability, but it's behavior is not
+consistent between graphical devices. The cursor position control is
+possible with \texttt{rgl} \citep{rgl} but this package is focused on
+3D, not the arbitrary dimension of tours. The level of interactivity
+would have been possible with the \texttt{cranvas} system
+\citep{cranvas} or \texttt{rJava} \citep{rJava} but neither has a
+graphics toolkit that currently operates across platforms. What we have
+chosen to do is to use the interactivity of Mathematica to experiment,
+and to implement as an animation in the \texttt{tourr} \citep{tourr}
+package. There is hope that better interactivity will be available for R
+soon, perhaps in javascript \citep[e.g.][\citet{detourr}]{langevitour},
+which might allow for the experimentation to develop new techniques.
 

### 2023-01-19 paper/paper.Rmd
fig 5 modified

@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
-\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo. The two models have similar boundaries except that the RF is more of a step function.}
+\centerline{\includegraphics[width=0.95\textwidth]{figures/slice1_p.pdf}}
+%\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
+%\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
+%\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
+\caption{Shifting the slice center allows for exploring the boundary in different subspaces.
+Here the center has been shifted out along the axis of bill depth (icon display at right represents slice center position).
+This region has no Gentoo.
+The two models have similar boundaries except that the RF is has a step shift in the middle.}
 \label{slice1p}

### 2023-01-19 paper/paper.Rmd
fig 5 and 6 modified

@@
 
-We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in Figure \ref{slice1p}.
+We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in top row of Figure \ref{slice1p}.
 We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
@@
 \caption{Shifting the slice center allows for exploring the boundary in different subspaces.
-Here the center has been shifted out along the axis of bill depth (icon display at right represents slice center position).
-This region has no Gentoo.
-The two models have similar boundaries except that the RF is has a step shift in the middle.}
+Here the center has been shifted along the axis of `bd` (icon display at right represents slice center position).
+(Top row) Center is at large values of `bd`, showing a region with no Gentoo.
+The two models have similar boundaries except that the LDA is more linear.
+(Bottom row) Center is at small values of `bd`, reveals a region where the models differ substantially.
+The RF model carves some area for Adelie, but LDA would only predict to be Gentoo or Chinstrap here.
+LDA better matches the observed data in this region.}
 \label{slice1p}
@@
 \end{figure*}
+-->
 
-A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in Figure \ref{slice1m}.
+A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in the bottom row of Figure \ref{slice1p}.
 The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
@@
 
-Finally it is interesting to compare the slice views to the projection of the models seen in Fig.
+Finally, it is interesting to compare the slice views to the projection of the models seen in Fig.
 \ref{proj1} to better understand how the boundaries change along the `bd` direction and where the differences in the projections come from.

### 2023-01-19 paper/paper.tex
fig 5 and 6 modified

@@
 shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
-values of bill depth in Figure \ref{slice1p}. We can see that all three
-slices (the two models and the data) contain almost no points from the
-third class (Gentoo), and that the decision boundary between the two
-models is very similar.
+values of bill depth in top row of Figure \ref{slice1p}. We can see that
+all three slices (the two models and the data) contain almost no points
+from the third class (Gentoo), and that the decision boundary between
+the two models is very similar.
 
@@
 
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
-\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
-\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo.
-The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
-\label{slice1m}
-\end{figure*}
+A more interesting comparison is found for \(\mathbf{c}^{-}\), thus the
+slice localized towards low values of bill depth, shown in the bottom
+row of Figure \ref{slice1p}. The RF model (left) predicts all three
+species within this slice, with an interesting boundary for the third
+class (Gentoo). On the other hand the LDA model (middle) predominantly
+predicts the third class within the slice, this appears to be enforced
+through the linear structure of the model. Looking finally at the thick
+slice through the data we see that there are primarily observations from
+this class within the slice we can conclude that the two models have
+filled in the ``empty'' space (where we do not have any training
+observations) in very different ways and according to what we might
+expect given the model structure.
 
-A more interesting comparison is found for \(\mathbf{c}^{-}\), thus the
-slice localized towards low values of bill depth, shown in Figure
-\ref{slice1m}. The RF model (left) predicts all three species within
-this slice, with an interesting boundary for the third class (Gentoo).
-On the other hand the LDA model (middle) predominantly predicts the
-third class within the slice, this appears to be enforced through the
-linear structure of the model. Looking finally at the thick slice
-through the data we see that there are primarily observations from this
-class within the slice we can conclude that the two models have filled
-in the ``empty'' space (where we do not have any training observations)
-in very different ways and according to what we might expect given the
-model structure.
-
-Finally it is interesting to compare the slice views to the projection
+Finally, it is interesting to compare the slice views to the projection
 of the models seen in Fig. \ref{proj1} to better understand how the

### 2023-01-19 review/response.Rmd
updated response
added new plots

@@
 
-\textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny}
+\textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny/rgl.}
 
@@
 Also since is the first time you have mentioned Mathematica you should include the reference here (Wolfram reference appears in page 4 line 26 for first time)
+
 \textcolor{blue}{done}
@@
 
-\textcolor{blue}{definition of $V_m$ added but have not addressed the second part of the comment}
+\textcolor{blue}{definition of $V_m$ added but have not addressed the second part of the comment TODO}
 

### 2023-01-19 paper/paper.Rmd
added new plots

@@
 
-The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in Mathematica [@Mathematica].
+The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in Mathematica.
 The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
@@
 Most of the data, though, is first constructed using R, especially the classification boundaries.
-If new technology for interactive graphics becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.
+If new technology for interactive graphics with precise cursor control becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.
 

### 2023-01-19 paper/paper.tex
added new plots

@@
 algebra capabilities, and relatively new interactive graphics interface,
-available in Mathematica \citep{Mathematica}. The components
-corresponding to \(V_m\) are directly controlled by cursor movement,
-which updates row \(m\) of \(A\). The updated matrix is then
-orthonormalised.
+available in Mathematica. The components corresponding to \(V_m\) are
+directly controlled by cursor movement, which updates row \(m\) of
+\(A\). The updated matrix is then orthonormalised.
 
@@
 using R, especially the classification boundaries. If new technology for
-interactive graphics becomes available in R it would be useful to create
-a tighter coupling of models and visualization to allow exploring and
-comparing fits.
+interactive graphics with precise cursor control becomes available in R
+it would be useful to create a tighter coupling of models and
+visualization to allow exploring and comparing fits.
 

### 2023-01-19 review/response.Rmd
handling various reviewer requests

@@
  XXX If use new slice center control for fig 7, need to make a note of it here, updated the mathematica notebook and check text and captions correct.
+
+ XXX References need to be consistent, all initials or all full names, and all sentence case or title case on titles.
 
@@
 
-\textcolor{blue}{Colour in figures has been changed to work in black and white as well as colour. Captions have been adjusted accordingly. TODO}
+\textcolor{blue}{Colour in figures has been changed to work in black and white as well as colour. Captions have been adjusted accordingly.}
 
@@
 
-\textcolor{blue}{see text in blue page 3, is this adequate?}
+\textcolor{blue}{Done}
 
@@
 
-\textcolor{blue}{Modified}
+\textcolor{blue}{Modified accordingly.}
 
@@
 
-\textcolor{blue}{This is better explained}
+\textcolor{blue}{This is better explained.}
 
@@
 
-\textcolor{blue}{done}
+\textcolor{blue}{Done}
 
@@
 
+\textcolor{blue}{Links added}
+
 - Pg 2, line 23: "A video illustrating the capabilities is available through video library of ASA Statistical Graphics Section (2022)."
 Add "the" before video
-\textcolor{blue}{done}
+
+\textcolor{blue}{Done}
 
@@
 
-\textcolor{blue}{definition of $V_m$ added but have not addressed the second part of the comment TODO}
+\textcolor{blue}{Definition of $V_m$ added and Figure 1 has been annotated to make the algorithm easier to understand.
+Controlling more than one variable is possible with the R implementation and this has been noted in the text.}
 
@@
 
+\textcolor{blue}{Interpreting this to be Section 2.2, not line 2.2. Annotation has been added to existing figure.}
+
 Section 3: will be useful to include a schematic diagram illustrating the interactivity
+
+\textcolor{blue}{We're not sure what this refers to.
+The additional details and explanations of the algorithm are provided in the appendix.
+Additional refinement, and figure annotations, were made in the previous sections.}
 
@@
 This part is difficult to follow, it is more clear after reading the application Section because you have included some videos but I think you need to include something here to help the reader.
+
+\textcolor{blue}{Additional explanation, clearer notation, and annotation of the plot has been done to help here.}
 
@@
 
+\textcolor{blue}{Additional text has been added to explain why this is difficult in R/Shiny/rgl.}
+
 Section 4: Page 10 line 36: Figure 3 summarizes (space)
 
+\textcolor{blue}{Thanks}
+
 Section 5: page 15 line 26 "If new technology for interactive graphics became available"  be more specific about the needed technology in R to implement the manual controls as you did with Mathematica
+
+\textcolor{blue}{Done}

### 2023-01-19 paper/paper.Rmd
handling various reviewer requests

@@
 A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
-Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections.
+Tours involve views of high-dimensional  data with low-dimensional projections.
 In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
-Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
+(Various videos of tours, e.g.
+@dataviewer, @gt-pp-video, on the @ASA22 illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-(A video illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
+(A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
 After Asimov there have been many tour developments, which are summarized in @lee2021.
@@
 
-A manual tour allows the user to alter the coefficients of one (or more) variables contributing to a $d$ dimensional projection.
+A manual tour allows the user to alter the coefficients of one (or more) variables, of $p$, contributing to a $d$-dimensional projection.
 The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be \textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.} A method to update the values of the component of the controlled variable $V_m$ is then needed.
@@
 This algorithm will produce the changes to a projection as illustrated in Figure \ref{fig:manualsequence}.
-The controlled variable, $V_m$, corresponds to the black line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (orange dot).
+The controlled variable, $V_m$, corresponds to the dark line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (dot).
 Changes in the other components happen as a result of the orthonormalisation, but are uncontrolled.
 
-```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
-The dot (orange) indicates the chosen values for the controlled variable.
+The algorithm will also allow for more than one variable to be controlled. This is more difficult to implement interactively, but is easy with scripting, as is available in the R implementation.
+
+```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (dark) is changed using unconstrained orthonormalisation.
+The dot indicates the chosen values for the controlled variable, $V_m$.
 It can be seen that the actual axis does not precisely match the chosen position, but it is close."}
@@
 plot_basis(A, vchange, vcol="black")
+text(A[vchange, 1], A[vchange, 2], expression(V[m]), pos=3, col="orange")
+points(A[vchange, 1], A[vchange, 2], pch=16, col="orange")
 
@@
   plot_basis(Anew, vchange, vcol="black")
+  text(Anew[vchange, 1], Anew[vchange, 2], expression(V[m]), pos=3, col="orange")
   points(exact[1], exact[2], pch=16, col="orange")
@@
 
-\textcolor{blue}{The Mathematica implementation relies on the inbuilt Mathematica function  `LocatorPane`.
-This function creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
+\textcolor{blue}{The Mathematica implementation relies on the inbuilt function  `LocatorPane`.
+This function creates a region on the screen where the position of the cursor is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
 We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
-Control objects, such as sliders, locator panes, and input fields, are used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that variable will be updated.}
+User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale and point size.}
 
@@
 
-The R package `tourr` provides numerous types of tours.
-An additional function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.
+The R package `tourr` provides numerous types of tours, displayed as animations.
+
+The new function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values.
 It is applicable for any projection dimension, $d$.
+The script interface has one advantage over an interactive interface, in that two or more variables can be controlled simultaneously.
+
 Changing slice center manually can be accomplished with the new function, `manual_slice()`.
@@
 It slices from the center of the data out to one side, back to the center, out to the opposite side, and then back to the center.
-The appendix contains a discussion of a possible visual guide for the slice center location.
+<!--The appendix contains a discussion of a possible visual guide for the slice center location.-->
 
@@
 - Appendix with additional details about manual controls, and Mathematica functions.
+- The `Rmd` file that generated the paper, containing the R code to produce Figure 1.
 - Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2023-01-19 paper/paper.tex
handling various reviewer requests

@@
 space to create dynamic data visualizations called \emph{tours}. Tours
-involve views of high-dimensional (\(p\)) data with low-dimensional
-(\(d\)) projections. In his original paper on the grand tour,
-\citet{As85} provided several algorithms for tour paths that could
-theoretically show the viewer the data \emph{from all sides}. Prior to
-Asimov's work, there were numerous preparatory developments including
-\citet{tukey}'s PRIM-9. PRIM-9 had user-controlled rotations on
-coordinate axes, allowing one to manually tour through low-dimensional
-projections. (A video illustrating the capabilities is available through
-\textcolor{blue}{the} video library of \citet{ASA22}.) Steering through
-all possible projections is impossible, unlike Asimov's tours which
-allows one to quickly see many, many different projections. After Asimov
-there have been many tour developments, which are summarized in
-\citet{lee2021}.
+involve views of high-dimensional data with low-dimensional projections.
+In his original paper on the grand tour, \citet{As85} provided several
+algorithms for tour paths that could theoretically show the viewer the
+data \emph{from all sides}. (Various videos of tours, e.g.
+\citet{dataviewer}, \citet{gt-pp-video}, on the \citet{ASA22} illustrate
+this concept.) Prior to Asimov's work, there were numerous preparatory
+developments including \citet{tukey}'s PRIM-9. PRIM-9 had
+user-controlled rotations on coordinate axes, allowing one to manually
+tour through low-dimensional projections. (A video \citep{PRIM9-video}
+illustrating the capabilities is available through \textcolor{blue}{the}
+video library of \citet{ASA22}.) Steering through all possible
+projections is impossible, unlike Asimov's tours which allows one to
+quickly see many, many different projections. After Asimov there have
+been many tour developments, which are summarized in \citet{lee2021}.
 
@@
 A manual tour allows the user to alter the coefficients of one (or more)
-variables contributing to a \(d\) dimensional projection. The initial
-ingredients are an orthonormal basis (\(A_{p\times d}\)) defining the
-projection of the data, and a variable id (\(m \in \{1, ..., p\}\))
-specifying which coefficient will be
+variables, of \(p\), contributing to a \(d\)-dimensional projection. The
+initial ingredients are an orthonormal basis (\(A_{p\times d}\))
+defining the projection of the data, and a variable id
+(\(m \in \{1, ..., p\}\)) specifying which coefficient will be
 \textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.}
@@
 in Figure \ref{fig:manualsequence}. The controlled variable, \(V_m\),
-corresponds to the black line, and sequential changes to row \(m\) of
-\(A\) can be seen to roughly follow a specified position (orange dot).
-Changes in the other components happen as a result of the
-orthonormalisation, but are uncontrolled.
+corresponds to the dark line, and sequential changes to row \(m\) of
+\(A\) can be seen to roughly follow a specified position (dot). Changes
+in the other components happen as a result of the orthonormalisation,
+but are uncontrolled.
+
+The algorithm will also allow for more than one variable to be
+controlled. This is more difficult to implement interactively, but is
+easy with scripting, as is available in the R implementation.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation.
-The dot (orange) indicates the chosen values for the controlled variable.
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (dark) is changed using unconstrained orthonormalisation.
+The dot indicates the chosen values for the controlled variable, $V_m$.
 It can be seen that the actual axis does not precisely match the chosen position, but it is close.}\label{fig:manualsequence}
@@
 
-The R package \texttt{tourr} provides numerous types of tours. An
-additional function, \texttt{radial\_tour()} has been constructed that
-will generate a sequence of projections that will decrease the
-coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
-values. It is applicable for any projection dimension, \(d\). Changing
-slice center manually can be accomplished with the new function,
-\texttt{manual\_slice()}. This changes the value of the center point
-(\(\mathbf{c}\)) of the slice, along a selected variable axis. It slices
-from the center of the data out to one side, back to the center, out to
-the opposite side, and then back to the center. The appendix contains a
-discussion of a possible visual guide for the slice center location.
+The R package \texttt{tourr} provides numerous types of tours, displayed
+as animations.
+
+The new function, \texttt{radial\_tour()} has been constructed that will
+generate a sequence of projections that will decrease the coordinates
+for \(V_m\) to \(\boldsymbol{0}\) and back to the original values. It is
+applicable for any projection dimension, \(d\). The script interface has
+one advantage over an interactive interface, in that two or more
+variables can be controlled simultaneously.
+
+Changing slice center manually can be accomplished with the new
+function, \texttt{manual\_slice()}. This changes the value of the center
+point (\(\mathbf{c}\)) of the slice, along a selected variable axis. It
+slices from the center of the data out to one side, back to the center,
+out to the opposite side, and then back to the center.
 
@@
 \item
+  The \texttt{Rmd} file that generated the paper, containing the R code
+  to produce Figure 1.
+\item
   Animations illustrating the manual tour and slicing, matching the

### 2023-01-19 paper/appendix.Rmd
description of slice centre guide in software added to appendix

@@
 
+\textcolor{blue}{The slice centre position is controlled via numerical input along each axis in the original command `SliceDynamic`.
+This provides numerical feedback on the position of the centre.
+For easier navigation but without the numerical feedback, an implementation of the slice centre control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
+
 The package includes the additional functions not used in the example application: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.

### 2023-01-19 paper/appendix.tex
description of slice centre guide in software added to appendix

@@
 
+\textcolor{blue}{The slice centre position is controlled via numerical input along each axis in the original command `SliceDynamic`.
+This provides numerical feedback on the position of the centre.
+For easier navigation but without the numerical feedback, an implementation of the slice centre control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
+
 The package includes the additional functions not used in the example

### 2023-01-19 review/response.Rmd
more details on slice center guide

@@
 
+\textcolor{blue}{Noted}
+
 2- I will download and run your code as part of the next round of reviews.
@@
 Before resubmitting please run your code in a "clean" environment as issues such as required packages that aren't loaded or calls to files that don't exist are common and can delay publication (I'm not saying these exist in your case, just asking you to be aware!).
+
+\textcolor{blue}{Done - TODO check this}
 
@@
 
+\textcolor{blue}{Done}
+
 **Associate Editor**
@@
     - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat diﬀicult to assess the software side of this paper.
+
+\textcolor{blue}{Done.
+Previously we provided the public github repo URL so that all files to reproduce could be downloaded.
+Here we have included all the files in the supplementary zip.
+XXX TODO Please check the list of Supplementary material in the paper.}
+
 - Grammar: There are a lot of minor grammatical issues with this paper.
@@
 The ones that I noticed as I read through the first section are detailed in minor issues below, but I am trying to focus more on the content than the style, so I will try to not nit-pick through the rest of the paper and trust that the authors can use the online tools.
+
+\textcolor{blue}{Done. Apologies for the grammatical errors.}
+
 - Cite your software packages appropriately - they're tools you depend on to do your work and should be cited in the same way papers are cited.
+
+\textcolor{blue}{Done. They were all cited, but not by our error, at the first mention.}
 
@@
 
+\textcolor{blue}{Done.}
+
     – Page 3, line 32: Cite R
+
+\textcolor{blue}{Done.}
 
@@
 
-\textcolor{blue}{Done - the citations were there, but not placed at the first mention.}
+\textcolor{blue}{Done.}
 
@@
 
-\textcolor{blue}{Done}
+\textcolor{blue}{Done.}
 
@@
 
-\textcolor{blue}{Links added}
+\textcolor{blue}{Links added.}
 
@@
 
+\textcolor{blue}{More explanation has been added, and more details in the tourr implementation section.}
+
 Section 1 line 29 "The software section describes mathematica package"  mathematica should be capitalized as you did  through the paper in most cases.
@@
 
-\textcolor{blue}{Thanks}
+\textcolor{blue}{Thanks.}
 
@@
 
-\textcolor{blue}{Done}
+\textcolor{blue}{Done.}

### 2023-01-19 paper/paper.Rmd
more details on slice center guide

@@
 This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration into a later study, as shown in the applications below.
-
-Details about the implementation and usage instructions are given in the Appendix.
-
-<!-- 
-The following functions are implemented in the Mathematica package `mmtour.wl`:
-
-- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
-colour(=Automatic)]` with arguments
-
-    - `data`: A list of data matrices (they must all have the same dimensionality)
-    - ``projmat`: A projection matrix which describes the projection plane. This is the 
-initial projection and does not need to be orthonormal (XXX is this right?).
-    - ``xRange`: A list which details the range of the horizontal axis.
-    - ``yRange`: A list which details the range of the vertical axis.
-    - ``colour`: A list of graphics directives to be applied to each data matrix.
-
-There are problems with the default values for x and y ranges and colour. It would be 
-better to fix the ranges and colour based on the data as well as the origin of axes 
-and the size of the points.
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**
-
-The format for the printed projection matrix can be improved also. Maybe fix the 
-number of digits and the position of the label.
-
-- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)
-
-    - `data`: A data matrix
-    - `projmat`: The initial projection matrix
-    - `centrepoint`: The initial position of the slice
-    - `height`: The initial height of the slice
-    - `heightRange`: The range of heights
-    - `ptSize1`: The size of the points that exist within the slice
-    - `ptSize2`: The size of the points that exist outside of the slice
-
-This function requires some auxiliary functions. It does not accept more than one data set.
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**
-
-It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes
-
-- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments
-
-- `data`: A data matrix with an extra grouping or clustering column (numerical) 
-- `projmat`: The initial projection matrix
-- `centrepoint`: The initial position of the slice
-- `height`: The initial height of the slice
-- `heightRange`: The range of heights
-- `ptSize1`: The size of the points that exist within the slice
-- `ptSize2`: The size of the points that exist outside of the slice
-
-**Does this need to be a separate function from SlicePlot? Fix 2**
-
-**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**
-
-**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**
-
-**I have found it useful to input different data sets rather than have this function sort them automatically using the last column.
-One data set with a flag column at the end, is best!
-Is the code too complicated?
-Fix 5**
-
-- `Projected2DSliderPlot[]`
-
-This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalized to arbitrary data if we want to keep it.
-
-**For Alex to do**
-
-- **We would need a screenshot (and video) for each of these.**
-- **Fix 1, 2, 3, 4, 5**
-- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
-    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing).
-It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
-    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
-    - **Can we have some heuristic to decide the range for the slice thickness slider?
-In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting.
-My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
-    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
-    - **pdfsense data: can we use random samples instead of the first 10k rows?
-Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
-    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
-    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
-    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1.
-Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection.
-We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables.
-Much less clear, maybe some indication of smaller x2 and larger x3 for this.**
-    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
-    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further.
-Now interpretation becomes tricky.
-Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables).
-For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
-    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**
--->
+More details about the implementation and usage instructions are given in the Appendix.
 
@@
 We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in top row of Figure \ref{slice1p}.
+The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.
 We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
@@
 A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in the bottom row of Figure \ref{slice1p}.
-The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
+(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.) The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
 On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
@@
 - The `Rmd` file that generated the paper, containing the R code to produce Figure 1.
+- Data files for the examples.
 - Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2023-01-19 paper/paper.tex
more details on slice center guide

@@
 identified as interesting in the manual exploration into a later study,
-as shown in the applications below.
-
-Details about the implementation and usage instructions are given in the
-Appendix.
+as shown in the applications below. More details about the
+implementation and usage instructions are given in the Appendix.
 
@@
 shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
-values of bill depth in top row of Figure \ref{slice1p}. We can see that
-all three slices (the two models and the data) contain almost no points
-from the third class (Gentoo), and that the decision boundary between
-the two models is very similar.
+values of bill depth in top row of Figure \ref{slice1p}. The star glyph
+at right indicates the slice center value, which shows that it is higher
+on \texttt{bd} and at the middle value for the other variables. This
+slice center guide is as proposed in \citet{condviz2}, and more details
+can be found in the Appendix. We can see that all three slices (the two
+models and the data) contain almost no points from the third class
+(Gentoo), and that the decision boundary between the two models is very
+similar.
 
@@
 slice localized towards low values of bill depth, shown in the bottom
-row of Figure \ref{slice1p}. The RF model (left) predicts all three
-species within this slice, with an interesting boundary for the third
-class (Gentoo). On the other hand the LDA model (middle) predominantly
-predicts the third class within the slice, this appears to be enforced
-through the linear structure of the model. Looking finally at the thick
-slice through the data we see that there are primarily observations from
-this class within the slice we can conclude that the two models have
-filled in the ``empty'' space (where we do not have any training
-observations) in very different ways and according to what we might
-expect given the model structure.
+row of Figure \ref{slice1p}. (The star glyph at right indicates the
+slice center value, which shows that it is lower on \texttt{bd} and at
+the middle value for the other variables.) The RF model (left) predicts
+all three species within this slice, with an interesting boundary for
+the third class (Gentoo). On the other hand the LDA model (middle)
+predominantly predicts the third class within the slice, this appears to
+be enforced through the linear structure of the model. Looking finally
+at the thick slice through the data we see that there are primarily
+observations from this class within the slice we can conclude that the
+two models have filled in the ``empty'' space (where we do not have any
+training observations) in very different ways and according to what we
+might expect given the model structure.
 
@@
 \item
+  Data files for the examples.
+\item
   Animations illustrating the manual tour and slicing, matching the

### 2023-01-20 paper/appendix.Rmd

@@
 plot_basis(A, vchange, vcol="black")
+text(A[vchange, 1], A[vchange, 2], expression(V[m]), pos=3, col="orange")
 
@@
   points(exact[1], exact[2], pch=16, col="orange")
+  text(Anew[vchange, 1], Anew[vchange, 2], expression(V[m]), pos=3, col="orange")
+
 }

### 2023-01-20 paper/paper.Rmd

@@
 We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in top row of Figure \ref{slice1p}.
-The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
-This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.
-We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
+\textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
 
@@
 A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in the bottom row of Figure \ref{slice1p}.
-(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.) The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
+\textcolor{blue}{(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.)} The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
 On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.

### 2023-01-20 paper/paper.tex

--- 2023-01-19 paper/paper.tex
@@
 shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
-values of bill depth in top row of Figure \ref{slice1p}. We can see that
-all three slices (the two models and the data) contain almost no points
-from the third class (Gentoo), and that the decision boundary between
-the two models is very similar.
+values of bill depth in top row of Figure \ref{slice1p}.
+\textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.}
+We can see that all three slices (the two models and the data) contain
+almost no points from the third class (Gentoo), and that the decision
+boundary between the two models is very similar.
 
@@
 slice localized towards low values of bill depth, shown in the bottom
-row of Figure \ref{slice1p}. The RF model (left) predicts all three
-species within this slice, with an interesting boundary for the third
-class (Gentoo). On the other hand the LDA model (middle) predominantly
-predicts the third class within the slice, this appears to be enforced
-through the linear structure of the model. Looking finally at the thick
-slice through the data we see that there are primarily observations from
-this class within the slice we can conclude that the two models have
-filled in the ``empty'' space (where we do not have any training
-observations) in very different ways and according to what we might
-expect given the model structure.
+row of Figure \ref{slice1p}.
+\textcolor{blue}{(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.)}
+The RF model (left) predicts all three species within this slice, with
+an interesting boundary for the third class (Gentoo). On the other hand
+the LDA model (middle) predominantly predicts the third class within the
+slice, this appears to be enforced through the linear structure of the
+model. Looking finally at the thick slice through the data we see that
+there are primarily observations from this class within the slice we can
+conclude that the two models have filled in the ``empty'' space (where
+we do not have any training observations) in very different ways and
+according to what we might expect given the model structure.
 

### 2023-01-27 paper/appendix.Rmd
grammarly changes

--- 2022-10-04 paper/appendix.Rmd
@@
 
-The problem with the new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.
+The problem with the new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalization will change them.
 
@@
 If $d>2$ the process would be sequentially repeated in the same manner that Gram-Schmidt is applied sequentially to orthormalise the columns of a matrix.
-If $d=1$ no orthonormalisation is needed, and the projection vector would simply need to be normalized after each adjustment.
+If $d=1$ no orthonormalization is needed, and the projection vector would simply need to be normalized after each adjustment.
 
-<!-- Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.
+<!-- Need to think about how the values for the selected variable are made exact, that is constrained orthonormalization.
 
-Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?-->
+Need to think about checks, and error catching, maybe in the implementation section. If the selected variable has a 0 coefficient, will that generate an orthonormalization error?-->
 
@@
 # Iterate and save images
-# Unconstrained orthonormalisation
+# Unconstrained orthonormalization
 Anew <- A
@@
 Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$.
-Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix.
+Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee the orthonormality of the new projection matrix.
 One possible algorithm to achieve this is
@@
 \item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
-\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
+\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
 \end{enumerate}
@@
 However, it renders projections in a way that is not continuous and may be distracting.
-This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.
+This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalization.
 
@@
 We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection.
-In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
+In that case, we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$.
 Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor.
-By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user.
+By adjusting the normalization procedure we can also ensure that the length remains as selected manually by the user.
 
@@
 ```{r anchornav, fig.width=10, fig.height=3, out.width="100%", fig.cap="Visual guide for the slice center.
-Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+Variable axes are displayed in polar coordinates, where the center corresponds to the minimum value and the outer end corresponds to the maximum value.
 The position of the center corresponds to the dark polygon.
-If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis."}
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B and C show its position when moving the center along one axis."}
 par(pty="s", mfrow=c(1,3))
@@
 
+\textcolor{blue}{The slice center position is controlled via numerical input along each axis in the original command `SliceDynamic`.
+This provides numerical feedback on the position of the center.
+For easier navigation but without the numerical feedback, an implementation of the slice center control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
+
 The package includes the additional functions not used in the example application: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.
@@
 This function differs from `SliceDynamic` in that it looks at only one data set but displays both the points inside and outside the slice.
-The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, center point and zoom level.
+The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, center point, and zoom level.
 The function and its arguments are given below:
@@
 
-- `data`: A data matrix containing only one group. In this case the data matrix should not contain column labels.
+- `data`: A data matrix containing only one group. In this case, the data matrix should not contain column labels.
 - `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.

### 2023-01-27 paper/appendix.tex
grammarly changes

--- 2022-10-04 paper/appendix.tex
@@
 The problem with the new simple method (Algorithm 1) is that the precise
-values for \(V_m\) cannot be specified because the orthonormalisation
+values for \(V_m\) cannot be specified because the orthonormalization
 will change them.
@@
 that Gram-Schmidt is applied sequentially to orthormalise the columns of
-a matrix. If \(d=1\) no orthonormalisation is needed, and the projection
+a matrix. If \(d=1\) no orthonormalization is needed, and the projection
 vector would simply need to be normalized after each adjustment.
@@
 constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
-to guarantee orthonormality of the new projection matrix. One possible
-algorithm to achieve this is
+to guarantee the orthonormality of the new projection matrix. One
+possible algorithm to achieve this is
 
@@
 the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
-orthogonalisation.
+orthogonalization.
 
@@
 through rotation within the \(d\)-dimensional hyperplane defined by the
-new projection. In that case we would again start from \(A^*\) as above,
-and then rotate the basis such that the first direction is aligned with
-the selected variable \(m\). Applying Gram-Schmid within that basis
-definition will not alter the direction for the first vector, thus the
-direction of \(m\) remains fixed to what was selected with the cursor.
-By adjusting the normalization procedure we can also ensure that the
-lenght remains as selcted manually by the user.
+new projection. In that case, we would again start from \(A^*\) as
+above, and then rotate the basis such that the first direction is
+aligned with the selected variable \(m\). Applying Gram-Schmid within
+that basis definition will not alter the direction for the first vector,
+thus the direction of \(m\) remains fixed to what was selected with the
+cursor. By adjusting the normalization procedure we can also ensure that
+the length remains as selected manually by the user.
 
@@
 \includegraphics[width=1\linewidth]{appendix_files/figure-latex/anchornav-1} \caption{Visual guide for the slice center.
-Variable axes are displayed in polar coordinates, where the center corresponds to minimum value and outer end corresponds to maximum value.
+Variable axes are displayed in polar coordinates, where the center corresponds to the minimum value and the outer end corresponds to the maximum value.
 The position of the center corresponds to the dark polygon.
-If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B, C show it's position when moving the center along one axis.}\label{fig:anchornav}
+If the center is at the center of the data, this will be displayed as a regular polygon (A), and plots B and C show its position when moving the center along one axis.}\label{fig:anchornav}
 \end{figure}
@@
 making 2 invisible.
+
+\textcolor{blue}{The slice center position is controlled via numerical input along each axis in the original command `SliceDynamic`.
+This provides numerical feedback on the position of the center.
+For easier navigation but without the numerical feedback, an implementation of the slice center control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
 
@@
 the user to specify the size of both sets of points (inside and outside)
-along with the slice height, center point and zoom level. The function
+along with the slice height, center point, and zoom level. The function
 and its arguments are given below:
@@
 \item
-  \texttt{data}: A data matrix containing only one group. In this case
+  \texttt{data}: A data matrix containing only one group. In this case,
   the data matrix should not contain column labels.

### 2023-01-27 paper/paper.Rmd
grammarly changes

@@
 The Mathematica package as well as example notebooks are provided, which contain functions enabling the user to experiment with these new manual controls, with one specifically for exploring regions and boundaries produced by classification models.
-The advantage of Mathematica is its linear algebra capabilities, and interactive cursor location controls.
+The advantage of Mathematica is its linear algebra capabilities and interactive cursor location controls.
 Some limited implementation has also been made available in the R package \tt{tourr}.
@@
 (Various videos of tours, e.g.
-@dataviewer, @gt-pp-video, on the @ASA22 illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
+@dataviewer, @gt-pp-video, on the @ASA22, illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-(A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections.
-After Asimov there have been many tour developments, which are summarized in @lee2021.
+(A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allow one to quickly see many, many different projections.
+After Asimov, there have been many tour developments, which are summarized in @lee2021.
 
@@
 One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour.
-@cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, or out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
+@cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value.
 @spyrison_spinifex_2020 refined this algorithm and implemented it to generate radial tour animation sequences.
@@
 <!-- Purpose -->
-Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection.
+Manual controls are especially useful for assessing the sensitivity of structure to particular elements of the projection.
 There are many places where it is useful.
 In exploratory data analysis, where one sees clusters in a projection, one may ask whether some variables can be removed from the projection without affecting the clustering.
-For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance.
+For interpreting models, one can reduce or increase a variable's contribution, to examine the variable importance.
 Having the user interact with a projection is extremely valuable for understanding high-dimensional data.
-However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, (2) extending to higher dimensional control is difficult.
+However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, and (2) extending to higher dimensional control is difficult.
 
-Another potentially useful manual control, is to allow the user to choose the position of the center of a slice.
+Another potentially useful manual control is to allow the user to choose the position of the center of a slice.
 The slice tour was introduced in @slicetour.
@@
 The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers.
-The software section describes a Mathematica [@Mathematica] package that is used for the application, and describes the interactive environment that would be desirable within R [@rref] as new technology becomes available.
+The software section describes a Mathematica [@Mathematica] package that is used for the application and describes the interactive environment that would be desirable within R [@rref] as new technology becomes available.
 The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the applications.
@@
 Movements of a cursor are recorded and converted into changes in the values of $V_m$ thus changing the displayed 2D projection.
-The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.
+The algorithm also provided constraints to horizontal, vertical, radial, or angular motions only.
 The construction of the manipulation space overly complicates the manual controls, especially when considering possible techniques that will apply to arbitrary $d$.
@@
 \centerline{\includegraphics[width=0.8\textwidth]{figures/manip_space.pdf}}
-\caption{Original contruction of the manual tour designed for 2D projections and created a 3D space from which to utilise track ball controls to change it's contribution.
+\caption{Original construction of the manual tour designed for 2D projections and created a 3D space from which to utilize track ball controls to change its contribution.
 (Figure 3 from Cook and Buja (1997).)}
@@
 
-The new approach emerged from experiments on the tour using the linear algebra capabilities, and relatively new interactive graphics interface, available in Mathematica.
+The new approach emerged from experiments on the tour using linear algebra capabilities, and a relatively new interactive graphics interface, available in Mathematica.
 The components corresponding to $V_m$ are directly controlled by cursor movement, which updates row $m$ of $A$.
-The updated matrix is then orthonormalised.
+The updated matrix is then orthonormalized.
 
@@
 <!--
-Note: I removed the angular bracets since cdot denotes scalar product already (also used like this below for the slice definitions)
+Note: I removed the angular brackets since cdot denotes scalar product already (also used like this below for the slice definitions)
 -->
@@
 The controlled variable, $V_m$, corresponds to the dark line, and sequential changes to row $m$ of $A$ can be seen to roughly follow a specified position (dot).
-Changes in the other components happen as a result of the orthonormalisation, but are uncontrolled.
+Changes in the other components happen as a result of the orthonormalization but are uncontrolled.
 
@@
 
-```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where contribution of one variable is controlled (dark) is changed using unconstrained orthonormalisation.
+```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization.
 The dot indicates the chosen values for the controlled variable, $V_m$.
@@
 
-The problem with the new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalisation will change them.
+The problem with the new simple method is that it is not faithful to the precise values for $V_m$ because the orthonormalization will change them.
 Even though these changes are for the most part imperceptible, one may wish to avoid them and there are numerous ways that this can be enforced, a few are detailed in the Appendix.
-These primarily differ in how the remaining variables are adjusted during orthonormalisation.
+These primarily differ in how the remaining variables are adjusted during orthonormalization.
 
@@
 To better explore the space we combine the manual controls for the projection with manual controls for slicing.
-A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].
+A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space, and the slice thickness $h$ [@slicetour].
 A data point is inside the slice if its orthogonal distance from the projection plane (passing through the center point) is below the thickness $h$.
@@
 
-In addition it is also useful to interactively change the slice thickness $h$ (also called the slice radius), in particular to find the preferred value for exploring the input data.
+In addition, it is also useful to interactively change the slice thickness $h$ (also called the slice radius), in particular, to find the preferred value for exploring the input data.
 For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with
@@
 The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it.
-In practice, the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.
+In practice, the closest axis will be selected, and the dragging results in tracking the mouse position which will update the projection in small steps, such that no further interpolation is required.
 
@@
 Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness, $h$ (named `height` in the Mathematica interface), and a numeric input to specify the center point $c$ explicitly.
-The display will update based on these input values, in particular the view will jump to a new slice when $c$ is changed.<!-- (though an interpolation might also be useful).-->
+The display will update based on these input values, in particular, the view will jump to a new slice when $c$ is changed.<!-- (though an interpolation might also be useful).-->
 
 \textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing.
-However, the `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system.
-It lacks capability to implement the primary tool needed for a manual tour, that is the tracking of the cursor position in small increments.
-In theory, the historical `locator()` function of base R should provide the necessary capability, but it's behavior is not consistent between graphical devices.
+However, the `plotly` package is designed to add to `ggplot2` graphics, not to be a programmable interactive graphics system.
+It lacks the capability to implement the primary tool needed for a manual tour, which is the tracking of the cursor position in small increments.
+In theory, the historical `locator()` function of base R should provide the necessary capability, but its behavior is not consistent between graphical devices.
 The cursor position control is possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary dimension of tours.
 The level of interactivity would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that currently operates across platforms.
-What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package.
+What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement the resulting ideas as an animation in the `tourr` [@tourr] package.
 There is hope that better interactivity will be available for R soon, perhaps in javascript [e.g.
@@
 We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
-User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale and point size.}
+User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale, and point size.}
 
@@
 
-Changing slice center manually can be accomplished with the new function, `manual_slice()`.
+Changing the slice center manually can be accomplished with the new function, `manual_slice()`.
 This changes the value of the center point ($\mathbf{c}$) of the slice, along a selected variable axis.
@@
 
-Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice, similarly to how the manual tour was implemented here.
+Potential adjustments for future work include better visualization of the current center point, potentially with interactive graphics to move the slice, similar to how the manual tour was implemented here.
 An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.
@@
 
-```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
+```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data.
+The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."}
 library(tidyverse)
@@
 Figure \ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
-There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo).
-The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
+There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap, and Gentoo).
+The scatterplot matrix shows that the three species appear to be likely separable and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.
 The steps for exploring boundaries in this example are as follows:
@@
 And, interestingly, this particular projection of the original data shows very distinct clusters of the three species.
-That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
+That means the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.
 
@@
 %\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
-\caption{A projection identified using the manual tour, because it shows differences in classification boundaries induced by different models.
-A RF model produces blocky, rectangular, boundaries, while the LDA model produces oblique linear boundaries.
-For comparison, in this projection the data shows the three groups are almost perfectly separable.}
+\caption{A projection identified using the manual tour because it shows differences in classification boundaries induced by different models.
+An RF model produces blocky, rectangular, boundaries, while the LDA model produces oblique linear boundaries.
+For comparison, in this projection, the data shows the three groups are almost perfectly separable.}
 \label{proj1}
@@
 
-This example can be reproduced with the code in penguins_exploring_manually.nb and a run through is shown in the video at https://vimeo.com/747585410.
+This example can be reproduced with the code in penguins_exploring_manually.nb and a run-through is shown in the video at https://vimeo.com/747585410.
 
@@
 For both models we look at a thin ($h=0.5$) slice through the center, $\mathbf{c}^0 = (0,0,0,0)$.
-At  first we explore how changing the projection away from $A_1$ can help with understanding the boundary better.
+At first, we explore how changing the projection away from $A_1$ can help with understanding the boundary better.
 For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.
@@
 The RF is almost straight in this view, as something we might expect from a tree model where splitting occurs on single variables only.
-However, it is straight in a combination of the second and third variable (`bd` and `fl`).
-By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl` it is possible to cut on a value where most of the Gentoo penguins are different from the other two species for any tree, and likely only one is needed.
+However, it is straight in a combination of the second and third variables (`bd` and `fl`).
+By examining the scatterplot matrix, we can understand how this boundary was built: on each of `bd` and `fl`, it is possible to cut on a value where most of the Gentoo penguins are different from the other two species for any tree, and likely only one is needed.
 Thus the forest construction is providing a sample of trees that use one variable or the other variable to split, producing the blocky boundary on the combination of the two.
 
-In contrast, the LDA boundary has an oblique split between the two, and roughly divides the space into three similarly sized areas.
-It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary half-way between the group means.
+In contrast, the LDA boundary has an oblique split between the two and roughly divides the space into three similarly sized areas.
+It is almost like a textbook illustration of how LDA works for two variables (2D) where by assuming the clusters have equal variance-covariance, it places a boundary halfway between the group means.
 
@@
 \ref{slice1}, top-left).
-This is similar for the LDA model (Fig.
+This is similar to the LDA model (Fig.
 \ref{slice1}, top-right), but again the linearity results in different boundaries and thus differences in how the classes overlap.
@@
 The LDA boundary is more oblique between Gentoo and Chinstrap than the RF, which is too close to the Chinstrap group.
-On the other hand between Adelie and Gentoo, the RF is better situated about half-way between the two groups.
+On the other hand between Adelie and Gentoo, the RF is better situated about halfway between the two groups.
 Interestingly, the RF boundaries, which can be highly nonlinear, albeit blocky, are relatively linear for this problem.
@@
 Finally, to determine which model describes the boundaries better, compare them with the $A_2$ projection of the data (bottom row of Figure \ref{slice1}).
-Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group.
+Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest have greedily grasped any one of many places to make a split to separate the group.
 It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data.
@@
 
-This example can be reproduced with the code in `penguins_slicing_through_center.nb` and a run through is shown in the first part of the video at https://vimeo.com/747590472.
+This example can be reproduced with the code in `penguins_slicing_through_center.nb` and a run-through is shown in the first part of the video at https://vimeo.com/747590472.
 
@@
 
-We start by a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in top row of Figure \ref{slice1p}.
+We start with a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in the top row of Figure \ref{slice1p}.
 \textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
-This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo), and that the decision boundary between the two models is very similar.
+This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.
 
@@
 Here the center has been shifted along the axis of `bd` (icon display at right represents slice center position).
-(Top row) Center is at large values of `bd`, showing a region with no Gentoo.
+(Top row) The center is at large values of `bd`, showing a region with no Gentoo.
 The two models have similar boundaries except that the LDA is more linear.
-(Bottom row) Center is at small values of `bd`, reveals a region where the models differ substantially.
+(Bottom row) The center is at small values of `bd`, revealing a region where the models differ substantially.
 The RF model carves some area for Adelie, but LDA would only predict to be Gentoo or Chinstrap here.
@@
 \textcolor{blue}{(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.)} The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
-On the other hand the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
+On the other hand, the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
 Looking finally at the thick slice through the data we see that there are primarily observations from this class within the slice we can conclude that the two models have filled in the "empty" space (where we do not have any training observations) in very different ways and according to what we might expect given the model structure.
@@
 
-This example can be reproduced with the code in penguins_shifting_slice_center.nb and a run through is shown in the second part of the video at https://vimeo.com/747590472.
+This example can be reproduced with the code in penguins_shifting_slice_center.nb and a run-through is shown in the second part of the video at https://vimeo.com/747590472.
 
@@
 This short note has described new technology for manually interacting with tours.
-The manual control of projection coefficients is most useful for assessing variable importance to perceived structure, but can be generally used for steering a viewer through high-dimensional space.
+The manual control of projection coefficients is most useful for assessing variable importance to the perceived structure but can be generally used for steering a viewer through high-dimensional space.
 Changing the center of a slice manually enables exploring the space orthogonal to a projection, specifically in the direction of a single variable.
-This is a new tool, and as shown in the example, when used with the slice tour can be very useful for understanding boundaries of classifiers.
+This is a new tool, and as shown in the example, when used with the slice tour can be very useful for understanding the boundaries of classifiers.
 This is also likely a useful method for examining low-dimensional non-linear manifolds, and also functions of multiple parameters.
@@
 
-Where else might it be useful to use mathematica
+Where else might it be useful to use Mathematica
 
@@
 The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
-The scatterplot matrix of the penguins data was produced by the GGally [@GGally] package built on ggplot2 [@ggplot2] graphics.
+The scatterplot matrix of the penguins' data was produced by the GGally [@GGally] package built on ggplot2 [@ggplot2] graphics.
 We thank the Institute of Statistics, BOKU, for their hospitality while part of this work was conducted.

### 2023-01-27 paper/paper.tex
grammarly changes

@@
 for exploring regions and boundaries produced by classification models.
-The advantage of Mathematica is its linear algebra capabilities, and
+The advantage of Mathematica is its linear algebra capabilities and
 interactive cursor location controls. Some limited implementation has
@@
 data \emph{from all sides}. (Various videos of tours, e.g.
-\citet{dataviewer}, \citet{gt-pp-video}, on the \citet{ASA22} illustrate
-this concept.) Prior to Asimov's work, there were numerous preparatory
-developments including \citet{tukey}'s PRIM-9. PRIM-9 had
+\citet{dataviewer}, \citet{gt-pp-video}, on the \citet{ASA22},
+illustrate this concept.) Prior to Asimov's work, there were numerous
+preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
 user-controlled rotations on coordinate axes, allowing one to manually
@@
 video library of \citet{ASA22}.) Steering through all possible
-projections is impossible, unlike Asimov's tours which allows one to
-quickly see many, many different projections. After Asimov there have
+projections is impossible, unlike Asimov's tours which allow one to
+quickly see many, many different projections. After Asimov, there have
 been many tour developments, which are summarized in \citet{lee2021}.
@@
 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space,
-allowing the user to select any variable axis, and rotate it into, or
-out of, or around the projection through horizontal, vertical, oblique,
+allowing the user to select any variable axis, and rotate it into, out
+of, or around the projection through horizontal, vertical, oblique,
 radial or angular changes in value. \citet{spyrison_spinifex_2020}
@@
 
-Manual controls are especially useful for assessing sensitivity of
+Manual controls are especially useful for assessing the sensitivity of
 structure to particular elements of the projection. There are many
@@
 interpreting models, one can reduce or increase a variable's
-contribution to examine the variable importance. Having the user
+contribution, to examine the variable importance. Having the user
 interact with a projection is extremely valuable for understanding
@@
 the pre-processing of creating a manipulation space overly complicates
-the algorithm, (2) extending to higher dimensional control is difficult.
+the algorithm, and (2) extending to higher dimensional control is
+difficult.
 
-Another potentially useful manual control, is to allow the user to
-choose the position of the center of a slice. The slice tour was
-introduced in \citet{slicetour}. It operates by converting the
-projection plane into a slice, by removing or de-emphasizing points that
-are further than a fixed orthogonal distance from the plane. The
-projection plane is usually thought of as passing through the center of
-the data. Manual control would allow the user to change the position of
-the center point, by shifting it along a coordinate axis, while keeping
-the orientation of the projection plane fixed. The purpose would be to
-explore how or if the shape of the data, in the space orthogonal to the
-projection, changes as one gets away from the center. It would also
-allow the user to interactively decide on the thickness of the slice.
+Another potentially useful manual control is to allow the user to choose
+the position of the center of a slice. The slice tour was introduced in
+\citet{slicetour}. It operates by converting the projection plane into a
+slice, by removing or de-emphasizing points that are further than a
+fixed orthogonal distance from the plane. The projection plane is
+usually thought of as passing through the center of the data. Manual
+control would allow the user to change the position of the center point,
+by shifting it along a coordinate axis, while keeping the orientation of
+the projection plane fixed. The purpose would be to explore how or if
+the shape of the data, in the space orthogonal to the projection,
+changes as one gets away from the center. It would also allow the user
+to interactively decide on the thickness of the slice.
 
@@
 The software section describes a Mathematica \citep{Mathematica} package
-that is used for the application, and describes the interactive
+that is used for the application and describes the interactive
 environment that would be desirable within R \citep{rref} as new
@@
 displayed 2D projection. The algorithm also provided constraints to
-horizontal, vertical, radial or angular motions only. The construction
+horizontal, vertical, radial, or angular motions only. The construction
 of the manipulation space overly complicates the manual controls,
@@
 
-The new approach emerged from experiments on the tour using the linear
-algebra capabilities, and relatively new interactive graphics interface,
-available in Mathematica. The components corresponding to \(V_m\) are
-directly controlled by cursor movement, which updates row \(m\) of
-\(A\). The updated matrix is then orthonormalised.
+The new approach emerged from experiments on the tour using linear
+algebra capabilities, and a relatively new interactive graphics
+interface, available in Mathematica. The components corresponding to
+\(V_m\) are directly controlled by cursor movement, which updates row
+\(m\) of \(A\). The updated matrix is then orthonormalized.
 
@@
 \(A\) can be seen to roughly follow a specified position (dot). Changes
-in the other components happen as a result of the orthonormalisation,
-but are uncontrolled.
+in the other components happen as a result of the orthonormalization but
+are uncontrolled.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (dark) is changed using unconstrained orthonormalisation.
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization.
 The dot indicates the chosen values for the controlled variable, $V_m$.
@@
 The problem with the new simple method is that it is not faithful to the
-precise values for \(V_m\) because the orthonormalisation will change
+precise values for \(V_m\) because the orthonormalization will change
 them. Even though these changes are for the most part imperceptible, one
@@
 enforced, a few are detailed in the Appendix. These primarily differ in
-how the remaining variables are adjusted during orthonormalisation.
+how the remaining variables are adjusted during orthonormalization.
 
@@
 data that is defined by a projection, a center point that is anchoring
-it in the high-dimensional space and the slice thickness \(h\)
+it in the high-dimensional space, and the slice thickness \(h\)
 \citep{slicetour}. A data point is inside the slice if its orthogonal
@@
 
-In addition it is also useful to interactively change the slice
-thickness \(h\) (also called the slice radius), in particular to find
+In addition, it is also useful to interactively change the slice
+thickness \(h\) (also called the slice radius), in particular, to find
 the preferred value for exploring the input data. For guidance the
@@
 position of that axis by dragging it. In practice, the closest axis will
-be selected, and the dragging results in tracking of the mouse position
+be selected, and the dragging results in tracking the mouse position
 which will update the projection in small steps, such that no further
@@
 \(c\) explicitly. The display will update based on these input values,
-in particular the view will jump to a new slice when \(c\) is changed.
+in particular, the view will jump to a new slice when \(c\) is changed.
 
@@
 \texttt{shiny} \citep{shinym} with \texttt{plotly} \citep{plotly}
-drawing. However, the \texttt{plotly} package is designed to add on to
+drawing. However, the \texttt{plotly} package is designed to add to
 \texttt{ggplot2} graphics, not to be a programmable interactive graphics
-system. It lacks capability to implement the primary tool needed for a
-manual tour, that is the tracking of the cursor position in small
+system. It lacks the capability to implement the primary tool needed for
+a manual tour, which is the tracking of the cursor position in small
 increments. In theory, the historical \texttt{locator()} function of
-base R should provide the necessary capability, but it's behavior is not
+base R should provide the necessary capability, but its behavior is not
 consistent between graphical devices. The cursor position control is
@@
 chosen to do is to use the interactivity of Mathematica to experiment,
-and to implement as an animation in the \texttt{tourr} \citep{tourr}
-package. There is hope that better interactivity will be available for R
-soon, perhaps in javascript \citep[e.g.][\citet{detourr}]{langevitour},
-which might allow for the experimentation to develop new techniques.
+and to implement the resulting ideas as an animation in the
+\texttt{tourr} \citep{tourr} package. There is hope that better
+interactivity will be available for R soon, perhaps in javascript
+\citep[e.g.][\citet{detourr}]{langevitour}, which might allow for the
+experimentation to develop new techniques.
 
@@
 
-Changing slice center manually can be accomplished with the new
+Changing the slice center manually can be accomplished with the new
 function, \texttt{manual\_slice()}. This changes the value of the center
@@
 
-\caption{Scatterplot matrix of the (standardised) penguins data.
-The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
+\caption{Scatterplot matrix of the (standardized) penguins data.
+The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
 \end{figure}
@@
 (\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
-measuring the size of the penguins from three species (Adelie, Chinstrap
-and Gentoo). The scatterplot matrix shows that the three species appear
-to be likely separable, and that at least the Gentoo can be
-distinguished from the other two species when \texttt{bd} is paired with
-\texttt{fl} or \texttt{bm}. The steps for exploring boundaries in this
-example are as follows:
+measuring the size of the penguins from three species (Adelie,
+Chinstrap, and Gentoo). The scatterplot matrix shows that the three
+species appear to be likely separable and that at least the Gentoo can
+be distinguished from the other two species when \texttt{bd} is paired
+with \texttt{fl} or \texttt{bm}. The steps for exploring boundaries in
+this example are as follows:
 
@@
 particular projection of the original data shows very distinct clusters
-of the three species. That means, the obscuring of the boundaries
-between groups for both of the models is driven by what is happening in
-the orthogonal space to the plane of the selected projection.
+of the three species. That means the obscuring of the boundaries between
+groups for both of the models is driven by what is happening in the
+orthogonal space to the plane of the selected projection.
 
@@
 This example can be reproduced with the code in
-penguins\_exploring\_manually.nb and a run through is shown in the video
+penguins\_exploring\_manually.nb and a run-through is shown in the video
 at \url{https://vimeo.com/747585410}.
@@
 projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
-through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first we explore
+through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first, we explore
 how changing the projection away from \(A_1\) can help with
@@
 However, it is straight in a combination of the second and third
-variable (\texttt{bd} and \texttt{fl}). By examining the scatterplot
+variables (\texttt{bd} and \texttt{fl}). By examining the scatterplot
 matrix, we can understand how this boundary was built: on each of
-\texttt{bd} and \texttt{fl} it is possible to cut on a value where most
+\texttt{bd} and \texttt{fl}, it is possible to cut on a value where most
 of the Gentoo penguins are different from the other two species for any
@@
 
-In contrast, the LDA boundary has an oblique split between the two, and
+In contrast, the LDA boundary has an oblique split between the two and
 roughly divides the space into three similarly sized areas. It is almost
@@
 where by assuming the clusters have equal variance-covariance, it places
-a boundary half-way between the group means.
+a boundary halfway between the group means.
 
@@
 Figure \ref{slice1}). Gentoo is more separated from the other two in
-this projection, and one can imagine that the trees in the forest has
+this projection, and one can imagine that the trees in the forest have
 greedily grasped any one of many places to make a split to separate the
@@
 This example can be reproduced with the code in
-\texttt{penguins\_slicing\_through\_center.nb} and a run through is
+\texttt{penguins\_slicing\_through\_center.nb} and a run-through is
 shown in the first part of the video at
@@
 
-We start by a comparison of the two models and the data distribution
+We start with a comparison of the two models and the data distribution
 shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
-values of bill depth in top row of Figure \ref{slice1p}.
+values of bill depth in the top row of Figure \ref{slice1p}.
 \textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
@@
 We can see that all three slices (the two models and the data) contain
-almost no points from the third class (Gentoo), and that the decision
+almost no points from the third class (Gentoo) and that the decision
 boundary between the two models is very similar.
@@
 The RF model (left) predicts all three species within this slice, with
-an interesting boundary for the third class (Gentoo). On the other hand
+an interesting boundary for the third class (Gentoo). On the other hand,
 the LDA model (middle) predominantly predicts the third class within the
@@
 This example can be reproduced with the code in
-penguins\_shifting\_slice\_center.nb and a run through is shown in the
+penguins\_shifting\_slice\_center.nb and a run-through is shown in the
 second part of the video at \url{https://vimeo.com/747590472}.
@@
 with tours. The manual control of projection coefficients is most useful
-for assessing variable importance to perceived structure, but can be
+for assessing variable importance to the perceived structure but can be
 generally used for steering a viewer through high-dimensional space.
@@
 variable. This is a new tool, and as shown in the example, when used
-with the slice tour can be very useful for understanding boundaries of
-classifiers. This is also likely a useful method for examining
+with the slice tour can be very useful for understanding the boundaries
+of classifiers. This is also likely a useful method for examining
 low-dimensional non-linear manifolds, and also functions of multiple
@@
 \citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
-matrix of the penguins data was produced by the GGally \citep{GGally}
+matrix of the penguins' data was produced by the GGally \citep{GGally}
 package built on ggplot2 \citep{ggplot2} graphics. We thank the

### 2023-01-30 supp/R/example/penguins.R
Suggested structure for supplementary material

@@ -1,0 +1,53 @@
+library(tidyverse)
+library(palmerpenguins)
+library(classifly)
+library(MASS)
+library(tourr)
+library(randomForest)
+library(here)
+
+penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
+  drop_na() %>%
+  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x)))
+
+
+penguins_lda <- classifly(penguins_data, species ~ . , lda)
+penguins_lda[,1:4] <- apply(penguins_lda[,1:4], 2,  function(x) (x-mean(x))/sd(x))
+lda_samples <- penguins_lda %>%
+  as_tibble() %>%
+  filter(.TYPE == "simulated") %>%
+  mutate(species_flag = as.numeric(as.factor(species)))
+
+animate_slice(lda_samples[,1:4], col = lda_samples$species, v_rel = 0.01)
+
+write_csv(lda_samples[, c(1:4,6,12)], file = here("../../data/penguins_lda.csv"))
+
+set.seed(21)
+penguins_rf <- classifly(penguins_data, species ~ . , randomForest)
+penguins_rf[,1:4] <- apply(penguins_rf[,1:4], 2,  function(x) (x-mean(x))/sd(x))
+rf_samples <- penguins_rf %>%
+  as_tibble() %>%
+  filter(.TYPE == "simulated") %>%
+  mutate(species_flag = as.numeric(as.factor(species)))
+
+animate_slice(rf_samples[,1:4], col = rf_samples$species, v_rel = 0.01)
+
+write_csv(rf_samples[, c(1:4,6, 12)], file = here("../../data/penguins_rf.csv"))
+
+# Plots for paper
+library(GGally)
+penguins_rf <- read_csv(here("../../data/penguins_rf.csv"))
+penguins_lda <- read_csv(here("../../data/penguins_lda.csv"))
+ggscatmat(penguins_data, columns = 2:5, col="species")
+ggscatmat(penguins_rf, columns = 1:4, col="species", alpha=0.1)
+ggscatmat(penguins_lda, columns = 1:4, col="species", alpha=0.1)
+
+
+# saving the data to look at in mathematica
+penguins_data %>%
+  dplyr::select(c(2,3,4,5,1)) %>%
+  mutate(species_flag =
+           as.numeric(factor(species,
+                             levels = # making sure we have the same flags
+                               levels(as.factor(rf_samples$species))))) %>%
+  write_csv(file = here("../../data/penguins.csv"))

### 2023-01-30 supp/R/paper/appendix.R

@@ -1,0 +1,102 @@
+## ----setup, include=FALSE--------------------------------------------------------------------
+knitr::opts_chunk$set(
+  echo = FALSE, 
+  message = FALSE,
+  warning = FALSE)
+
+
+## ----libraries-------------------------------------------------------------------------------
+# Load libraries
+library(tourr)
+library(plotrix)
+library(here)
+
+
+## ----othermethod, out.width="100%", width=12, height=4, fig.cap="Manual controls with algorithm 2. The precise location of the axis is maintained."----
+source(here("../src/plot_basis.R"))
+source(here("../src/linear_alg.R"))
+
+# Base plot set up
+plot.new()
+par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0),
+    mfrow=c(1,4))
+
+# Create manual tour example
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+  
+# Now set a new position
+vchange <- 3
+eps <- 0.1
+plot_basis(A, vchange, vcol="black")
+text(A[vchange, 1], A[vchange, 2], expression(V[m]), pos=3, col="orange")
+
+# j is the component that should be kept fixed
+orthonormalize_rotated <- function(A, j){
+  m1 <- A[j, 1]
+  m2 <- A[j, 2]
+  c <- 1 / (sqrt(1 + (m2/m1)^2))
+  s <- (m2/m1) / (sqrt(1 + (m2/m1)^2))
+  rot <- matrix(c(c, -s, s, c), nrow = 2)
+  rot_inv <- matrix(c(c, s, -s, c), nrow = 2)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot %*% A[i,]
+  }
+  A[,1] <- normalize_w_fixed(A[,1], j)
+  A <- tourr::orthonormalise(A)
+  for(i in 1:nrow(A)){
+    A[i,] <- rot_inv %*% A[i,]
+    }
+  A
+}
+
+# normalize a vector while keeping component j fixed
+normalize_w_fixed <- function(v, j){
+  c <- v[j]
+  k <- sqrt((1-c^2) / sum(v[-j]^2))
[…]

### 2023-01-30 supp/R/paper/paper.R

@@ -1,0 +1,134 @@
+## ----setup, include=FALSE--------------------------------------------------------------------
+knitr::opts_chunk$set(
+  echo = FALSE, 
+  message = FALSE,
+  warning = FALSE)
+
+
+## ----libraries-------------------------------------------------------------------------------
+# Load libraries
+library(tourr)
+library(plotrix)
+
+
+## ----manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization. The dot indicates the chosen values for the controlled variable, $V_m$. It can be seen that the actual axis does not precisely match the chosen position, but it is close."----
+source("../src/plot_basis.R")
+source("../src/linear_alg.R")
+
+# Base plot set up
+plot.new()
+par(pty="s", xaxt="n", yaxt="n", bty="n",
+      omi=c(0,0,0,0), mar=c(0,0,0,0),
+    mfrow=c(2,4))
+
+# Create manual tour example
+p <- 4
+d <- 2
+set.seed(24)
+A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
+colnames(A) <- c("P1", "P2")
+rownames(A) <- paste0("V", 1:p)
+A <- tourr::orthonormalise(A)
+  
+# Now set a new position
+vchange <- 3
+eps <- 0.1
+plot_basis(A, vchange, vcol="black")
+text(A[vchange, 1], A[vchange, 2], expression(V[m]), pos=3, col="orange")
+points(A[vchange, 1], A[vchange, 2], pch=16, col="orange")
+
+# Iterate and save images
+# Unconstrained orthonormalisation
+Anew <- A
+for (i in 1:3) {
+  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+  exact <- Anew[vchange,]
+  Anew <- tourr::orthonormalise(Anew)
+  plot_basis(Anew, vchange, vcol="black")
+  text(Anew[vchange, 1], Anew[vchange, 2], expression(V[m]), pos=3, col="orange")
+  points(exact[1], exact[2], pch=16, col="orange")
+}
+
+# Iterate and save images
+# Constrained orthonormalisation
+#Anew <- A
+#plot_basis(A, vchange, vcol="black")
+#for (i in 1:3) {
+#  Anew[vchange,] <- A[vchange,]*(1-i*eps)
+#  exact <- Anew[vchange,]
[…]

### 2023-01-30 supp/R/src/manual.R

@@ -1,0 +1,37 @@
+
+
+# Plot basis
+#png("mtour_plots/mtour0.png", width=250, height=250)
+#plot_basis(A)
+#dev.off()
+
+#A[vchange,] <- A[vchange,]*(1-eps)
+#A <- tourr::orthonormalise(A)
+#plot_basis(A)
+
+
+
+# Utility function to plot a basis
+# Need to generalise to arbitrary pxd
+plot_basis <- function(A, v=NULL) {
+  # Set plot canvas
+  plot(c(-1.1, 1.1), c(-1.1, 1.1),
+       type="n", xlab="", ylab="", axes=FALSE)
+
+  # draw circle
+  plotrix::draw.circle(0,0,1)
+
+  # label axes
+  text(A[,1], A[,2], rownames(A), col="grey50")
+  if (!is.null(v))
+    text(A[v,1], A[v,2], rownames(A), col="orange")
+
+
+  # draw axes
+  A <- cbind(A, xctr=rep(0, nrow(A)), yctr=rep(0, nrow(A)))
+  segments(A[,3], A[,4], A[,1], A[,2], lwd=3)
+  if (!is.null(v))
+    segments(A[v,3], A[v,4], A[v,1], A[v,2],
+             col="orange", lwd=3)
+
+}

### 2023-01-31 supp/R/paper/paper.R
readme and project files for the supp folder

@@ -1,6 +1,6 @@
 ## ----setup, include=FALSE--------------------------------------------------------------------
 knitr::opts_chunk$set(
-  echo = FALSE, 
+  echo = FALSE,
   message = FALSE,
   warning = FALSE)
 
@@ -9,11 +9,11 @@ knitr::opts_chunk$set(
 # Load libraries
 library(tourr)
 library(plotrix)
-
+library(here)
 
 ## ----manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization. The dot indicates the chosen values for the controlled variable, $V_m$. It can be seen that the actual axis does not precisely match the chosen position, but it is close."----
-source("../src/plot_basis.R")
-source("../src/linear_alg.R")
+source(here("../src/plot_basis.R"))
+source(here("../src/linear_alg.R"))
 
 # Base plot set up
 plot.new()
@@ -29,7 +29,7 @@ A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
 colnames(A) <- c("P1", "P2")
 rownames(A) <- paste0("V", 1:p)
 A <- tourr::orthonormalise(A)
-  
+
 # Now set a new position
 vchange <- 3
 eps <- 0.1
@@ -74,7 +74,7 @@ penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
          "fl"="flipper_length_mm",
          "bm"="body_mass_g")
 #ggscatmat(penguins_data, columns = 2:5, col="species") +
-# library(paletteer) 
+# library(paletteer)
 # rcartocolor::ag_Sunset
 # rcartocolor::ag_GrnYl
 # RColorBrewer::Spectral
@@ -84,49 +84,49 @@ penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
 clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
 penguins_diag <- tibble(x=1, y=c(3,2,1),
      species=levels(penguins_data$species))
-pscat <- ggpairs(penguins_data, columns = 2:5, 
-          mapping = ggplot2::aes(color = species, 
+pscat <- ggpairs(penguins_data, columns = 2:5,
+          mapping = ggplot2::aes(color = species,
                                  fill = species),
           upper = list(continuous = "points"),
[…]

### 2023-02-01 supp/R/example/penguins.R
commented out the write lines in R code

@@ -20,7 +20,8 @@ lda_samples <- penguins_lda %>%
 
 animate_slice(lda_samples[,1:4], col = lda_samples$species, v_rel = 0.01)
 
-write_csv(lda_samples[, c(1:4,6,12)], file = here("../../data/penguins_lda.csv"))
+# Write the result, if necessary
+# write_csv(lda_samples[, c(1:4,6,12)], file = here("../../data/penguins_lda.csv"))
 
 set.seed(21)
 penguins_rf <- classifly(penguins_data, species ~ . , randomForest)
@@ -32,7 +33,8 @@ rf_samples <- penguins_rf %>%
 
 animate_slice(rf_samples[,1:4], col = rf_samples$species, v_rel = 0.01)
 
-write_csv(rf_samples[, c(1:4,6, 12)], file = here("../../data/penguins_rf.csv"))
+# Write the result, if necessary
+# write_csv(rf_samples[, c(1:4,6, 12)], file = here("../../data/penguins_rf.csv"))
 
 # Plots for paper
 library(GGally)
@@ -44,10 +46,10 @@ ggscatmat(penguins_lda, columns = 1:4, col="species", alpha=0.1)
 
 
 # saving the data to look at in mathematica
-penguins_data %>%
-  dplyr::select(c(2,3,4,5,1)) %>%
-  mutate(species_flag =
-           as.numeric(factor(species,
-                             levels = # making sure we have the same flags
-                               levels(as.factor(rf_samples$species))))) %>%
-  write_csv(file = here("../../data/penguins.csv"))
+# penguins_data %>%
+#   dplyr::select(c(2,3,4,5,1)) %>%
+#   mutate(species_flag =
+#            as.numeric(factor(species,
+#                              levels = # making sure we have the same flags
+#                                levels(as.factor(rf_samples$species))))) %>%
+# write_csv(file = here("../../data/penguins.csv"))

### 2023-02-01 supp/R/paper/paper.R
cleaned up paper.R

@@ -49,19 +49,6 @@ for (i in 1:3) {
   points(exact[1], exact[2], pch=16, col="orange")
 }
 
-# Iterate and save images
-# Constrained orthonormalisation
-#Anew <- A
-#plot_basis(A, vchange, vcol="black")
-#for (i in 1:3) {
-#  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-#  exact <- Anew[vchange,]
-#  Anew <- orthonormalise_frozen(Anew, vchange)
-#  plot_basis(Anew, vchange, vcol="black")
-#  points(exact[1], exact[2], pch=16, col="orange")
-#}
-
-
 ## ----penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data. The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."----
 library(tidyverse)
 library(palmerpenguins)
@@ -73,14 +60,6 @@ penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
          "bd"="bill_depth_mm",
          "fl"="flipper_length_mm",
          "bm"="body_mass_g")
-#ggscatmat(penguins_data, columns = 2:5, col="species") +
-# library(paletteer)
-# rcartocolor::ag_Sunset
-# rcartocolor::ag_GrnYl
-# RColorBrewer::Spectral
-# clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
-# clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
-# clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
 clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
 penguins_diag <- tibble(x=1, y=c(3,2,1),
      species=levels(penguins_data$species))
@@ -89,8 +68,6 @@ pscat <- ggpairs(penguins_data, columns = 2:5,
                                  fill = species),
           upper = list(continuous = "points"),
           diag = list(continuous = "densityDiag")) +
-#  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
[…]

### 2023-02-02 paper/paper.Rmd
cleaned up paper.Rmd

@@
 }
-
-# Iterate and save images
-# Constrained orthonormalisation
-#Anew <- A
-#plot_basis(A, vchange, vcol="black")
-#for (i in 1:3) {
-#  Anew[vchange,] <- A[vchange,]*(1-i*eps)
-#  exact <- Anew[vchange,]
-#  Anew <- orthonormalise_frozen(Anew, vchange)
-#  plot_basis(Anew, vchange, vcol="black")
-#  points(exact[1], exact[2], pch=16, col="orange")
-#}
 ```
@@
          "bm"="body_mass_g")
-#ggscatmat(penguins_data, columns = 2:5, col="species") +
-# library(paletteer) 
-# rcartocolor::ag_Sunset
-# rcartocolor::ag_GrnYl
-# RColorBrewer::Spectral
-# clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
-# clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
-# clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
 clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
@@
           diag = list(continuous = "densityDiag")) +
-#  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
-#  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
   scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +

### 2023-02-07 review/response.Rmd
note to check added
Getting ready to re-submit

@@
 
- XXX References need to be consistent, all initials or all full names, and all sentence case or title case on titles.
+ XXX References need to be consistent, all initials or all full names, and all sentence case or title case on titles. - Done
+
+ XXX Check that slicing explanation mentions scaling the data! - where should this go?
 
@@
     - Data files should be uploaded along with code files (they don't seem to be part of supplemental material?)
-    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat diﬀicult to assess the software side of this paper.
+    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat difficult to assess the software side of this paper.
 
@@
 XXX TODO Please check the list of Supplementary material in the paper.}
+\textcolor{red}{I think it is complete, but maybe can be structured better, to discuss}
 
@@
 
-\textcolor{blue}{Done. They were all cited, but not by our error, at the first mention.}
+\textcolor{blue}{Done. They were all cited, but not, by our error, at the first mention.}
 
@@
 \textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny/rgl.}  
+\textcolor{red}{Maybe have a single reply to all these points? It seems the blue text color in the paper is not there for all changes (paragraph on shiny is fully new?).}
 

### 2023-02-07 paper/appendix.Rmd
Getting ready to re-submit

@@
 
-For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}.
+For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.
+\ref{matrices}.
 Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$.

### 2023-02-07 paper/appendix.tex
Getting ready to re-submit

@@
 of \(O\) (an orthonormal basis for the p-dimensional space), formed by
-its first two columns as illustrated in
-Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
-for the p-dimensional space is given by
+its first two columns as illustrated in Eq. \ref{matrices}. Whereas
+orthonormality of the basis for the p-dimensional space is given by
 \(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the

### 2023-02-07 paper/paper.Rmd
Getting ready to re-submit

@@
 In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
-(Various videos of tours, e.g.
-@dataviewer, @gt-pp-video, on the @ASA22, illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
+\textcolor{blue}{(Various videos of tours, e.g.} @dataviewer, @gt-pp-video, on the @ASA22, \textcolor{blue}{illustrate this concept.)} Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
@@
 1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
-2. Change values in row $m$, for example, if $d=2$ gives 
+2. Change values in row $m$, for example, if $d=2$ 
 $$
@@
     ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1$.
+
+\textcolor{red}{a2 should be normalised after the second step? Fig 1: this is now showing tracking, with the yellow dot moving in each image?}
 
@@
 \textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing.
-However, the `plotly` package is designed to add to `ggplot2` graphics, not to be a programmable interactive graphics system.
+However, the `plotly` package is designed to add to `ggplot2` [@ggplot2] graphics, not to be a programmable interactive graphics system.
 It lacks the capability to implement the primary tool needed for a manual tour, which is the tracking of the cursor position in small increments.
@@
 @langevitour, @detourr], which might allow for the experimentation to develop new techniques.
+
+\textcolor{red}{plotly does 3D graphs that can be dragged as well, maybe need to clarify further? Are there any references that document shortcomings? Maybe this paragraph should be a subsection?}
 
@@
 The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice center.
+\textcolor{red}{insert sentence about new slice position display here.}
 The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided.
@@
 ```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data.
-The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."}
+The three species are reasonably different in size, with Gentoo distinguished from the other two on bill depth relative to flipper length and body mass."}
 library(tidyverse)
@@
 We start with a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in the top row of Figure \ref{slice1p}.
-\textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+\textcolor{blue}{The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
 This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.
@@
 Mathematica provided a useful sandbox to experiment with the ideas presented in this paper. 
-Most of the data, though, is first constructed using R, especially the classification boundaries.
+Most of the data, though, is first constructed using R, especially the classification boundaries. \textcolor{red}{what should this say?}
 If new technology for interactive graphics with precise cursor control becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.
@@
 - Appendix with additional details about manual controls, and Mathematica functions.
-- The `Rmd` file that generated the paper, containing the R code to produce Figure 1.
 - Data files for the examples.
+- R code to reproduce the example data, Figures 1 and 2 in the paper and Figures 1 and 2 in the appendix.
 - Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2023-02-07 paper/paper.tex
Getting ready to re-submit

@@
 algorithms for tour paths that could theoretically show the viewer the
-data \emph{from all sides}. (Various videos of tours, e.g.
-\citet{dataviewer}, \citet{gt-pp-video}, on the \citet{ASA22},
-illustrate this concept.) Prior to Asimov's work, there were numerous
-preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
-user-controlled rotations on coordinate axes, allowing one to manually
-tour through low-dimensional projections. (A video \citep{PRIM9-video}
-illustrating the capabilities is available through \textcolor{blue}{the}
-video library of \citet{ASA22}.) Steering through all possible
-projections is impossible, unlike Asimov's tours which allow one to
-quickly see many, many different projections. After Asimov, there have
-been many tour developments, which are summarized in \citet{lee2021}.
+data \emph{from all sides}.
+\textcolor{blue}{(Various videos of tours, e.g.} \citet{dataviewer},
+\citet{gt-pp-video}, on the \citet{ASA22},
+\textcolor{blue}{illustrate this concept.)} Prior to Asimov's work,
+there were numerous preparatory developments including \citet{tukey}'s
+PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes,
+allowing one to manually tour through low-dimensional projections. (A
+video \citep{PRIM9-video} illustrating the capabilities is available
+through \textcolor{blue}{the} video library of \citet{ASA22}.) Steering
+through all possible projections is impossible, unlike Asimov's tours
+which allow one to quickly see many, many different projections. After
+Asimov, there have been many tour developments, which are summarized in
+\citet{lee2021}.
 
@@
 \item
-  Change values in row \(m\), for example, if \(d=2\) gives \[
+  Change values in row \(m\), for example, if \(d=2\) \[
   A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
@@
 \end{enumerate}
+
+\textcolor{red}{a2 should be normalised after the second step? Fig 1: this is now showing tracking, with the yellow dot moving in each image?}
 
@@
 drawing. However, the \texttt{plotly} package is designed to add to
-\texttt{ggplot2} graphics, not to be a programmable interactive graphics
-system. It lacks the capability to implement the primary tool needed for
-a manual tour, which is the tracking of the cursor position in small
-increments. In theory, the historical \texttt{locator()} function of
-base R should provide the necessary capability, but its behavior is not
-consistent between graphical devices. The cursor position control is
-possible with \texttt{rgl} \citep{rgl} but this package is focused on
-3D, not the arbitrary dimension of tours. The level of interactivity
-would have been possible with the \texttt{cranvas} system
-\citep{cranvas} or \texttt{rJava} \citep{rJava} but neither has a
-graphics toolkit that currently operates across platforms. What we have
-chosen to do is to use the interactivity of Mathematica to experiment,
-and to implement the resulting ideas as an animation in the
-\texttt{tourr} \citep{tourr} package. There is hope that better
-interactivity will be available for R soon, perhaps in javascript
-\citep[e.g.][\citet{detourr}]{langevitour}, which might allow for the
-experimentation to develop new techniques.
+\texttt{ggplot2} \citep{ggplot2} graphics, not to be a programmable
+interactive graphics system. It lacks the capability to implement the
+primary tool needed for a manual tour, which is the tracking of the
+cursor position in small increments. In theory, the historical
+\texttt{locator()} function of base R should provide the necessary
+capability, but its behavior is not consistent between graphical
+devices. The cursor position control is possible with \texttt{rgl}
+\citep{rgl} but this package is focused on 3D, not the arbitrary
+dimension of tours. The level of interactivity would have been possible
+with the \texttt{cranvas} system \citep{cranvas} or \texttt{rJava}
+\citep{rJava} but neither has a graphics toolkit that currently operates
+across platforms. What we have chosen to do is to use the interactivity
+of Mathematica to experiment, and to implement the resulting ideas as an
+animation in the \texttt{tourr} \citep{tourr} package. There is hope
+that better interactivity will be available for R soon, perhaps in
+javascript \citep[e.g.][\citet{detourr}]{langevitour}, which might allow
+for the experimentation to develop new techniques.
+
+\textcolor{red}{plotly does 3D graphs that can be dragged as well, maybe need to clarify further? Are there any references that document shortcomings? Maybe this paragraph should be a subsection?}
 
@@
 a slider controls the slice thickness; and there is an input field,
-which changes the slice center. The user can also change the appearance
-of the plot by zooming into the center or changing the point size with
-the sliders provided. It is worth noting that this zoom works best with
-data that has been centered and scaled. The projected data can be
-displayed to contrast it with the slice by ticking a box. The explicit
-projection matrix can also be displayed via another checkbox. This is
-especially useful when we wish to import a projection that was
-identified as interesting in the manual exploration into a later study,
-as shown in the applications below. More details about the
-implementation and usage instructions are given in the Appendix.
+which changes the slice center.
+\textcolor{red}{insert sentence about new slice position display here.}
+The user can also change the appearance of the plot by zooming into the
+center or changing the point size with the sliders provided. It is worth
+noting that this zoom works best with data that has been centered and
+scaled. The projected data can be displayed to contrast it with the
+slice by ticking a box. The explicit projection matrix can also be
+displayed via another checkbox. This is especially useful when we wish
+to import a projection that was identified as interesting in the manual
+exploration into a later study, as shown in the applications below. More
+details about the implementation and usage instructions are given in the
+Appendix.
 
@@
 \caption{Scatterplot matrix of the (standardized) penguins data.
-The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
+The three species are reasonably different in size, with Gentoo distinguished from the other two on bill depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
 \end{figure}
@@
 values of bill depth in the top row of Figure \ref{slice1p}.
-\textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+\textcolor{blue}{The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
 This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.}
@@
 presented in this paper. Most of the data, though, is first constructed
-using R, especially the classification boundaries. If new technology for
-interactive graphics with precise cursor control becomes available in R
-it would be useful to create a tighter coupling of models and
-visualization to allow exploring and comparing fits.
+using R, especially the classification boundaries.
+\textcolor{red}{what should this say?} If new technology for interactive
+graphics with precise cursor control becomes available in R it would be
+useful to create a tighter coupling of models and visualization to allow
+exploring and comparing fits.
 
@@
 \item
-  The \texttt{Rmd} file that generated the paper, containing the R code
-  to produce Figure 1.
+  Data files for the examples.
 \item
-  Data files for the examples.
+  R code to reproduce the example data, Figures 1 and 2 in the paper and
+  Figures 1 and 2 in the appendix.
 \item

### 2023-02-08 review/response.Rmd
Working on the response

@@
 ```
-
- XXX If use new slice center control for fig 7, need to make a note of it here, updated the mathematica notebook and check text and captions correct.
-
- XXX References need to be consistent, all initials or all full names, and all sentence case or title case on titles. - Done
-
- XXX Check that slicing explanation mentions scaling the data! - where should this go?
 
@@
 
-\textcolor{blue}{Done - TODO check this}
+\textcolor{blue}{Done}
 

### 2023-02-08 review/response.Rmd
Changes working on review

@@
 
-\textcolor{blue}{Done.
-Previously we provided the public github repo URL so that all files to reproduce could be downloaded.
-Here we have included all the files in the supplementary zip.
-XXX TODO Please check the list of Supplementary material in the paper.}
-\textcolor{red}{I think it is complete, but maybe can be structured better, to discuss}
+\textcolor{blue}{Done. Previously we provided the public github repo URL so that all files to reproduce could be downloaded. Here we have included all the files in the supplementary zip.}
 
@@
 
-\textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny/rgl.}  
-\textcolor{red}{Maybe have a single reply to all these points? It seems the blue text color in the paper is not there for all changes (paragraph on shiny is fully new?).}
+\textcolor{blue}{Revised accordingly.}
 

### 2023-02-08 paper/paper.Rmd
removed an empty line
note on standardiisation
Changes working on review

@@
 with $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$, $\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
+Although not required, manual control of slices works more understandably if the data is first standardized.
 
@@
 
-The source material and animated gifs for this paper are available at [https://github.com/uschiLaa/mmtour](https://github.com/uschiLaa/mmtour).
-
-The supplementary materials include:
+The source material and animated gifs for this paper are available at [https://github.com/uschiLaa/mmtour](https://github.com/uschiLaa/mmtour). The supplementary materials include:
 
 - The Mathematica source code defining the new functions in mmtour.wl.
-- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
+- Two Mathematica notebooks with the code used in the application (corresponding to the three subsections).
 - Appendix with additional details about manual controls, and Mathematica functions.
@@
 - R code to reproduce the example data, Figures 1 and 2 in the paper and Figures 1 and 2 in the appendix.
-- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.
+- Animations illustrating the manual tour and slicing are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2023-02-08 paper/paper.tex
removed an empty line
note on standardiisation
Changes working on review

@@
 and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
-matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\).
+matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\). Although not
+required, manual control of slices works more understandably if the data
+is first standardized.
 
@@
 The source material and animated gifs for this paper are available at
-\url{https://github.com/uschiLaa/mmtour}.
-
-The supplementary materials include:
+\url{https://github.com/uschiLaa/mmtour}. The supplementary materials
+include:
 
@@
 \item
-  Three Mathematica notebooks with the code used in the application
+  Two Mathematica notebooks with the code used in the application
   (corresponding to the three subsections).
@@
 \item
-  Animations illustrating the manual tour and slicing, matching the
-  static figures in the paper. These are also available at
-  \url{https://vimeo.com/747585410} and
+  Animations illustrating the manual tour and slicing are also available
+  at \url{https://vimeo.com/747585410} and
   \url{https://vimeo.com/747590472}.

### 2023-02-08 paper/paper.Rmd
added plotly to rgl part

@@
 In theory, the historical `locator()` function of base R should provide the necessary capability, but its behavior is not consistent between graphical devices.
-The cursor position control is possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary dimension of tours.
+The cursor position control is possible for 3D plots with `rgl` [@rgl] and `plotly` but these not the arbitrary dimension needed for tours.
 The level of interactivity would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that currently operates across platforms.
@@
 @langevitour, @detourr], which might allow for the experimentation to develop new techniques.
-
-\textcolor{red}{plotly does 3D graphs that can be dragged as well, maybe need to clarify further? Are there any references that document shortcomings? Maybe this paragraph should be a subsection?}
 

### 2023-02-08 paper/paper.tex
added plotly to rgl part

@@
 capability, but its behavior is not consistent between graphical
-devices. The cursor position control is possible with \texttt{rgl}
-\citep{rgl} but this package is focused on 3D, not the arbitrary
-dimension of tours. The level of interactivity would have been possible
-with the \texttt{cranvas} system \citep{cranvas} or \texttt{rJava}
-\citep{rJava} but neither has a graphics toolkit that currently operates
-across platforms. What we have chosen to do is to use the interactivity
-of Mathematica to experiment, and to implement the resulting ideas as an
-animation in the \texttt{tourr} \citep{tourr} package. There is hope
-that better interactivity will be available for R soon, perhaps in
-javascript \citep[e.g.][\citet{detourr}]{langevitour}, which might allow
-for the experimentation to develop new techniques.
-
-\textcolor{red}{plotly does 3D graphs that can be dragged as well, maybe need to clarify further? Are there any references that document shortcomings? Maybe this paragraph should be a subsection?}
+devices. The cursor position control is possible for 3D plots with
+\texttt{rgl} \citep{rgl} and \texttt{plotly} but these not the arbitrary
+dimension needed for tours. The level of interactivity would have been
+possible with the \texttt{cranvas} system \citep{cranvas} or
+\texttt{rJava} \citep{rJava} but neither has a graphics toolkit that
+currently operates across platforms. What we have chosen to do is to use
+the interactivity of Mathematica to experiment, and to implement the
+resulting ideas as an animation in the \texttt{tourr} \citep{tourr}
+package. There is hope that better interactivity will be available for R
+soon, perhaps in javascript \citep[e.g.][\citet{detourr}]{langevitour},
+which might allow for the experimentation to develop new techniques.
 

### 2023-02-08 paper/appendix.Rmd
Last few changes

@@
 
-\textcolor{blue}{The slice center position is controlled via numerical input along each axis in the original command `SliceDynamic`.
+The slice center position is controlled via numerical input along each axis in the original command `SliceDynamic`.
 This provides numerical feedback on the position of the center.
-For easier navigation but without the numerical feedback, an implementation of the slice center control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
+For easier navigation but without the numerical feedback, an implementation of the slice center control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.
 

### 2023-02-08 paper/appendix.tex
Last few changes

@@
 
-\textcolor{blue}{The slice center position is controlled via numerical input along each axis in the original command `SliceDynamic`.
-This provides numerical feedback on the position of the center.
-For easier navigation but without the numerical feedback, an implementation of the slice center control via a visual display like the one in Figure \ref{fig:anchornav} is provided in `SliceDynamicCC`.}
+The slice center position is controlled via numerical input along each
+axis in the original command \texttt{SliceDynamic}. This provides
+numerical feedback on the position of the center. For easier navigation
+but without the numerical feedback, an implementation of the slice
+center control via a visual display like the one in Figure
+\ref{fig:anchornav} is provided in \texttt{SliceDynamicCC}.
 

### 2023-02-08 paper/paper.Rmd
Last few changes

@@
 <!-- Background -->
-From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, \textcolor{blue}{methods which work for arbitrary dimensions have more general utility}.
+From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, methods which work for arbitrary dimensions have more general utility.
 A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*.
@@
 In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*.
-\textcolor{blue}{(Various videos of tours, e.g.} @dataviewer, @gt-pp-video, on the @ASA22, \textcolor{blue}{illustrate this concept.)} Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
+(Various videos of tours, e.g.
+@dataviewer, @gt-pp-video, on the @ASA22, illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9.
 PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections.
-(A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allow one to quickly see many, many different projections.
+(A video [@PRIM9-video] illustrating the capabilities is available through the video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allow one to quickly see many, many different projections.
 After Asimov, there have been many tour developments, which are summarized in @lee2021.
@@
 A manual tour allows the user to alter the coefficients of one (or more) variables, of $p$, contributing to a $d$-dimensional projection.
-The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be \textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.} A method to update the values of the component of the controlled variable $V_m$ is then needed.
+The initial ingredients are an orthonormal basis ($A_{p\times d}$) defining the projection of the data, and a variable id ($m \in \{1, ..., p\}$) specifying which coefficient will be changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.
+A method to update the values of the component of the controlled variable $V_m$ is then needed.
 
@@
 3. Orthonormalise $A^*$, using Gram-Schmidt. 
-    i. Normalise $\boldsymbol{a}^*_1$ and $\boldsymbol{a}^*_2$.
+    i. Normalise $\boldsymbol{a}^*_1$.
     ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1$.
-
-\textcolor{red}{a2 should be normalised after the second step? Fig 1: this is now showing tracking, with the yellow dot moving in each image?}
+    iii. Normalise $\boldsymbol{a}^*_2$.
 
@@
 
-```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization.
+```{r manualsequence, out.width="100%", width=12, height=4, fig.cap="Sequence of projections where the contribution of one variable is controlled (dark).
 The dot indicates the chosen values for the controlled variable, $V_m$.
+The change is tracking the chosen value after unconstrained orthonormalization.
 It can be seen that the actual axis does not precisely match the chosen position, but it is close."}
@@
 The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example).
-This display should be able to track the mouse position and adjust the projection based on the \textcolor{blue}{user's cursor position}.
+This display should be able to track the mouse position and adjust the projection based on the user's cursor position.
 The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it.
@@
 
-The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized and \textcolor{blue}{re-displayed}.
+The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized and re-displayed.
 A second, linked display shows the projected data in sync with the updates from the axis display.
@@
 
-\textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing.
+Within R, the natural implementation for an interactive interface would use `shiny` [@shinym] with `plotly` [@plotly] drawing.
 However, the `plotly` package is designed to add to `ggplot2` [@ggplot2] graphics, not to be a programmable interactive graphics system.
@@
 
-\textcolor{blue}{The Mathematica implementation relies on the inbuilt function  `LocatorPane`.
+The Mathematica implementation relies on the inbuilt function  `LocatorPane`.
 This function creates a region on the screen where the position of the cursor is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
 We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
-User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale, and point size.}
+User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale, and point size.
 
@@
 After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualization appears on the right.
-The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice center.
-\textcolor{red}{insert sentence about new slice position display here.}
+The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field or an interactive star glyph, to change the slice center.
 The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided.
@@
 We start with a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in the top row of Figure \ref{slice1p}.
-\textcolor{blue}{The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
-This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.
+The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
+This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.
+We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.
 
@@
 A more interesting comparison is found for $\mathbf{c}^{-}$, thus the slice localized towards low values of bill depth, shown in the bottom row of Figure \ref{slice1p}.
-\textcolor{blue}{(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.)} The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
+(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.) The RF model (left) predicts all three species within this slice, with an interesting boundary for the third class (Gentoo).
 On the other hand, the LDA model (middle) predominantly predicts the third class within the slice, this appears to be enforced through the linear structure of the model.
@@
 Mathematica provided a useful sandbox to experiment with the ideas presented in this paper.
-Most of the data, though, is first constructed using R, especially the classification boundaries. \textcolor{red}{what should this say?}
-If new technology for interactive graphics with precise cursor control becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.
+However, we have used R for the statistical data analysis and modeling in our example, as is the standard approach in the community.
+Since the proposed visualization should ideally be integrated within this workflow, the new methods should also be implemented as interactive graphics accessible from R once new technology with precise cursor control becomes available.
 
@@
 - The Mathematica source code defining the new functions in mmtour.wl.
-- Two Mathematica notebooks with the code used in the application (corresponding to the three subsections).
+- Two Mathematica notebooks with the code used in the application.
 - Appendix with additional details about manual controls, and Mathematica functions.

### 2023-02-08 paper/paper.tex
Last few changes

@@
 From a statistical perspective 3D is a rare data dimension, so unlike in
-most 3D rotation computer graphics applications,
-\textcolor{blue}{methods which work for arbitrary dimensions have more general utility}.
-A good approach is to show projections from an arbitrary dimensional
-space to create dynamic data visualizations called \emph{tours}. Tours
-involve views of high-dimensional data with low-dimensional projections.
-In his original paper on the grand tour, \citet{As85} provided several
-algorithms for tour paths that could theoretically show the viewer the
-data \emph{from all sides}.
-\textcolor{blue}{(Various videos of tours, e.g.} \citet{dataviewer},
-\citet{gt-pp-video}, on the \citet{ASA22},
-\textcolor{blue}{illustrate this concept.)} Prior to Asimov's work,
-there were numerous preparatory developments including \citet{tukey}'s
-PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes,
-allowing one to manually tour through low-dimensional projections. (A
-video \citep{PRIM9-video} illustrating the capabilities is available
-through \textcolor{blue}{the} video library of \citet{ASA22}.) Steering
-through all possible projections is impossible, unlike Asimov's tours
-which allow one to quickly see many, many different projections. After
-Asimov, there have been many tour developments, which are summarized in
-\citet{lee2021}.
+most 3D rotation computer graphics applications, methods which work for
+arbitrary dimensions have more general utility. A good approach is to
+show projections from an arbitrary dimensional space to create dynamic
+data visualizations called \emph{tours}. Tours involve views of
+high-dimensional data with low-dimensional projections. In his original
+paper on the grand tour, \citet{As85} provided several algorithms for
+tour paths that could theoretically show the viewer the data \emph{from
+all sides}. (Various videos of tours, e.g. \citet{dataviewer},
+\citet{gt-pp-video}, on the \citet{ASA22}, illustrate this concept.)
+Prior to Asimov's work, there were numerous preparatory developments
+including \citet{tukey}'s PRIM-9. PRIM-9 had user-controlled rotations
+on coordinate axes, allowing one to manually tour through
+low-dimensional projections. (A video \citep{PRIM9-video} illustrating
+the capabilities is available through the video library of
+\citet{ASA22}.) Steering through all possible projections is impossible,
+unlike Asimov's tours which allow one to quickly see many, many
+different projections. After Asimov, there have been many tour
+developments, which are summarized in \citet{lee2021}.
 
@@
 defining the projection of the data, and a variable id
-(\(m \in \{1, ..., p\}\)) specifying which coefficient will be
-\textcolor{blue}{changed: the $m^{th}$ row of $A_{p\times d}$ which we denote by $V_m$.}
-A method to update the values of the component of the controlled
-variable \(V_m\) is then needed.
+(\(m \in \{1, ..., p\}\)) specifying which coefficient will be changed:
+the \(m^{th}\) row of \(A_{p\times d}\) which we denote by \(V_m\). A
+method to update the values of the component of the controlled variable
+\(V_m\) is then needed.
 
@@
   \item
-    Normalise \(\boldsymbol{a}^*_1\) and \(\boldsymbol{a}^*_2\).
+    Normalise \(\boldsymbol{a}^*_1\).
   \item
     \(\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1\).
+  \item
+    Normalise \(\boldsymbol{a}^*_2\).
   \end{enumerate}
 \end{enumerate}
-
-\textcolor{red}{a2 should be normalised after the second step? Fig 1: this is now showing tracking, with the yellow dot moving in each image?}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where the contribution of one variable is controlled (dark) is changed using unconstrained orthonormalization.
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where the contribution of one variable is controlled (dark).
 The dot indicates the chosen values for the controlled variable, $V_m$.
+The change is tracking the chosen value after unconstrained orthonormalization.
 It can be seen that the actual axis does not precisely match the chosen position, but it is close.}\label{fig:manualsequence}
@@
 able to track the mouse position and adjust the projection based on the
-\textcolor{blue}{user's cursor position}. The user can select one axis
-(corresponding to one variable) by clicking on it, and then adjust the
-position of that axis by dragging it. In practice, the closest axis will
-be selected, and the dragging results in tracking the mouse position
-which will update the projection in small steps, such that no further
-interpolation is required.
+user's cursor position. The user can select one axis (corresponding to
+one variable) by clicking on it, and then adjust the position of that
+axis by dragging it. In practice, the closest axis will be selected, and
+the dragging results in tracking the mouse position which will update
+the projection in small steps, such that no further interpolation is
+required.
 
@@
 current projection matrix, which will then be orthonormalized and
-\textcolor{blue}{re-displayed}. A second, linked display shows the
-projected data in sync with the updates from the axis display.
+re-displayed. A second, linked display shows the projected data in sync
+with the updates from the axis display.
 
@@
 
-\textcolor{blue}{Within R, the natural implementation for an interactive interface would use}
-\texttt{shiny} \citep{shinym} with \texttt{plotly} \citep{plotly}
+Within R, the natural implementation for an interactive interface would
+use \texttt{shiny} \citep{shinym} with \texttt{plotly} \citep{plotly}
 drawing. However, the \texttt{plotly} package is designed to add to
@@
 
-\textcolor{blue}{The Mathematica implementation relies on the inbuilt function  `LocatorPane`.
-This function creates a region on the screen where the position of the cursor is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.
-We then use the graphics function `ListPlot` embedded in dynamic objects via  `DynamicModule` to produce the display.
-User interface objects, such as sliders, locator panes, and input fields, are used to provide user control of slice thickness, center, scale, and point size.}
+The Mathematica implementation relies on the inbuilt function
+\texttt{LocatorPane}. This function creates a region on the screen where
+the position of the cursor is captured and then converted to input that
+updates the graphics functions, enabling the manual navigation of the
+tour. We then use the graphics function \texttt{ListPlot} embedded in
+dynamic objects via \texttt{DynamicModule} to produce the display. User
+interface objects, such as sliders, locator panes, and input fields, are
+used to provide user control of slice thickness, center, scale, and
+point size.
 
@@
 of the slice via the locator pane, which changes the projection matrix;
-a slider controls the slice thickness; and there is an input field,
-which changes the slice center.
-\textcolor{red}{insert sentence about new slice position display here.}
-The user can also change the appearance of the plot by zooming into the
-center or changing the point size with the sliders provided. It is worth
-noting that this zoom works best with data that has been centered and
-scaled. The projected data can be displayed to contrast it with the
-slice by ticking a box. The explicit projection matrix can also be
-displayed via another checkbox. This is especially useful when we wish
-to import a projection that was identified as interesting in the manual
-exploration into a later study, as shown in the applications below. More
-details about the implementation and usage instructions are given in the
-Appendix.
+a slider controls the slice thickness; and there is an input field or an
+interactive star glyph, to change the slice center. The user can also
+change the appearance of the plot by zooming into the center or changing
+the point size with the sliders provided. It is worth noting that this
+zoom works best with data that has been centered and scaled. The
+projected data can be displayed to contrast it with the slice by ticking
+a box. The explicit projection matrix can also be displayed via another
+checkbox. This is especially useful when we wish to import a projection
+that was identified as interesting in the manual exploration into a
+later study, as shown in the applications below. More details about the
+implementation and usage instructions are given in the Appendix.
 
@@
 shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
-values of bill depth in the top row of Figure \ref{slice1p}.
-\textcolor{blue}{The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables.
-This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.}
-We can see that all three slices (the two models and the data) contain
-almost no points from the third class (Gentoo) and that the decision
-boundary between the two models is very similar.
+values of bill depth in the top row of Figure \ref{slice1p}. The star
+glyph at the right indicates the slice center value, which shows that it
+is higher on \texttt{bd} and at the middle value for the other
+variables. This slice center guide is as proposed in \citet{condviz2},
+and more details can be found in the Appendix. We can see that all three
+slices (the two models and the data) contain almost no points from the
+third class (Gentoo) and that the decision boundary between the two
+models is very similar.
 
@@
 slice localized towards low values of bill depth, shown in the bottom
-row of Figure \ref{slice1p}.
-\textcolor{blue}{(The star glyph at right indicates the slice center value, which shows that it is lower on `bd` and at the middle value for the other variables.)}
-The RF model (left) predicts all three species within this slice, with
-an interesting boundary for the third class (Gentoo). On the other hand,
-the LDA model (middle) predominantly predicts the third class within the
-slice, this appears to be enforced through the linear structure of the
-model. Looking finally at the thick slice through the data we see that
-there are primarily observations from this class within the slice we can
-conclude that the two models have filled in the ``empty'' space (where
-we do not have any training observations) in very different ways and
-according to what we might expect given the model structure.
+row of Figure \ref{slice1p}. (The star glyph at right indicates the
+slice center value, which shows that it is lower on \texttt{bd} and at
+the middle value for the other variables.) The RF model (left) predicts
+all three species within this slice, with an interesting boundary for
+the third class (Gentoo). On the other hand, the LDA model (middle)
+predominantly predicts the third class within the slice, this appears to
+be enforced through the linear structure of the model. Looking finally
+at the thick slice through the data we see that there are primarily
+observations from this class within the slice we can conclude that the
+two models have filled in the ``empty'' space (where we do not have any
+training observations) in very different ways and according to what we
+might expect given the model structure.
 
@@
 Mathematica provided a useful sandbox to experiment with the ideas
-presented in this paper. Most of the data, though, is first constructed
-using R, especially the classification boundaries.
-\textcolor{red}{what should this say?} If new technology for interactive
-graphics with precise cursor control becomes available in R it would be
-useful to create a tighter coupling of models and visualization to allow
-exploring and comparing fits.
+presented in this paper. However, we have used R for the statistical
+data analysis and modeling in our example, as is the standard approach
+in the community. Since the proposed visualization should ideally be
+integrated within this workflow, the new methods should also be
+implemented as interactive graphics accessible from R once new
+technology with precise cursor control becomes available.
 

### 2023-02-08 paper/paper.tex
Small fix in the paper

@@
 devices. The cursor position control is possible for 3D plots with
-\texttt{rgl} \citep{rgl} and \texttt{plotly} but these not the arbitrary
+\texttt{rgl} \citep{rgl} and \texttt{plotly} but not for the arbitrary
 dimension needed for tours. The level of interactivity would have been
@@
 \item
-  Two Mathematica notebooks with the code used in the application
-  (corresponding to the three subsections).
+  Two Mathematica notebooks with the code used in the application.
 \item
